### 2020-04-07 abm tutorial.Rmd
added all updated files

---
title: "Simulation Models of Cultural Evolution"
author: "Alex Mesoudi"
---

# Model 1: Unbiased transmission

Here we will simulate a case where $N$ individuals each possess one of two cultural traits, denoted $A$ and $B$. Each generation, the $N$ agents are replaced with $N$ new agents. Each new agent picks a member of the previous generation at random and copies their cultural trait. This is known as unbiased oblique cultural transmission: unbiased because traits are copied entirely at random, and oblique because one generation learns from the previous non-overlapping generation (as opposed to horizontal cultural transmission where individuals copy members of the same generation, and vertical cultural transmission where offspring copy their parents - our model is way too simple to have actual parents and offspring though).

We are interested in tracking the proportion of individuals who possess trait $A$ over successive generations. We will call this proportion $p$. We could also track the proportion who possess trait $B$, but this will always be $1 - p$ given that everyone who does not possess $A$ must possess $B$, so there is really no need. For example, if 70% of the population have trait $A$, then $p = 0.7$. The remaining 30% must have trait $B$, which is a proportion $1 - p = 1 - 0.7 = 0.3$.

The output of the model will be a plot showing $p$ over all generations up to the last generation, which we will call $t_{max}$. Generations (aka timesteps) are denoted by $t$, where generation one is $t = 1$, generation two is $t = 2$, up to the last generation $t = t_{max}$. These could correspond to biological generations, but could equally be 'cultural generations' (or learning episodes), which might be much shorter.

First we need to specify the fixed parameters of the model. These are $N$ (the number of individuals) and $t_{max}$ (the number of generations). Let's start with $N = 100$ and $t_{max} = 200$.

```{r}
N <- 100
t_max <- 200
```

Run this code snippet using the green 'play' triangle in the top right of the snippet. If you have the Environment pane visible on the right, you should be able to see $N$ and $t_{max}$ appear, with their assigned values. The Environment pane is useful for keeping track of active variables and their current values.

Now we need to create our agents. These will be stored in a dataframe, a commonly-used data format in R. The only information we need to keep about our agents is their cultural trait ($A$ or $B$). Hence we need a dataframe that is $N$ rows long, with a single column for the trait. We'll call this dataframe *agent*. Initially, we will give each agent either an $A$ or $B$ at random, using the **sample** command.

```{r}
agent <- data.frame(trait = sample(c("A","B"), N, replace = TRUE), stringsAsFactors = FALSE)
```

Here, *trait* specifies the name of the sole variable/column in the *agent* dataframe. This is filled using **sample**. The first part of the **sample** command lists the elements to pick at random, in our case, the traits $A$ and $B$. The second part gives the number of times to pick, in our case $N$ times, once for each agent. The final part says to replace or reuse the elements after they've been picked (otherwise there would only be one copy of $A$ and one copy of $B$, so we could only give two agents traits before running out). The argument about stringsAsFactors tells R that the *trait* variable is not a factor. A factor is a categorical variable with set levels. This is not a problem now, but can cause unexpected problems later on. If in doubt, avoid factors.

We can see the first few lines of *agent* using the **head** command, to check it worked:

```{r}
head(agent)
```

As expected, there is a single column called *trait* containing $A$s and $B$s. Note that **head** only shows the first 6 rows for brevity; the other 94 are not shown. The numbers on the left are automatically generated in a dataframe as row labels, but in our case we can think of them as the agents' id numbers (agent 1, agent 2,..., agent 100).

A specific agents' trait can be retrieved using standard dataframe indexing. For example, agent 4's trait can be retrieved using:

```{r}
agent$trait[4]
```

This should match the fourth row in the **head** output above.

We also need a dataframe to track the trait frequency $p$ in each generation. This will have a single column with $t_{max}$ rows, one for each generation. We'll call this dataframe *output*, because it is the output of the model. At this stage we don't know what $p$ will be in each generation, so for now let's fill the *output* dataframe with lots of NAs, which is R's symbol for Not Available, or missing value. We can use the **rep** (repeat) command to repeat NA $t_{max}$ times.

```{r}
output <- data.frame(p = rep(NA, t_max))
```

Here, $p$ is the name of the sole variable/column in the *output* dataframe, and it's filled entirely with NAs. We're using NA rather than, say, zero, because zero could be misinterpreted as $p = 0$, which would mean that all agents have trait $B$. This would be misleading, because at the moment we haven't yet calculated $p$, so it's nonexistent, rather than zero.

We can, however, fill in the first value of $p$ for our already-created first generation of agents, held in *agent*. The command below sums the number of $A$s in *agent* and divides by $N$ to get a proportion out of 1 rather than an absolute number. It then puts this proportion in the first slot of $p$ in *output*, the one for the first generation, $t = 1$. We can again use **head** to check it worked.

```{r}
output$p[1] <- sum(agent$trait == "A") / N
head(output)
```

This first $p$ value should be approximately 0.5: maybe not exactly, because we have a finite and relatively small population size. Analogously, flipping a coin 100 times will not always give exactly 50 heads and 50 tails. Sometimes it would give 51 heads, sometimes 52 heads, sometimes 48. Similarly, sometimes we will have 51 $A$s, sometimes 48, etc.

Now we need to iterate our population over $t_{max}$ generations. In each generation, we need to:

* copy the current agents to a separate dataframe called *previous_agent* to use as demonstrators for the new agents; this allows us to implement oblique transmission with its non-overlapping generations, rather than mixing up the generations and getting in a muddle

* create a new generation of agents, each of whose trait is picked at random from the *previous_agent* dataframe

* calculate $p$ for this new generation and store it in the appropriate slot in *output*

To iterate, we'll use a for-loop, using $t$ to track the generation. We've already done generation 1 so we'll start at generation 2. The random picking of models is done with **sample** again, but this time picking from the traits held in *previous_agent*. Note that I've added comments briefly explaining what each line does. This is perhaps superfluous in a tutorial like this, but it's always good practice. Code often gets cut-and-pasted into other places and loses its context. Explaining what each line does lets other people - and a future, forgetful you - know what's going on.

```{r}
for (t in 2:t_max) {

  previous_agent <- agent  # copy agent dataframe to previous_agent dataframe

  agent <- data.frame(trait = sample(previous_agent$trait, N, replace = TRUE), stringsAsFactors = FALSE)  # randomly copy from previous generation's agents

  output$p[t] <- sum(agent$trait == "A") / N  # get p and put it into the output slot for this generation t

}
```

Now we should have 200 values of $p$ stored in *output*, one for each generation. Let's plot them.

```{r}
plot(output$p, type = 'l', ylab = "p, proportion of agents with trait A", xlab = "generation", ylim = c(0,1), main = paste("N =", N))
```

Note the title of the graph, which gives the value of $N$ for this simulation. It's easy to create plots and forget what the parameters were. It's therefore a good idea to include it somewhere on the graph. There's no need to include our other parameter, $t_{max}$, because that can be seen from the x-axis scale.

Unbiased transmission, or random copying, is by definition random, so different runs of this simulation will generate different plots. If you rerun all the code you'll get something different again. It probably starts off hovering around 0.5 (the approximate starting value of $p$), and might go to 0 or 1 at some point. At $p = 0$ there are no $A$s and every agent possesses $B$. At $p = 1$ there are no $B$s and every agent possesses $A$. This is a typical feature of cultural drift, analogous to genetic drift: in small populations, with no selection or other directional processes operating, traits can be lost purely by chance.

Ideally we would like to repeat the simulation to explore this idea in more detail, perhaps changing some of the parameters. For example, if we increase $N$, are we more or less likely to lose one of the traits? With our code scattered about in chunks, it is hard to quickly repeat the simulation. Instead we can wrap it all up in a function, like so:

```{r}
UnbiasedTransmission <- function (N, t_max) {

  agent <- data.frame(trait = sample(c("A","B"), N, replace = TRUE), stringsAsFactors = FALSE)

  output <- data.frame(p = rep(NA, t_max))

  output$p[1] <- sum(agent$trait == "A") / N

  for (t in 2:t_max) {

    previous_agent <- agent  # copy agent to previous_agent dataframe

    agent <- data.frame(trait = sample(previous_agent$trait, N, replace = TRUE))  # randomly copy from previous generation

    output$p[t] <- sum(agent$trait == "A") / N  # get p and put it into output slot for this generation t

  }

  plot(output$p, type = 'l', ylab = "p, proportion of agents with trait A", xlab = "generation", ylim = c(0,1), main = paste("N =", N))

}
```

This is just all of the code snippets that we already ran above, but all within a function with parameters $N$ and $t_{max}$ as arguments to the function. Nothing will happen when you run the above code, because all you've done is define the function, not actually run it. The point is that we can now call the function in one go, easily changing the values of $N$ and $t_{max}$. Let's try first with the same values of $N$ and $t_{max}$ as before, to check it works. You can re-run the code below several times, to see different dynamics. It should be different every time.

```{r}
UnbiasedTransmission(N = 100, t_max = 200)
```

Now let's try changing the parameters. The following code re-runs the simulation with a much larger $N$.

```{r}
UnbiasedTransmission(N = 10000, t_max = 200)
```

You should see much less fluctuation. Rarely in a population of $N = 10000$ will either trait go to fixation.

Wrapping a simulation in a function like this is good because we can easily re-run it with just a single command. However, it's a bit laborious to manually re-run it. Say we wanted to re-run the simulation 10 times with the same parameter values to see how many times $A$ goes to fixation, and how many times $B$ goes to fixation. Currently, we'd have to manually run the **UnbiasedTransmission** function 10 times and record somewhere else what happened in each run. It would be better to automatically re-run the simulation several times and plot each run as a separate line on the same plot. We could also add a line showing the mean value of $p$ across all runs.

Let's use a new parameter $r_{max}$ to specify the number of independent runs, and use another for-loop to cycle over the $r_{max}$ runs. We also need to expand the number of columns in the *output* dataframe, because we need one column of $p$ values for each of the runs. Let's rewrite the **UnbiasedTransmission** function to handle multiple runs.

```{r}
UnbiasedTransmission <- function (N, t_max, r_max) {

  output <- as.data.frame(matrix(NA, t_max, r_max))  # create a matrix with t_max rows and r_max columns, filled with NAs, then convert to data.frame

  names(output) <- paste("run", 1:r_max, sep="")  # purely cosmetic: rename the columns with run1, run2 etc.

  for (r in 1:r_max) {

    agent <- data.frame(trait = sample(c("A","B"), N, replace = TRUE), stringsAsFactors = FALSE)  # create first generation

    output[1,r] <- sum(agent$trait == "A") / N  # add first generation's p to first row of column r

    for (t in 2:t_max) {

      previous_agent <- agent  # copy agent to previous_agent dataframe

      agent <- data.frame(trait = sample(previous_agent$trait, N, replace = TRUE))  # randomly copy from previous generation

      output[t,r] <- sum(agent$trait == "A") / N  # get p and put it into output slot for this generation t and run r

    }

  }

  # first plot a thick line for the mean p
  plot(rowMeans(output), type = 'l', ylab = "p, proportion of agents with trait A", xlab = "generation", ylim = c(0,1), lwd = 3, main = paste("N =", N))

  for (r in 1:r_max) {

    lines(output[,r], type = 'l')  # add lines for each run, up to r_max

  }

  output  # export data from function
}
```

There are a few changes here. First, we created *output* initially as a matrix, and then immediately converted it into a dataframe. This seems weird, but is just because it is easier to create a multi-row data structure using the **matrix** command than the **dataframe** command. The next command is purely cosmetic, and re-names the columns of *outcome* as run1, run2 etc.

Then we set up our $r$ loop, which executes once for each run. The code is mostly the same as before, except that we now use the "[row,column]" notation to put $p$ into the right place in *output*. The row is $t$ as before, and now the column is $r$. The **plot** command is also changed to handle multiple runs: first we plot the mean as a thick line, then we add one plotted line for each run.

Finally, the **UnbiasedTransmission** function now ends with the *output* dataframe. This means that this dataframe will be exported from the function when it is run. This can be useful for storing data from simulations wrapped in functions, otherwise that data is lost after the function is executed. In the function call below, the raw data from the simulation is put into a dataframe called *data_model1*, as a record of what happened. Run it now to display the new plot.

```{r}
data_model1 <- UnbiasedTransmission(N = 100, t_max = 200, r_max = 5)
```

You should be able to see five independent runs of our simulation shown as regular thin lines, along with a thicker line showing the mean of these lines. Some runs have probably gone to 0 or 1, and the mean should be somewhere in between. The data is stored in *data_model1*, which we can inspect with **head**:

```{r}
head(data_model1)
```

Now let's run the updated **UnbiasedTransmission** model with $N = 10000$, to compare with $N = 100$.

```{r}
data_model1 <- UnbiasedTransmission(N = 10000, t_max = 200, r_max = 5)
```

The mean line should be almost exactly at $p = 0.5$ now, with the five independent runs fairly close to it.

Let's add one final modification. So far the starting frequencies of $A$ and $B$ have been the same, roughly 0.5 each. But what if we were to start at different initial frequencies of $A$ and $B$? Say, $p = 0.2$ or $p = 0.9$? Would unbiased transmission keep $p$ at these initial values, or would it go to $p = 0.5$ as we have found so far?

To find out, we can add another parameter, $p_0$, which specifies the initial probability of drawing an $A$ rather than a $B$ in the first generation. Previously this was always $p_0 = 0.5$, but in the new function below we add it to the **sample** function to weight the initial allocation of traits in $t = 1$.

```{r}
UnbiasedTransmission <- function (N, p_0, t_max, r_max) {

  output <- as.data.frame(matrix(NA,t_max,r_max))  # create a matrix with t_max rows and r_max columns, filled with NAs, then convert to data.frame

  names(output) <- paste("run", 1:r_max, sep="")  # purely cosmetic: rename the columns with run1, run2 etc.

  for (r in 1:r_max) {

    agent <- data.frame(trait = sample(c("A","B"), N, replace = TRUE, prob = c(p_0,1-p_0)), stringsAsFactors = FALSE)  # create first generation

    output[1,r] <- sum(agent$trait == "A") / N  # add first generation's p to first row of column r

    for (t in 2:t_max) {

      previous_agent <- agent  # copy agent to previous_agent dataframe

      agent <- data.frame(trait = sample(previous_agent$trait, N, replace = TRUE))  # randomly copy from previous generation

      output[t,r] <- sum(agent$trait == "A") / N  # get p and put it into output slot for this generation t and run r

    }

  }

  # first plot a thick line for the mean p
  plot(rowMeans(output), type = 'l', ylab = "p, proportion of agents with trait A", xlab = "generation", ylim = c(0,1), lwd = 3, main = paste("N =", N))

  for (r in 1:r_max) {

    lines(output[,r], type = 'l')  # add lines for each run, up to r_max

  }

  output  # export data from function
}
```

The only change here is the addition of $p_0$ as a definable parameter for the function, and the prob argument in the **sample** command. The prob argument gives the probability of picking each option, in our case $A$ and $B$, in the first generation. The probability of $A$ is now $p_0$, and the probability of $B$ is now $1 - p_0$. Let's see what happens with a different value of $p_0$.

```{r}
data_model1 <- UnbiasedTransmission(N = 10000, p_0 = 0.2, t_max = 200, r_max = 5)
```

With $p_0 = 0.2$, trait frequencies stay at $p = 0.2$. Unbiased transmission is truly non-directional: it maintains trait frequencies at whatever they were in the previous generation, barring random fluctuations caused by small population sizes.

***

### Summary of Model 1

Even this extremely simple model provides some valuable insights. First, unbiased transmission does not in itself change trait frequencies. As long as populations are large, trait frequencies remain the same.

Second, the smaller the population size, the more likely traits are to be lost by chance. This is a basic insight from population genetics, known there as genetic drift, but it can also be applied to cultural evolution. Many studies have tested (and some supported) the idea that population size and other demographic factors can shape cultural diversity.

Furthermore, generating expectations about cultural change under simple assumptions like random cultural drift can be useful for detecting non-random patterns like selection: if we don't have a baseline, we won't know selection or other directional processes when we see them.

In Model 1 we have introduced several programming techniques that will be useful in later models. We've seen how to use dataframes to hold characteristics of agents, how to use loops to cycle through generations and simulation runs, how to use **sample** to pick randomly from sets of elements, how to wrap simulations in functions to easily re-run them with different parameter values, how to plot the results of simulations, and how to store the output of the simulation in a dataframe that persists after the simulation function is run.

***

## Analytical Appendix

If $p$ is the frequency of $A$ in one generation, we are interested in calculating $p'$, the frequency of $A$ in the next generation under the assumption of unbiased transmission. Each new individual in the next generation picks a demonstrator at random from amongst the previous generation. The demonstrator will have $A$ with probability $p$. The frequency of $A$ in the next generation, then, is simply the frequency of $A$ in the previous generation:

$$p' = p                \hspace{30 mm}(1.1)$$

Equation 1.1 simply says that under unbiased transmission there is no change in $p$ over time. If, as we assumed above, the initial value of $p$ in a particular population is $p_0$, then the equilibrium value of $p$, $p^*$, at which there is no change in $p$ over time, is just $p_0$. 
We can plot this recursion, to recreate the final simulation plot above:
```{r}
p_0 <- 0.2
t_max <- 200
p <- rep(NA, t_max)
p[1] <- p_0
for (i in 2:t_max) {
  p[i] <- p[i-1]
}
plot(p, type = 'l', ylab = "p, proportion of agents with trait A", xlab = "generation", ylim = c(0,1), lwd = 3)
```
Don't worry, it gets more complicated than this in later chapters. The key point here is that analytical (or deterministic) models assume infinite populations (note there is no $N$ in the above recursion) and no stochasticity. Simulations with very large populations should give the same results as analytical models. Basically, the closer we can get in stochastic models to the assumption of infinite populations, the closer the match to infinite-population deterministic models. Deterministic models give the ideal case; stochastic models permit more realistic dynamics based on finite populations.
More generally, creating deterministic recursion-based models can be a good way of verifying simulation models, and vice versa: if the same dynamics occur in both agent-based and recursion-based models, then we can be more confident that those dynamics are genuine and not the result of a programming error or mathematical mistake. 
***

### 2020-04-07 Trial abm.R

@@ -1,0 +1,35 @@
+abmtrial <- function (N, t_max,resourceOpen,resourceProtected) {
+  
+  agent<-data.frame(behavior = rep("C",N),stringsAsFactors = F)
+  
+  
+  
+  output <- data.frame(p = rep(NA, t_max))
+  
+  output$p[1] <- sum(agent$trait == "C") / N
+  
+  for (t in 2:t_max) {
+    
+    previous_agent <- agent  # copy agent to previous_agent dataframe
+    
+    prev_resourceOpen <- resourceOpen
+    prev_resourceProtected <-resourceProtected
+    
+    resourceOpen <- (resourceOpen - nrow(agent$behavior == "C"))
+    resourceProtected <- (resourceProtected - nrow(agent$behavior == "D"))
+    
+    probCoop <- 0.5
+    
+    agent <- data.frame(behavior = sample(c("C","D"), N, replace = TRUE,
+                                          prob=c(probCoop,1-probCoop)),stringsAsFactors = "F")  # randomly copy from previous generation
+    
+    output$p[t] <- sum(agent$behavior == "C") / N  # get p and put it into output slot for this generation t
+    
+  }
+  
+ # plot(output$p, type = 'l', ylab = "p, proportion of agents cooperating", xlab = "generation", ylim = c(0,1), main = paste("N =", N))
+  plot(output$resourceOpen, type = 'l', ylab = "Amount of common pool resource", xlab = "generation", main = paste("N =", N))
+  
+}
+
+abmtrial(55,25,200,20)

### 2020-04-07 club goods model - 14.11.19.R

@@ -350,7 +350,7 @@ mutation_toggle				<- TRUE		# toggles strategy mutation on/off
 main_graphic_toggle 		<- TRUE		# toggles main graphical output on/off
 stock_graphic_toggle		<- TRUE		# toggles resource stock lines on/off	
 preference_graphic_toggle	<- TRUE		# toggles harvest pref lines on/off
-restriction_graphic_toggle	<- FALSE		# toggles restriction votes lines on/off
+restriction_graphic_toggle	<- TRUE		# toggles restriction votes lines on/off
 stint_graphic_toggle		<- TRUE		# toggles implementation of restriction 
 										# lines on/off
 harvest_graphic_toggle		<- TRUE		# toggles total harvest lines on/off
@@ -377,7 +377,7 @@ sqrt_num_group = 5
 
 num_group = sqrt_num_group ^ 2			# number of groups. 
 
-num_year = 100							# simulation length
+num_year = 500						# simulation length
 
 
 r = 0.9                              	# rate of resource reproduction in a 

### 2020-04-07 club goods model - Modified_MC.R

@@ -1,0 +1,1547 @@
+# 	Model of club goods with positive externalities
+#	cMLS Fishery Model
+ 
+# Karthik Panchanathan and Tim Waring
+# 09.04.14
+ 
+#-----------------------> 
+# NOTES
+
+# Things to consider for later
+
+	# 	Capture figures (maybe 2 per run - one for groups, one for the pop)
+	#	for each run through the simulation (i.e., each replicate and 
+	#	combination of parameter values). Display the parameter values on
+	#	the graphic, so people can easily see what it being modeled. Refer
+	#	to the different parameters with letters and subscripts. This can
+	#	be something that can be toggled on/off for publication. The title
+	#	of the figure should be the row number from the data frame which
+	#	houses parameter value combinations.
+	
+	#	I need to understand the literature on bio-economics. Maybe start 
+	#	with Clark's book. We need to think about how our cultural and
+	#	group-structured approach adds to the current literature on
+	#	fisheries. What is the current state-of-the-art? What, if anything,
+	#	do we gain by thinking about multi-level selection and/or cultural
+	#	evolution.
+
+	#	Consider changing our voting rules. If a vote is passed, maybe the 
+	#	restriction lasts for X years, rather than re-voting each year.
+	
+	# 	Vary group size. N=10, 20, 50, 100. With bigger groups, it will be
+	#	much harder to drift to an institution.
+	
+	#	Change the parameter input structure. Have this drawn either from
+	#	a CSV file or build the matrix right inside the simulation. 
+	#	Either way, we can eliminate the bulky loop structure (i.e., one
+	#	loop variable for each loop parameter. Instead, we construct a
+	#	data frame housing all the parameter combinations. Then we just
+	#	loop through the rows of this data frame -- one loop.
+	
+	#	Clean up and modularize the code. 
+		
+		#	Can we separate functions into separate files? Going along 
+		#	with this, we might put in different options for functions.
+		#	For example, for "reproduction", we might allow for (1) an
+		#	Allee effect, (2) logistic growth, (3) a Ricker model.
+		
+		#	Have a call to the function that gathers the parameter values.
+		
+		#	The "main program" file should be really minimal. The user
+		#	shouldn't be overwhelmed.
+		
+		#	Put notes and version history in separate files.
+		
+		#	With all of this, how do you store old files? Maybe no date
+		#	stamp on the most recent file. When a new file is created, put
+		#	today's date on the last file?
+
[…]

### 2020-04-15 Ashley Lagrou VIP timeline.Rmd
misc stuff

---
title: "Ashley Lagrou VIP Plan"
author: "Matt Clark"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

```

Schedule
======

| Week| Focus |Deliverable
:---:|:---:|:---
4/6/2020 | Protocol | Finished stove efficiency protocol
4/13/2020 | Data Simulation| Simulated dataset
4/20/2020 | Plotting | Social Connectivity vs stove outcome plots
4/27/2020 | Write-up / Catch-up| Brief write-up of results
5/4/2020 | Lightening talk | Lightening talk

Weekly Specifics
======

4/6/2020 - Protocol
------

What you have now is a really great start. It would be useful however to organize the document into a more readable format and provide some more detail. i would suggest that you go back and read the methods of the McCracken and Smith paper that I sent you awhile back. On pages 740 - 741 they provide a level of methodological detail that would be good for this protocol. I also think somewhere in there they conclude that the water boiling test alone is virtually useless (maybe I'm wrong), so just double check that.

4/13/2020 - Data Simulation
------

Often in science, we simulate the data that we *think* we will collect in the field before we actually do it. The reason for doing this is that you can build in the effects that you expect to see and make sure that your statistics can actually detect them. For a complicated, messy project like this one, data simulation is especially important. You'll be taking the first crack at the simulation for this project.

There are a few key functions that I rely on really heavily for data simulation.
Here's a primer on each of them.

# Sample
```{r}
x<-1:10
sample(x,3,replace=TRUE) #randomly selects a specified number of observations out of a vector
x<-c("John","Jacob","Jackson","Jeffrey")
sample(x,3,replace=TRUE) 
```

# Rep
```{r}
rep(2,10) #replecated the first argument, the number of times of the second

rep("I Love Learning R!",10)
```

# pulling from a distribution (rpois, rnorm, rbeta, rbinom, runif, rgamma)
```{r}
rpois(10,5) #pull 10 samples from a poisson distribution with a mean of 5. Poisson gives positive integers

hist(rpois(1000,5))
```
rnorm, rbeta, rbinom, runif, rgamma sample data randomly from a normal, beta, binomial, uniform, and gamma distributions respectively. Your choice of distribution will matter greatly here. If you don't know what one of them is, you can Google it or ask. They all take slightly different arguments, so make sure to use the ?command in r to see how to use each.

Here's some code to get you started. 
```{r}
#Make a vector of stove Id's
namevec<-vector()
for (i in 1:100){ #paste "ID" onto the numbers 1 through 100
  namevec[i]<-paste0("ID",i)}

#could we make these variables in any better way?
dat<-data.frame(stoveID=namevec, #stove ID comes from the name vector we just made
  copiedFrom=c(rep("Original",10),sample(namevec[1:10],90,replace = T)), #10 original stoves, 90 copied stoves randomly selected
  location = rep(NA,100), #not sure how to fill this quite yet
  VisSimModel = rbeta(100,35,75), #visiual similarity.
  BoilTime = sort(rnorm(100,10,2)),
  WaterVol = runif(100,15,20),
  WoodStart = sort(runif(100,300,600)),
  WoodEnd = sort(runif(100,50,250)),
  MonthBuilt= sample(1:12,100,replace=T),
  YearBuilt= c(rep(2015,5),rep(2017,5),sample(2016:2020,90,replace=T)),
  UsePerDay = sort(rpois(100,2)),
  UseTotalEstimate = rep(NA,100),
  stdPmEsmssions = rnorm(100,0,1),
  GeogDistModel = sort(rgamma(100,75,6)),
  SocialDistModel = c(rep(1,55),rep(2,30),rep(3,15)),
  GeneticDistModel = rep(NA,100),
  GenderBuilder = sample(c(rep("Female",8),rep("Male",2)),100,replace=T),
  GenderModel =sample(c(rep("Female",8),rep("Male",2)),100,replace=T),
  Use = sample(c("Commercial","Family","Both"),100,replace=T))

```

Based on your knowledge of how the fuel efficiency tests will work, I want you to modify this code above to be more reflective of what the completed datasheet may actually look like.

4/20/2020 - Plotting
------

Plotting is an important first step to understanding your data. It should always be your first stop once you have clean data in hand.

We will be using the ggplot2 package in R for all of our plotting
```{r}
library(ggplot2)
```

### 2020-04-15 AdvancedVisualizationMapping.R

@@ -1,0 +1,7 @@
+library(raster)
+x<-getData('worldclim',var = 'prec',res = 0.5, lon=39.74,lat=-5.313)
+plot(x)
+
+
+tzProtect<-read.csv("~/Pemba Project/TanzaniaProtectedAreas.csv")
+tzProtect

### 2020-04-15 ProtectedAreasABM.R

@@ -1,0 +1,56 @@
+
+
+abm<-function(#Specified parameters
+  Individuals=100, #number of total resource users in a population
+  
+  TotalResourceUnit=10000, #total available resource units
+  
+  PercProtected=0.2, #percent of the total resource that's in a protected area
+  
+  CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
+  
+  TimeSteps=200){
+  
+  ## Set parameters
+  PercWorking= 1-PercProtected #percent of resource in a working landscape
+  TotalResourceWorking = PercWorking*TotalResourceUnit #number of resources in the working landscape starting
+  TotalResourceProtected = PercProtected*TotalResourceUnit #number of resources in the protected landscape starting
+  CoopNumStart= CoopPercStart*Individuals #number of individuals cooperating fully at t0
+  
+  DefNumStart<- (1-CoopPercStart)*Individuals #number of individuals defecting at t0
+
+  PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart+1,1,2))
+  PercTimeWorking = (1-PercTimeProtected)
+  
+  agents<-data.frame(PercTimeProtected,
+                     PercTimeWorking,
+                     PayoffProtectedLastTime = rep(NA,Individuals),
+                     PayoffWorkingLastTime= rep(NA,Individuals))
+  
+  for ( i in 1:nrow(agents)){
+    agent2<-agents[i,]
+    agents[i,3]<- rbinom(1,TotalResourceProtected/DefNumStart,agent2$PercTimeProtected)
+    agents[i,4]<-rbinom(1,TotalResourceWorking/Individuals,agent2$PercTimeWorking)}
+  
+  agents$PayoffTotalLastTime<- agents$PayoffProtectedLastTime + agents$PayoffWorkingLastTime
+   
+  
+  
+  #this is currently broken because the protected area followers will always beat out the defecters
+  # Maybe this will be ok? 
+  #might change iver time
+  #think through how the resource extraction effects each pool 
+  
+  return(agents)
+  
+  
+}
+abm()
+
+
+
+
+
+
+
+

### 2020-04-15 RichardEvolutionSocialLearningAbm.R

@@ -1,0 +1,107 @@
+# basic evolution social learning
+# Rogers' model
+# Q: what is the benefit of social learning to a species?
+
+# "Roughgarden method" - after Joan Roughgarden
+# three kinds of assumptions
+# (1) what is the population structure?
+# (2) individual states
+# (3) life cycle
+
+# (1) population structure
+# finite population, well-mixed (everyone can interact with everyone else)
+# population all experiences same environment
+# environment changes periodically
+
+# (2) individual states
+# heritable states: learning strategies - 
+# (a) individual learning: pay a personal cost to figure out what to do - c is cost, s is chance you figure out the right thing
+# (b) social learning: pay no cost! copy a random adult and maybe get a good behavior
+# b is benefit of doing the right thing
+
+# (3) life cycle
+# (a) birth
+# (b) environmental change
+# (c) babies learn
+# (d) all adults die
+
+sim_rogers <- function( 
+  num_generations = 1e3,
+  population_size = 200,
+  c = 1,                 # cost of individual learning
+  s = 1,                 # chance learn right thing
+  b = 2,                 # benefit of right thing
+  w0 = 10,               # baseline fitness
+  u = 0.1,               # prob environment changes
+  mu = 0.001			   # mutation rate
+) 
+{
+  
+  # initialize the population
+  population_strategies <- sample( 1:2 , size=population_size , replace=TRUE )
+  population_behavior <- rep( 0 , population_size )
+  
+  # initialize history
+  history <- matrix( NA , nrow=num_generations , ncol=4 )
+  
+  # loop over generations
+  for ( t in 1:num_generations ) {
+    
+    # record history
+    # freq of individual learners
+    history[t,1] <- sum(population_strategies==1)/population_size
+    # freq of social learners
+    history[t,2] <- sum(population_strategies==2)/population_size
+    # freq of right behavior
+    history[t,3] <- sum(population_behavior==1)/population_size
+    
+    # reproduction / birth
+    # compute fitness of each adult
+    fitness <- w0 + (population_behavior==1)*b - (population_strategies==1)*c
+    # record fitness
+    history[t,4] <- mean(fitness)
+    # produce babies in proportion to fitness
[…]

### 2020-04-15 ProtectedAreasABM.R
First draft of PA abm

@@ -5,11 +5,13 @@ abm<-function(#Specified parameters
   
   TotalResourceUnit=10000, #total available resource units
   
-  PercProtected=0.2, #percent of the total resource that's in a protected area
+  PercProtected=0.3, #percent of the total resource that's in a protected area
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
-  TimeSteps=200){
+  TimeSteps=200,
+  ResourceRegenerationPerTimeStep=1.5,
+  harvestMax=25){
   
   ## Set parameters
   PercWorking= 1-PercProtected #percent of resource in a working landscape
@@ -17,35 +19,136 @@ abm<-function(#Specified parameters
   TotalResourceProtected = PercProtected*TotalResourceUnit #number of resources in the protected landscape starting
   CoopNumStart= CoopPercStart*Individuals #number of individuals cooperating fully at t0
   
-  DefNumStart<- (1-CoopPercStart)*Individuals #number of individuals defecting at t0
-
-  PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart+1,1,2))
-  PercTimeWorking = (1-PercTimeProtected)
+  DefNumStart= (1-CoopPercStart)*Individuals #number of individuals defecting at t0
+  PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart+1,1,2)) #percent of their foraging time each indv spends in the PA
+  PercTimeWorking = (1-PercTimeProtected)#percent of their foraging time each indv spends in the working landscape
   
   agents<-data.frame(PercTimeProtected,
                      PercTimeWorking,
                      PayoffProtectedLastTime = rep(NA,Individuals),
-                     PayoffWorkingLastTime= rep(NA,Individuals))
+                     PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
   
-  for ( i in 1:nrow(agents)){
-    agent2<-agents[i,]
-    agents[i,3]<- rbinom(1,TotalResourceProtected/DefNumStart,agent2$PercTimeProtected)
-    agents[i,4]<-rbinom(1,TotalResourceWorking/Individuals,agent2$PercTimeWorking)}
+  ProtectPerDefect<-as.integer(TotalResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
[…]

### 2020-04-15 ProtectedAreasABM.R
No structural changes. Next steps should be to start model in degreded environment. Starting at carrying capacity for resources is unrealistic.

@@ -9,7 +9,7 @@ abm<-function(#Specified parameters
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
-  TimeSteps=200,
+  TimeSteps=20,
   ResourceRegenerationPerTimeStep=1.5,
   harvestMax=25){
   
@@ -148,7 +148,7 @@ mtext(paste("Individuals=",Individuals,
  return(output) 
 }
 
-abm(harvestMax = 30)
+abm(harvestMax = 40,ResourceRegenerationPerTimeStep = 1.8,TimeSteps = 50)
 
 
 

### 2020-04-15 ABM_Output/app.R
web app

@@ -1,0 +1,223 @@
+
+
+
+library(shiny)
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+
+    # Application title
+    titlePanel("AbmOutput"),
+
+    # Sidebar with a slider input for number of bins     
+    sidebarLayout(
+        sidebarPanel(
+            sliderInput("individuals",
+                        "Number of individuals:",
+                        min = 10,
+                        max = 1000,
+                        value = 100),
+           sliderInput("totResources","Total Resources Max",
+                       min = 100,max=100000,value=10000),
+           sliderInput("timeSteps",
+                       "Time Steps",
+                       min = 10,
+                       max = 1000,
+                       value = 100),
+           sliderInput("percProtect",
+                       "Percent of Resources Protected",
+                       min = 0.01,
+                       max = 0.99,
+                       value = 0.20),
+           sliderInput("startCooperators",
+                       "Starting Obligatory Cooperators",
+                       min = 0.1,
+                       max = 0.95,
+                       value = 0.9),
+           sliderInput("resourceRegen",
+                       "Resource Regeneration Rate",
+                       min = 1.1,
+                       max = 1.9,
+                       value = 1.5),
+           sliderInput("maxHarvest",
+                       "Max individual harvest per time step",
+                       min = 10,
+                       max = 100,
+                       value = 30)
+        ),
+
+        # Show a plot of the generated distribution
+        mainPanel(
+           plotOutput("distPlot")
+        )
+    )
+)
+
+# Define server logic required to draw a histogram
+server <- function(input, output) {
+    
+    abm<-function(#Specified parameters
+        Individuals=100, #number of total resource users in a population
+        
+        TotalResourceUnit=10000, #total available resource units
+        
+        PercProtected=0.3, #percent of the total resource that's in a protected area
+        
+        CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
+        
+        TimeSteps=20,
+        ResourceRegenerationPerTimeStep=1.5,
+        harvestMax=25){
+        
+        ## Set parameters
+        PercWorking= 1-PercProtected #percent of resource in a working landscape
[…]

### 2020-04-15 ABM_Output/app.R
some plotting adjustments

@@ -2,6 +2,7 @@
 
 
 library(shiny)
+library(ggplot2)
 
 # Define UI for application that draws a histogram
 ui <- fluidPage(
@@ -189,11 +190,25 @@ server <- function(input, output) {
             
         }
         
-        par(mfrow=c(3,1))
-        plot(output$percCCProtect, type = 'l',col="green")
-        plot(output$percCCWorking, type = 'l',col="brown")
-        plot(output$meanPayoff, type = 'l',col="red")
-        mtext("Output", side=3, outer=TRUE, line=-3,cex=0.8)
+        
+        p1<-ggplot(data=output,aes(x=timeStep))+
+            geom_line(aes(y=percCCProtect),size=3,color="#b2df8a")+
+            geom_line(aes(y=percCCWorking),size=3,color="#1f78b4")+
+            ylim(0, 1.0)+theme_classic()+ylab("Percent Carrying Capacity")+
+            scale_colour_manual(name = 'Area', 
+                                values =c('#b2df8a'='#b2df8a','#1f78b4'='#1f78b4'), labels = c('Protected Area','Working Landscape'))+
+            ggtitle("Resource Degredation")
+        
+        
+        p2<-ggplot(data=output,aes(x=timeStep))+
+            geom_line(aes(y=meanPayoff),size=3,color="#993404")+
+            theme_classic()+ylab("Resource Units")+
+            ggtitle("Mean Individual Payoff")
+        
+        
+        x<-gridExtra::grid.arrange(p1,p2,ncol=1)
+        
+        return(x) 
         
         
         

### 2020-04-15 ProtectedAreasABM.R

@@ -133,19 +133,36 @@ output$meanPayoff[t]  <- mean(agents$PayoffTotalLastTime)
 
 }
 
-par(mfrow=c(3,1))
-plot(output$percCCProtect, type = 'l',col="green")
-plot(output$percCCWorking, type = 'l',col="brown")
-plot(output$meanPayoff, type = 'l',col="red")
-mtext(paste("Individuals=",Individuals,
-            "  Total Resources=",TotalResourceUnit,
-            "  Percent Protected=",PercProtected,
-            "  Max Harvest=",harvestMax,
-            "  Resource Regeneration=",ResourceRegenerationPerTimeStep,
-            "  Percent Cooperating at start",CoopPercStart ), side=3, outer=TRUE, line=-3,cex=0.8)
+#par(mfrow=c(3,1))
+#plot(output$percCCProtect, type = 'l',col="green")
+#plot(output$percCCWorking, type = 'l',col="brown")
+#plot(output$meanPayoff, type = 'l',col="red")
+#mtext(paste("Individuals=",Individuals,
+ #           "  Total Resources=",TotalResourceUnit,
+  #          "  Percent Protected=",PercProtected,
+   #         "  Max Harvest=",harvestMax,
+    #        "  Resource Regeneration=",ResourceRegenerationPerTimeStep,
+     #       "  Percent Cooperating at start",CoopPercStart ), side=3, outer=TRUE, line=-3,cex=0.8)
+
+library(ggplot2)
+p1<-ggplot(data=output,aes(x=timeStep))+
+  geom_line(aes(y=percCCProtect),size=3,color="#b2df8a")+
+  geom_line(aes(y=percCCWorking),size=3,color="#1f78b4")+
+  ylim(0, 1.0)+theme_classic()+ylab("Percent Carrying Capacity")+
+  scale_colour_manual(name = 'Area', 
+                      values =c('#b2df8a'='#b2df8a','#1f78b4'='#1f78b4'), labels = c('Protected Area','Working Landscape'))+
+  ggtitle("Resource Degredation")
+  
+
+p2<-ggplot(data=output,aes(x=timeStep))+
+  geom_line(aes(y=meanPayoff),size=3,color="#993404")+
+  theme_classic()+ylab("Resource Units")+
+  ggtitle("Mean Individual Payoff")
 
 
- return(output) 
+x<-gridExtra::grid.arrange(p1,p2,ncol=1)
+
[…]

### 2020-04-16 ABM_Output/app.R
fixed the starting defectors bug

@@ -74,10 +74,9 @@ server <- function(input, output) {
         PercWorking= 1-PercProtected #percent of resource in a working landscape
         TotalResourceWorking = PercWorking*TotalResourceUnit #number of resources in the working landscape starting
         TotalResourceProtected = PercProtected*TotalResourceUnit #number of resources in the protected landscape starting
-        CoopNumStart= CoopPercStart*Individuals #number of individuals cooperating fully at t0
-        
-        DefNumStart= (1-CoopPercStart)*Individuals #number of individuals defecting at t0
-        PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart+1,1,2)) #percent of their foraging time each indv spends in the PA
+        CoopNumStart= as.integer(CoopPercStart*Individuals) #number of individuals cooperating fully at t0
+        DefNumStart= Individuals - CoopNumStart#number of individuals defecting at t0
+        PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart,1,2)) #percent of their foraging time each indv spends in the PA
         PercTimeWorking = (1-PercTimeProtected)#percent of their foraging time each indv spends in the working landscape
         
         agents<-data.frame(PercTimeProtected,

### 2020-04-16 ProtectedAreasABM.R

@@ -17,10 +17,9 @@ abm<-function(#Specified parameters
   PercWorking= 1-PercProtected #percent of resource in a working landscape
   TotalResourceWorking = PercWorking*TotalResourceUnit #number of resources in the working landscape starting
   TotalResourceProtected = PercProtected*TotalResourceUnit #number of resources in the protected landscape starting
-  CoopNumStart= CoopPercStart*Individuals #number of individuals cooperating fully at t0
-  
-  DefNumStart= (1-CoopPercStart)*Individuals #number of individuals defecting at t0
-  PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart+1,1,2)) #percent of their foraging time each indv spends in the PA
+  CoopNumStart= as.integer(CoopPercStart*Individuals) #number of individuals cooperating fully at t0
+  DefNumStart= Individuals - CoopNumStart#number of individuals defecting at t0
+  PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart,1,2)) #percent of their foraging time each indv spends in the PA
   PercTimeWorking = (1-PercTimeProtected)#percent of their foraging time each indv spends in the working landscape
   
   agents<-data.frame(PercTimeProtected,
@@ -133,39 +132,23 @@ output$meanPayoff[t]  <- mean(agents$PayoffTotalLastTime)
 
 }
 
-#par(mfrow=c(3,1))
-#plot(output$percCCProtect, type = 'l',col="green")
-#plot(output$percCCWorking, type = 'l',col="brown")
-#plot(output$meanPayoff, type = 'l',col="red")
-#mtext(paste("Individuals=",Individuals,
- #           "  Total Resources=",TotalResourceUnit,
-  #          "  Percent Protected=",PercProtected,
-   #         "  Max Harvest=",harvestMax,
-    #        "  Resource Regeneration=",ResourceRegenerationPerTimeStep,
-     #       "  Percent Cooperating at start",CoopPercStart ), side=3, outer=TRUE, line=-3,cex=0.8)
-
-library(ggplot2)
-p1<-ggplot(data=output,aes(x=timeStep))+
-  geom_line(aes(y=percCCProtect),size=3,color="#b2df8a")+
-  geom_line(aes(y=percCCWorking),size=3,color="#1f78b4")+
[…]

### 2020-04-16 ABM_Output/app.R
added starting degredation variable to abm

@@ -18,8 +18,10 @@ ui <- fluidPage(
                         min = 10,
                         max = 1000,
                         value = 100),
-           sliderInput("totResources","Total Resources Max",
+           sliderInput("totCareCapac","Total Carrying Capacity Landscape",
                        min = 100,max=100000,value=10000),
+           sliderInput("StartPercCarryingCapacity","Starting Percent Carrying Capacity",
+                       min=0.1,max=1.0,value=0.75),
            sliderInput("timeSteps",
                        "Time Steps",
                        min = 10,
@@ -57,10 +59,13 @@ ui <- fluidPage(
 # Define server logic required to draw a histogram
 server <- function(input, output) {
     
+    
     abm<-function(#Specified parameters
         Individuals=100, #number of total resource users in a population
         
-        TotalResourceUnit=10000, #total available resource units
+        TotalCarryingCapacity=10000, #total available resource units
+        
+        StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
         
         PercProtected=0.3, #percent of the total resource that's in a protected area
         
@@ -72,8 +77,8 @@ server <- function(input, output) {
         
         ## Set parameters
         PercWorking= 1-PercProtected #percent of resource in a working landscape
-        TotalResourceWorking = PercWorking*TotalResourceUnit #number of resources in the working landscape starting
-        TotalResourceProtected = PercProtected*TotalResourceUnit #number of resources in the protected landscape starting
+        TotalResourceWorking = PercWorking*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the working landscape starting
+        TotalResourceProtected = PercProtected*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the protected landscape starting
         CoopNumStart= as.integer(CoopPercStart*Individuals) #number of individuals cooperating fully at t0
         DefNumStart= Individuals - CoopNumStart#number of individuals defecting at t0
         PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart,1,2)) #percent of their foraging time each indv spends in the PA
@@ -221,15 +226,18 @@ server <- function(input, output) {
         # generate bins based on input$bins from ui.R
        individuals<-input$individuals
        totRecources<-input$totResources
[…]

### 2020-04-16 ProtectedAreasABM.R

@@ -3,7 +3,9 @@
 abm<-function(#Specified parameters
   Individuals=100, #number of total resource users in a population
   
-  TotalResourceUnit=10000, #total available resource units
+  TotalCarryingCapacity=10000, #total available resource units
+  
+  StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
   
   PercProtected=0.3, #percent of the total resource that's in a protected area
   
@@ -15,8 +17,8 @@ abm<-function(#Specified parameters
   
   ## Set parameters
   PercWorking= 1-PercProtected #percent of resource in a working landscape
-  TotalResourceWorking = PercWorking*TotalResourceUnit #number of resources in the working landscape starting
-  TotalResourceProtected = PercProtected*TotalResourceUnit #number of resources in the protected landscape starting
+  TotalResourceWorking = PercWorking*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the working landscape starting
+  TotalResourceProtected = PercProtected*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the protected landscape starting
   CoopNumStart= as.integer(CoopPercStart*Individuals) #number of individuals cooperating fully at t0
   DefNumStart= Individuals - CoopNumStart#number of individuals defecting at t0
   PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart,1,2)) #percent of their foraging time each indv spends in the PA
@@ -137,7 +139,7 @@ plot(output$percCCProtect, type = 'l',col="green")
 plot(output$percCCWorking, type = 'l',col="brown")
 plot(output$meanPayoff, type = 'l',col="red")
 mtext(paste("Individuals=",Individuals,
-            "  Total Resources=",TotalResourceUnit,
+            "  Total Resources=",TotalCarryingCapacity,
             "  Percent Protected=",PercProtected,
           "  Max Harvest=",harvestMax,
            "  Resource Regeneration=",ResourceRegenerationPerTimeStep,
@@ -148,7 +150,7 @@ mtext(paste("Individuals=",Individuals,
  return(output) 
 }
 
-abm(harvestMax = 40,ResourceRegenerationPerTimeStep = 1.5,TimeSteps = 50,CoopPercStart = 0.4)
+abm(harvestMax = 40,ResourceRegenerationPerTimeStep = 1.5,TimeSteps = 50,CoopPercStart = 0.4,StartPercCarryingCapacity = 0.5)
 
 
 

### 2020-04-21 ProtectedAreasABM.R
fixed starting % CC bug

@@ -17,8 +17,10 @@ abm<-function(#Specified parameters
   
   ## Set parameters
   PercWorking= 1-PercProtected #percent of resource in a working landscape
-  TotalResourceWorking = PercWorking*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the working landscape starting
-  TotalResourceProtected = PercProtected*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the protected landscape starting
+  TotalCCResourceProtected=PercProtected*TotalCarryingCapacity #total resource units in carrying capacity of protected area
+  TotalCCResourceWorking=PercWorking*TotalCarryingCapacity #total resource units in carrying capacity of working area
+  StartResourceWorking = TotalCCResourceWorking*StartPercCarryingCapacity #number of resources in the working landscape starting
+  StartResourceProtected = TotalCCResourceProtected*StartPercCarryingCapacity #number of resources in the protected landscape starting
   CoopNumStart= as.integer(CoopPercStart*Individuals) #number of individuals cooperating fully at t0
   DefNumStart= Individuals - CoopNumStart#number of individuals defecting at t0
   PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart,1,2)) #percent of their foraging time each indv spends in the PA
@@ -29,8 +31,8 @@ abm<-function(#Specified parameters
                      PayoffProtectedLastTime = rep(NA,Individuals),
                      PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
   
-  ProtectPerDefect<-as.integer(TotalResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
-  WorkingPerTotal<-as.integer(TotalResourceWorking/Individuals) #working landscape resources per individual
+  ProtectPerDefect<-as.integer(StartResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
+  WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
   
[…]

### 2020-04-21 CookstoveDataSimulation.R
Stuff for ashley

@@ -38,7 +38,7 @@ dat<-data.frame(stoveID=namevec,
   YearBuilt= c(rep(2015,5),rep(2017,5),sample(2016:2020,90,replace=T)),
   UsePerDay = sort(rpois(100,2)),
   UseTotalEstimate = rep(NA,100),
-  stdPmEsmssions = rnorm(100,0,1),
+  stdPmEmissions = rnorm(100,0,1),
   GeogDistModel = sort(rgamma(100,75,6)),
   SocialDistModel = c(rep(1,55),rep(2,30),rep(3,15)),
   GeneticDistModel = rep(NA,100),
@@ -47,5 +47,35 @@ dat<-data.frame(stoveID=namevec,
   Use = sample(c("Commercial","Family","Both"),100,replace=T))
 
 head(dat)
-
 write.csv(dat,file="SimulatedStoveData.csv")
+
+
+dat$totwoodburned<-dat$WoodStart-dat$WoodEnd
+
+library(ggplot2)
+
+ggplot(data=dat,aes(x=as.character(YearBuilt), y=stdPmEmissions))+
+  geom_boxplot()+
+  geom_jitter(color="purple")
+
+ggplot(data=dat,aes(x=VisSimModel, y=stdPmEmissions))+
+  geom_line()
+
+
+ggplot(data=dat,aes(x=YearBuilt, y=stdPmEmissions))+
+  geom_smooth(method="lm",se=F)
+#PM emissions by the year built. 
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2020-04-22 app.R
changed the name to App so it runs in mobatxerm

@@ -77,8 +77,10 @@ server <- function(input, output) {
         
         ## Set parameters
         PercWorking= 1-PercProtected #percent of resource in a working landscape
-        TotalResourceWorking = PercWorking*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the working landscape starting
-        TotalResourceProtected = PercProtected*TotalCarryingCapacity*StartPercCarryingCapacity #number of resources in the protected landscape starting
+        TotalCCResourceProtected=PercProtected*TotalCarryingCapacity #total resource units in carrying capacity of protected area
+        TotalCCResourceWorking=PercWorking*TotalCarryingCapacity #total resource units in carrying capacity of working area
+        StartResourceWorking = TotalCCResourceWorking*StartPercCarryingCapacity #number of resources in the working landscape starting
+        StartResourceProtected = TotalCCResourceProtected*StartPercCarryingCapacity #number of resources in the protected landscape starting
         CoopNumStart= as.integer(CoopPercStart*Individuals) #number of individuals cooperating fully at t0
         DefNumStart= Individuals - CoopNumStart#number of individuals defecting at t0
         PercTimeProtected = c(rep(0,CoopNumStart),rbeta(DefNumStart,1,2)) #percent of their foraging time each indv spends in the PA
@@ -89,8 +91,8 @@ server <- function(input, output) {
                            PayoffProtectedLastTime = rep(NA,Individuals),
                            PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
         
-        ProtectPerDefect<-as.integer(TotalResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
-        WorkingPerTotal<-as.integer(TotalResourceWorking/Individuals) #working landscape resources per individual
+        ProtectPerDefect<-as.integer(StartResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
+        WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
         
         for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
[…]

### 2020-04-23 ProtectedAreasABM.R
added at cooperation % plot and changed the layout a bit

@@ -136,10 +136,11 @@ output$meanPayoff[t]  <- mean(agents$PayoffTotalLastTime)
 
 }
 
-par(mfrow=c(3,1))
+par(mfrow=c(4,1))
 plot(output$percCCProtect, type = 'l',col="green")
 plot(output$percCCWorking, type = 'l',col="brown")
 plot(output$meanPayoff, type = 'l',col="red")
+plot(output$meanTimeWorking, type = 'l',col="purple")
 mtext(paste("Individuals=",Individuals,
             "  Total Resources=",TotalCarryingCapacity,
             "  Percent Protected=",PercProtected,
@@ -152,7 +153,6 @@ mtext(paste("Individuals=",Individuals,
  return(output) 
 }
 
-abm(harvestMax = 1,ResourceRegenerationPerTimeStep = 1.5,TimeSteps = 50,CoopPercStart = 0.5,StartPercCarryingCapacity = 0.2)
 
 
 abm()

### 2020-04-23 app.R

@@ -1,60 +1,64 @@
-
-
-
 library(shiny)
 library(ggplot2)
+library(shinythemes)
 
 # Define UI for application that draws a histogram
-ui <- fluidPage(
 
-    # Application title
-    titlePanel("AbmOutput"),
+ui<-shinyUI(navbarPage("Protected Areas ABM",
+                       theme = shinytheme("flatly"),
+                       tabPanel("AbmOutput",
+                                # plot the map
+                                fluidRow(column(12,
+                                                h1("AbmOutput"),
+                                                p("Instructions Go Here"),
+                                                br(),
+                                                strong("Important information goes here"))),
+                                
+                                hr(),
+                                fluidRow(sidebarPanel(width = 3,
+                                                      
+                                                      helpText("Help Text"),
+                                                      sliderInput("individuals",
+                                                                  "Number of individuals:",
+                                                                  min = 10,
+                                                                  max = 1000,
+                                                                  value = 100),
+                                                      sliderInput("totCareCapac","Total Carrying Capacity Landscape",
+                                                                  min = 100,max=100000,value=10000),
+                                                      sliderInput("StartPercCarryingCapacity","Starting Percent Carrying Capacity",
+                                                                  min=0.1,max=1.0,value=0.75),
+                                                      sliderInput("timeSteps",
+                                                                  "Time Steps",
+                                                                  min = 10,
+                                                                  max = 1000,
+                                                                  value = 100),
+                                                      sliderInput("percProtect",
+                                                                  "Percent of Resources Protected",
+                                                                  min = 0.01,
+                                                                  max = 0.99,
+                                                                  value = 0.20),
+                                                      sliderInput("startCooperators",
+                                                                  "Starting Obligatory Cooperators",
[…]

### 2020-05-05 ProtectedAreasABM.R
changed plots to ggplot

@@ -1,4 +1,4 @@
-
+library(ggplot2)
 
 abm<-function(#Specified parameters
   Individuals=100, #number of total resource users in a population
@@ -136,18 +136,37 @@ output$meanPayoff[t]  <- mean(agents$PayoffTotalLastTime)
 
 }
 
-par(mfrow=c(4,1))
-plot(output$percCCProtect, type = 'l',col="green")
-plot(output$percCCWorking, type = 'l',col="brown")
-plot(output$meanPayoff, type = 'l',col="red")
-plot(output$meanTimeWorking, type = 'l',col="purple")
-mtext(paste("Individuals=",Individuals,
-            "  Total Resources=",TotalCarryingCapacity,
-            "  Percent Protected=",PercProtected,
-          "  Max Harvest=",harvestMax,
-           "  Resource Regeneration=",ResourceRegenerationPerTimeStep,
-            "  Percent Cooperating at start",CoopPercStart ), side=3, outer=TRUE, line=-3,cex=0.8)
+#par(mfrow=c(4,1))
+#plot(output$percCCProtect, type = 'l',col="green")
+#plot(output$percCCWorking, type = 'l',col="brown")
+#plot(output$meanPayoff, type = 'l',col="red")
+#plot(output$meanTimeWorking, type = 'l',col="purple")
+#mtext(paste("Individuals=",Individuals,
+ #           "  Total Resources=",TotalCarryingCapacity,
+  #          "  Percent Protected=",PercProtected,
+   #       "  Max Harvest=",harvestMax,
+    #       "  Resource Regeneration=",ResourceRegenerationPerTimeStep,
+     #       "  Percent Cooperating at start",CoopPercStart ), side=3, outer=TRUE, line=-3,cex=0.8)
+p1<-ggplot(data=output,aes(x=timeStep))+
+  geom_line(aes(y=percCCProtect),size=3,color="#b2df8a")+
+  geom_line(aes(y=percCCWorking),size=3,color="#1f78b4")+
+  ylim(0, 1.0)+theme_classic()+ylab("Percent Carrying Capacity")+
+  scale_colour_manual(name = 'Area', 
+                      values =c('#b2df8a'='#b2df8a','#1f78b4'='#1f78b4'), labels = c('Protected Area','Working Landscape'))+
+  ggtitle("Resource Degredation")
 
+
+p2<-ggplot(data=output,aes(x=timeStep))+
[…]

### 2020-05-08 Ashley Lagrou VIP timeline.Rmd
Stuff for ashley
cleaning up

@@
 
-Plotting is an important first step to understanding your data. It should always be your first stop once you have clean data in hand.
+We will use the package ggplot2 to plot the data that you have collected/extrapolated.
 
-We will be using the ggplot2 package in R for all of our plotting
+If you remember from before, the ggplot() function follows the [grammar of graphics](https://towardsdatascience.com/a-comprehensive-guide-to-the-grammar-of-graphics-for-effective-visualization-of-multi-dimensional-1f92b4ed4149), a theory for how to visualize data that relies on stacking layers on top of each other.
+
+When you create ggplot() code, the top most code will be the first layer created.
+For our purposed, we will always want this to be the axes of the figure.
+We will do this by specifying the data and the x and y axes.
+Remember that the x and y axes need to be wrapped in an aestetic function aes().
+This just tells ggplot to look for these things in the dataset that we specified.
+
 ```{r}
-library(ggplot2)
+library(ggplot2) #load the package
+
+ggplot(data= dat, aes(x=GeogDistModel,y=BoilTime))
+
 ```
+
+You'll see that this just "sets the stage" for the rest of the plot. We haven't told it what kind of plot we want yet.
+
+Next, we will layer the data on top of this background.
+
+```{r}
+ggplot(data= dat, aes(x=GeogDistModel,y=BoilTime))+
+  geom_point()
+```
+
+Is this the best way to visualize this simple dataset? probably not. Can you think of  better way?
+
+There is a ton of flexibility in ggplot2. See if you can make this plot better.
+
+Once you do that, see if you can plot your data that you created by extrapolating from the fee data we had.
+
+I like to hand draw my plots first so I know what I'm aiming for.
+
+If you want help and inspiration check out the [Cookbook for ggplot](http://www.cookbook-r.com/Graphs/) and the [ggplot gallery](https://www.r-graph-gallery.com/).

### 2020-05-08 AdvancedVisualizationMapping.R

@@ -1,7 +1,85 @@
 library(raster)
-x<-getData('worldclim',var = 'prec',res = 0.5, lon=39.74,lat=-5.313)
-plot(x)
+library(RColorBrewer)
+prec<-getData('worldclim',var = 'prec',res = 0.5, lon=39.74,lat=-5.313)
+tmin<-getData('worldclim',var = 'tmin',res = 0.5, lon=39.74,lat=-5.313)
+tmax<-getData('worldclim',var = 'tmax',res = 0.5, lon=39.74,lat=-5.313)
+bio<-getData('worldclim',var = 'bio',res = 0.5, lon=39.74,lat=-5.313)
+save(bio,file="bio.rda")
+
+writeRaster(bio,filename="bio.tif")
+
+bio = stack("bio.tif")
+
+scale.parameter = 0.3  # scaling paramter. less than 1 is zooming in, more than 1 zooming out. 
+xshift = -7 # Shift to right in map units. 
+yshift = 10  # Shift to left in map units. 
+original.bbox = bbox(prec) # Pass bbox of your Spatial* Object. 
+# Just copy-paste the following
+edges = original.bbox
+
+edges[1, ] <- (edges[1, ] - mean(edges[1, ])) * scale.parameter + mean(edges[1, 
+                                                                             ]) + xshift
+edges[2, ] <- (edges[2, ] - mean(edges[2, ])) * scale.parameter + mean(edges[2, 
+                                                                             ]) + yshift
+
+#Monthly average plots
+months<-c( "January", "February", "March", "April", 
+                                     "May", "June", "July", "August", "September", "October", 
+                                     "November", "December")
+
+names(tmin)<-months
+names(tmax)<-months
+names(prec)<-months
+
+tmin<-tmin/10
+tmax<-tmax/10
+
+prec.pal <- brewer.pal(n = 7, name = "Blues")
+spplot(prec$January, 
+       col.regions = prec.pal, 
+       cuts = 6,main="Eastern Tanzania & Zanzibar Precipitation",
+       sub="Average Precipitation from 1950 - 2000", xlim = edges[1, ], ylim = edges[2, ])
 
 
-tzProtect<-read.csv("~/Pemba Project/TanzaniaProtectedAreas.csv")
-tzProtect
+t.pal <- rev(brewer.pal(n = 7, name = "RdYlBu"))
[…]

### 2020-05-13 ProtectedAreasABM.R
added resource mobility

@@ -12,8 +12,9 @@ abm<-function(#Specified parameters
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
   TimeSteps=20,
-  ResourceRegenerationPerTimeStep=1.5,
-  harvestMax=25){
+  ResourceRegenerationPerTimeStep=1.15,
+  harvestMax=25,
+  ProbOfMobility=0.2){
   
   ## Set parameters
   PercWorking= 1-PercProtected #percent of resource in a working landscape
@@ -105,11 +106,25 @@ agents<-data.frame(PercTimeProtected,   #new agent dataframe to fill
 #New resource pools to pull from
 #make sure they dont regenerate past their carrying capacity
 
-NewProtectedResourcesTotal<-output$NumResourcesProtect[t-1] * ResourceRegenerationPerTimeStep
+NewProtectedResourcesTotal<-round(output$NumResourcesProtect[t-1] * ResourceRegenerationPerTimeStep,digits=0)
 NewProtectedResourcesTotal<-ifelse(NewProtectedResourcesTotal<=  TotalCCResourceProtected,NewProtectedResourcesTotal, TotalCCResourceProtected)
-NewWorkingResourcesTotal<-output$NumResourcesWorking[t-1] * ResourceRegenerationPerTimeStep
+NewWorkingResourcesTotal<-round(output$NumResourcesWorking[t-1] * ResourceRegenerationPerTimeStep,digits=0)
 NewWorkingResourcesTotal<-ifelse(NewWorkingResourcesTotal<=TotalCCResourceWorking,NewWorkingResourcesTotal,TotalCCResourceWorking)
 
+##Resource mobility
+LeaveWorking<-rbinom(1,NewWorkingResourcesTotal,ProbOfMobility) #number of resources which leave the protected area
+LeaveProtected<-rbinom(1,NewProtectedResourcesTotal,ProbOfMobility)#number of resources which leave the working area
+
+#do the accounting on entering vs leaving individuals
+NewWorkingResourcesTotal<-NewWorkingResourcesTotal-LeaveWorking+LeaveProtected
+NewProtectedResourcesTotal<-NewProtectedResourcesTotal-LeaveProtected+LeaveWorking
+
+#Make sure they dont go over the CC again
[…]

### 2020-05-13 app.R

@@ -26,7 +26,7 @@ ui<-shinyUI(navbarPage("Protected Areas ABM",
                                                       sliderInput("totCareCapac","Total Carrying Capacity Landscape",
                                                                   min = 100,max=100000,value=10000),
                                                       sliderInput("StartPercCarryingCapacity","Starting Percent Carrying Capacity",
-                                                                  min=0.1,max=1.0,value=0.75),
+                                                                  min=0.01,max=1.0,value=0.65),
                                                       sliderInput("timeSteps",
                                                                   "Time Steps",
                                                                   min = 10,
@@ -34,24 +34,29 @@ ui<-shinyUI(navbarPage("Protected Areas ABM",
                                                                   value = 100),
                                                       sliderInput("percProtect",
                                                                   "Percent of Resources Protected",
-                                                                  min = 0.01,
+                                                                  min = 0.00,
                                                                   max = 0.99,
-                                                                  value = 0.20),
+                                                                  value = 0.4),
                                                       sliderInput("startCooperators",
                                                                   "Starting Obligatory Cooperators",
-                                                                  min = 0.1,
-                                                                  max = 0.95,
+                                                                  min = 0.0,
+                                                                  max = 0.99,
                                                                   value = 0.9),
                                                       sliderInput("resourceRegen",
                                                                   "Resource Regeneration Rate",
-                                                                  min = 1.1,
-                                                                  max = 1.9,
+                                                                  min = 1.01,
+                                                                  max = 1.99,
                                                                   value = 1.5),
+                                                      sliderInput("ProbOfMobility",
+                                                                  "Probability of resource mobility",
+                                                                  min = 0.00,
[…]

### 2020-05-13 ProtectedAreasABM.R
Added learning strategies "success bias" & "conformist". Maybe add individual next? Leader following?

@@ -11,6 +11,8 @@ abm<-function(#Specified parameters
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
+  LearningStrategy = "Success Bias", #options are Success Bias & Conformist
+  
   TimeSteps=20,
   ResourceRegenerationPerTimeStep=1.15,
   harvestMax=25,
@@ -41,12 +43,12 @@ abm<-function(#Specified parameters
     agents[i,4]<-ifelse(agents[i,4]>harvestMax,harvestMax,agents[i,4])
     agents[i,3]<- ifelse(agents[i,4]<= harvestMax,rbinom(1,ProtectPerDefect,agent2$PercTimeProtected),0) #if they didnt get the max harvest, then 
     agents[i,3]<-ifelse(agents[i,3]>harvestMax-agents[i,4],harvestMax-agents[i,4],agents[i,3])
-    } #see what they get from the working landscape first
+  } #see what they get from the working landscape first
   
   
   
   agents$PayoffTotalLastTime<- agents$PayoffProtectedLastTime + agents$PayoffWorkingLastTime #total payoff is the sum of what they get from both areas
-   
+  
   
   output<-data.frame(timeStep = 1:TimeSteps, #create df to fill with the outputs from each time step
                      meanTimeProtect = rep(NA,TimeSteps),
@@ -56,132 +58,138 @@ abm<-function(#Specified parameters
                      NumResourcesWorking = rep(NA,TimeSteps),
                      percCCWorking = rep(NA,TimeSteps),
                      meanPayoff = rep(NA,TimeSteps))
-
-output$meanTimeProtect[1] <- mean(agents$PercTimeProtected)  #fill it with the outouts from time 1
-output$meanTimeWorking[1]  <- mean(agents$PercTimeWorking) 
-output$NumResourcesProtect[1] <- StartResourceProtected-sum(agents$PayoffProtectedLastTime) 
-output$NumResourcesWorking[1] <- StartResourceWorking-sum(agents$PayoffWorkingLastTime) 
-output$percCCProtect[1]  <- output$NumResourcesProtect[1]/TotalCCResourceProtected
-output$percCCWorking[1]  <- output$NumResourcesWorking[1]/TotalCCResourceWorking
-output$meanPayoff[1]  <- mean(agents$PayoffTotalLastTime) 
-
   
-
-for (t in 2:TimeSteps){  #run the abm for all o fthe time steps
[…]

### 2020-05-13 app.R

@@ -23,6 +23,9 @@ ui<-shinyUI(navbarPage("Protected Areas ABM",
                                                                   min = 10,
                                                                   max = 1000,
                                                                   value = 100),
+                                                      selectInput("LearningStrategy","Learning Strategy",
+                                                                  c("Success Bias" = "Success Bias",
+                                                                    "Conformist"= "Conformist")),
                                                       sliderInput("totCareCapac","Total Carrying Capacity Landscape",
                                                                   min = 100,max=100000,value=10000),
                                                       sliderInput("StartPercCarryingCapacity","Starting Percent Carrying Capacity",
@@ -74,11 +77,13 @@ server <- function(input, output) {
         
         TotalCarryingCapacity=10000, #total available resource units
         
-        StartPercCarryingCapacity = 0.6, #amount of resources available in the landscape at the start in proportion to CC
+        StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
         
-        PercProtected=0.00, #percent of the total resource that's in a protected area
+        PercProtected=0.3, #percent of the total resource that's in a protected area
         
         CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
+        
+        LearningStrategy = "Success Bias", #options are Success Bias & Conformist
         
         TimeSteps=20,
         ResourceRegenerationPerTimeStep=1.15,
@@ -146,22 +151,33 @@ server <- function(input, output) {
             LastTimePercWorking<-LastTimeAgents$PercTimeWorking  #pull out the previous working foraging times as vector
             
             PercTimeProtected<-rep(NA,length(PercTimeProtected))  #nullify the previous perc time protected vector so we can fill it later and be confident the values changed
-            for (j in 1:nrow(LastTimeAgents)){  
-                ThisAgent<-LastTimeAgents[j,]      #modify the foraging strategy of each agent individualls
-                OtherAgents<-LastTimeAgents[-j,]   #dataframe of ther agents for them to copy from
-                ThisAgentPayoff<-ThisAgent$PayoffTotalLastTime   #payoff last time period for this agent
[…]

### 2020-05-15 ResourceMobilityABMManuscript.rmd
started intro in a markdown doc

---
title: "Resource Mobility & Protected Areas Efficacy"
author: "Matt Clark"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

# Abstract

# Introduction 
-Protected areas are widely used. Both marine and terrestrial
-Provide a source of species and should be adopted widely based on theory (beneficial "technology")
-They are not always adopted, especially over time.
-Adoption/cooperation is a key factor which influences the success (social & ecological) of any given area. 
-People will adopt/cooperate more readily if adoption/cooperation yeilds immediate, noticable benefits to them
-It makes sense logically that protected areas will provide more immediate benefits to individuals if the resources are highly mobile
-While it makes sense logically, we need to formally show this. 
-We need to formally show it both in terms of the spatial/ecological aspect, but also in terms of the social adoption component. 
-Extending this hypothesis, we think that highly mobile resources should require a smaller refuge, especially if these resources are quickly regnerating.
-Because these systems consist of two interacting groups (both with subgroups) all of which are evolving every time step, equation based models can not accurately depic the system. Differential equation for example try to find an equilibrium point in a changing system, which is not the goal of this study.
-Using a dynamic model will allow us test our logic and intuition about how these system works.

-p1 Resource mobility will increase the efficacy (social & ecological) of protected areas
-p2 Protected area size will also increase efficacy 
-p3 Resource mobility and PA size will complement each other 
-p4 Resource regeneration rate will decrease necessary mobility and PA size
-p5 Different learning strategies will affect these dynamics, but the overall predictions will hold true.

### additional?
-We have a database of all protected areas in the world (~kinda)
-Check if there's any correlation between % of resource protected and resource mobility

### 2020-06-01 Abm_MultipleReplications_NOTDONE.R
It's running! Now i need to figure out how to best save and display the results

@@ -1,0 +1,209 @@
+abm<-function(#Specified parameters
+  
+  Runs=3,
+  Individuals=100, #number of total resource users in a population
+  
+  TotalCarryingCapacity=10000, #total available resource units
+  
+  StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
+  
+  PercProtected=0.3, #percent of the total resource that's in a protected area
+  
+  CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
+  
+  LearningStrategy = "Success Bias", #options are Success Bias & Conformist
+  
+  TimeSteps=20,
+  ResourceRegenerationPerTimeStep=1.15,
+  harvestMax=25,
+  ProbOfMobility=0.2){
+  
+  ## Create output dfs
+  
+  outputmeanTimeProtect<-as.data.frame(matrix(NA,TimeSteps,Runs)) #output df for mean time harvesting from protected area
+  names(outputmeanTimeProtect)<- paste("Run", 1:Runs, sep="")
+  
+  outputmeanTimeWorking<-as.data.frame(matrix(NA,TimeSteps,Runs))#output df for mean time harvesting from working area
+  names(outputmeanTimeWorking)<- paste("Run", 1:Runs, sep="")
+  
+  outputNumResourcesProtect<-as.data.frame(matrix(NA,TimeSteps,Runs))#output df for resources in protected area
+  names(outputNumResourcesProtect)<- paste("Run", 1:Runs, sep="")
+  
+  outputpercCCProtect<-as.data.frame(matrix(NA,TimeSteps,Runs))#output df for % of CC in protected area
+  names(outputpercCCProtect)<- paste("Run", 1:Runs, sep="")
+  
+  outputNumResourcesWorking<-as.data.frame(matrix(NA,TimeSteps,Runs))#output df for resources in working area
+  names(outputNumResourcesWorking)<- paste("Run", 1:Runs, sep="")
+  
+  outputpercCCWorking<-as.data.frame(matrix(NA,TimeSteps,Runs))#output df for % of CC in working area
+  names(outputpercCCWorking)<- paste("Run", 1:Runs, sep="")
+  
+  outputmeanPayoff<-as.data.frame(matrix(NA,TimeSteps,Runs)) #output df for mean payoff
+  names(outputmeanPayoff)<- paste("Run", 1:Runs, sep="")
+  
+  
+  
+  for (r in 1:Runs){
+    PercWorking= 1-PercProtected #percent of resource in a working landscape
[…]

### 2020-06-02 MultipleReplicationsNoPlots.R
Multiple replications with mean outputs stored as a df running.

@@ -182,44 +182,8 @@ abm<-function(#Specified parameters
       outputmeanPayoff[t,r]  <- mean(agents$PayoffTotalLastTime)
       
     }
-    par(mfrow=c(2,2))
     
-    #mean payoff
-    plot(rowMeans(outputmeanPayoff), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,harvestMax), lwd = 3, main = "Mean Payoff")
     
-    for (j in 1:Runs) {  
-      
-      lines(outputmeanPayoff[,j], type = 'l')  # add lines for each run, up to r_max
-      
-    }
-    
-    #% CC protected
-    plot(rowMeans(outputmeanTimeWorking), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,1), lwd = 3, main = "Average time in working (coop)")
-    
-    for (j in 1:Runs) {  
-      
-      lines(outputmeanTimeWorking[,j], type = 'l')  # add lines for each run, up to r_max
-      
-    }
-    
-    #% CC protected
-    plot(rowMeans(outputpercCCProtect), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,1), lwd = 3, main = "Percent CC Protected")
-    
-    for (j in 1:Runs) {  
-      
-      lines(outputpercCCProtect[,j], type = 'l')  # add lines for each run, up to r_max
-      
-    }
-    
-    #% CC working
-    plot(rowMeans(outputpercCCWorking), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,1), lwd = 3, main = "Percent CC Working")
-    
-    for (j in 1:Runs) {  
-      
-      lines(outputpercCCWorking[,j], type = 'l')  # add lines for each run, up to r_max
-      
-    }
-
     
   }
   outputmeanPayoff$Timestep<-c(1:TimeSteps)
@@ -255,13 +219,9 @@ FillData<-expand.grid(PercProtected=seq(0.1,0.9,0.05), ResourceMobility=seq(0.1,
 FillData$ProtectedCC<-rep(NA,nrow(FillData))
 
 for (i in 1:nrow(FillData)){
-abm(Runs=10,Individuals = 100,
-    PercProtected = FillData[i,]$PercProtected, 
-    ProbOfMobility = FillData[i,]$ResourceMobility)
[…]

### 2020-06-02 app.R

@@ -112,8 +112,8 @@ server <- function(input, output) {
         for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
             agent2<-agents[i,]   #of resources there per individual searching there, and a random draw with a chance of success equal to time spent
             agents[i,4]<-rbinom(1,WorkingPerTotal,agent2$PercTimeWorking)
-            agents[i,4]<-ifelse(agents[i,4]>harvestMax,harvestMax,agents[i,4])
-            agents[i,3]<- ifelse(agents[i,4]<= harvestMax,rbinom(1,ProtectPerDefect,agent2$PercTimeProtected),0) #if they didnt get the max harvest, then 
+            agents[i,4]<-ifelse(agents[i,4]>=harvestMax,harvestMax,agents[i,4])
+            agents[i,3]<- ifelse(agents[i,4]< harvestMax,rbinom(1,ProtectPerDefect,agent2$PercTimeProtected),0) #if they didnt get the max harvest, then 
             agents[i,3]<-ifelse(agents[i,3]>harvestMax-agents[i,4],harvestMax-agents[i,4],agents[i,3])
         } #see what they get from the working landscape first
         
@@ -218,8 +218,8 @@ server <- function(input, output) {
             for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
                 agent2<-agents[i,]   #of resources there per individual searching there, and a random draw with a chance of success equal to time spent
                 agents[i,4]<-rbinom(1,WorkingPerTotal,agent2$PercTimeWorking)
-                agents[i,4]<-ifelse(agents[i,4]>harvestMax,harvestMax,agents[i,4])
-                agents[i,3]<- ifelse(agents[i,4]<= harvestMax,rbinom(1,ProtectPerDefect,agent2$PercTimeProtected),0) #if they didnt get the max harvest, then 
+                agents[i,4]<-ifelse(agents[i,4]>=harvestMax,harvestMax,agents[i,4])
+                agents[i,3]<- ifelse(agents[i,4]< harvestMax,rbinom(1,ProtectPerDefect,agent2$PercTimeProtected),0) #if they didnt get the max harvest, then 
                 agents[i,3]<-ifelse(agents[i,3]>harvestMax-agents[i,4],harvestMax-agents[i,4],agents[i,3])
             } #see what they get from the working landscape first
             

### 2020-06-02 ResourceMobilityABMManuscript.rmd
stared model description

@@
 -Check if there's any correlation between % of resource protected and resource mobility
+
+### Model Description
+*Starting conditions
+Our agent based model represents a simplistic world where there is one homogenous group of agents who harvest resources from a landscape.
+This landscape is zoned into two patches, working and protected.
+Resources move between these patches at each time step with a probability designated in the model setup.
+At time = 0, the protected and working landscapes are both at an equal, designated percent carrying capacity.
+At each time step, agents aim to maximize their harvest up until a designated harvest max.
+At time = 0, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree (spend a small portion of therir time harvesting from the protected landscape).
+
+*Agent attributes and behavior
+
+*Environment attribtes and behavior
+
+*Key Processes
+At time step one, all agents harvest first spend their initial percentage of effort harvesting from the working landscape.
+If agents reach harvest max, they do not harvest from the protected landscape.
+If they do not reach harvest max, they will spend the rest of their effort harvesting from the protected area.
+They will harvest up to the harvest max if possible.
+The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape.
+If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies.
+If one agent preformed worse than another in the first time step, the underpreforming agent will adopt a strategy in-between their pervious strategy and the strategy of the agent who outpreformed them.
+If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group.
+This process repeats for all time steps.
+
+*Outputs

### 2020-06-02 MultipleReplicationsNoPlots.R
Added a simple counter so we can track model progress

@@ -223,5 +223,6 @@ for (i in 1:nrow(FillData)){
       PercProtected = FillData[i,]$PercProtected, 
       ProbOfMobility = FillData[i,]$ResourceMobility)
   FillData[i,3]<-apply(FullOutput$percCCProtect[75,1:10],1,mean)  
-  
+  print(nrow(FillData))
+  print(i)
 }

### 2020-06-02 Abm_MultipleReplications_NOTDONE.R
Fixed the NA problem. Small decimals were carrying into the rbinom call

@@ -7,7 +7,7 @@ abm<-function(#Specified parameters
   
   StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
   
-  PercProtected=0.3, #percent of the total resource that's in a protected area
+  PercProtected=0.15, #percent of the total resource that's in a protected area
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
@@ -16,7 +16,7 @@ abm<-function(#Specified parameters
   TimeSteps=75,
   ResourceRegenerationPerTimeStep=1.5,
   harvestMax=15,
-  ProbOfMobility=0.2){
+  ProbOfMobility=0.15){
   
   ## Create output dfs
   
@@ -64,7 +64,7 @@ abm<-function(#Specified parameters
     
     for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
       agent2<-agents[i,]   #of resources there per individual searching there, and a random draw with a chance of success equal to time spent
-      agents[i,4]<-rbinom(1,WorkingPerTotal,agent2$PercTimeWorking)
+      agents[i,4]<-rbinom(1,as.integer(WorkingPerTotal),agent2$PercTimeWorking)
       agents[i,4]<-ifelse(agents[i,4]>=harvestMax,harvestMax,agents[i,4])
       agents[i,3]<- ifelse(agents[i,4]< harvestMax,rbinom(1,ProtectPerDefect,agent2$PercTimeProtected),0) #if they didnt get the max harvest, then 
       agents[i,3]<-ifelse(agents[i,3]>harvestMax-agents[i,4],harvestMax-agents[i,4],agents[i,3])
@@ -145,7 +145,7 @@ abm<-function(#Specified parameters
       
       ##Resource mobility
       LeaveWorking<-rbinom(1,NewWorkingResourcesTotal,ProbOfMobility) #number of resources which leave the protected area
-      LeaveProtected<-rbinom(1,NewProtectedResourcesTotal,ProbOfMobility)#number of resources which leave the working area
+      LeaveProtected<-rbinom(1,as.integer(NewProtectedResourcesTotal),ProbOfMobility)#number of resources which leave the working area
       
       #do the accounting on entering vs leaving individuals
       NewWorkingResourcesTotal<-NewWorkingResourcesTotal-LeaveWorking+LeaveProtected
[…]

### 2020-06-02 MultipleReplicationsNoPlots.R

@@ -7,7 +7,7 @@ abm<-function(#Specified parameters
   
   StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
   
-  PercProtected=0.3, #percent of the total resource that's in a protected area
+  PercProtected=0.15, #percent of the total resource that's in a protected area
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
@@ -16,7 +16,7 @@ abm<-function(#Specified parameters
   TimeSteps=75,
   ResourceRegenerationPerTimeStep=1.5,
   harvestMax=15,
-  ProbOfMobility=0.2){
+  ProbOfMobility=0.15){
   
   ## Create output dfs
   
@@ -64,7 +64,7 @@ abm<-function(#Specified parameters
     
     for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
       agent2<-agents[i,]   #of resources there per individual searching there, and a random draw with a chance of success equal to time spent
-      agents[i,4]<-rbinom(1,WorkingPerTotal,agent2$PercTimeWorking)
+      agents[i,4]<-rbinom(1,as.integer(WorkingPerTotal),agent2$PercTimeWorking)
       agents[i,4]<-ifelse(agents[i,4]>=harvestMax,harvestMax,agents[i,4])
       agents[i,3]<- ifelse(agents[i,4]< harvestMax,rbinom(1,ProtectPerDefect,agent2$PercTimeProtected),0) #if they didnt get the max harvest, then 
       agents[i,3]<-ifelse(agents[i,3]>harvestMax-agents[i,4],harvestMax-agents[i,4],agents[i,3])
@@ -145,7 +145,7 @@ abm<-function(#Specified parameters
       
       ##Resource mobility
       LeaveWorking<-rbinom(1,NewWorkingResourcesTotal,ProbOfMobility) #number of resources which leave the protected area
-      LeaveProtected<-rbinom(1,NewProtectedResourcesTotal,ProbOfMobility)#number of resources which leave the working area
+      LeaveProtected<-rbinom(1,as.integer(NewProtectedResourcesTotal),ProbOfMobility)#number of resources which leave the working area
       
[…]

### 2020-06-02 MultipleReplicationsNoPlots.R
ready to run on HES bowl

@@ -214,21 +214,22 @@ abm(Runs=5)
 #fill a dataframe with the means
 
 
-FillData<-expand.grid(PercProtected=seq(0.1,0.3,0.05), ResourceMobility=seq(0.1,0.3,0.05))
+FillData<-expand.grid(PercProtected=seq(0.05,0.95,0.05), ResourceMobility=seq(0.05,0.95,0.05))
 FillData$ProtectedCC<-rep(NA,nrow(FillData))
 FillData$WorkingCC<-rep(NA,nrow(FillData))
 FillData$MeanPayoff<-rep(NA,nrow(FillData))
 FillData$meanTimeWorking<-rep(NA,nrow(FillData))
 
-##na's being produces sometimes??? why?????
+
+#Script for running on HES Bowl computer
 for (i in 1:nrow(FillData)){
-  abm(Runs=5,Individuals = 100,TimeSteps = 50,harvestMax = 17,
+  abm(Runs=50,Individuals = 100,TimeSteps = 100,harvestMax = 17,
       PercProtected = FillData[i,]$PercProtected, 
       ProbOfMobility = FillData[i,]$ResourceMobility)
-  FillData[i,3]<-apply(FullOutput$percCCProtect[50,1:5],1,mean) 
-  FillData[i,4]<-apply(FullOutput$percCCWorking[50,1:5],1,mean)  
-  FillData[i,5]<-apply(FullOutput$meanPayoff[50,1:5],1,mean)  
-  FillData[i,6]<-apply(FullOutput$meanTimeWorking[50,1:5],1,mean)   #[Timesteps,1:Runs]
+  FillData[i,3]<-apply(FullOutput$percCCProtect[100,1:5],1,mean) 
+  FillData[i,4]<-apply(FullOutput$percCCWorking[100,1:5],1,mean)  
+  FillData[i,5]<-apply(FullOutput$meanPayoff[100,1:5],1,mean)  
+  FillData[i,6]<-apply(FullOutput$meanTimeWorking[100,1:50],1,mean)   #[Timesteps,1:Runs]
   print(nrow(FillData))
   print(i)
 }

### 2020-06-03 MultipleReplicationsNoPlots.R
heatmaps from hes bowl model runs

@@ -5,7 +5,7 @@ abm<-function(#Specified parameters
   
   TotalCarryingCapacity=10000, #total available resource units
   
-  StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
+  StartPercCarryingCapacity = 0.65, #amount of resources available in the landscape at the start in proportion to CC
   
   PercProtected=0.15, #percent of the total resource that's in a protected area
   
@@ -220,29 +220,47 @@ FillData$WorkingCC<-rep(NA,nrow(FillData))
 FillData$MeanPayoff<-rep(NA,nrow(FillData))
 FillData$meanTimeWorking<-rep(NA,nrow(FillData))
 
-
-#Script for running on HES Bowl computer
+##na's being produces sometimes??? why?????
 for (i in 1:nrow(FillData)){
-  abm(Runs=50,Individuals = 100,TimeSteps = 100,harvestMax = 17,
+  abm(Runs=3,Individuals = 100,TimeSteps = 100,harvestMax = 17,
       PercProtected = FillData[i,]$PercProtected, 
       ProbOfMobility = FillData[i,]$ResourceMobility)
-  FillData[i,3]<-apply(FullOutput$percCCProtect[100,1:5],1,mean) 
-  FillData[i,4]<-apply(FullOutput$percCCWorking[100,1:5],1,mean)  
-  FillData[i,5]<-apply(FullOutput$meanPayoff[100,1:5],1,mean)  
-  FillData[i,6]<-apply(FullOutput$meanTimeWorking[100,1:50],1,mean)   #[Timesteps,1:Runs]
+  FillData[i,3]<-apply(FullOutput$percCCProtect[100,1:3],1,mean) 
+  FillData[i,4]<-apply(FullOutput$percCCWorking[100,1:3],1,mean)  
+  FillData[i,5]<-apply(FullOutput$meanPayoff[100,1:3],1,mean)  
+  FillData[i,6]<-apply(FullOutput$meanTimeWorking[100,1:3],1,mean)   #[Timesteps,1:Runs]
   print(nrow(FillData))
   print(i)
 }
 
+library(ggplot2)
+library(viridis)
+ggplot(data=FillData, mapping=aes(x=PercProtected, y=ResourceMobility, fill= ProtectedCC)) + 
+  geom_tile()+scale_fill_viridis()
 
[…]

### 2020-06-03 heatmaps.R

@@ -1,0 +1,45 @@
+PaSize = seq(0,1,0.01)
+ResourceMobility =seq(0,1,0.01)
+fakeData<-expand.grid(X=PaSize, Y=ResourceMobility)
+fakeData$DegredationWorking<-sort(rbeta(nrow(fakeData), 2,2))
+names(fakeData)<-c("PaSize","ResourceMobility","DegredationWorking")
+library(ggplot2)
+library(viridis)
+ggplot(data=fakeData, mapping=aes(x=PaSize, y=ResourceMobility, fill= DegredationWorking)) + 
+            geom_tile()+scale_fill_viridis()
+
+
+
+initOutputSuccess<-read.csv("~/Pemba Project/ModelResultsSuccessBias.csv")
+initOutputConform<-read.csv("~/Pemba Project/ModelResultsConformityBias.csv")
+
+
+ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= ProtectedCC)) + 
+  geom_tile()+scale_fill_viridis()+ggtitle("Success Bias -- Protected CC")
+
+ggplot(data=initOutputConform, mapping=aes(x=PercProtected, y=ResourceMobility, fill= ProtectedCC)) + 
+  geom_tile()+scale_fill_viridis()+ggtitle("Conform Bias -- Protected CC")
+
+
+ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= WorkingCC)) + 
+  geom_tile()+scale_fill_viridis()+ggtitle("Success Bias -- Working CC")
+
+ggplot(data=initOutputConform, mapping=aes(x=PercProtected, y=ResourceMobility, fill= WorkingCC)) + 
+  geom_tile()+scale_fill_viridis()+ggtitle("Conform Bias -- Working CC")
+
+ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= MeanPayoff)) + 
+  geom_tile()+scale_fill_viridis()+ggtitle("Success Bias -- Mean Payoff")
+
+ggplot(data=initOutputConform, mapping=aes(x=PercProtected, y=ResourceMobility, fill= MeanPayoff)) + 
+  geom_tile()+scale_fill_viridis()+ggtitle("Conform Bias -- Mean Payoff")
+
[…]

### 2020-06-03 MultipleReplicationsNoPlots.R
updated files from hes bowl

@@ -222,13 +222,13 @@ FillData$meanTimeWorking<-rep(NA,nrow(FillData))
 
 ##na's being produces sometimes??? why?????
 for (i in 1:nrow(FillData)){
-  abm(Runs=3,Individuals = 100,TimeSteps = 100,harvestMax = 17,
+  abm(Runs=20,Individuals = 100,TimeSteps = 100,harvestMax = 17,
       PercProtected = FillData[i,]$PercProtected, 
       ProbOfMobility = FillData[i,]$ResourceMobility)
-  FillData[i,3]<-apply(FullOutput$percCCProtect[100,1:3],1,mean) 
-  FillData[i,4]<-apply(FullOutput$percCCWorking[100,1:3],1,mean)  
-  FillData[i,5]<-apply(FullOutput$meanPayoff[100,1:3],1,mean)  
-  FillData[i,6]<-apply(FullOutput$meanTimeWorking[100,1:3],1,mean)   #[Timesteps,1:Runs]
+  FillData[i,3]<-apply(FullOutput$percCCProtect[100,1:20],1,mean) 
+  FillData[i,4]<-apply(FullOutput$percCCWorking[100,1:20],1,mean)  
+  FillData[i,5]<-apply(FullOutput$meanPayoff[100,1:20],1,mean)  
+  FillData[i,6]<-apply(FullOutput$meanTimeWorking[100,1:20],1,mean)   #[Timesteps,1:Runs]
   print(nrow(FillData))
   print(i)
 }
@@ -248,13 +248,13 @@ FillData$meanTimeWorking<-rep(NA,nrow(FillData))
 
 ##na's being produces sometimes??? why?????
 for (i in 1:nrow(FillData)){
-  abm(Runs=3,Individuals = 100,TimeSteps = 100,harvestMax = 17,LearningStrategy = "Conformist",
+  abm(Runs=20,Individuals = 100,TimeSteps = 100,harvestMax = 17,LearningStrategy = "Conformist",
       PercProtected = FillData[i,]$PercProtected, 
       ProbOfMobility = FillData[i,]$ResourceMobility)
-  FillData[i,3]<-apply(FullOutput$percCCProtect[100,1:3],1,mean) 
-  FillData[i,4]<-apply(FullOutput$percCCWorking[100,1:3],1,mean)  
-  FillData[i,5]<-apply(FullOutput$meanPayoff[100,1:3],1,mean)  
[…]

### 2020-06-04 ResourceMobilityABMManuscript.rmd
ready to run on HES bowl
a ton of small changes

@@
 -Adoption/cooperation is a key factor which influences the success (social & ecological) of any given area. 
--People will adopt/cooperate more readily if adoption/cooperation yeilds immediate, noticable benefits to them
+-People will adopt/cooperate more readily if adoption/cooperation yields immediate, noticeable benefits to them
 -It makes sense logically that protected areas will provide more immediate benefits to individuals if the resources are highly mobile
@@
 -We need to formally show it both in terms of the spatial/ecological aspect, but also in terms of the social adoption component. 
--Extending this hypothesis, we think that highly mobile resources should require a smaller refuge, especially if these resources are quickly regnerating.
--Because these systems consist of two interacting groups (both with subgroups) all of which are evolving every time step, equation based models can not accurately depic the system.
+-Extending this hypothesis, we think that highly mobile resources should require a smaller refuge, especially if these resources are quickly regenerating.
+-Because these systems consist of two interacting groups (both with subgroups) all of which are evolving every time step, equation based models can not accurately depict the system.
 Differential equation for example try to find an equilibrium point in a changing system, which is not the goal of this study.
@@
 
+-When to use permeable working landscapes and when to use sectioned off protected areas?
+
 ### Model Description
-*Starting conditions
-Our agent based model represents a simplistic world where there is one homogenous group of agents who harvest resources from a landscape.
+* Starting conditions
+Our agent based model represents a simplistic world where there is one homogeneous group of agents who harvest resources from a landscape.
 This landscape is zoned into two patches, working and protected.
@@
 At each time step, agents aim to maximize their harvest up until a designated harvest max.
-At time = 0, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree (spend a small portion of therir time harvesting from the protected landscape).
+At time = 0, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree (spend a small portion of their time harvesting from the protected landscape).
 
-*Agent attributes and behavior
+The independent variables of interest are: Resource mobility, protected area size, learning strategy, and resource regeneration rate.
 
-*Environment attribtes and behavior
+* Agent attributes and behavior
+Agents "choose" to spend a given percent of their time harvesting from the protected area (defecting) and a percent of their time harvesting from the working landscape (defecting).
+Full defection is when the average percent of time agents spend in the protected area is equal to the percent of the landscape protected (i.e.
+they do not treat the protected area any differently than the working landscape).
 
-*Key Processes
+* Environment attributes and behavior
+At every time step, both areas lost the number of resources that agents harvest from them.
+The areas then regenerate resources according to the regeneration rate per time step and lastly, the resources move from one area to the other with probability p pulled from a binomial distribution.
+
+* Key Processes
 At time step one, all agents harvest first spend their initial percentage of effort harvesting from the working landscape.
@@
 The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape.
+Resources move between the protected and working landscapes with a probability proportional to designated probability of mobility.
 If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies.
-If one agent preformed worse than another in the first time step, the underpreforming agent will adopt a strategy in-between their pervious strategy and the strategy of the agent who outpreformed them.
+If one agent preformed worse than another in the first time step, the under preforming agent will adopt a strategy in-between their previous strategy and the strategy of the agent who outperformed them.
 If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group.
@@
 
-*Outputs
+* Outputs
+The dependent variables of interest are: Cooperative behavior, protected area degradation, working landscape degradation, and mean payoff to individuals.
+
+We can run any number of runs for any combination of variables.

### 2020-06-04 Abm_MultipleReplicationsPlots.R

@@ -1,11 +1,11 @@
-abmnp<-function(#Specified parameters
+abm<-function(#Specified parameters
   
   Runs=3,
   Individuals=100, #number of total resource users in a population
   
   TotalCarryingCapacity=10000, #total available resource units
   
-  StartPercCarryingCapacity = 0.65, #amount of resources available in the landscape at the start in proportion to CC
+  StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
   
   PercProtected=0.15, #percent of the total resource that's in a protected area
   
@@ -182,8 +182,44 @@ abmnp<-function(#Specified parameters
       outputmeanPayoff[t,r]  <- mean(agents$PayoffTotalLastTime)
       
     }
+    par(mfrow=c(2,2))
     
+    #mean payoff
+    plot(rowMeans(outputmeanPayoff), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,harvestMax), lwd = 3, main = "Mean Payoff")
     
+    for (j in 1:Runs) {  
+      
+      lines(outputmeanPayoff[,j], type = 'l')  # add lines for each run, up to r_max
+      
+    }
+    
+    #% CC protected
+    plot(rowMeans(outputmeanTimeWorking), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,1), lwd = 3, main = "Average time in working (coop)")
+    
+    for (j in 1:Runs) {  
+      
+      lines(outputmeanTimeWorking[,j], type = 'l')  # add lines for each run, up to r_max
+      
+    }
+    
+    #% CC protected
+    plot(rowMeans(outputpercCCProtect), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,1), lwd = 3, main = "Percent CC Protected")
+    
+    for (j in 1:Runs) {  
+      
+      lines(outputpercCCProtect[,j], type = 'l')  # add lines for each run, up to r_max
+      
+    }
+    
+    #% CC working
+    plot(rowMeans(outputpercCCWorking), type = 'l', ylab = "y label", xlab = "x label", ylim=c(0,1), lwd = 3, main = "Percent CC Working")
+    
+    for (j in 1:Runs) {  
+      
+      lines(outputpercCCWorking[,j], type = 'l')  # add lines for each run, up to r_max
+      
+    }
[…]

### 2020-06-04 ConceptMapPlot.R

@@ -1,0 +1,73 @@
+dat<-data.frame(var=c(rep("Protected area resources",20),rep("Working landscape resources",20),rep("Mean harvest",20)),Percentage=sort(rbeta(60,1,2)))
+
+
+summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
+                      conf.interval=.95, .drop=TRUE) {
+  library(plyr)
+  
+  # New version of length which can handle NA's: if na.rm==T, don't count them
+  length2 <- function (x, na.rm=FALSE) {
+    if (na.rm) sum(!is.na(x))
+    else       length(x)
+  }
+  
+  # This does the summary. For each group's data frame, return a vector with
+  # N, mean, and sd
+  datac <- ddply(data, groupvars, .drop=.drop,
+                 .fun = function(xx, col) {
+                   c(N    = length2(xx[[col]], na.rm=na.rm),
+                     mean = mean   (xx[[col]], na.rm=na.rm),
+                     sd   = sd     (xx[[col]], na.rm=na.rm)
+                   )
+                 },
+                 measurevar
+  )
+  
+  # Rename the "mean" column    
+  datac <- rename(datac, c("mean" = measurevar))
+  
+  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+  
+  # Confidence interval multiplier for standard error
+  # Calculate t-statistic for confidence interval: 
+  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+  datac$ci <- datac$se * ciMult
+  
+  return(datac)
+}
+
+dat<-summarySE(dat, measurevar="Percentage", groupvars=c("var"))
+
+install.packages("extrafont")
+library(extrafont)
+font_import()
+loadfonts(device="win")       #Register fonts for Windows bitmap output
+fonts()  
+mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+                plot.title = element_text( size=18, color="black",face="bold"),
+                axis.title.x = element_text( size=18),
+                axis.title.y = element_text( size=18),
+                axis.text=(element_text(color="black", size=14)),
+                legend.title = element_text(colour="black", size=18),
[…]

### 2020-06-04 KeyOutputs.R

@@ -1,0 +1,168 @@
+
+
+#using 0.6 as the starting % cc. REsults show that other parameters have greatest effect from 0.5 to 0.7
+
+
+FillData<-expand.grid(PercProtected=seq(0.05,0.95,0.05), ResourceMobility=c(0.01,0.5,0.99))
+FillData$ProtectedCC<-rep(NA,nrow(FillData))
+FillData$WorkingCC<-rep(NA,nrow(FillData))
+FillData$MeanPayoff<-rep(NA,nrow(FillData))
+FillData$meanTimeWorking<-rep(NA,nrow(FillData))
+
+
+for (i in 1:nrow(FillData)){
+  abmnp(Runs=5,Individuals = 100,TimeSteps = 50,harvestMax = 15,StartPercCarryingCapacity = 0.6,
+      PercProtected = FillData[i,]$PercProtected, 
+      ProbOfMobility = FillData[i,]$ResourceMobility)
+  FillData[i,3]<-apply(FullOutput$percCCProtect[50,1:5],1,mean) 
+  FillData[i,4]<-apply(FullOutput$percCCWorking[50,1:5],1,mean)  
+  FillData[i,5]<-apply(FullOutput$meanPayoff[50,1:5],1,mean)  
+  FillData[i,6]<-apply(FullOutput$meanTimeWorking[50,1:5],1,mean)   #[Timesteps,1:Runs]
+  print(nrow(FillData))
+  print(i)
+}
+
+Res1_succ<-FillData
+write.csv(Res1_succ,file="rename1.csv")
+Res1_succ<-read.csv("~/Pemba Project/rename1.csv")
+Res1_succ$ResourceMobility<-ifelse(Res1_succ$ResourceMobility ==0.01, "NonMobile",
+                           ifelse(Res1_succ$ResourceMobility==0.50,"ModerateMobility","HighlyMobile"))
+
+p1<-ggplot(Res1_succ,aes(x=PercProtected,y=ProtectedCC,color=ResourceMobility))+
+  geom_point()+geom_line()+theme_classic()+mytheme
+p2<-ggplot(Res1_succ,aes(x=PercProtected,y=WorkingCC,color=ResourceMobility))+
+  geom_point()+geom_line()+theme_classic()+mytheme
+p3<-ggplot(Res1_succ,aes(x=PercProtected,y=MeanPayoff,color=ResourceMobility))+
[…]

### 2020-06-04 SensitivityAnalyses.R

@@ -1,0 +1,64 @@
+#sensitivity analyses for harvest max
+fillData<-expand.grid(ResourceMobility=seq(0.1,0.9,0.1), harvmax=seq(10,90,10))
+fillData$ProtectedCC<-rep(NA,nrow(fillData))
+fillData$WorkingCC<-rep(NA,nrow(fillData))
+fillData$MeanPayoff<-rep(NA,nrow(fillData))
+fillData$meanTimeWorking<-rep(NA,nrow(fillData))
+
+
+for (i in 1:nrow(fillData)){
+  abmnp(Runs=2,Individuals = 100,TimeSteps = 50,harvestMax = fillData[i,]$harvmax,
+      PercProtected = 0.35,
+      ProbOfMobility = fillData[i,]$ResourceMobility)
+  fillData[i,3]<-apply(FullOutput$percCCProtect[50,1:2],1,mean) 
+  fillData[i,4]<-apply(FullOutput$percCCWorking[50,1:2],1,mean)  
+  fillData[i,5]<-apply(FullOutput$meanPayoff[50,1:2],1,mean)  
+  fillData[i,6]<-apply(FullOutput$meanTimeWorking[50,1:2],1,mean)   #[Timesteps,1:Runs]
+  print(nrow(fillData))
+  print(i)
+}
+
+
+ggplot(data=fillData, mapping=aes(x=ResourceMobility, y=harvmax, fill= ProtectedCC)) + 
+  geom_tile()+scale_fill_viridis()
+ggplot(data=fillData, mapping=aes(x=ResourceMobility, y=harvmax, fill= WorkingCC)) + 
+  geom_tile()+scale_fill_viridis()
+ggplot(data=fillData, mapping=aes(x=ResourceMobility, y=harvmax, fill= MeanPayoff)) + 
+  geom_tile()+scale_fill_viridis()
+ggplot(data=fillData, mapping=aes(x=ResourceMobility, y=harvmax, fill= meanTimeWorking)) + 
+  geom_tile()+scale_fill_viridis()
+
+
+
+#narrow in
+
+
+
+fillData<-expand.grid(ResourceMobility=seq(0.1,0.9,0.2), harvmax=seq(10,18,2))
+fillData$ProtectedCC<-rep(NA,nrow(fillData))
+fillData$WorkingCC<-rep(NA,nrow(fillData))
+fillData$MeanPayoff<-rep(NA,nrow(fillData))
[…]

### 2020-06-04 heatmaps.R

@@ -14,31 +14,37 @@ initOutputSuccess<-read.csv("~/Pemba Project/ModelResultsSuccessBias.csv")
 initOutputConform<-read.csv("~/Pemba Project/ModelResultsConformityBias.csv")
 
 
-ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= ProtectedCC)) + 
+p1<-ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= ProtectedCC)) + 
   geom_tile()+scale_fill_viridis()+ggtitle("Success Bias -- Protected CC")
 
 ggplot(data=initOutputConform, mapping=aes(x=PercProtected, y=ResourceMobility, fill= ProtectedCC)) + 
   geom_tile()+scale_fill_viridis()+ggtitle("Conform Bias -- Protected CC")
 
 
-ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= WorkingCC)) + 
+p2<-ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= WorkingCC)) + 
   geom_tile()+scale_fill_viridis()+ggtitle("Success Bias -- Working CC")
 
 ggplot(data=initOutputConform, mapping=aes(x=PercProtected, y=ResourceMobility, fill= WorkingCC)) + 
   geom_tile()+scale_fill_viridis()+ggtitle("Conform Bias -- Working CC")
 
-ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= MeanPayoff)) + 
+p3<-ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= MeanPayoff)) + 
   geom_tile()+scale_fill_viridis()+ggtitle("Success Bias -- Mean Payoff")
 
 ggplot(data=initOutputConform, mapping=aes(x=PercProtected, y=ResourceMobility, fill= MeanPayoff)) + 
   geom_tile()+scale_fill_viridis()+ggtitle("Conform Bias -- Mean Payoff")
 
-ggplot(data=initOutputSuccess, mapping=aes(x=PercProtected, y=ResourceMobility, fill= meanTimeWorking)) + 
[…]

### 2020-09-29 DissertationProposal/DissertationProposal.Rmd
proposal rmd

--- 
title: Cultural Evolution of Conservation Behavior at the Intersection of Working
  Landscapes and Protected Areas
author: "Matt Clark"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    extra_dependencies: "subfig"
  bookdown::pdf_document2:
    toc: yes
  bookdown::html_document2:
    toc: yes
fontsize: 12pt
link-citations: yes
csl: AmJBot.csl
subtitle: Dissertation Proposal for Committee Review
bibliography: [References/Dissertationproposal.bib]
---
```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)

```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)

## Caching code
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')
```

# Summary
Area-based conservation is critical for stemming global biodiversity loss. As such, each year billions of dollars are spent to protect key areas. It is common however, for these areas to fail to reach their stated conservation goals. The most chronic causes of conservation project failure are the human dimensions. Conversely, historic conservation successes are also attributed to widespread human adoption of conservation behaviors. Research examining how and why conservation practices are adopted has been quite limited; suffering from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes. This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin social adoption of conservation practices in and around protected areas. Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.

1. Develop a dynamic model to generate predictions regarding if and how conservation norms can emerge and persist via competition between groups.
2. Empirically test theoretical predictions of group conservation adoption as a function of group competition and individual costs and benefits. 
3. Using the ongoing spread of an improved efficiency cookstove design, specifically examine how the nature of interpersonal relationships may modify conservation behaviors as they spread within groups.

Globally, we have expended a tremendous amount of resources to identify conservation goals which will deliver the greatest benefits to biodiversity and ecosystem function. For these benefits to come to fruition however, conservation initiatives must be adopted at a rate which exceeds ecological degradation. This research will bring together cultural evolutionary science and ecology to identify the key mechanisms of conservation adoption.

# Introduction
## Phenomenon
Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources. Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources (GFN, 2019). Today, on average, we degenerate these resources at a rate 56% faster than regeneration (Lin et al., 2020). This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining (Beltratti et al., 1993). Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity (Regev et al., 1998).

The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity (Haines-Young & Potschin, 2010, Redford, 1995). As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target (Parrish et al., 2003). Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (UNEP-WCMC & IUCN 2020). These area-based conservation initiatives are from hereon referred to as protected areas.

Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local human populations (@adams_biodiversity_2004) (Leverington et al., 2010). While high-profile protected area successes are salient in the conservation narrative, these areas also commonly fail to reach their intended biodiversity targets. The ubiquity of these failures is highly speculative as unsuccessful protected area initiatives go largely unreported (Bottrill et al., 2011; Sutherland et al., 2004). Despite a high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations (Catalano et al., 2019).

Community adoption of protected area regulations may emerge and persist as a group norm, emerge only for a short period of time, or may never emerge within a given group (Rogers, 2010; Simon et al., 2012). These behaviors are characterized as individually costly, yet beneficial at the group level. Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success (Bennett et al., 2017). This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present (Ostrom, 1990; Richerson & Boyd, 2001). What is less understood however, is the emergence and initial adoption of conservation behaviors.

Research on the diffusion of conservation behaviors suffers from a lack of generalizability. Case studies tend to be highly context dependent and offer little mechanistic insight into the emergence and spread of these behaviors. Excitingly, parallel to the development of conservation social science, the field of evolutionary anthropology has developed the theory of cultural group selection as a mechanistic explanation for the adoption of individually costly, group beneficial behaviors.

## Primer on cultural group selection
The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations (Boyd & Richerson, 2010). Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not. As an individual’s behavior is flexible, individuals will either migrate from less to more successful groups or directly copy the cooperative behavior of others in visibly successful cooperative groups (Henrich, 2004). This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection. In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall (McElreath & Boyd, 2008; Okasha, 2004; Price, 1972).

Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors (Boyd et al., 2011; Lehmann et al., 2008). Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically. Conservation behaviors are fundamentally different from generic cooperative behaviors as they operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates. As such, cultural group selection theory must be extended to include these dynamics in order to be effectively applied to conservation.

Applying cultural group selection to the adoption of conservation behaviors
Protected areas are an ideal microcosm for studying and applying cultural group selection. Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals (Adams et al., 2004). While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice (Naughton-Treves et al., 2005). Agrawal and Redford (2006) argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful. Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection. Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.

## Fidelity of behavioral transmission
The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission. In situations where behaviors do emerge and spread within a population, as the behavior is transmitted along the chain of individuals, it may be modified considerably to the point where the behavioral outcome is unrecognizable (Lewis & Laland, 2012; McGuigan et al., 2011). This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves. Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator (King & Cowlishaw, 2007; Muthukrishna et al., 2016). These insights are crucial for practitioners aiming to implement long-lasting, effective conservation projects, but must first be tested empirically in a conservation context where an individual’s access to a behavioral demonstration is realistic and outcomes of adopted behaviors have real impacts for individuals, communities, and the environment.

## Dissertation objectives
The overarching question this research seeks to answer is: What are the underlying social mechanisms which cause conservation behaviors to emerge and spread efficiently within populations?

-Objective 1: Create a dynamic, agent-based model which extends theories of cultural group selection into the contexts of conservation. Formally test the intuition that between group imitation can lead to the spread of community wide protected area adoption whereas solitary groups may fail to cooperate under identical environmental constraints such as resource degradation, regeneration, and mobility.

-Objective 2: Calculate theoretical adoption probabilities of protected areas known as Community Forest Management Agreements (CoFMAs) for each ward on Pemba Island, Tanzania. Probabilities are based on the salience of group cultural identities and the costs and benefits of CoFMAs to individuals and groups. Compare theoretical predictions to the presence or absence of CoFMAs in reality.

-Objective 3: Quantitatively examine how social and physical access to a behavioral demonstration during behavioral transmission affects both the copying fidelity as well as the performance of copied behavior. This project will make use of the ongoing spread of fuel efficient cookstove technology on Pemba Island, Tanzania.

# Chapter 1
## Background

# Chapter 2

# Chapter 3

### 2020-09-30 DissertationProposal/DissertationProposal.Rmd
tons of edits

@@
 output:
+  bookdown::pdf_document2:
+    toc: yes
   pdf_document:
@@
     extra_dependencies: "subfig"
-  bookdown::pdf_document2:
-    toc: yes
   bookdown::html_document2:
@@
 As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target (Parrish et al., 2003).
-Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (UNEP-WCMC & IUCN 2020).
+Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) (UNEP-WCMC & IUCN 2020).
 These area-based conservation initiatives are from hereon referred to as protected areas.
+
+```{r paNumber, echo=FALSE, fig.cap="Number of newly established  no-take protected areas reported by the IUCN 2020", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/Number.png")
+```
 
@@
 
--Objective 1: Create a dynamic, agent-based model which extends theories of cultural group selection into the contexts of conservation.
+* Objective 1: Create a dynamic, agent-based model which extends theories of cultural group selection into the contexts of conservation.
 Formally test the intuition that between group imitation can lead to the spread of community wide protected area adoption whereas solitary groups may fail to cooperate under identical environmental constraints such as resource degradation, regeneration, and mobility.
 
--Objective 2: Calculate theoretical adoption probabilities of protected areas known as Community Forest Management Agreements (CoFMAs) for each ward on Pemba Island, Tanzania.
+* Objective 2: Calculate theoretical adoption probabilities of protected areas known as Community Forest Management Agreements (CoFMAs) for each ward on Pemba Island, Tanzania.
 Probabilities are based on the salience of group cultural identities and the costs and benefits of CoFMAs to individuals and groups.
@@
 
--Objective 3: Quantitatively examine how social and physical access to a behavioral demonstration during behavioral transmission affects both the copying fidelity as well as the performance of copied behavior.
+* Objective 3: Quantitatively examine how social and physical access to a behavioral demonstration during behavioral transmission affects both the copying fidelity as well as the performance of copied behavior.
 This project will make use of the ongoing spread of fuel efficient cookstove technology on Pemba Island, Tanzania.
 
-# Chapter 1
+# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Area-Based Conservation. 
 ## Background
+Over the last 20-30 years, both the number and size of new no-take protected areas created each year have increased steadily (Figures \@ref(fig:paNumber) & \@ref(fig:paSize)) (UNEP-WCMC & IUCN 2020).
+During this time period, the stated mission of most protected areas has shifted as well.
+Due to the overlap between biodiversity and low income populations, as well as the price and availability of land, protected areas are largely adjacent to, or even contain, low income communities (Fisher & Christopher, 2007; Naughton-Treves et al., 2005).
+With this recognition, their objectives now commonly include community development goals alongside conservation targets.
 
-# Chapter 2
+```{r paSize, echo=FALSE, fig.cap="Size of newly established no-take protected areas reported by the IUCN 2020", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/Size.png")
+```
 
-# Chapter 3
+Joint conservation and community development initiatives are a popular prescription for combating climate change and global biodiversity loss (Tallis et al., 2008).
+These dual-objective projects attempt to mitigate concerns of environmental neocolonialism and resource disenfranchisement of local populations (Hall & Tucker, 2004; Langholz & Lassoie, 2001).
+Importantly, protected areas which explicitly seek to bolster the livelihoods of local populations actually show greater conservation success than those that do not (Fisher & Christopher, 2007).
+Protected areas then represent a considerable opportunity for “win-win” projects which support global conservation and poverty alleviation (Adams et al., 2004).
+
+Protected areas, while well intentioned, often fail to meet both their stated social and ecological goals (Christie, 2004; Watson et al., 2014).
+In fact, a study of protected areas in South Asia found them to be ecologically equivalent to adjacent unprotected lands (Clark et al., 2013).
+It’s unsurprising that the success of protected areas in reaching both social and ecological targets has been linked to the degree of local support and adherence to regulations (Bennett & Dearden, 2014; Christie, 2005).
+
+The degree of local buy-in for protected areas is not random.
+The scale of implementation is highly correlated with the percentage of individuals who adopt protected area regulations.
+A 2019 study demonstrated that nationally implemented initiatives had 84-100% buy-in from local populations, while locally established areas saw adoption from 0.13 - 51% of potential adopters (Mills et al., 2019).
+These findings are particularly worrying given a global decline in government support for land conservation (Watson et al., 2014).
+
+With the recognition that local adoption is a difficult and necessary component of conservation success, human dimensions have become a major focus of conservation research (Bennett et al., 2017).
+This body of work has largely relied on case studies of conservation projects to draw correlations between social phenomena and conservation outcomes.
+As such, findings from conservation social science tend to offer little mechanistic insights for how conservation behaviors spread within and between groups.
+Recently however, theories from cultural group selection have been identified as a potentially useful tool to better elucidate the mechanisms which underpin local buy-in for conservation initiatives (Brooks et al., 2018).
+
+Cultural group selection theorizes that individuals will adopt behaviors which adhere to protected area regulations if the group benefits of doing so allow their cultural group to outcompete other groups (Rindos et al., 1985).
+Under this framework, multiple competing groups are actually required for community adoption of protected area regulations to emerge (Richerson et al., 2016).
+
+While the theory and principles of cultural group selection are relatively well established, they are yet to be tested mechanistically in the context of protected area adoption.
+With this in mind, this project aims to demonstrate how the presence of multiple competing groups can facilitate the adoption of adherence to protected areas.
+In this chapter I will develop a dynamic, agent-based model to test the logic of how protected area norms can emerge under very simple sets of rules in the presence of multiple competing groups, but fail to emerge for solitary groups under identical environmental contexts.
+
+### Question:
+Can between group competition provoke protected area adoption where solitary groups may fail to adopt?
+
+### Predictions:
+* P1: Adoption of protected area adherence will only emerge under the presence of multiple competing groups.
+
+* P2: Where it already exists, protected area adherence will perpetuate in populations even in solitary groups given group conformity.
+
+* P3: Greater degrees of resource mobility will increase the frequency in which the adoption of protected area adherence emerges in the community(ies).
+
+## Research Design
+
+An agent-based modeling framework is ideal for demonstrating the theory of cultural group selection regarding protected area norms, because it allows such group level properties to emerge out of simple decision rules for each individual (Railsback & Grimm, 2012; Wilensky & Rand, 2015).
+This design mirrors our real-world understanding of the important system components.
+
+### Model Description
+
+```{r conceptDiagram, echo=FALSE, fig.cap="Concept diagram of agent-based model scheduling under single group dynamics", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/ConceptDiagram.png")
+```
+
+## Preliminary Work
+
+```{r abmOutput, echo=FALSE, fig.cap="Fifteen example runs of the current, single-group agent based model", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/ABM_Output.png")
+```
+
+## Expected Products
+I plan to communicate the findings of this project both to the academic community and to non-scientists.
+The insights gained from this model will be published in a peer reviewed, academic journal such as Environmental Modeling and Software.
+This paper will show logically how we expect that protected areas function in the context of community level cooperative conservation.
+An example of how this model may inform implementation of protected areas for various resource mobilities or starting resource degradation is displayed in figure (XXX).
+
+To make this model readily available to a general audience however, I will update the online application provided above to better reflect the final version of the model.
+In this application I will also embed a graphical display of the model function as a .gif that displays key outputs at each timestep of the model.
+This will allow users to better internalize the implications of the model and to better apply them to real world conservation.
+
+# Chapter 2: Effects of inter-group resource competition and selection on community adoption of costly conservation policy.
+
+## Background
+Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection (Molleman et al., 2013).
+One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use (Waring et al., 2017).
+Under this prediction, selection acts both on individuals and on groups of individuals.
+When the ratio of group benefit to individual cost of cooperation outweighs the individual fitness consequences of cooperation, those behaviors will be favored by selection.
+Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups (Boyd & Richerson, 2010; Henrich, 2004).
+
+While these predictions are theoretically sound, they must be tested empirically.
+Two such studies have begun to test the basis of cultural group selection in the natural world.
+The first demonstrated that culture offers greater fodder for group selection than do genetics.
+Researchers estimated the total variance in both cultural and genetic traits within and between groups of individuals.
+The proportion of variance in an allele or trait found between groups is referred to as genetic Fst or cultural Fst respectively.
+Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits (Bell et al., 2009).
+
+A second empirical study tested the prediction that large scale human cooperation is in fact a product of competition between cultural groups.
+Researchers created pair-wise estimates of cultural Fst between 12 ethnic groups.
+These estimates of cultural differentiation were combined with a survey of individual willingness to cooperate with members of the other groups.
+Results of this study showed a strong association between cultural similarity and willingness to cooperate.
+This research indicates that the presence of variable cultural groups influences within-group cooperative behavior (Handley & Mathew, 2020).
+
+This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly.
+Given the theoretical capacity for cultural group selection to provide solutions to pressing environmental issues of collective action, these predictions must be validated experimentally so that they might be implemented effectively.
+
+Conservation via protected areas represents an ideal study system for testing the environmental applications of cultural group selection in the real world.
+Ignoring the regional to global scale ecosystem service benefits of protected areas, the units of the individual costs and group benefits of protected area adoption are directly comparable (i.e.
+fuelwood growth/use).
+This comparability will allow us to use established methods of measuring multilevel selection pressures to quantitatively predict if protected areas are likely to be selected for in each group.
+Finally, this project will compare the calculated selection pressures for protected area adoption from each group against the observed adoption of community managed protected areas.
+
+## Research Design
+
+## Preliminary Work
+This project benefits from preliminary work done by our collaborators at the Max Planck Institute for Evolutionary Anthropology (MPI).
+MPI researchers conducted a preliminary qualitative survey and coalition of oral histories with the assistance of Department of Forestry staff to confirm anecdotally that competition between wards has led to CoFMA adoption at the ward level.
+These researchers have also visited wards across the island and recorded which wards have adopted CoFMAs and which have not.
+These adoption characteristics will be compared against the adoption characteristics recorded in my work as a form of data validation.
+
+## Expected Outputs
+The primary outputs of data collection and analysis in this project will produce binary predictions of protected area adoption for each study ward, as well as the calculated strength of selection for protected area adoption for each study ward.
+I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of community managed protected areas.
+(Figure \@ref(fig:aelPress)) shows an example of what the results of this study might look like.
+The findings of this study will be submitted for publication in a peer reviewed, academic journal.
+
+```{r selPress, echo=FALSE, fig.cap="Simulated data showing protected area adoption vs calculated selection pressure for group adoption of protected areas.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch3/SelPressureAdopt.png")
+```
+
+# Chapter 3: Social influences on transmission fidelity in household level diffusion of beneficial technology.
+
+## Background
+Forests sequester 2 billion tons of atmospheric CO2 each year, while also providing fuel to 40% of the world’s rural poor — fuelwood produces more renewable energy than solar, wind, and hydroelectric sources combined (SOFO, 2018).
+Household fuel production and global carbon sequestration are often seen as conflicting forest benefits.
+Solutions to mitigate this tradeoff include expanding markets for non-timber forest products and technological advances like solar panels (McShane et al., 2011).
+Given limited funding, these projects are not intended to solve problems in their entirety, but rather to seed innovations that catalyze broad change.
+Many of these initiatives, however, fail to spread and produce long-term benefits for local livelihoods or forest conservation because they do not account for underlying social dynamics (Keppel et al., 2012; Oates, 1999; Stringer & Reed, 2007).
+
+This project will examine the spread of fuel-efficient cookstoves on Pemba Island, Tanzania (see Figure \@ref(fig:stovePic)).
+An improved stove design which increases fuel efficiency by up to 50%, was seeded in three communities on the island in 2015 (Lung & Espira, 2019; Sakthivadivel & Iniyan, 2018).
+The improved design has been subsequently copied throughout the communities and is thought to reduce deforestation and the time individuals — primarily women and children — must spend harvesting fuelwood.
+
+```{r stovePic, echo=FALSE, fig.cap="Example of an improved, fuel efficient cookstove on Pemba Island, Tanzania.
+Image from Same Mwanga Environmental Conservation Advisory Organization", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch3/cookstove.jpg")
+```
+
+While this project will inform effective implementation of household level conservation projects on the ground, it will also provide important insights for understanding behavioral adoption more generally.
+While the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially (Lyons et al., 2007; McGuigan et al., 2011).
+In a foundational study, researchers show that learned behavior is reconstructed and assimilated with an individual’s prior experiences, rather than replicated directly (Bartlett, 1932).
+Behavioral reconstruction creates variation in copied behaviors (Horner et al., 2006).
+Replication variation is thought to be greatly influenced by environmental factors, including social context (Mesoudi & Whiten, 2008).
+
+Both theoretical and experimental studies have examined the relationship between transmission fidelity and environmental influences.
+Theoretical work suggests that increasing transmission fidelity should be associated with greater reliance on social learning (Boyd & Henrich, 2002; Nakahashi et al., 2012).
+This theory has also been validated experimentally (Muthukrishna et al., 2016).
+When coupled with the theoretical prediction that social learning should increase with greater access to behavioral demonstrators, it follows logically that transmission fidelity should increase with imitator access to the behavioral model (King & Cowlishaw, 2007).
+
+While transmission fidelity can vary significantly given the context, it is often unclear what effect this variation has on the outcomes of imitated behavior.
+Theoretical models have shown that even small changes in transmission fidelity can have strong impacts on population-level behavioral outcomes (Lewis & Laland, 2012).
+Alternatively, some behaviors require only very low-fidelity replication to achieve identical outcomes (Enquist et al., 2010).
+In fact, it is likely that if a behavior is under strong environmental constraints then individual learning should help correct errors in the transmission process.
+
+Currently, we do not yet understand how various forms of access to behavioral demonstration may affect transmission fidelity.
+Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes (Mesoudi & Whiten, 2008).
+This research will directly measure the impact of geographic, genetic, and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.
+
+In summary, conserving forest cover is critical for the global carbon balance and for the well-being of many rural communities.
+Conservation projects often fail due to a lack of consideration of the social processes underlying adoption.
+With few studies that measure transmission fidelity in a naturalistic setting, where outcomes have real economic and health impacts - we have an opportunity to explore how environmental constraints and social learning interact when outcomes matter.
+This project seeks to further investigate how and why behaviors are modified during the transmission process and how modification (both enrounious and intentional) affect performance targets.
+
+### Question
+What effect do geographic distance, genetic relatedness, and social relatedness have on transmission fidelity? Secondly, what is the relationship between transmission fidelity and behavioral outcomes?
+
+### Predictions
+* P1: Using geographic distance, genetic relatedness, and social network distance as proxies for access to behavioral demonstrators, we expect transmission fidelity to increase with increasing access.
+
+* P2: As access to behavioral demonstration decreases, we expect to see differential changes in the fidelity of replicated behavior and the behavioral performance target of interest.
+
+* P3: Given the high degree of environmental influence on a behavior, we expect to observe a greater change in the fidelity of transmission than in the behavioral performance target, correlative with decreasing access to behavioral demonstrators.
+
+## Research Design
+
+### Approach
+This project will take advantage of an ongoing initiative to spread fuel-efficient cookstoves on Pemba Island, Tanzania.
+This initiative has funded the building of a small number of improved cookstoves across communities which have subsequently been voluntarily copied by other community members.
+To answer the research questions above, I will collect data on the visual similarity of all copied cookstoves and the models they were copied from to assess direct transmission fidelity.
+I will also measure the fuel efficiency of the cookstoves as the behavioral outcomes of interest.
+
+### Methodological Specifics
+I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig XXX).
+With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.
+
+```{r mapp, echo=FALSE, fig.cap="Satellite map of Pemba Island, Tanzania.
+Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch3/BadPembaMap.png")
+```
+
+The survey instrument used for this research will elicit information regarding the relationship between the builder of each cookstove and the individual they copied their stove design from.
+Specifically, this survey tool will be used to gather data on geographic distance, genetic relatedness, and social relatedness.
+I will also use this survey to collect data on the decision making processes individuals exhibit when deciding to build a fuel-efficient cookstove.
+
+While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics such as the exterior dimensions and size of the openings.
+I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from.
+To measure stove fuel efficiency, I will follow the standard water boiling test protocol (DeFoort et al., 2009; McCracken & Smith, 1998).
+Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test.
+
+The key research questions for this project will be assessed using multilevel regression operationalized in a Bayesian context.
+I will compare standardized parameter estimates, to determine which factors (genetic relatedness, geographic distance, social relatedness) most strongly contribute to copying fidelity in this study system.
+Similar analyses will be used to determine which factors affect cookstove primary outcome variables (fuel efficiency and particulate matter production).
+
+This project will contribute to scientific reproducibility in multiple ways.
+First, this work will be preregistered with the Open Science Framework.
+To confirm the analytical plan in advance of preregistration, I will fit the statistical models proposed above to simulated data and verify that the models recover the parameters known to have been generated in the simulated data.
+Secondarily, once all data for this project has been collected and the analyses are complete, all non-confidential data and R code used for this project will be made publically available through a GitHub repository shared by all team members.
+
+## Preliminary Work
+In preparation for a cancelled field season in the Summer of 2020, I created a variety of forms and documents to facilitate data collection. These include:
+* A household livelihood and stove adoption survey.  
+* A datasheet for cookstove fuel efficiency measurements.
+* A fuel efficiency measurement protocol. 
+I also simulated data that matches the data would be collected by the multiple instruments.
+These data were used to test the ability of my proposed statistical methods to recover the signal of demonstrator access on behavioral transmission fidelity and behavioral outcomes.
+
+## Expected Outputs
+The regression outcomes from this analysis will be the primary result presented in the publication of this research.
+Figures  \@ref(fig:csp1) and  \@ref(fig:csp2) show significant and insignificant theoretical outcomes that we might observe.
+The findings from this paper will also be shared broadly as a white paper describing how the insights from this research can help shape future conservation and development initiatives which distribute fuel efficient cookstoves in rural areas.
+
+```{r csp1, echo=FALSE, fig.cap="Simulated outputs showing a possible relationship between stove copying fidelity and weekly communication between the demonstrator and copier", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch3/CSP1.png")
+```
+
+```{r csp2, echo=FALSE, fig.cap="Simulated outputs showing a possible relationship between stove copying fidelity and geographic distance between the demonstrator and copier", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch3/CSP2.png")
+```
+# Conclusion
+
+# References

### 2020-09-30 DissertationProposal/Figures/Ch1/Ch1_Plots.R
figures

@@ -1,0 +1,60 @@
+x<-read.csv("~/Pemba_Project/WDPA_Jul2020-csv.csv")
+#let's cut down the data so that it's faster to work with
+dat<-x[,c(12,13,15,17,18,20)]
+dat$MARINE<-as.character(dat$MARINE)
+dat$MARINE<-dplyr::recode(dat$MARINE, "0" = "Terrestrial", 
+                          "1" = "Coastal","2"="Marine", .default = NA_character_)
+
+
+
+library(tidyverse)
+library(scales)
+library(extrafont)
+
+p2<-dat%>%filter(NO_TK_AREA>0)%>%filter(STATUS_YR>1970)%>%group_by(STATUS_YR)%>%
+  mutate(avgSize=mean(NO_TK_AREA))%>%
+  ggplot(.,aes(x=STATUS_YR,y=avgSize))+
+  geom_jitter(shape=21,size=1.5,stroke=1.5,fill="#a6bddb",
+              color="#0570b0",alpha=0.75,width=1.5,height=3000)+
+  geom_smooth(se=F,color="black",size=2)+
+  theme_classic()+
+  scale_y_continuous(label=comma)+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        plot.title = element_text( size=14, color="black",face="bold"),
+        axis.title.x = element_text( size=12),
+        axis.title.y = element_text( size=12),
+        axis.text=(element_text(color="black", size=12)),text=element_text(family="Calibri"),
+        legend.title = element_text(colour="black", size=12),
+        legend.text = element_text( size = 12),
+        legend.position = "none")+
+  xlab("Year Established")+ylab("No-Take Conservation Area (Km2)")+ggtitle("New No-Take Area Size")
+
+p1<-dat%>%filter(NO_TAKE =="All")%>%filter(STATUS_YR>1970)%>%group_by(STATUS_YR)%>%
+  count()%>%
+  ggplot(.,aes(x=STATUS_YR,y=n))+
+  geom_point(shape=21,size=1.5,stroke=1.5,fill="#a6bddb",
+             color="#0570b0",alpha=0.75)+
+  geom_smooth(se=F,color="black",size=2,method="lm")+
+  theme_classic()+
+  scale_y_continuous(label=comma)+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        plot.title = element_text( size=14, color="black",face="bold"),
+        axis.title.x = element_text( size=12),
[…]

### 2020-09-30 DissertationProposal/Figures/Ch3/Ch3_plots.R

@@ -1,0 +1,78 @@
+
+
+SP<-rnorm(100,0,3)
+xb<-0+1.5*SP
+p<-1/(1+exp(-xb))
+y <- rbinom(n = 100, size = 1, prob = p)
+
+data<-data.frame(COF=y,SelPres=SP)
+
+p<-ggplot(data,aes(x=SelPres,y=COF))+
+  geom_point(shape=21,size=1.5,stroke=1.5,fill="#a6bddb",
+             color="#0570b0",alpha=0.75)+
+  geom_smooth(se=F,color="black",size=2)+
+  theme_classic()+
+  scale_y_continuous(breaks=c(0,1))+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        plot.title = element_text( size=14, color="black",face="bold"),
+        axis.title.x = element_text( size=12),
+        axis.title.y = element_text( size=12),
+        axis.text=(element_text(color="black", size=12)),
+        legend.title = element_text(colour="black", size=12),
+        legend.text = element_text( size = 12),
+        legend.position = "none")+
+  xlab("Calculated Selection Pressure for \n Costly Conservation Adoption")+
+  ylab("Protected Area Observed \n(Y/N)")+ggtitle("Simulated Selection Pressure vs.\nProtected Area Adoption")
+
+png(filename = "DissertationProposal/Figures/Ch3/SelPressureAdopt.png", pointsize=10, width=900, height=480, res=200)
+p
+dev.off()
+
+
+data<-data.frame(CF=sort(rbeta(100,1,2)),WC=sort(rpois(100,7)),GD=rpois(100,5))
+
+CSP1<-ggplot(data,aes(x=WC,y=CF))+
+  geom_point(shape=21,size=1.5,stroke=1.5,fill="#a6bddb",
+             color="#0570b0",alpha=0.75)+
+  geom_smooth(se=F,color="black",size=2,method="lm")+
+  theme_classic()+
+  scale_y_continuous(label=comma)+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        plot.title = element_text( size=14, color="black",face="bold"),
+        axis.title.x = element_text( size=12),
+        axis.title.y = element_text( size=12),
+        axis.text=(element_text(color="black", size=12)),
+        legend.title = element_text(colour="black", size=12),
+        legend.text = element_text( size = 12),
+        legend.position = "none")+
[…]

### 2020-09-30 DissertationProposal/DissertationProposal.Rmd
figuring out refs

@@
 subtitle: Dissertation Proposal for Committee Review
-bibliography: [References/Dissertationproposal.bib]
+bibliography: [References/Dissertationproposal.bib,References/NonZoteroRefs.bib]
 ---
@@
 While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice (Naughton-Treves et al., 2005).
-Agrawal and Redford (2006) argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful.
+Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful.
 Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection.

### 2020-09-30 DissertationProposal/Figures/Ch1/Ch1_Plots.R

@@ -58,3 +58,7 @@ p2
 dev.off()
 
 
+
+
+
+

### 2020-09-30 DissertationProposal/DissertationProposal.Rmd
finished cites for now

@@
 Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources.
-Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources (GFN, 2019).
-Today, on average, we degenerate these resources at a rate 56% faster than regeneration (Lin et al., 2020).
-This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining (Beltratti et al., 1993).
-Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity (Regev et al., 1998).
+Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources [@network2019global].
+Today, on average, we degenerate these resources at a rate 56% faster than regeneration [@network2020global].
+This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining [@beltratti_sustainable_1993].
+Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity [@regev_biological_1998].
 
-The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity (Haines-Young & Potschin, 2010, Redford, 1995).
-As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target (Parrish et al., 2003).
-Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) (UNEP-WCMC & IUCN 2020).
+The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity [@haines2010links; @redford1995human].
+As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target [@parrish_are_2003].
+Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) [@iucn2020protectedplanet].
 These area-based conservation initiatives are from hereon referred to as protected areas.
@@
 
-Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local human populations (@adams_biodiversity_2004) (Leverington et al., 2010).
+Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local human populations [@adams_biodiversity_2004; @leverington_global_2010].
 While high-profile protected area successes are salient in the conservation narrative, these areas also commonly fail to reach their intended biodiversity targets.
-The ubiquity of these failures is highly speculative as unsuccessful protected area initiatives go largely unreported (Bottrill et al., 2011; Sutherland et al., 2004).
-Despite a high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations (Catalano et al., 2019).
+The ubiquity of these failures is highly speculative as unsuccessful protected area initiatives go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004].
+Despite a high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].
 
-Community adoption of protected area regulations may emerge and persist as a group norm, emerge only for a short period of time, or may never emerge within a given group (Rogers, 2010; Simon et al., 2012).
+Community adoption of protected area regulations may emerge and persist as a group norm, emerge only for a short period of time, or may never emerge within a given group [@rogers_diffusion_2010; @simon_win-win_2012].
 These behaviors are characterized as individually costly, yet beneficial at the group level.
-Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success (Bennett et al., 2017).
-This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present (Ostrom, 1990; Richerson & Boyd, 2001).
+Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017].
+This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present [@ostrom_governing_1990; @richerson2001evolution].
 What is less understood however, is the emergence and initial adoption of conservation behaviors.
@@
 ## Primer on cultural group selection
-The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations (Boyd & Richerson, 2010).
+The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations [@boyd_transmission_2010].
 Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not.
-As an individual’s behavior is flexible, individuals will either migrate from less to more successful groups or directly copy the cooperative behavior of others in visibly successful cooperative groups (Henrich, 2004).
+As an individual’s behavior is flexible, individuals will either migrate from less to more successful groups or directly copy the cooperative behavior of others in visibly successful cooperative groups [@henrich_cultural_2004].
 This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection.
-In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall (McElreath & Boyd, 2008; Okasha, 2004; Price, 1972).
+In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall [@mcelreath_mathematical_2008; @okasha_multilevel_2004; @price1972extension].
 
-Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors (Boyd et al., 2011; Lehmann et al., 2008).
+Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors [@boyd_rapid_2011; @lehmann_cultural_2008].
 Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically.
@@
 Protected areas are an ideal microcosm for studying and applying cultural group selection.
-Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals (Adams et al., 2004).
-While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice (Naughton-Treves et al., 2005).
+Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals [@adams_biodiversity_2004].
+While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice [@naughton-treves_role_2005].
 Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful.
@@
 The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission.
-In situations where behaviors do emerge and spread within a population, as the behavior is transmitted along the chain of individuals, it may be modified considerably to the point where the behavioral outcome is unrecognizable (Lewis & Laland, 2012; McGuigan et al., 2011).
+In situations where behaviors do emerge and spread within a population, as the behavior is transmitted along the chain of individuals, it may be modified considerably to the point where the behavioral outcome is unrecognizable [@lewis_transmission_2012; @mcguigan_over-imitation_2011].
 This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves.
-Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator (King & Cowlishaw, 2007; Muthukrishna et al., 2016).
+Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator [@king2007use; @muthukrishna_when_2016].
 These insights are crucial for practitioners aiming to implement long-lasting, effective conservation projects, but must first be tested empirically in a conservation context where an individual’s access to a behavioral demonstration is realistic and outcomes of adopted behaviors have real impacts for individuals, communities, and the environment.
@@
 ## Background
-Over the last 20-30 years, both the number and size of new no-take protected areas created each year have increased steadily (Figures \@ref(fig:paNumber) & \@ref(fig:paSize)) (UNEP-WCMC & IUCN 2020).
+Over the last 20-30 years, both the number and size of new no-take protected areas created each year have increased steadily (Figures \@ref(fig:paNumber) & \@ref(fig:paSize)) [@iucn2020protectedplanet].
 During this time period, the stated mission of most protected areas has shifted as well.
-Due to the overlap between biodiversity and low income populations, as well as the price and availability of land, protected areas are largely adjacent to, or even contain, low income communities (Fisher & Christopher, 2007; Naughton-Treves et al., 2005).
+Due to the overlap between biodiversity and low income populations, as well as the price and availability of land, protected areas are largely adjacent to, or even contain, low income communities [@fisher_poverty_2007; @naughton-treves_role_2005].
 With this recognition, their objectives now commonly include community development goals alongside conservation targets.
@@
 
-Joint conservation and community development initiatives are a popular prescription for combating climate change and global biodiversity loss (Tallis et al., 2008).
-These dual-objective projects attempt to mitigate concerns of environmental neocolonialism and resource disenfranchisement of local populations (Hall & Tucker, 2004; Langholz & Lassoie, 2001).
-Importantly, protected areas which explicitly seek to bolster the livelihoods of local populations actually show greater conservation success than those that do not (Fisher & Christopher, 2007).
-Protected areas then represent a considerable opportunity for “win-win” projects which support global conservation and poverty alleviation (Adams et al., 2004).
+Joint conservation and community development initiatives are a popular prescription for combating climate change and global biodiversity loss [@tallis_ecosystem_2008].
+These dual-objective projects attempt to mitigate concerns of environmental neocolonialism and resource disenfranchisement of local populations [@hall_tourism_2004; @langholz_perils_2001].
+Importantly, protected areas which explicitly seek to bolster the livelihoods of local populations actually show greater conservation success than those that do not [@fisher_poverty_2007].
+Protected areas then represent a considerable opportunity for “win-win” projects which support global conservation and poverty alleviation [@adams_biodiversity_2004].
 
-Protected areas, while well intentioned, often fail to meet both their stated social and ecological goals (Christie, 2004; Watson et al., 2014).
-In fact, a study of protected areas in South Asia found them to be ecologically equivalent to adjacent unprotected lands (Clark et al., 2013).
-It’s unsurprising that the success of protected areas in reaching both social and ecological targets has been linked to the degree of local support and adherence to regulations (Bennett & Dearden, 2014; Christie, 2005).
+Protected areas, while well intentioned, often fail to meet both their stated social and ecological goals [@christie_marine_2004; @watson_performance_2014].
+In fact, a study of protected areas in South Asia found them to be ecologically equivalent to adjacent unprotected lands [@clark_protected_2013].
+It’s unsurprising that the success of protected areas in reaching both social and ecological targets has been linked to the degree of local support and adherence to regulations [@bennett_why_2014; @christie_observed_2005].
 
@@
 The scale of implementation is highly correlated with the percentage of individuals who adopt protected area regulations.
-A 2019 study demonstrated that nationally implemented initiatives had 84-100% buy-in from local populations, while locally established areas saw adoption from 0.13 - 51% of potential adopters (Mills et al., 2019).
-These findings are particularly worrying given a global decline in government support for land conservation (Watson et al., 2014).
+A 2019 study demonstrated that nationally implemented initiatives had 84-100% buy-in from local populations, while locally established areas saw adoption from 0.13 - 51% of potential adopters [@mills_how_2019].
+These findings are particularly worrying given a global decline in government support for land conservation [@watson_performance_2014].
 
-With the recognition that local adoption is a difficult and necessary component of conservation success, human dimensions have become a major focus of conservation research (Bennett et al., 2017).
+With the recognition that local adoption is a difficult and necessary component of conservation success, human dimensions have become a major focus of conservation research [@bennett_conservation_2017].
 This body of work has largely relied on case studies of conservation projects to draw correlations between social phenomena and conservation outcomes.
 As such, findings from conservation social science tend to offer little mechanistic insights for how conservation behaviors spread within and between groups.
-Recently however, theories from cultural group selection have been identified as a potentially useful tool to better elucidate the mechanisms which underpin local buy-in for conservation initiatives (Brooks et al., 2018).
+Recently however, theories from cultural group selection have been identified as a potentially useful tool to better elucidate the mechanisms which underpin local buy-in for conservation initiatives [@brooks_applying_2018].
 
-Cultural group selection theorizes that individuals will adopt behaviors which adhere to protected area regulations if the group benefits of doing so allow their cultural group to outcompete other groups (Rindos et al., 1985).
-Under this framework, multiple competing groups are actually required for community adoption of protected area regulations to emerge (Richerson et al., 2016).
+Cultural group selection theorizes that individuals will adopt behaviors which adhere to protected area regulations if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985].
+Under this framework, multiple competing groups are actually required for community adoption of protected area regulations to emerge [@richerson_cultural_2016].
 
@@
 ## Background
-Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection (Molleman et al., 2013).
-One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use (Waring et al., 2017).
+Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection [@molleman_cultural_2013].
+One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use [@waring_coevolution_2017].
 Under this prediction, selection acts both on individuals and on groups of individuals.
 When the ratio of group benefit to individual cost of cooperation outweighs the individual fitness consequences of cooperation, those behaviors will be favored by selection.
-Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups (Boyd & Richerson, 2010; Henrich, 2004).
+Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups [@boyd_transmission_2010; @henrich_cultural_2004].
 
@@
 The proportion of variance in an allele or trait found between groups is referred to as genetic Fst or cultural Fst respectively.
-Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits (Bell et al., 2009).
+Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits [@bell_culture_2009].
 
@@
 Results of this study showed a strong association between cultural similarity and willingness to cooperate.
-This research indicates that the presence of variable cultural groups influences within-group cooperative behavior (Handley & Mathew, 2020).
+This research indicates that the presence of variable cultural groups influences within-group cooperative behavior [@handley_human_2020].
 
@@
 I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of community managed protected areas.
-(Figure \@ref(fig:aelPress)) shows an example of what the results of this study might look like.
+(Figure \@ref(fig:selPress)) shows an example of what the results of this study might look like.
 The findings of this study will be submitted for publication in a peer reviewed, academic journal.
@@
 ## Background
-Forests sequester 2 billion tons of atmospheric CO2 each year, while also providing fuel to 40% of the world’s rural poor — fuelwood produces more renewable energy than solar, wind, and hydroelectric sources combined (SOFO, 2018).
+Forests sequester 2 billion tons of atmospheric CO2 each year, while also providing fuel to 40% of the world’s rural poor — fuelwood produces more renewable energy than solar, wind, and hydroelectric sources combined [@muller2018state].
 Household fuel production and global carbon sequestration are often seen as conflicting forest benefits.
-Solutions to mitigate this tradeoff include expanding markets for non-timber forest products and technological advances like solar panels (McShane et al., 2011).
+Solutions to mitigate this tradeoff include expanding markets for non-timber forest products and technological advances like solar panels [@mcshane_hard_2011].
 Given limited funding, these projects are not intended to solve problems in their entirety, but rather to seed innovations that catalyze broad change.
-Many of these initiatives, however, fail to spread and produce long-term benefits for local livelihoods or forest conservation because they do not account for underlying social dynamics (Keppel et al., 2012; Oates, 1999; Stringer & Reed, 2007).
+Many of these initiatives, however, fail to spread and produce long-term benefits for local livelihoods or forest conservation because they do not account for underlying social dynamics [@keppel_conservation_2012; @oates_myth_1999; @stringer_land_2007].
 
 This project will examine the spread of fuel-efficient cookstoves on Pemba Island, Tanzania (see Figure \@ref(fig:stovePic)).
-An improved stove design which increases fuel efficiency by up to 50%, was seeded in three communities on the island in 2015 (Lung & Espira, 2019; Sakthivadivel & Iniyan, 2018).
+An improved stove design which increases fuel efficiency by up to 50%, was seeded in three communities on the island in 2015 [@lung_large-scale_2019; @sakthivadivel2018experimental].
 The improved design has been subsequently copied throughout the communities and is thought to reduce deforestation and the time individuals — primarily women and children — must spend harvesting fuelwood.
@@
 While this project will inform effective implementation of household level conservation projects on the ground, it will also provide important insights for understanding behavioral adoption more generally.
-While the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially (Lyons et al., 2007; McGuigan et al., 2011).
-In a foundational study, researchers show that learned behavior is reconstructed and assimilated with an individual’s prior experiences, rather than replicated directly (Bartlett, 1932).
-Behavioral reconstruction creates variation in copied behaviors (Horner et al., 2006).
-Replication variation is thought to be greatly influenced by environmental factors, including social context (Mesoudi & Whiten, 2008).
+While the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially [@lyons_hidden_2007; @mcguigan_over-imitation_2011].
+In a foundational study, researchers show that learned behavior is reconstructed and assimilated with an individual’s prior experiences, rather than replicated directly [@bartlett1932remembering].
+Behavioral reconstruction creates variation in copied behaviors [@horner_faithful_2006].
+Replication variation is thought to be greatly influenced by environmental factors, including social context [@mesoudi_multiple_2008].
 
 Both theoretical and experimental studies have examined the relationship between transmission fidelity and environmental influences.
-Theoretical work suggests that increasing transmission fidelity should be associated with greater reliance on social learning (Boyd & Henrich, 2002; Nakahashi et al., 2012).
-This theory has also been validated experimentally (Muthukrishna et al., 2016).
-When coupled with the theoretical prediction that social learning should increase with greater access to behavioral demonstrators, it follows logically that transmission fidelity should increase with imitator access to the behavioral model (King & Cowlishaw, 2007).
+Theoretical work suggests that increasing transmission fidelity should be associated with greater reliance on social learning [@boyd_modeling_2002; @nakahashi_adaptive_2012].
+This theory has also been validated experimentally [@muthukrishna_when_2016].
+When coupled with the theoretical prediction that social learning should increase with greater access to behavioral demonstrators, it follows logically that transmission fidelity should increase with imitator access to the behavioral model [@king_when_2007].
 
 While transmission fidelity can vary significantly given the context, it is often unclear what effect this variation has on the outcomes of imitated behavior.
-Theoretical models have shown that even small changes in transmission fidelity can have strong impacts on population-level behavioral outcomes (Lewis & Laland, 2012).
-Alternatively, some behaviors require only very low-fidelity replication to achieve identical outcomes (Enquist et al., 2010).
+Theoretical models have shown that even small changes in transmission fidelity can have strong impacts on population-level behavioral outcomes [@lewis_transmission_2012].
+Alternatively, some behaviors require only very low-fidelity replication to achieve identical outcomes [@enquist_one_2010].
 In fact, it is likely that if a behavior is under strong environmental constraints then individual learning should help correct errors in the transmission process.
@@
 Currently, we do not yet understand how various forms of access to behavioral demonstration may affect transmission fidelity.
-Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes (Mesoudi & Whiten, 2008).
+Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes [@mesoudi_multiple_2008].
 This research will directly measure the impact of geographic, genetic, and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.
@@
 ### Methodological Specifics
-I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig XXX).
+I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)).
 With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.
@@
 I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from.
-To measure stove fuel efficiency, I will follow the standard water boiling test protocol (DeFoort et al., 2009; McCracken & Smith, 1998).
+To measure stove fuel efficiency, I will follow the standard water boiling test protocol [@defoort2009stove; @mccracken1998emissions].
 Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test.
@@
 In preparation for a cancelled field season in the Summer of 2020, I created a variety of forms and documents to facilitate data collection. These include:
+
 * A household livelihood and stove adoption survey.
+
 * A datasheet for cookstove fuel efficiency measurements.
+
 * A fuel efficiency measurement protocol.
+
 I also simulated data that matches the data would be collected by the multiple instruments.

### 2020-09-30 DissertationProposal/DissertationProposal.Rmd
Figures, equations, and plots. Mostly equations

@@
 
-```{r abmOutput, echo=FALSE, fig.cap="Fifteen example runs of the current, single-group agent based model", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+```{r abmOutput, echo=FALSE, fig.cap="Fifteen example runs of the current, single group agent-based model", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
 knitr::include_graphics("Figures/Ch1/ABM_Output.png")
@@
 This paper will show logically how we expect that protected areas function in the context of community level cooperative conservation.
-An example of how this model may inform implementation of protected areas for various resource mobilities or starting resource degradation is displayed in figure (XXX).
+An example of how this model may inform implementation of protected areas for various resource mobilities or starting resource degradation is displayed in figures \@ref(fig:abmSim1) & \@ref(fig:abmSim2).
 
@@
 
+```{r abmSim1, echo=FALSE, fig.cap="Simulated agent-based model output of conservation adoption under single group dynamics.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/abmSim1.png")
+```
+
+```{r abmSim2, echo=FALSE, fig.cap="Simulated agent-based model output of conservation adoption under multiple group dynamics.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/abmSim2.png")
+```
 # Chapter 2: Effects of inter-group resource competition and selection on community adoption of costly conservation policy.
@@
 ## Research Design
+
+Inline LaTeX equations can be written in a pair of dollar signs using the LaTeX syntax, e.g., $f(k) = {n \choose k} p^{k} (1-p)^{n-k}$  math expressions of the display style can be written in a pair of double dollar signs, e.g.,
+
+$$f(k) = {n \choose k} p^{k} (1-p)^{n-k}$$, and the output looks like this:
+
+Cultural fst equation
+
+$$F_{STx} = \frac{\frac{n_{i}}{n_{i}+n_{j}} (p_{i}-\bar{p})^2 + \frac{n_{j}}{n_{i}+n_{j}} (p_{j}-\bar{p})^2}{\bar{p}(1-\bar{p})}$$
+
+Rogers inequality
+$$\frac{c_{w}}{b_{p}} < \frac{F_{ST}}{(1-F_{ST})}$$
+
+Price equation
+$$ \bar{W} \Delta \bar{Z} = \text{cov}(W_{g},Z_{g}) + E[\text{cov}(W_{ig},Z_{ig})]$$
 
@@
 # Conclusion
+[@dudley2018essential]
 

### 2020-09-30 DissertationProposal/Figures/Ch1/Ch1_Plots.R

@@ -58,7 +58,63 @@ p2
 dev.off()
 
 
+PaSize = seq(0,1,0.01)
+ResourceMobility =seq(0,1,0.01)
+fakeData<-expand.grid( Y=ResourceMobility,X=PaSize)
+names(fakeData)<-c("PaSize","ResourceMobility")
+fakeData$Adoption<-rep(NA,nrow(fakeData))
+
+for (i in 1:nrow(fakeData)){
+  
+  fakeData[i,]$Adoption<-ifelse(  fakeData[i,]$PaSize>0.2&fakeData[i,]$PaSize<0.9&fakeData[i,]$ResourceMobility>0.45&fakeData[i,]$ResourceMobility<1,
+                                  rbeta(1,2,2),rbeta(1,2,5))}
+
+for (i in 1:nrow(fakeData)){
+  fakeData[i,]$Adoption<-ifelse(fakeData[i,]$PaSize>0.4&fakeData[i,]$PaSize<0.7&fakeData[i,]$ResourceMobility>0.65&fakeData[i,]$ResourceMobility<0.9,
+                                rbeta(1,5,1),fakeData[i,]$Adoption)}
 
 
+names(fakeData)<-c("PaSize","ResourceMobility","Adoption")
+
+p2<-ggplot(data=fakeData, mapping=aes(x=PaSize, y=ResourceMobility, fill= Adoption)) + 
+  geom_tile()+scale_fill_viridis()+theme_classic()+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        plot.title = element_text( size=14, color="black",face="bold"),
+        axis.title.x = element_text( size=12),
+        axis.title.y = element_text( size=12),
+        axis.text=(element_text(color="black", size=12)),
+        legend.title = element_text(colour="black", size=12),
+        legend.text = element_text( size = 12))+
+  xlab("Percent of Resource Protected")+ylab("Probability of Resource Mobility")+ggtitle("Multiple Groups")
+
+png(filename = "DissertationProposal/Figures/Ch1/abmSim2.png", pointsize=10, width=900, height=680, res=200)
+p2
+dev.off()
 
 
+for (i in 1:nrow(fakeData)){
+  
[…]

### 2020-10-01 DissertationProposal/DissertationProposal.Rmd
Ch2 methods updated

@@
 
-While these predictions are theoretically sound, they must be tested empirically.
+While these predictions are theoretically sound, they must be tested empirically [@kline2018designing].
 Two such studies have begun to test the basis of cultural group selection in the natural world.
@@
 Researchers estimated the total variance in both cultural and genetic traits within and between groups of individuals.
-The proportion of variance in an allele or trait found between groups is referred to as genetic Fst or cultural Fst respectively.
+The proportion of variance in an allele or trait found between groups is referred to as genetic $F_{ST}$ or cultural $F_{ST}$ respectively.
 Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits [@bell_culture_2009].
@@
 A second empirical study tested the prediction that large scale human cooperation is in fact a product of competition between cultural groups.
-Researchers created pair-wise estimates of cultural Fst between 12 ethnic groups.
+Researchers created pair-wise estimates of cultural $F_{ST}$ between 12 ethnic groups.
 These estimates of cultural differentiation were combined with a survey of individual willingness to cooperate with members of the other groups.
@@
 This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly.
-Given the theoretical capacity for cultural group selection to provide solutions to pressing environmental issues of collective action, these predictions must be validated experimentally so that they might be implemented effectively.
+Given the theoretical capacity for cultural group selection to provide solutions to pressing environmental issues of collective action, these predictions must be validated experimentally so that they can be implemented effectively.
 
@@
 
+### Question: 
+Are predictions of costly conservation adoption derived from cultural group selection theory significant predictors of real world Community Forest Management Agreements in forest adjacent wards on Pemba Island, Tanzania?
+
+### Predictions
+* P1: In wards where cultural group selection theory predicts that group selection should favor the adoption of community managed forested areas, we will observe significantly more real world adoption than in wards where adoption is not predicted to be favored by selection.
+
+* P2: If significant cultural differentiation between wards is not present in the population, wards will not adopt Community Forest Management Agreements without external influences such nonprofit conservation initiatives.
+
 ## Research Design
+Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level.
+These group level selection pressures are what we hypothesize to drive the adoption of Community Forest Management Agreements (CoFMAs).
+Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are more opaque.
+For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.
 
-Inline LaTeX equations can be written in a pair of dollar signs using the LaTeX syntax, e.g., $f(k) = {n \choose k} p^{k} (1-p)^{n-k}$  math expressions of the display style can be written in a pair of double dollar signs, e.g.,
-
-$$f(k) = {n \choose k} p^{k} (1-p)^{n-k}$$, and the output looks like this:
-
-Cultural fst equation
+I will conduct an in-person survey of randomly sampled forest users in each Pemban ward.
+This survey will elicit information on approximately 50 cultural norms ranging from crime, cultural markers, punishment, family dynamics, etc.
+Cultural $F_{ST}$ values will be calculated for each trait given the following formula:
 
@@
 
-Rogers inequality
+Where $i$ and $j$ are the two populations being compared, $n_{i}$ and $n_{j}$ represent the number of survey respondents for each group respectively.
+For trait $x$, $\bar{p}$ is the total frequency of trait $x$ in both populations, $p_{i}$ and $p_{j}$ are the trait frequency in each population.
+The equation can be interpreted as the variance in the relative frequency of trait $x$ between populations divided by the total variance in the trait across both populations.
+These values will be averaged across all traits to produce the total cultural $F_{ST}$ values between each pair of wards.
+
+In addition to measuring the between and within group cultural variance, the in-person survey will also elicit information regarding individual’s perceived costs/benefits of CoFMAs, as well as group costs/benefits of CoFMAs.
+We can then use the cultural $F_{ST}$ values, paired with perceived costs and benefits to assess whether or not group cooperation will be favored by selection using the Rogers model of group selection by selective migration, from hereon referred to as the Rogers Inequality.
+
+The Rogers Inequality model was initially designed to measure genetic evolution of groups via selective migration, but is equally applicable to, and has been adapted for cultural evolution [@bowles2006group].
+This model presents an inequality under which individually costly, yet group beneficial traits are favored by selection.
+The emergence of group cooperative norms will be favored by selection when the ratio of between-group variance ($F_{ST}$) to within-group variance ($1-F_{ST}$) exceeds the ratio of individual costs of a behavior ($C_{w}$) to group level population benefit ($B_{p}$).
+Thus, if the following inequality is true, then individually costly, group beneficial behavior is favored by selection [@kline2018designing].
+
 $$\frac{c_{w}}{b_{p}} < \frac{F_{ST}}{(1-F_{ST})}$$
 
-Price equation
-$$ \bar{W} \Delta \bar{Z} = \text{cov}(W_{g},Z_{g}) + E[\text{cov}(W_{ig},Z_{ig})]$$
+While the primary output of this model is a true/false estimate of whether the emergence of a costly cooperative norm is favored by selection, we can also use the ratio presented in the inequality as a measure of selection strength.
+For example, we will expect Community Forest Management Agreements to be more strongly selected for when the right-hand side of the inequality is far greater than the left-hand side compared to when the sides are close to equal.
 

### 2020-10-02 DissertationProposal/DissertationProposal.Rmd
model desc

@@
 ### Model Description
+In the most simple scenario, the agent based model represents a simplistic world where there is one homogeneous group of agents who harvest resources from a landscape.
+This landscape is divided into two types, working and protected.
+Resources move between these patches at each time step with some probability.
+At time = 0, the protected and working landscapes are both at an equal, designated percent carrying capacity.
+At each time step, agents aim to maximize their harvest up until a designated harvest max.
+At time = 0, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree, spending a small portion of their time harvesting from the protected landscape.
 
-```{r conceptDiagram, echo=FALSE, fig.cap="Concept diagram of agent-based model scheduling under single group dynamics", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+The model also allows for ecological variation during the initial specification.
+The variable ecological starting conditions are as follows: Resource mobility designates the probability that a resource unit will move from the protected landscape into the working landscape or vice versa at each time step.
+Protected area size demarcates what percentage of the available resource units are within the protected area.
+Percent carrying capacity sets the total number of harvestable resource units as a percent of the landscape carrying capacity.
+Lastly, the resource regeneration rate designates what percentage the resource regenerates at each time step.
+
+The environment also responds to agent behavior.
+At every time step, both the protected and working landscapes lose exactly the number of resources that agents harvest from them.
+The areas then regenerate resources according to the regeneration rate per time step and lastly, the resources move from one area to the other with the probability specified during model initiation.
+
+Agents behave according to a specific set of rules.
+Agents “choose” to spend a given percent of their time harvesting from the protected area (defecting) and a percent of their time harvesting from the working landscape (defecting).
+Full defection is when the average percent of time agents spend in the protected area is equal to the percent of the landscape protected (i.e.
+they do not treat the protected area any differently than the working landscape).
+Agents learn these behaviors according to their specific learning mechanism.
+The learning will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents.
+Behaviors may be modified at each time step by the key processes described below.
+
+For the given number of timesteps, the model follows a set of key processes.
+At time step one, all agents harvest first spend their initial assigned percentage of effort harvesting from the working landscape.
+If agents reach harvest max, they do not harvest from the protected landscape, even if they are assigned as having the ability to defect.
+If they do not reach harvest max, they will spend the rest of their effort harvesting from the protected area.
+They will harvest up to the harvest max if possible.
+The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape.
+Resources move between the protected and working landscapes with a probability proportional to designated probability of mobility.
+If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies.
+If one agent preformed worse than another in the first time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them.
+If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group.
+This process repeats for all time steps.
+This process is summarized conceptually in Fig \@ref(fig:conceptDiagram).
+
+In the multiple competing groups scenario, two groups with separate resource pools will follow the processes described above.
+However, at the end of each time step the groups will compare the average harvest rate each group.
+Whichever group has the highest average harvest rate for that timestep will report the average harvest times for the protected and working landscapes.
+A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group.
+
+In both single and competing group scenarios, the dependent variables of interest are as follows.: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step.
+Full defection is defined as all agents harvesting from the protected landscape for a proportion of time equal to the percent of the resource which is protected.
+Protected area degradation measures the percent carrying capacity of the protected landscape.
+Working landscape degradation measures the percent carrying capacity of the working landscape.
+Lastly, mean payoff to individuals measures the average payoff that agents receive at each time step.
+
+```{r conceptDiagram, echo=FALSE, fig.cap="Concept diagram of agent-based model scheduling under single group dynamics", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
 knitr::include_graphics("Figures/Ch1/ConceptDiagram.png")
@@
 ## Preliminary Work
+I have developed a model which follows the above processes for one group with one resource pool.
+An interactive application which performs one run of that model can is found [at this link](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
+
+This model currently assigns one learning strategy to all agents in the population.
+Users can select either successes-biased or conformity-biased learning to be applied.
+Future development will allow users to specify the percent of the population employing each learning strategy as specified in the methods above.
+
+Environmental conditions can be set to provide neither an over abundance of resources nor a restrictive scarcity.
+Under these conditions, we see conformity biases maintain starting group norms around harvest behaviors.
+Of greater interest is that under success biased learning, we observe that protected areas enhance agent wellbeing (average harvest) initially, but defection eventually spreads and causes a failure of the protected area.
+With this in-mind, the next step for this research is to introduce a second group dynamic which will allow us to test the logic of cultural group selection for costly conservation behaviors under these conditions.
+Example outputs from this model can be seen in Fig \@ref(fig:abmOutput).
 
@@
 ```{r stovePic, echo=FALSE, fig.cap="Example of an improved, fuel efficient cookstove on Pemba Island, Tanzania.
-Image from Same Mwanga Environmental Conservation Advisory Organization", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+Image from Same Mwanga Environmental Conservation Advisory Organization", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
 knitr::include_graphics("Figures/Ch3/cookstove.jpg")
@@
 ```{r mapp, echo=FALSE, fig.cap="Satellite map of Pemba Island, Tanzania.
-Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "75%", fig.align='center', dpi=72, fig.pos= "H"}
 knitr::include_graphics("Figures/Ch3/BadPembaMap.png")
@@
 # Conclusion
-[@dudley2018essential]
+This research seeks to make clear how and why conservation behaviors succeed or fail to spread efficiently within a population.
+Moreover, the aim of this dissertation is to provide generalizable insights which can be applied to conservation initiatives in a variety of environments and population structures.
+A key advance made by this work is to link the fields of conservation social science and cultural group selection.
+Cultural group selection theory makes important predictions about how cooperative behaviors like conservation adoption may emerge and spread within populations, but has not yet been tested in a conservation context explicitly.
+Similarly, conservation social science incorporates environmental phenomena like resource mobility and regeneration, but largely omits generalizable insights about human behavioral evolution.
+By linking these two fields, we are greatly increasing the applicability of both bodies of work.
+
+This dissertation will achieve this goal in three progressive steps.
+The first will formalize theory around cultural group selection in conservation contexts.
+This step is critical as conservation is a fundamentally different system from where these theories were originally developed.
+The second step will empirically validate these theoretical insights by comparing calculated cultural group selection pressures to real-world adoption of Community Forest Management Agreements on Pemba Island, Tanzania.
+Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities, to specifically examine how behavioral efficiency may be altered by social contexts (e.g.
+communication frequency, genetic relatedness) as a behavior spreads throughout a population.
+All together, this research will allow us to better understand and possibly predict the emergence and effective spread of conservation norms within a population.
+
+Human wellbeing is definitively tied to the persistence of renewable resources and ecosystem services.
+We are however, degrading these resources more rapidly than they can regenerate; reducing the stock of available resources while notably also slowing further regeneration.
+Conservation scientists have identified focal areas which, if preserved, may provide a disproportionate benefit to ecosystem function and offer a lifeline for human wellbeing.
+Simply put, we understand the environmental targets that we must achieve to ensure continued availability of renewable resources and we understand how conservation can achieve those targets.
+What we do not understand is how to effectively implement these required conservation measures.
+Reaching global biodiversity goals such as those set by the UN Convention on Biological Diversity will require “big and bold” conservation initiatives to be adopted at every level of the human population [@convention2010strategic; @dudley2018essential].
+Currently however, even small-scale, local conservation measures often fail to gain traction and effectively run aground.
+This research will provide crucial theory and empirical evidence for how we can better shape and implement effective conservation initiatives at a variety of scales and in a variety of environmental conditions.
 

### 2020-10-05 DissertationProposal/DissertationProposal.Rmd
abm parameter table

@@
 Lastly, mean payoff to individuals measures the average payoff that agents receive at each time step.
+
+```{r tabAbmVars, echo=F, fig.align='center'}
+abmVars <- data.frame(Parameters = c("Individuals","Total Carrying Capacity", "Percent Carrying Capacity", 
+                                     "Percent Protected", "Starting Cooperators", "Time Steps", "Max Harvest",
+                                     "Resource Regeneration",  "Resource Mobility","Groups","Learning Strategy", 
+                                     "Percent Cooperation", "Average Payoff", "Protected Area Degredation",
+                                     "Working Landscape Degredation"), 
+                      Type = c("Starting Condition","Starting Condition","Starting Condition",
+                               "Starting Condition","Starting Condition","Starting Condition",
+                               "Starting Condition","Key Independent Variable","Key Independent Variable",
+                               "Key Independent Variable","Key Independent Variable",
+                               "Outcome","Outcome","Outcome","Outcome"), 
+                      Description = linebreak(c("Number of agents in simulation",
+                                      "Maximum number of resources\nallowed on landscape",
+                                      "Percent of carrying capacity at\ninitialization",
+                                      "Percent of landscape in\nprotected area",
+                                      "Starting percent of agents\nfully cooperating",
+                                      "Number of time steps\nsimulation will run for",
+                                      "Maximum number of resources\nagents will harvest during\none time step",
+                                      "The percent of available resources\nwhich will regenerate at each time\nstep in both protected and\nworking landscapes",
+                                      "The probability that any available\nresource will move from the\n protected to working landscape\nor vice versa",
+                                      "Solitary or multiple\ncompeting groups",
+                                      "Assigns a percent of the population\nto success-biased imitation and\nthe remaining population to\nconformist-biased imitation",
+                                      "Percent of total population\nharvest time spent in\nthe working landscape","Mean harvest for all agents\nat each time step",
+                                      "Working landscape percent of\ntotal carrying capacity",
+                                      "Protected area percent of\ntotal carrying capacity")))
+knitr::kable(abmVars, caption = "Description of all model parameters.",escape=F) %>% 
+  kable_styling(c("striped",full_width=F))%>% row_spec(0,bold=TRUE)
+
+```
 

### 2020-10-06 DissertationProposal/DissertationProposal.Rmd
Done before committee review I guess....

@@
 As such, each year billions of dollars are spent to protect key areas.
-It is common however, for these areas to fail to reach their stated conservation goals.
+It is common, however, for these areas to fail to reach their stated conservation goals.
 The most chronic causes of conservation project failure are the human dimensions.
 Conversely, historic conservation successes are also attributed to widespread human adoption of conservation behaviors.
-Research examining how and why conservation practices are adopted has been quite limited; suffering from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes.
+Research examining how and why conservation practices are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes.
 This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin social adoption of conservation practices in and around protected areas.
@@
 1. Develop a dynamic model to generate predictions regarding if and how conservation norms can emerge and persist via competition between groups.
+
 2. Empirically test theoretical predictions of group conservation adoption as a function of group competition and individual costs and benefits.
+
 3.
@@
 
-Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local human populations [@adams_biodiversity_2004; @leverington_global_2010].
-While high-profile protected area successes are salient in the conservation narrative, these areas also commonly fail to reach their intended biodiversity targets.
-The ubiquity of these failures is highly speculative as unsuccessful protected area initiatives go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004].
-Despite a high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].
+Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local communities [@adams_biodiversity_2004; @leverington_global_2010].
+While high-profile protected area successes are salient in the conservation narrative, protected areas also commonly fail to reach their intended biodiversity targets.
+The ubiquity of these failures is somewhat speculative as unsuccessful protected areas go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004].
+Regardless of the high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].
 
-Community adoption of protected area regulations may emerge and persist as a group norm, emerge only for a short period of time, or may never emerge within a given group [@rogers_diffusion_2010; @simon_win-win_2012].
-These behaviors are characterized as individually costly, yet beneficial at the group level.
+Community norms around protected area adherence are characterized as individually costly, yet beneficial at the group level.
+Substantial variation exists in the degree to which communities adhere to protected area regulations; some communities develop strong and lasting conservation norms, while in others these norms degrade over time or may fail to emerge at all [@rogers_diffusion_2010; @simon_win-win_2012].
 Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017].
@@
 Case studies tend to be highly context dependent and offer little mechanistic insight into the emergence and spread of these behaviors.
-Excitingly, parallel to the development of conservation social science, the field of evolutionary anthropology has developed the theory of cultural group selection as a mechanistic explanation for the adoption of individually costly, group beneficial behaviors.
+Outside of conservation science, the field of evolutionary anthropology has devoted considerable effort to understanding how and why individuals adopt costly cooperative behaviors.
+Of particular relevance to understanding group adherence to protected area regulations, the theory of cultural group selection offers a mechanistic explanation for the emergence and spread of individually costly group norms.
 
@@
 
-Applying cultural group selection to the adoption of conservation behaviors
+## Applying cultural group selection to the adoption of conservation behaviors
 Protected areas are an ideal microcosm for studying and applying cultural group selection.
@@
 ## Background
-Over the last 20-30 years, both the number and size of new no-take protected areas created each year have increased steadily (Figures \@ref(fig:paNumber) & \@ref(fig:paSize)) [@iucn2020protectedplanet].
-During this time period, the stated mission of most protected areas has shifted as well.
-Due to the overlap between biodiversity and low income populations, as well as the price and availability of land, protected areas are largely adjacent to, or even contain, low income communities [@fisher_poverty_2007; @naughton-treves_role_2005].
-With this recognition, their objectives now commonly include community development goals alongside conservation targets.
+Cultural group selection theorizes that individuals will adopt cooperative behavior if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985].
+Under this framework, the presence of multiple competing groups drives community cooperative norms to emerge under a much broader set of conditions compared to solitary groups [@richerson_cultural_2016].
+Protected area conservation is a cooperative phenomenon where individuals sacrifice some degree of their own harvest potential in order to benefit the group as a whole.
+Therefore, insights from cultural group selection may help conservation scientists better understand and predict community adherence to protected area regulations [@brooks_applying_2018].
 
-```{r paSize, echo=FALSE, fig.cap="Size of newly established no-take protected areas reported by the IUCN 2020", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch1/Size.png")
-```
+While cultural group selection theory offers mechanistic insights for the emergence of cooperative norms in groups broadly, its applicability to conservation specifically is unknown.
+Conservation has many aspects which make it unique compared to cooperation in the abstract.
+In conservation, an individual’s behavior directly affects their environment, which may in turn become altered in a way which shapes future individual decision making.
+Feedbacks between individual behavior and resource availability are further complicated by resource characteristics such as mobility and regeneration rate.
+These specifics make conservation a fundamentally different system from general cooperation, thus we must extend cultural group selection theory considering conservation explicitly.
 
-Joint conservation and community development initiatives are a popular prescription for combating climate change and global biodiversity loss [@tallis_ecosystem_2008].
-These dual-objective projects attempt to mitigate concerns of environmental neocolonialism and resource disenfranchisement of local populations [@hall_tourism_2004; @langholz_perils_2001].
-Importantly, protected areas which explicitly seek to bolster the livelihoods of local populations actually show greater conservation success than those that do not [@fisher_poverty_2007].
-Protected areas then represent a considerable opportunity for “win-win” projects which support global conservation and poverty alleviation [@adams_biodiversity_2004].
-
-Protected areas, while well intentioned, often fail to meet both their stated social and ecological goals [@christie_marine_2004; @watson_performance_2014].
-In fact, a study of protected areas in South Asia found them to be ecologically equivalent to adjacent unprotected lands [@clark_protected_2013].
-It’s unsurprising that the success of protected areas in reaching both social and ecological targets has been linked to the degree of local support and adherence to regulations [@bennett_why_2014; @christie_observed_2005].
-
-The degree of local buy-in for protected areas is not random.
-The scale of implementation is highly correlated with the percentage of individuals who adopt protected area regulations.
-A 2019 study demonstrated that nationally implemented initiatives had 84-100% buy-in from local populations, while locally established areas saw adoption from 0.13 - 51% of potential adopters [@mills_how_2019].
-These findings are particularly worrying given a global decline in government support for land conservation [@watson_performance_2014].
-
-With the recognition that local adoption is a difficult and necessary component of conservation success, human dimensions have become a major focus of conservation research [@bennett_conservation_2017].
-This body of work has largely relied on case studies of conservation projects to draw correlations between social phenomena and conservation outcomes.
-As such, findings from conservation social science tend to offer little mechanistic insights for how conservation behaviors spread within and between groups.
-Recently however, theories from cultural group selection have been identified as a potentially useful tool to better elucidate the mechanisms which underpin local buy-in for conservation initiatives [@brooks_applying_2018].
-
-Cultural group selection theorizes that individuals will adopt behaviors which adhere to protected area regulations if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985].
-Under this framework, multiple competing groups are actually required for community adoption of protected area regulations to emerge [@richerson_cultural_2016].
-
-While the theory and principles of cultural group selection are relatively well established, they are yet to be tested mechanistically in the context of protected area adoption.
-With this in mind, this project aims to demonstrate how the presence of multiple competing groups can facilitate the adoption of adherence to protected areas.
-In this chapter I will develop a dynamic, agent-based model to test the logic of how protected area norms can emerge under very simple sets of rules in the presence of multiple competing groups, but fail to emerge for solitary groups under identical environmental contexts.
+This project aims to logically test the applicability of cultural group selection to explain the emergence and spread of protected area adoption.
+In this chapter I will develop a dynamic, agent-based simulation of individuals harvesting resources from a landscape under a variety of environmental constraints.
+I will then use this simulation to test under what circumstances adherence to protected area regulations emerges as a group norm within solitary and multiple, competing groups.
+This model will generate qualitative predictions about system outcomes given a very simple set of assumptions.
 
 ### Question:
-Can between group competition provoke protected area adoption where solitary groups may fail to adopt?
+Under what environmental contexts might the mechanisms of cultural group selection influence individual decision making to generate lasting group norms around the adherence to protected area regulations?
 
 ### Predictions:
-* P1: Adoption of protected area adherence will only emerge under the presence of multiple competing groups.
+* P1: Groups will adopt protected area regulations in both the solitary and multiple competing groups scenarios under specific environmental contexts.
 
-* P2: Where it already exists, protected area adherence will perpetuate in populations even in solitary groups given group conformity.
-
-* P3: Greater degrees of resource mobility will increase the frequency in which the adoption of protected area adherence emerges in the community(ies).
+* P2: In the presence of multiple competing groups, the environmental contexts which generate group adoption of protected area regulations will be much more broad compared to solitary groups.
 
@@
 
-An agent-based modeling framework is ideal for demonstrating the theory of cultural group selection regarding protected area norms, because it allows such group level properties to emerge out of simple decision rules for each individual (Railsback & Grimm, 2012; Wilensky & Rand, 2015).
+An agent-based modeling framework is ideal for demonstrating the theory of cultural group selection regarding protected area norms, because it allows such group level properties to emerge out of simple decision rules for each individual [@railsback_agent-based_2012; @wilensky_introduction_2015].
 This design mirrors our real-world understanding of the important system components.
@@
 Resources move between these patches at each time step with some probability.
-At time = 0, the protected and working landscapes are both at an equal, designated percent carrying capacity.
+During initialization, the protected and working landscapes are both at an equal, designated percent carrying capacity.
 At each time step, agents aim to maximize their harvest up until a designated harvest max.
-At time = 0, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree, spending a small portion of their time harvesting from the protected landscape.
+During initialization, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree, spending a small portion of their time harvesting from the protected landscape.
 
@@
 Percent carrying capacity sets the total number of harvestable resource units as a percent of the landscape carrying capacity.
-Lastly, the resource regeneration rate designates what percentage the resource regenerates at each time step.
-
-The environment also responds to agent behavior.
-At every time step, both the protected and working landscapes lose exactly the number of resources that agents harvest from them.
-The areas then regenerate resources according to the regeneration rate per time step and lastly, the resources move from one area to the other with the probability specified during model initiation.
-
-Agents behave according to a specific set of rules.
-Agents “choose” to spend a given percent of their time harvesting from the protected area (defecting) and a percent of their time harvesting from the working landscape (defecting).
-Full defection is when the average percent of time agents spend in the protected area is equal to the percent of the landscape protected (i.e.
-they do not treat the protected area any differently than the working landscape).
-Agents learn these behaviors according to their specific learning mechanism.
-The learning will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents.
-Behaviors may be modified at each time step by the key processes described below.
-
-For the given number of timesteps, the model follows a set of key processes.
-At time step one, all agents harvest first spend their initial assigned percentage of effort harvesting from the working landscape.
-If agents reach harvest max, they do not harvest from the protected landscape, even if they are assigned as having the ability to defect.
-If they do not reach harvest max, they will spend the rest of their effort harvesting from the protected area.
-They will harvest up to the harvest max if possible.
-The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape.
-Resources move between the protected and working landscapes with a probability proportional to designated probability of mobility.
-If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies.
-If one agent preformed worse than another in the first time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them.
-If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group.
-This process repeats for all time steps.
-This process is summarized conceptually in Fig \@ref(fig:conceptDiagram).
-
-In the multiple competing groups scenario, two groups with separate resource pools will follow the processes described above.
-However, at the end of each time step the groups will compare the average harvest rate each group.
-Whichever group has the highest average harvest rate for that timestep will report the average harvest times for the protected and working landscapes.
-A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group.
-
-In both single and competing group scenarios, the dependent variables of interest are as follows.: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step.
-Full defection is defined as all agents harvesting from the protected landscape for a proportion of time equal to the percent of the resource which is protected.
-Protected area degradation measures the percent carrying capacity of the protected landscape.
-Working landscape degradation measures the percent carrying capacity of the working landscape.
-Lastly, mean payoff to individuals measures the average payoff that agents receive at each time step.
+Lastly, the resource regeneration rate designates what percentage the resource regenerates at each time step (see Table \@ref(tab:tabAbmVars)).
 
@@
 
+The environment also responds to agent behavior.
+At every time step, both the protected and working landscapes lose exactly the number of resources that agents harvest from them.
+The areas then regenerate resources according to the regeneration rate per time step and lastly, the resources move from one area to the other with the probability specified during model initiation.
+
+Agents behave according to a specific set of rules.
+Agents choose to spend a given percent of their time harvesting from the protected area (defecting) and a percent of their time harvesting from the working landscape (cooperating).
+Full defection is when the average percent of time agents spend in the protected area is equal to the percent of the landscape protected (i.e.
+they do not treat the protected area any differently than the working landscape).
+
+Agents learn their behaviors according to one of two possible learning mechanisms.
+Agents who learn via success-biased learning will preferentially copy the behavior of more successful agents when they encounter them.
+Agents who learn via conformist-biased learning will disproportionately copy the behavior of other agents they see as being in the majority, regardless of the success of those agents.
+The learning strategy of the overall population will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents.
+Behaviors may be modified at each time step by the key processes described below.
+
+For the given number of timesteps, the model follows a set of key processes.
+At time step one, all agents first spend their initial assigned percentage of effort harvesting from the working landscape.
+If agents reach the max harvest amount, they do not harvest from the protected landscape, even if they are assigned as having the ability to defect.
+If they do not reach harvest max, they will spend the rest of their effort harvesting from the protected area.
+They will harvest up to the harvest max if possible.
+The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape.
+Resources move between the protected and working landscapes with a probability proportional to designated probability of mobility (see Table \@ref(tab:tabAbmVars)).
+If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies.
+If one agent performs worse than another in the first time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them.
+If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group.
+This process repeats for all time steps.
+This process is summarized conceptually in Fig \@ref(fig:conceptDiagram).
+
 ```{r conceptDiagram, echo=FALSE, fig.cap="Concept diagram of agent-based model scheduling under single group dynamics", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
@@
 
+In the multiple competing groups scenario, two groups with separate resource pools will follow the processes described above.
+However, at the end of each time step the groups will compare the average harvest rate each group.
+Whichever group has the highest average harvest rate for that timestep will report the average harvest times for the protected and working landscapes.
+A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group (see Table \@ref(tab:tabAbmVars)).
+
+In both single and competing group scenarios, the dependent variables of interest are as follows: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step.
+Full defection is defined as all agents harvesting from the protected landscape for a proportion of time equal to the percent of the resource which is protected.
+Protected area degradation measures the percent carrying capacity of the protected landscape.
+Working landscape degradation measures the percent carrying capacity of the working landscape.
+Lastly, mean payoff to individuals measures the average payoff that agents receive at each time step.
+
+Overall, the goal of this model is to test how environmental conditions affect cultural group selection mechanisms.
+Cultural group selection may lead to community adoption of sustainable norms, where solitary groups fail to adopt cooperative behaviors.
+With this model, we can test under what environmental conditions this is true and what the implications are for community wellbeing and environmental protection.
+
 ## Preliminary Work
 I have developed a model which follows the above processes for one group with one resource pool.
-An interactive application which performs one run of that model can is found [at this link](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
+An interactive application which performs one run of that model can is found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
 
@@
 This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly.
-Given the theoretical capacity for cultural group selection to provide solutions to pressing environmental issues of collective action, these predictions must be validated experimentally so that they can be implemented effectively.
+Insights from cultural group selection may provide solutions to pressing environmental challenges of collective action.
+If validated experimentally, principles of cultural group selection may allow conservation scientists to better predict the viability of conservation initiatives.
 
@@
 ### Question: 
-Are predictions of costly conservation adoption derived from cultural group selection theory significant predictors of real world Community Forest Management Agreements in forest adjacent wards on Pemba Island, Tanzania?
+Does cultural group selection theory accurately predict real world Community Forest Management Agreements in forest adjacent wards on Pemba Island, Tanzania?
 
@@
 
-* P2: If significant cultural differentiation between wards is not present in the population, wards will not adopt Community Forest Management Agreements without external influences such nonprofit conservation initiatives.
+* P2: If significant cultural differentiation between wards is not present in the population, wards will not adopt Community Forest Management Agreements without external influences such as nonprofit conservation initiatives.
 
@@
 These group level selection pressures are what we hypothesize to drive the adoption of Community Forest Management Agreements (CoFMAs).
-Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are more opaque.
+Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are blurred.
 For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.
@@
 Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes [@mesoudi_multiple_2008].
-This research will directly measure the impact of geographic, genetic, and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.
+This research will directly measure the impact of geographic and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.
 
@@
 ### Question
-What effect do geographic distance, genetic relatedness, and social relatedness have on transmission fidelity? Secondly, what is the relationship between transmission fidelity and behavioral outcomes?
+What effect do geographic distance and social relatedness have on transmission fidelity? Secondly, what is the relationship between transmission fidelity and behavioral outcomes?
 
 ### Predictions
-* P1: Using geographic distance, genetic relatedness, and social network distance as proxies for access to behavioral demonstrators, we expect transmission fidelity to increase with increasing access.
+* P1: Transmission fidelity will increase with increased access to behavioral demonstrators through geographic proximity and/or social connectedness.
 
-* P2: As access to behavioral demonstration decreases, we expect to see differential changes in the fidelity of replicated behavior and the behavioral performance target of interest.
-
-* P3: Given the high degree of environmental influence on a behavior, we expect to observe a greater change in the fidelity of transmission than in the behavioral performance target, correlative with decreasing access to behavioral demonstrators.
+* P2: As access to behavioral demonstration decreases, the replication fidelity of non-functional stove traits will decrease strongly relative to functional traits.
 
@@
 ### Methodological Specifics
-I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)).
+I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi (aka Before Ningeze Forest) wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)).
 With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.
@@
 The survey instrument used for this research will elicit information regarding the relationship between the builder of each cookstove and the individual they copied their stove design from.
-Specifically, this survey tool will be used to gather data on geographic distance, genetic relatedness, and social relatedness.
+Specifically, this survey tool will be used to gather data on geographic distance and communication frequency.
+Geographic distance will be measured according to the shortest walking distance between the two stoves.
+Communication frequency will be measured in terms of weekly communication and taken as a measure of social connectedness between the original stove builder and the copying individual.
 I will also use this survey to collect data on the decision making processes individuals exhibit when deciding to build a fuel-efficient cookstove.
 
-While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics such as the exterior dimensions and size of the openings.
-I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from.
 To measure stove fuel efficiency, I will follow the standard water boiling test protocol [@defoort2009stove; @mccracken1998emissions].
 Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test.
+Fuel efficiency will be calculated using the following thermal efficiency equation:
+
+$$n_{th} = \frac{C_{p}*m_{wi}*(T_{f}-T_{i})+H_{v}*(m_{wi}-m_{wf})}{(F_{i}-F_{f})(1-\frac{M}{100})*LHV_{wood}-(F_{i}-F_{f})*\frac{M}{100}*(C_{p}*(T_{b}-T_{a})+H_{v})-LHV_{char}*C_{c}} $$
+
+where $C_{p}$ represents the heat capacity of water (4.186 J/g-K), $m_{wi}$ is the initial mass of water used, $m_{wf}$ is the mass or remaining water, $T_{i}$ and $T_{f}$ are the initial and final water temperatures respectively, and $H_{v}$ represents the enthalpy of vaporization of water (2260 J/g).
+$F_{i}$ and $F_{f}$ represent the initial and final masses of fuel used.
+$M$ is the percent moisture content of the fuelwood used, $LHV_{wood}$ and $LHV_{char}$ are the lower heating values of wood and charcoal respectively, $T_{a}$ is the ambient temperature, $T_{b}$ is the local boiling temperature of water, and $C_{c}$ is the mass of the charcoal created during the test [@defoort2009stove].
+
+While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics.
+I will categorize these characteristics as functional (stove size) or nonfunctional (decorative qualities).
+I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from in both functional and nonfunctional characteristics.
 
 The key research questions for this project will be assessed using multilevel regression operationalized in a Bayesian context.
-I will compare standardized parameter estimates, to determine which factors (genetic relatedness, geographic distance, social relatedness) most strongly contribute to copying fidelity in this study system.
+I will compare standardized parameter estimates, to determine which factors (geographic distance or social relatedness) most strongly contribute to copying fidelity in this study system.
 Similar analyses will be used to determine which factors affect cookstove primary outcome variables (fuel efficiency and particulate matter production).
@@
 ## Preliminary Work
-In preparation for a cancelled field season in the Summer of 2020, I created a variety of forms and documents to facilitate data collection. These include:
+In preparation for a cancelled field season in the Summer of 2020, I created a suite of forms and documents to facilitate data collection. These include:
 
@@
 Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities, to specifically examine how behavioral efficiency may be altered by social contexts (e.g.
-communication frequency, genetic relatedness) as a behavior spreads throughout a population.
+communication frequency) as a behavior spreads throughout a population.
 All together, this research will allow us to better understand and possibly predict the emergence and effective spread of conservation norms within a population.

### 2020-10-06 DissertationProposal/Figures/Ch1/Ch1_Plots.R

@@ -82,7 +82,7 @@ p2<-ggplot(data=fakeData, mapping=aes(x=PaSize, y=ResourceMobility, fill= Adopti
         plot.title = element_text( size=14, color="black",face="bold"),
         axis.title.x = element_text( size=12),
         axis.title.y = element_text( size=12),
-        axis.text=(element_text(color="black", size=12)),
+        axis.text=(element_text(color="black", size=12)),text=element_text(family="Calibri"),
         legend.title = element_text(colour="black", size=12),
         legend.text = element_text( size = 12))+
   xlab("Percent of Resource Protected")+ylab("Probability of Resource Mobility")+ggtitle("Multiple Groups")
@@ -94,12 +94,12 @@ dev.off()
 
 for (i in 1:nrow(fakeData)){
   
-  fakeData[i,]$Adoption<-ifelse(  fakeData[i,]$PaSize>0.2&fakeData[i,]$PaSize<0.9&fakeData[i,]$ResourceMobility>0.45&fakeData[i,]$ResourceMobility<1,
-                                  rbeta(1,1,5),rbeta(1,1,5))}
+  fakeData[i,]$Adoption<-ifelse(  fakeData[i,]$PaSize>0.35&fakeData[i,]$PaSize<0.75&fakeData[i,]$ResourceMobility>0.6&fakeData[i,]$ResourceMobility<0.85,
+                                  rbeta(1,2,2),rbeta(1,2,5))}
 
 for (i in 1:nrow(fakeData)){
-  fakeData[i,]$Adoption<-ifelse(fakeData[i,]$PaSize>0.4&fakeData[i,]$PaSize<0.7&fakeData[i,]$ResourceMobility>0.65&fakeData[i,]$ResourceMobility<0.9,
-                                rbeta(1,1,5),fakeData[i,]$Adoption)}
+  fakeData[i,]$Adoption<-ifelse(fakeData[i,]$PaSize>0.45&fakeData[i,]$PaSize<0.6&fakeData[i,]$ResourceMobility>0.65&fakeData[i,]$ResourceMobility<0.75,
+                                rbeta(1,5,1),fakeData[i,]$Adoption)}
 
 
 names(fakeData)<-c("PaSize","ResourceMobility","Adoption")
@@ -110,7 +110,7 @@ p1<-ggplot(data=fakeData, mapping=aes(x=PaSize, y=ResourceMobility, fill= Adopti
         plot.title = element_text( size=14, color="black",face="bold"),
[…]

### 2020-10-06 DissertationProposal/Figures/Ch1/Ch1_Plots.R
small abm plot change

@@ -66,11 +66,11 @@ fakeData$Adoption<-rep(NA,nrow(fakeData))
 
 for (i in 1:nrow(fakeData)){
   
-  fakeData[i,]$Adoption<-ifelse(  fakeData[i,]$PaSize>0.2&fakeData[i,]$PaSize<0.9&fakeData[i,]$ResourceMobility>0.45&fakeData[i,]$ResourceMobility<1,
+  fakeData[i,]$Adoption<-ifelse(  fakeData[i,]$PaSize>0.2&fakeData[i,]$PaSize<0.9&fakeData[i,]$ResourceMobility>0.35&fakeData[i,]$ResourceMobility<0.85,
                                   rbeta(1,2,2),rbeta(1,2,5))}
 
 for (i in 1:nrow(fakeData)){
-  fakeData[i,]$Adoption<-ifelse(fakeData[i,]$PaSize>0.4&fakeData[i,]$PaSize<0.7&fakeData[i,]$ResourceMobility>0.65&fakeData[i,]$ResourceMobility<0.9,
+  fakeData[i,]$Adoption<-ifelse(fakeData[i,]$PaSize>0.4&fakeData[i,]$PaSize<0.7&fakeData[i,]$ResourceMobility>0.55&fakeData[i,]$ResourceMobility<0.7,
                                 rbeta(1,5,1),fakeData[i,]$Adoption)}
 
 

### 2020-10-06 DissertationProposal/DissertationProposal.Rmd
small changes.

@@
 
+\newpage
 # Summary
@@
 This research will bring together cultural evolutionary science and ecology to identify the key mechanisms of conservation adoption.
+\newpage
 
@@
 This project will make use of the ongoing spread of fuel efficient cookstove technology on Pemba Island, Tanzania.
+
+## Study System
+All empirical research for this dissertation will take place on Pemba Island, Tanzania.
+In 2014, Pemba Island had a reported population of 406,808 [@RGZ2014Zanzibar].
+This population is divided into wards of 2,000 to 6,000 individuals.
+The majority of Pemban wards are described as forest-dependent.
+The Department of Forests and Non-Renewable Natural Resources allows wards to enter into Community Forest Management Agreements (CoFMAs), which among other things, established forest protected areas and exclude outside individuals [@benjaminsen2014between].
+As of 2018, 18 Pemban wards had adopted CoFMAs [@andrews_cultural_2018].
+
+\newpage
 
@@
 ```
+\newpage
+
 # Chapter 2: Effects of inter-group resource competition and selection on community adoption of costly conservation policy.
@@
 
-Conservation via protected areas represents an ideal study system for testing the environmental applications of cultural group selection in the real world.
-Ignoring the regional to global scale ecosystem service benefits of protected areas, the units of the individual costs and group benefits of protected area adoption are directly comparable (i.e.
-fuelwood growth/use).
-This comparability will allow us to use established methods of measuring multilevel selection pressures to quantitatively predict if protected areas are likely to be selected for in each group.
-Finally, this project will compare the calculated selection pressures for protected area adoption from each group against the observed adoption of community managed protected areas.
+Protected areas are an ideal study system for testing the mechanisms of cultural group selection in real world conservation.
+When a community adopts a protected area, like the Community Forest Management Agreements observed on Pemba Island, Tanzania, individuals agree to forego a portion of their personal harvest potential for the benefit of the community.
+Under cultural group selection, we only expect this behavior to emerge when the proportion of dissimilar to similar cultural traits between a group and its neighbors outweighs the benefit/cost ratio of engaging in cooperative conservation.
+Plainly put, communities should be more likely to adopt CoFMAs if they are culturally dissimilar to their surrounding communities.
+This project will test this theory by evaluating the perceived individual costs and group benefits of CoFMAs, within and between group cultural similarities (cultural $F_{st}$), and actual adoption of CoFMAs in forest dependent wards on Pemba Island [@kline2018designing].
 
 ### Question: 
-Does cultural group selection theory accurately predict real world Community Forest Management Agreements in forest adjacent wards on Pemba Island, Tanzania?
+Do Pemban wards with stronger cultural differentiation from surrounding wards adopt CoFMAs at a higher rate than wards who are more culturally similar to their neighbors?
 
 ### Predictions
-* P1: In wards where cultural group selection theory predicts that group selection should favor the adoption of community managed forested areas, we will observe significantly more real world adoption than in wards where adoption is not predicted to be favored by selection.
+* P1: Cultural differentiation of Pemban wards will predict CoFMA adoption only when the perceived individual costs of CoFMAs are relatively low.
 
-* P2: If significant cultural differentiation between wards is not present in the population, wards will not adopt Community Forest Management Agreements without external influences such as nonprofit conservation initiatives.
+* P2: High levels of forest dependence and perceived costs of CoFMAs will prevent CoFMA adoption, even when groups are considerably culturally dissimilar to adjacent groups.
 
@@
 Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level.
-These group level selection pressures are what we hypothesize to drive the adoption of Community Forest Management Agreements (CoFMAs).
+These group level selection pressures are what we hypothesize to drive the adoption of CoFMAs.
 Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are blurred.
@@
 ```
+\newpage
 
@@
 ```
+\newpage
+
 # Conclusion
@@
 
+\newpage
+
 # References

### 2020-10-08 DissertationProposal/DissertationProposal.Rmd
small change
text edits for clarity

@@
 As such, each year billions of dollars are spent to protect key areas.
-It is common, however, for these areas to fail to reach their stated conservation goals.
-The most chronic causes of conservation project failure are the human dimensions.
-Conversely, historic conservation successes are also attributed to widespread human adoption of conservation behaviors.
-Research examining how and why conservation practices are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes.
-This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin social adoption of conservation practices in and around protected areas.
+It is common, however, for protected areas to fail to reach their stated conservation goals.
+The most chronic cause of conservation failure is a lack of local support and adoption of mandates.
+Conversely, historic protected area successes are also attributed to widespread adoption of conservation behaviors.
+Research examining how and why protected areas are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes.
+This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin local adoption of conservation practices in and around protected areas.
 Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.
 
-1. Develop a dynamic model to generate predictions regarding if and how conservation norms can emerge and persist via competition between groups.
+1. Develop a dynamic, theoretical model to generate predictions regarding if and how group-level adoption of protected areas can emerge and persist via competition between groups.
 
-2. Empirically test theoretical predictions of group conservation adoption as a function of group competition and individual costs and benefits.
+2. Empirically test theoretical predictions of protected area adoption as a function of group competition and perceived costs and benefits to local community members.
 
-3.
-Using the ongoing spread of an improved efficiency cookstove design, specifically examine how the nature of interpersonal relationships may modify conservation behaviors as they spread within groups.
+3. Using the ongoing spread of an improved efficiency cookstove design, examine how the nature of interpersonal relationships may modify conservation behaviors as they spread within groups.
 
@@
 Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not.
-As an individual’s behavior is flexible, individuals will either migrate from less to more successful groups or directly copy the cooperative behavior of others in visibly successful cooperative groups [@henrich_cultural_2004].
+As an individual’s behavior is flexible, individuals may migrate from less to more successful groups or directly copy the behaviors they observe in visibly successful cooperative groups [@henrich_cultural_2004].
 This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection.
@@
 
-This project aims to logically test the applicability of cultural group selection to explain the emergence and spread of protected area adoption.
+This project aims to logically test the applicability of cultural group selection to conservation.
+Specifically, this research will address how behavior-environment feedbacks and resource mobility and regeneration rates affect the emergence and spread of protected area adoption.
 In this chapter I will develop a dynamic, agent-based simulation of individuals harvesting resources from a landscape under a variety of environmental constraints.
@@
 ### Model Description
-In the most simple scenario, the agent based model represents a simplistic world where there is one homogeneous group of agents who harvest resources from a landscape.
+
+*Single group setup and basic function*
+
+In the most simple scenario, the agent-based model represents a simplistic world where there is one homogeneous group of agents who harvest resources from a landscape.
 This landscape is divided into two types, working and protected.
@@
 At every time step, both the protected and working landscapes lose exactly the number of resources that agents harvest from them.
-The areas then regenerate resources according to the regeneration rate per time step and lastly, the resources move from one area to the other with the probability specified during model initiation.
+The areas then regenerate resources according to the regeneration rate per time step and lastly, the resources move from one area to the other with the probability specified during model initialization.
 
@@
 Behaviors may be modified at each time step by the key processes described below.
+
+*Key processes*
 
@@
 ```
+*Multiple competing groups*
 
@@
 A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group (see Table \@ref(tab:tabAbmVars)).
+
+*Outcomes*
 
@@
 I will compare standardized parameter estimates, to determine which factors (geographic distance or social relatedness) most strongly contribute to copying fidelity in this study system.
-Similar analyses will be used to determine which factors affect cookstove primary outcome variables (fuel efficiency and particulate matter production).
+Similar analyses will be used to determine which factors, if any, are strongly associated with stove fuel efficiency.
 

### 2020-10-08 DissertationProposal/DissertationProposal.Rmd
DUN!

@@
 The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission.
-In situations where behaviors do emerge and spread within a population, as the behavior is transmitted along the chain of individuals, it may be modified considerably to the point where the behavioral outcome is unrecognizable [@lewis_transmission_2012; @mcguigan_over-imitation_2011].
+In situations where behaviors do emerge and spread within populations, behaviors are transmitted along a chain of individuals, where they may be modified considerably to the point where the behavioral outcome is unrecognizable [@lewis_transmission_2012; @mcguigan_over-imitation_2011].
 This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves.
@@
 The majority of Pemban wards are described as forest-dependent.
-The Department of Forests and Non-Renewable Natural Resources allows wards to enter into Community Forest Management Agreements (CoFMAs), which among other things, established forest protected areas and exclude outside individuals [@benjaminsen2014between].
+The Department of Forests and Non-Renewable Natural Resources allows wards to enter into Community Forest Management Agreements (CoFMAs), which among other things, establish forest protected areas and exclude outside individuals [@benjaminsen2014between].
 As of 2018, 18 Pemban wards had adopted CoFMAs [@andrews_cultural_2018].
@@
 This landscape is divided into two types, working and protected.
-Resources move between these patches at each time step with some probability.
+Resources move between these types at each time step with some probability.
 During initialization, the protected and working landscapes are both at an equal, designated percent carrying capacity.
@@
 
-In the multiple competing groups scenario, two groups with separate resource pools will follow the processes described above.
-However, at the end of each time step the groups will compare the average harvest rate each group.
-Whichever group has the highest average harvest rate for that timestep will report the average harvest times for the protected and working landscapes.
+In the multiple competing groups scenario, two groups with separate resource pools will each follow the processes described above.
+However, at the end of each time step the groups will compare the average harvest rates.
+We assume that the underperforming group can observe the average harvest strategy (time spent in the protected or working areas) of the better performing group.
 A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group (see Table \@ref(tab:tabAbmVars)).
@@
 
-In both single and competing group scenarios, the dependent variables of interest are as follows: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step.
+In both single and multiple group scenarios, the dependent variables of interest are as follows: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step.
 Full defection is defined as all agents harvesting from the protected landscape for a proportion of time equal to the percent of the resource which is protected.
@@
 I have developed a model which follows the above processes for one group with one resource pool.
-An interactive application which performs one run of that model can is found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
+An interactive application which performs one run of that model can be found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
 
@@
 Environmental conditions can be set to provide neither an over abundance of resources nor a restrictive scarcity.
-Under these conditions, we see conformity biases maintain starting group norms around harvest behaviors.
-Of greater interest is that under success biased learning, we observe that protected areas enhance agent wellbeing (average harvest) initially, but defection eventually spreads and causes a failure of the protected area.
+Under these conditions, we see conformity-biases maintain starting group norms around harvest behaviors.
+Of greater interest is that under success-biased learning, we observe that protected areas enhance agent wellbeing (average harvest) initially, but defection eventually spreads and causes a failure of the protected area as seen in Fig \@ref(fig:abmOutput).
 With this in-mind, the next step for this research is to introduce a second group dynamic which will allow us to test the logic of cultural group selection for costly conservation behaviors under these conditions.
-Example outputs from this model can be seen in Fig \@ref(fig:abmOutput).
 
-```{r abmOutput, echo=FALSE, fig.cap="Fifteen example runs of the current, single group agent-based model", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch1/ABM_Output.png")
+```{r abmOutput, echo=FALSE, fig.cap="Fifteen example runs of the current, single group agent-based model.
+CC indicates the carrying capacity of the protected and working landscapes respectively.
+This model has been specified to neither provide an over abundance or scarcity of resources.
+Agents all learn via a success-biased strategy.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/ABM_Output2.png")
 ```
@@
 This project benefits from preliminary work done by our collaborators at the Max Planck Institute for Evolutionary Anthropology (MPI).
-MPI researchers conducted a preliminary qualitative survey and coalition of oral histories with the assistance of Department of Forestry staff to confirm anecdotally that competition between wards has led to CoFMA adoption at the ward level.
+MPI researchers conducted a preliminary qualitative survey and collation of oral histories with the assistance of Department of Forests and Non-Renewable Natural Resources staff to confirm anecdotally that competition between wards has led to CoFMA adoption at the ward level.
 These researchers have also visited wards across the island and recorded which wards have adopted CoFMAs and which have not.
@@
 The primary outputs of data collection and analysis in this project will produce binary predictions of protected area adoption for each study ward, as well as the calculated strength of selection for protected area adoption for each study ward.
-I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of community managed protected areas.
-(Figure \@ref(fig:selPress)) shows an example of what the results of this study might look like.
+I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of CoFMAs.
+Figure \@ref(fig:selPress) shows an example of what the results of this study might look like.
 The findings of this study will be submitted for publication in a peer reviewed, academic journal.
@@
 While this project will inform effective implementation of household level conservation projects on the ground, it will also provide important insights for understanding behavioral adoption more generally.
-While the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially [@lyons_hidden_2007; @mcguigan_over-imitation_2011].
+Although the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially [@lyons_hidden_2007; @mcguigan_over-imitation_2011].
 In a foundational study, researchers show that learned behavior is reconstructed and assimilated with an individual’s prior experiences, rather than replicated directly [@bartlett1932remembering].
@@
 While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics.
-I will categorize these characteristics as functional (stove size) or nonfunctional (decorative qualities).
+I will categorize these characteristics as functional (e.g.
+stove size) or nonfunctional (e.g.
+decorative qualities).
 I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from in both functional and nonfunctional characteristics.
@@
 The second step will empirically validate these theoretical insights by comparing calculated cultural group selection pressures to real-world adoption of Community Forest Management Agreements on Pemba Island, Tanzania.
-Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities, to specifically examine how behavioral efficiency may be altered by social contexts (e.g.
+Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities to examine how behavioral efficiency may be altered by social contexts (e.g.
 communication frequency) as a behavior spreads throughout a population.

### 2020-10-08 DissertationProposal/Figures/Ch1/Ch1_Plots.R

@@ -45,7 +45,7 @@ p1<-dat%>%filter(NO_TAKE =="All")%>%filter(STATUS_YR>1970)%>%group_by(STATUS_YR)
         legend.title = element_text(colour="black", size=12),
         legend.text = element_text( size = 12),text=element_text(family="Calibri"),
         legend.position = "none")+
-  xlab("Year")+ylab("Number of Newly Established \n No-Take Conservation Areas")+ggtitle("Newly Established No-Take Conservation Areas")
+  xlab("Year")+ylab("Number of Newly Established \n No-Take Protected Areas")+ggtitle("Newly Established No-Take Protected Areas")
 
 
 

### 2020-11-09 Abm_MultipleReplicationsPlots.R
Fixed problem with working lands sucking protected resources

@@ -1,22 +1,22 @@
 abm<-function(#Specified parameters
   
-  Runs=3,
+  Runs=10,
   Individuals=100, #number of total resource users in a population
   
   TotalCarryingCapacity=10000, #total available resource units
   
-  StartPercCarryingCapacity = 0.7, #amount of resources available in the landscape at the start in proportion to CC
+  StartPercCarryingCapacity = 0.65, #amount of resources available in the landscape at the start in proportion to CC
   
-  PercProtected=0.15, #percent of the total resource that's in a protected area
+  PercProtected=0.4, #percent of the total resource that's in a protected area
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
   LearningStrategy = "Success Bias", #options are Success Bias & Conformist
   
-  TimeSteps=75,
+  TimeSteps=30,
   ResourceRegenerationPerTimeStep=1.5,
-  harvestMax=15,
-  ProbOfMobility=0.15){
+  harvestMax=17,
+  ProbOfMobility=0.2){
   
   ## Create output dfs
   
@@ -59,7 +59,7 @@ abm<-function(#Specified parameters
                        PayoffProtectedLastTime = rep(NA,Individuals),
                        PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
     
-    ProtectPerDefect<-as.integer(StartResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
+    ProtectPerDefect<-ifelse(DefNumStart ==0,0,as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
     WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
     
     for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
@@ -160,23 +160,28 @@ abm<-function(#Specified parameters
       
       #New resource pools to pull from
       #make sure they dont regenerate past their carrying capacity
-      
       NewProtectedResourcesTotal<-round(outputNumResourcesProtect[t-1,r] * ResourceRegenerationPerTimeStep,digits=0)
[…]

### 2020-11-09 ProtectedAreasABM.R

@@ -34,7 +34,7 @@ abm<-function(#Specified parameters
                      PayoffProtectedLastTime = rep(NA,Individuals),
                      PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
   
-  ProtectPerDefect<-as.integer(StartResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
+  ProtectPerDefect<-ifelse(DefNumStart ==0,0,as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
   WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
   
   for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
@@ -144,20 +144,32 @@ abm<-function(#Specified parameters
     
     
     #New resource pools to pull from
-    #make sure they dont regenerate past their carrying capacity
+    #Resources over the CC move with probability 'ProbOfMobility'
     
     NewProtectedResourcesTotal<-round(output$NumResourcesProtect[t-1] * ResourceRegenerationPerTimeStep,digits=0)
-    NewProtectedResourcesTotal<-ifelse(NewProtectedResourcesTotal<=  TotalCCResourceProtected,NewProtectedResourcesTotal, TotalCCResourceProtected)
+    ProtectedResourcesOverCC<-round(NewProtectedResourcesTotal-TotalCCResourceProtected,digits=0)
+    ProtectedResourcesOverCC<- ifelse(ProtectedResourcesOverCC<=  0,0, ProtectedResourcesOverCC)
+    
     NewWorkingResourcesTotal<-round(output$NumResourcesWorking[t-1] * ResourceRegenerationPerTimeStep,digits=0)
-    NewWorkingResourcesTotal<-ifelse(NewWorkingResourcesTotal<=TotalCCResourceWorking,NewWorkingResourcesTotal,TotalCCResourceWorking)
+    WorkingResourcesOverCC<-round(NewWorkingResourcesTotal-TotalCCResourceWorking,digits=0)
+    WorkingResourcesOverCC<- ifelse(WorkingResourcesOverCC<=  0,0, WorkingResourcesOverCC)
+    
     
     ##Resource mobility
-    LeaveWorking<-rbinom(1,NewWorkingResourcesTotal,ProbOfMobility) #number of resources which leave the protected area
[…]

### 2020-11-09 app.R

@@ -43,7 +43,7 @@ ui<-shinyUI(navbarPage("Protected Areas ABM",
                                                       sliderInput("startCooperators",
                                                                   "Starting Obligatory Cooperators",
                                                                   min = 0.0,
-                                                                  max = 0.99,
+                                                                  max = 1.0,
                                                                   value = 0.9),
                                                       sliderInput("resourceRegen",
                                                                   "Resource Regeneration Rate",
@@ -223,8 +223,16 @@ server <- function(input, output) {
             NewWorkingResourcesTotal<-ifelse(NewWorkingResourcesTotal<=TotalCCResourceWorking,NewWorkingResourcesTotal,TotalCCResourceWorking)
             
             ##Resource mobility
-            LeaveWorking<-rbinom(1,NewWorkingResourcesTotal,ProbOfMobility) #number of resources which leave the protected area
-            LeaveProtected<-rbinom(1,NewProtectedResourcesTotal,ProbOfMobility)#number of resources which leave the working area
+            
+            
+            #all resources mobile
+            #LeaveWorking<-rbinom(1,NewWorkingResourcesTotal,ProbOfMobility) #number of resources which leave the protected area
+            #LeaveProtected<-rbinom(1,NewProtectedResourcesTotal,ProbOfMobility)#number of resources which leave the working area
+            
+            #make it so only new resources can leave 
+            LeaveWorking<-rbinom(1,(NewWorkingResourcesTotal-output$NumResourcesWorking[t-1]),ProbOfMobility) #number of resources which leave the protected area
+            LeaveProtected<-rbinom(1,(NewProtectedResourcesTotal-output$NumResourcesProtect[t-1]),ProbOfMobility)#number of resources which leave the working area
+            
             
             #do the accounting on entering vs leaving individuals
             NewWorkingResourcesTotal<-NewWorkingResourcesTotal-LeaveWorking+LeaveProtected

### 2020-11-10 Abm_MultipleReplicationsPlots.R
small abm changes

@@ -275,20 +275,20 @@ abm<-function(#Specified parameters
   
 }
 
-abm(Runs=4,Individuals=100, #number of total resource users in a population
+abm(Runs=20,Individuals=100, #number of total resource users in a population
     
     TotalCarryingCapacity=10000, #total available resource units
     
     StartPercCarryingCapacity = 0.65, #amount of resources available in the landscape at the start in proportion to CC
     
-    PercProtected=0.7, #percent of the total resource that's in a protected area
+    PercProtected=0.5, #percent of the total resource that's in a protected area
     
-    CoopPercStart=1.0, #percent of individuals who start by following the rules at t0
+    CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
     
     LearningStrategy = "Success Bias", #options are Success Bias & Conformist
     
     TimeSteps=30,
-    ResourceRegenerationPerTimeStep=1.1,
+    ResourceRegenerationPerTimeStep=1.3,
     harvestMax=22,
     ProbOfMobility=0.99)
 

### 2020-11-10 MultipleReplicationsNoPlots.R

@@ -1,22 +1,22 @@
 abmnp<-function(#Specified parameters
   
-  Runs=3,
+  Runs=10,
   Individuals=100, #number of total resource users in a population
   
   TotalCarryingCapacity=10000, #total available resource units
   
   StartPercCarryingCapacity = 0.65, #amount of resources available in the landscape at the start in proportion to CC
   
-  PercProtected=0.15, #percent of the total resource that's in a protected area
+  PercProtected=0.4, #percent of the total resource that's in a protected area
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
   LearningStrategy = "Success Bias", #options are Success Bias & Conformist
   
-  TimeSteps=75,
+  TimeSteps=30,
   ResourceRegenerationPerTimeStep=1.5,
-  harvestMax=15,
-  ProbOfMobility=0.15){
+  harvestMax=17,
+  ProbOfMobility=0.2){
   
   ## Create output dfs
   
@@ -59,7 +59,7 @@ abmnp<-function(#Specified parameters
                        PayoffProtectedLastTime = rep(NA,Individuals),
                        PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
     
-    ProtectPerDefect<-as.integer(StartResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
+    ProtectPerDefect<-ifelse(DefNumStart ==0,0,as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
     WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
     
     for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
@@ -160,23 +160,28 @@ abmnp<-function(#Specified parameters
       
       #New resource pools to pull from
       #make sure they dont regenerate past their carrying capacity
-      
       NewProtectedResourcesTotal<-round(outputNumResourcesProtect[t-1,r] * ResourceRegenerationPerTimeStep,digits=0)
-      NewProtectedResourcesTotal<-ifelse(NewProtectedResourcesTotal<=  TotalCCResourceProtected,NewProtectedResourcesTotal, TotalCCResourceProtected)
[…]

### 2020-11-11 GoodPembaMap.R
Made a siiick map!

@@ -1,0 +1,78 @@
+library(sf)
+library(tidyverse)
+library(rmapshaper)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(ggspatial)
+
+worldmap <- ne_countries(scale = 'medium', type = 'map_units',
+                         returnclass = 'sf')
+africamap <- ne_countries(scale = 'medium', type = 'map_units',
+                         returnclass = 'sf',continent="africa")
+# have a look at these two columns only
+head(worldmap[c('name', 'continent')])
+
+
+ggplot() + geom_sf(data = worldmap) + theme_bw()
+ggplot() + geom_sf(data = africamap) + theme_bw()
+
+
+
+TZ <- africamap[africamap$name == 'Tanzania',]
+ggplot() + geom_sf(data = TZ) + theme_bw()
+
+
+#x<-read_sf("~/Pemba_Project/MapData/ne_110m_land.shp")
+Tanzania<-read_sf("~/Pemba_Project/MapData/TZA_adm0.shp")
+ggplot() + geom_sf(data = Tanzania) + theme_bw()
+
+EastAfrica <- st_crop(africamap, xmin = 32, xmax = 60,
+                          ymin = -26, ymax = 5)
+
+EAfrFin<-ggplot() + geom_sf(data = EastAfrica, fill="antiquewhite")+coord_sf(expand = FALSE)+ theme_bw()+
+  geom_rect(aes(xmin=39.5, xmax=40.1, ymin=-5.7, ymax=-4.5),fill="#b2df8a", color="#33a02c", alpha=0.5)+
+annotate(geom = "text", x = 38, y = 0.8, label = "Kenya", fontface = "italic", color = "grey22", size = 4)+
+  annotate(geom = "text", x = 35, y = -6, label = "Tanzania", fontface = "italic", color = "grey22", size = 4)+
+  annotate(geom = "text", x = 47, y = -18, label = "Madagascar", fontface = "italic", color = "grey22", size = 4)+
+  annotate(geom = "text", x = 37.5, y = -13, label = "Mozambique", fontface = "italic", color = "grey22", size = 4)+
+   theme( panel.background = element_rect(fill = "#c6dbef"),panel.grid.major = element_blank())+
+   annotation_scale(location = "bl", width_hint = 0.5) +ylab("")+xlab("")+
+  annotation_north_arrow(location = "bl", which_north = "true", 
[…]

### 2020-11-11 MultipleReplicationsNoPlots.R

@@ -59,7 +59,7 @@ abmnp<-function(#Specified parameters
                        PayoffProtectedLastTime = rep(NA,Individuals),
                        PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
     
-    ProtectPerDefect<-ifelse(DefNumStart ==0,0,as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
+    ProtectPerDefect<-ifelse(DefNumStart ==0,as.integer(StartResourceProtected),as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
     WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
     
     for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
@@ -239,4 +239,55 @@ abmnp<-function(#Specified parameters
 }
 
 abmnp(Runs=5,LearningStrategy = "Conformist")
+
+#fill a dataframe with the means
+
+FillData<-data.frame(PercProtected=seq(0.1,1,0.1),ResourceMobility=seq(0.1,1,0.1))
+FillData<-expand.grid(PercProtected=seq(0.1,1,0.1), ResourceMobility=seq(0.1,1,0.1))
+FillData$meanPayoff<-rep(NA,nrow(FillData))
+FillData%<>%mutate(PercProtected=round(PercProtected,digits=1))%>%mutate(ResourceMobility=round(ResourceMobility,digits=1))
+
+
+for (i in 1:nrow(FillData)){
+  abmnp(Runs=3,Individuals = 100,TotalCarryingCapacity=10000,
+      StartPercCarryingCapacity = 0.65,
+      PercProtected = FillData[i,]$PercProtected, 
+      ProbOfMobility = FillData[i,]$ResourceMobility,CoopPercStart=1.0, 
+      LearningStrategy = "Success Bias",
+      TimeSteps=30,
+      ResourceRegenerationPerTimeStep=1.3,
+      harvestMax=25)
+  FillData[i,3]<-apply(FullOutput$meanPayoff[30,1:3],1,mean)  
+  
+}
+
[…]

### 2020-11-11 DissertationProposal/DissertationProposal.Rmd
added good map. Should probably actually use this in the earlier intro part?

@@
 ### Methodological Specifics
-I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi (aka Before Ningeze Forest) wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)).
+I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)).
 With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.
@@
 Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "75%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch3/BadPembaMap.png")
+knitr::include_graphics("Figures/Ch3/MapLines.png")
 ```

### 2020-11-11 GoodPembaMap.R

@@ -35,11 +35,12 @@ annotate(geom = "text", x = 38, y = 0.8, label = "Kenya", fontface = "italic", c
   annotate(geom = "text", x = 35, y = -6, label = "Tanzania", fontface = "italic", color = "grey22", size = 4)+
   annotate(geom = "text", x = 47, y = -18, label = "Madagascar", fontface = "italic", color = "grey22", size = 4)+
   annotate(geom = "text", x = 37.5, y = -13, label = "Mozambique", fontface = "italic", color = "grey22", size = 4)+
-   theme( panel.background = element_rect(fill = "#c6dbef"),panel.grid.major = element_blank())+
+   theme( panel.background = element_rect(fill = "#c6dbef"),panel.grid.major = element_blank(),
+          axis.text=element_blank(), axis.ticks = element_blank())+
    annotation_scale(location = "bl", width_hint = 0.5) +ylab("")+xlab("")+
   annotation_north_arrow(location = "bl", which_north = "true", 
-                         pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), 
-                         style = north_arrow_fancy_orienteering)
+                         pad_x = unit(0.1, "in"), pad_y = unit(0.5, "in"), 
+                         style = north_arrow_fancy_orienteering)+theme(plot.margin=unit(c(0.5,-4,0.5,0),"cm"))
 
 
 
@@ -60,14 +61,14 @@ Pem_fin<-ggplot() + geom_sf(data = Pemba, fill="antiquewhite")+coord_sf()+ theme
             size = 4,
             fontface = "bold",
             hjust = c(0,-0.3,1.1),vjust = c(-1,0,0)) +ylab("")+xlab("")+
-  theme( panel.background = element_rect(fill = "#c6dbef"),panel.grid.major = element_blank())+
-  annotation_scale(location = "bl", width_hint = 0.5) 
+  theme( panel.background = element_rect(fill = "#c6dbef"),panel.grid.major = element_blank(),
+         axis.text=element_blank(), axis.ticks = element_blank())+
+  annotation_scale(location = "bl", width_hint = 0.5) +theme(plot.margin=unit(c(0.5,0,0.5,-4),"cm"))
   
 
 Map<-gridExtra::grid.arrange(EAfrFin,Pem_fin,nrow=1)
 
-EAfrFin
-+ geom_abline(intercept = 37, slope = -5)
+
 
 #alternative method. Sloppy right now
 library(grid)

### 2020-11-12 GoodPembaMap.R
inset map

@@ -21,6 +21,7 @@ ggplot() + geom_sf(data = africamap) + theme_bw()
 TZ <- africamap[africamap$name == 'Tanzania',]
 ggplot() + geom_sf(data = TZ) + theme_bw()
 
+#Two equal sized maps with reduced margins in between
 
 #x<-read_sf("~/Pemba_Project/MapData/ne_110m_land.shp")
 Tanzania<-read_sf("~/Pemba_Project/MapData/TZA_adm0.shp")
@@ -70,10 +71,51 @@ Map<-gridExtra::grid.arrange(EAfrFin,Pem_fin,nrow=1)
 
 
 
-#alternative method. Sloppy right now
+#alternative method. Inset map
 library(grid)
+Pemba <- st_crop(Tanzania, xmin = 39.1, xmax = 39.9,
+                 ymin = -5.6, ymax = -4.9)
+
+sites<-data.frame(Site=c("Makangale","Kifundi","Msituu mkuu"),
+                  lon=c(39.690753,39.710927,39.837124),
+                  lat=c(-4.906206,-4.940105,-4.997158))
+
+Pem_fin<-ggplot() + geom_sf(data = Pemba, fill="antiquewhite")+coord_sf(xlim = c(39.4,39.9))+ theme_bw()+
+  geom_point(data = sites, aes(x = lon, y = lat), size = 4, 
+             shape = 18) +
+  annotate(geom = "text", x = 39.77, y = -5.13, label = "Pemba Island", fontface = "italic", color = "grey22", size = 4)+
+  geom_text(data = sites, 
+            aes(x=lon, y=lat, label = Site),
+            size = 4,
+            fontface = "bold",
+            hjust = c(0,-0.3,1.1),vjust = c(-1,0,0)) +ylab("")+xlab("")+
+  theme( panel.background = element_rect(fill = "#c6dbef"),panel.grid.major = element_blank(),
+         axis.text=element_blank(), axis.ticks = element_blank()
+         )+
+  annotation_north_arrow(location = "bl", which_north = "true", 
+                         pad_x = unit(0.1, "in"), pad_y = unit(0.5, "in"), 
+                         style = north_arrow_fancy_orienteering)+
+  annotation_scale(location = "bl", width_hint = 0.5) +theme(plot.margin=unit(c(0.5,0,0.5,0),"cm"))
+
+EastAfrica <- st_crop(africamap, xmin = 32, xmax = 60,
[…]

### 2020-11-12 ProtectedAreasABM.R

@@ -34,7 +34,7 @@ abm<-function(#Specified parameters
                      PayoffProtectedLastTime = rep(NA,Individuals),
                      PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
   
-  ProtectPerDefect<-ifelse(DefNumStart ==0,0,as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
+  ProtectPerDefect<-ifelse(DefNumStart ==0,as.integer(StartResourceProtected),as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
   WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
   
   for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 

### 2020-11-12 GoodPembaMap.R
shifted pemba right slightly

@@ -80,7 +80,7 @@ sites<-data.frame(Site=c("Makangale","Kifundi","Msituu mkuu"),
                   lon=c(39.690753,39.710927,39.837124),
                   lat=c(-4.906206,-4.940105,-4.997158))
 
-Pem_fin<-ggplot() + geom_sf(data = Pemba, fill="antiquewhite")+coord_sf(xlim = c(39.4,39.9))+ theme_bw()+
+Pem_fin<-ggplot() + geom_sf(data = Pemba, fill="antiquewhite")+coord_sf(xlim = c(39.36,39.86))+ theme_bw()+
   geom_point(data = sites, aes(x = lon, y = lat), size = 4, 
              shape = 18) +
   annotate(geom = "text", x = 39.77, y = -5.13, label = "Pemba Island", fontface = "italic", color = "grey22", size = 4)+

### 2020-11-24 DissertationProposal/DissertationProposal.Rmd
Totally redid abm descriotion

@@
 library(extrafont)
-
 ```
@@
 
-*Single group setup and basic function*
+*Current Single Group Model - The Most Simple Case(s)*
 
-In the most simple scenario, the agent-based model represents a simplistic world where there is one homogeneous group of agents who harvest resources from a landscape.
-This landscape is divided into two types, working and protected.
-Resources move between these types at each time step with some probability.
-During initialization, the protected and working landscapes are both at an equal, designated percent carrying capacity.
-At each time step, agents aim to maximize their harvest up until a designated harvest max.
-During initialization, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree, spending a small portion of their time harvesting from the protected landscape.
+In the most basic form, the current agent-based model demonstrates repeated subtractions from a multiplicatively regenerating quantity of resources.
+In this context, there are no resource protections and therefore agents cannot modify their harvest strategies.
+We can write a simple analytical model to better denote these dynamics and determine interesting and appropriate starting conditions for more complex permutations of the agent-based model, as well as validate our intuitions as to how the model should behave.
+Given that the resources regenerate multiplicatively, any starting conditions which cause the resource pool to diminish after the first time step will further deplete the ever-shrinking resource pool with each additional time step.
+Specifically, any max harvest value which when multiplied by the number of agents is greater than the number of resources that regenerate in the first time step, will cause the resource pool to continually diminish over time and eventually reach zero.
+We can define the sustainable harvest threshold (${H_{sust}}$) for any given number of agents (${N}$), starting resources (${W_{r}}$), and regeneration rate (${R}$) as:
 
-The model also allows for ecological variation during the initial specification.
-The variable ecological starting conditions are as follows: Resource mobility designates the probability that a resource unit will move from the protected landscape into the working landscape or vice versa at each time step.
-Protected area size demarcates what percentage of the available resource units are within the protected area.
-Percent carrying capacity sets the total number of harvestable resource units as a percent of the landscape carrying capacity.
-Lastly, the resource regeneration rate designates what percentage the resource regenerates at each time step (see Table \@ref(tab:tabAbmVars)).
+$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)}{N} $$
+
+Figure \@ref(fig:FirstABM) shows the model behavior when the maximum harvest value is at the calculated sustainable harvest value for this scenario.
+We can see that the amount of resources harvested in each time step is below the amount regenerated, therefore we see no change in the available and harvested resources over time.
+
+```{r FirstABM, out.width = "100%",echo=FALSE, fig.align='center', dpi=72, fig.pos= "H",fig.cap="Resource quantity and mean harvest values over time for one run of the agent-based model at a maximum harvest value exactly at the calculated value for sustainable harvest.
+Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
+The regeneration factor at each time step is 1.20."}
+knitr::include_graphics("Figures/Ch1/abmP1a.png")
+```
+
+In contrast to the figure above, figure \@ref(fig:FirstABMb) shows the model behavior when the maximum harvest value is just one over the calculated sustainable harvest value.
+In this scenario, we can see that the resource pool becomes more depleted over time and the quantity of resources regenerated at each time step is reduced proportionally.
+
+```{r FirstABMb, out.width = "100%",echo=FALSE, fig.cap="Resource quantity and mean harvest vales over time for one run of the agent-based model at a maximum harvest valuue one above the calculated value for sustainable harvest.
+Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
+The regeneration factor at each time step is 1.20.", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/abmP1.png")
+```
+
+We can extend this model to explore how the sink/source dynamics between protected areas and adjacent working landscapes may provision more renewable resources than working landscapes alone.
+In this formation, we still have one homogenous group of agents who harvest regenerating resources from a landscape.
+This landscape now has some specified percentage sectioned off and inaccessible to the agents — the protected area.
+At every time step agents will attempt to harvest resources from the working landscape up until the specified maximum harvest amount.
+The outcome of this harvest attempt for each agent is the result of one random draw from a Binomial distribution where the number of trials is equal to the number of resources available on the working landscape divided by the number of individuals (i.e.
+that individual’s share of the available resources) and the probability of success is equal to the percent of an agent’s harvest time spent in the working landscape; in this scenario we are forcing obligatory cooperation so that percentage is always 100%.
+This will not always be the case as we build complexity in the model.
+
+After the agents harvest resources from the landscape, the resources in the working and protected areas will reproduce by the specified regeneration factor.
+If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e.
+not be available for agents in the following time step) or move to the other area (protected or working) with some specified probability also applied to a Binomial distribution.
+Resources which move to the adjacent area but exceed the carrying capacity will also die.
+If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the protected area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).
+
+$$P_{mob} = (R*P_{r})-P_{r}$$
+
+$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)+min\{P_{mob},W_{cc}-W_{r}\}}{N} $$
+
+Note that the sustainable harvest value is now dependent on the percent of the resources in the working landscape vs the protected area.
+We can run the model for every combination of protected area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (fig \@ref(fig:abm2)).
+In the figure below, we can see that up until the maximum sustainable harvest value for this scenario (16), agents will always reach the set maximum harvest value until the resources are no longer made available to them (i.e.
+all protected).
+Above the maximum sustainable harvest value, the size of the protected area becomes greatly important for determining average harvest payoff to agents.
+In this scenario where the resources regenerate at a relatively high rate (20%) and the probability of mobility is 100%, the sink/source dynamics allow agents to reach the greatest average harvest value when the protected area is large and therefore provisioning many resources to the ever depleting working landscape at every time step.
+
+```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) for many combinationa of maximum harvest values and protected area sizes.
+Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
+The regeneration factor at each time step is 1.20.", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/abmp2.png")
+```
+
+*Current Single Group Model - Incorporating Uncertainty*
+
+Up to this point, we have assumed that excess resources produced in the protected area will move to the working landscape, up until a specified carrying capacity, with 100% probability.
+Key insights from this model are gained from experimenting with different probabilities of resource mobility.
+To frame this in the context of real world protected areas, we can assume that insights from model runs with high probabilities of resource mobility may be applied to wide ranging organisms such as large aquatic animals, while model runs with low probabilities of resource mobility may be applied to relatively low vagility organisms such short or large-seeded trees [@thomson2011seed].
+
+The probability of resource mobility is operationalized in the model by using random draws from a Binomial distribution.
+For every resource unit regenerated over the carrying capacity of the protected area, we take one random binomial draw with the probability of success equal to the probability of resource mobility.
+With all other model specifications kept equal and a maximum harvest value set slightly above the maximum sustainable harvest value for a working landscape alone, we can explore the interaction between protected area size and resource mobility given obligatory cooperation.
+
+```{r abm3, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model.
+Outcomes are the result of many combinations of resource mobilities and protected area sizes.
+Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
+The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25.", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/abmp3.png")
+```
+
+*Current Single Group Model - Cooperation, Defection, & Social Learning*
+
+To apply this model to conservation adoption and different mechanisms of social learning, we can allow agents to alter their harvest strategies beyond obligatory cooperation around protected area regulations.
+During model initialization, we can set the percentage of agents who we will allow to spend a proportion of their harvest effort in the protected area (defecting).
+The proportion of effort an individual agent spends in the protected area is drawn from the beta distribution below (fig \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.
+
+```{r betastart, out.width = "75%",echo=FALSE, fig.align='center', fig.pos= "H", fig.cap="One thousand random draws from a beta distribution with the shape 1 and 2 parameters set at 1 and 2 respectively.
+This distribution is used to assign the harvest effort spent in the protected area for defecting agents during model initialization."}
+hist(rbeta(1000,1,2),main=NULL)
+```
+
+At each time step, all agents first harvest resources from the working landscape.
+The number of resources agents receive from the working landscape is determined by one draw from the binomial distribution where the number of trials is equal to the number of resources in the working landscape divided by the number of agents and the probability of success is equal to the percentage of effort that agent spends in the working landscape.
+Agents only harvest resources up to the specified maximum harvest value.
+
+After harvesting from the working landscape, agents who defect to some degree and did not already reach the maximum harvest value, spend the remainder of their harvest effort in the protected area.
+Following the same logic as harvest from the working landscape, the number of resources agents receive from the protected area is determined by one draw from the binomial distribution where the number of trials is equal to the number of resources in the protected area divided by the number of agents who defect in that time step and the probability of success is equal to the percentage of effort that agent spends in the protected area.
+
+Agents modify their behavior (the percentage of time they spend in each area) at each time step according to one of two learning strategies.
+If agents are following a success-biased learning strategy, they are paired with another agent at random at each time step and compare strategies.
+If one agent performs worse than another in the previous time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them.
+If agents are following a conformist-biased strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group.
+The learning strategy of the overall population is currently set to either success-biased or conformist-biased learning for all agents, but in future versions of the model, it will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents.
+The entire model process is visualized conceptually in the figure below and a table of all the model starting conditions, key variables, and outcomes can be found in Table \@ref(tab:tabAbmVars)).
+
+```{r conceptDiagram, echo=FALSE, fig.cap="Concept diagram of agent-based model scheduling under single group dynamics", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/ABMConceptDiagramFinal.png")
+```
+
+With all model parameters kept constant as in the output shown in (fig \@ref(fig:abm3)), we can run the model with just 5% of the agents assigned as having the possibility to defect at initialization and all agents modifying harvest strategies via success-biased learning.
+The average final time step outcomes from five runs of this model can be seen in (fig \@ref(fig:abm4)).
+
+```{r abm4, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model.
+Outcomes are the result of many combinations of resource mobilities and protected area sizes.
+Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
+The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25.
+Five percent of agents assigned to defect at initialization.
+Agents learn through success-biased learning.", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/abmp4.png")
+```
+Allowing agents to modify harvest strategies and harvest from the protected area greatly reduces the overall harvest outcomes.
+To better understand the underlying dynamics, we set the starting model parameters to the ideal conditions for maximum payoff as identified in (fig \@ref(fig:abm3)) and observe the model behavior at each time step.
+
+```{r abm5, out.width = "100%",echo=FALSE, fig.cap="Five runs of the agent-based model.
+Thin black lines show each run.
+Thick black lines show the average values across runs.
+Starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
+Eighty percent of resources are inside the protected area and the probability of resource mobility is 100 percent.
+The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25.
+Five percent of agents assigned to defect at initialization.
+Agents learn through success-biased learning.", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/ABM_Output.png")
+```
+
+Model dynamics show that over time, agents begin to harvest from the protected area as they copy more successful individuals.
+This behavior depletes the resource pool in the protected area and lowers the number of resources which regenerate and agents can harvest in the subsequent time step.
+Note that full defection is defined as agents spending an equal proportion of their effort in the protected area as the proportion protected (i.e.
+they treat the entire landscape equally).
+
+The results displayed above are for one homogenous group of agents harvesting resources from a common pool of resources.
+This model is useful for exploring the sink/source utility of protected areas to ensure long-term resource provisioning, especially in the contexts of social learning.
+To further explore this model, I have created an interactive application which performs one run of the model and can be found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
+Comparing these model runs to simple analytical models and our intuition about how the model should behave also allows us to ensure that the model is functioning properly.
+To address the research question of interest however, we will need to update the model to include multiple interacting groups.
 
@@
                                      "Resource Regeneration",  "Resource Mobility","Groups","Learning Strategy", 
-                                     "Percent Cooperation", "Average Payoff", "Protected Area Degredation",
-                                     "Working Landscape Degredation"), 
+                                     "Percent Cooperation", "Average Payoff", "Protected Area Degradation",
+                                     "Working Landscape Degradation"), 
                       Type = c("Starting Condition","Starting Condition","Starting Condition",
@@
                       Description = linebreak(c("Number of agents in simulation",
-                                      "Maximum number of resources\nallowed on landscape",
-                                      "Percent of carrying capacity at\ninitialization",
-                                      "Percent of landscape in\nprotected area",
-                                      "Starting percent of agents\nfully cooperating",
-                                      "Number of time steps\nsimulation will run for",
-                                      "Maximum number of resources\nagents will harvest during\none time step",
-                                      "The percent of available resources\nwhich will regenerate at each time\nstep in both protected and\nworking landscapes",
-                                      "The probability that any available\nresource will move from the\n protected to working landscape\nor vice versa",
-                                      "Solitary or multiple\ncompeting groups",
-                                      "Assigns a percent of the population\nto success-biased imitation and\nthe remaining population to\nconformist-biased imitation",
-                                      "Percent of total population\nharvest time spent in\nthe working landscape","Mean harvest for all agents\nat each time step",
-                                      "Working landscape percent of\ntotal carrying capacity",
-                                      "Protected area percent of\ntotal carrying capacity")))
-knitr::kable(abmVars, caption = "Description of all model parameters.",escape=F) %>% 
-  kable_styling(c("striped",full_width=F))%>% row_spec(0,bold=TRUE)
+                                      "Maximum number of resources allowed on landscape",
+                                      "Percent of carrying capacity at initialization",
+                                      "Percent of landscape in protected area",
+                                      "Starting percent of agents fully cooperating",
+                                      "Number of time steps simulation will run for",
+                                      "Maximum number of resources agents will harvest during one time step",
+                                      "The percent of available resources which will regenerate at each time step in both protected and working landscapes",
+                                      "The probability that any available resource will move from the  protected to working landscape or vice versa",
+                                      "Solitary or multiple competing groups",
+                                      "Assigns a percent of the population to success-biased imitation and the remaining population to conformist-biased imitation",
+                                      "Percent of total population harvest time spent in the working landscape","Mean harvest for all agents at each time step",
+                                      "Working landscape percent of total carrying capacity",
+                                      "Protected area percent of total carrying capacity")))
+kbl(abmVars,format="latex", caption = "Description of all model variables and outcomes",escape=F) %>% 
+  kable_classic_2(full_width=F)%>% row_spec(0,bold=TRUE)%>%
+    column_spec(1, width = "8em")%>%
+    column_spec(2, width = "8em")%>%
+  column_spec(3, width = "20em")
 
@@
 
-The environment also responds to agent behavior.
-At every time step, both the protected and working landscapes lose exactly the number of resources that agents harvest from them.
-The areas then regenerate resources according to the regeneration rate per time step and lastly, the resources move from one area to the other with the probability specified during model initialization.
+*Proposed Multi-Group Model*
 
-Agents behave according to a specific set of rules.
-Agents choose to spend a given percent of their time harvesting from the protected area (defecting) and a percent of their time harvesting from the working landscape (cooperating).
-Full defection is when the average percent of time agents spend in the protected area is equal to the percent of the landscape protected (i.e.
-they do not treat the protected area any differently than the working landscape).
-
-Agents learn their behaviors according to one of two possible learning mechanisms.
-Agents who learn via success-biased learning will preferentially copy the behavior of more successful agents when they encounter them.
-Agents who learn via conformist-biased learning will disproportionately copy the behavior of other agents they see as being in the majority, regardless of the success of those agents.
-The learning strategy of the overall population will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents.
-Behaviors may be modified at each time step by the key processes described below.
-
-*Key processes*
-
-For the given number of timesteps, the model follows a set of key processes.
-At time step one, all agents first spend their initial assigned percentage of effort harvesting from the working landscape.
-If agents reach the max harvest amount, they do not harvest from the protected landscape, even if they are assigned as having the ability to defect.
-If they do not reach harvest max, they will spend the rest of their effort harvesting from the protected area.
-They will harvest up to the harvest max if possible.
-The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape.
-Resources move between the protected and working landscapes with a probability proportional to designated probability of mobility (see Table \@ref(tab:tabAbmVars)).
-If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies.
-If one agent performs worse than another in the first time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them.
-If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group.
-This process repeats for all time steps.
-This process is summarized conceptually in Fig \@ref(fig:conceptDiagram).
-
-```{r conceptDiagram, echo=FALSE, fig.cap="Concept diagram of agent-based model scheduling under single group dynamics", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch1/ConceptDiagram.png")
-```
-*Multiple competing groups*
-
-In the multiple competing groups scenario, two groups with separate resource pools will each follow the processes described above.
+In the multiple interacting groups scenario, two groups with separate resource pools will each follow the processes described above.
 However, at the end of each time step the groups will compare the average harvest rates.
@@
 A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group (see Table \@ref(tab:tabAbmVars)).
-
-*Outcomes*
-
-In both single and multiple group scenarios, the dependent variables of interest are as follows: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step.
-Full defection is defined as all agents harvesting from the protected landscape for a proportion of time equal to the percent of the resource which is protected.
-Protected area degradation measures the percent carrying capacity of the protected landscape.
-Working landscape degradation measures the percent carrying capacity of the working landscape.
-Lastly, mean payoff to individuals measures the average payoff that agents receive at each time step.
 
@@
 
-## Preliminary Work
-I have developed a model which follows the above processes for one group with one resource pool.
-An interactive application which performs one run of that model can be found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
-
-This model currently assigns one learning strategy to all agents in the population.
-Users can select either successes-biased or conformity-biased learning to be applied.
-Future development will allow users to specify the percent of the population employing each learning strategy as specified in the methods above.
-
-Environmental conditions can be set to provide neither an over abundance of resources nor a restrictive scarcity.
-Under these conditions, we see conformity-biases maintain starting group norms around harvest behaviors.
-Of greater interest is that under success-biased learning, we observe that protected areas enhance agent wellbeing (average harvest) initially, but defection eventually spreads and causes a failure of the protected area as seen in Fig \@ref(fig:abmOutput).
-With this in-mind, the next step for this research is to introduce a second group dynamic which will allow us to test the logic of cultural group selection for costly conservation behaviors under these conditions.
-
-```{r abmOutput, echo=FALSE, fig.cap="Fifteen example runs of the current, single group agent-based model.
-CC indicates the carrying capacity of the protected and working landscapes respectively.
-This model has been specified to neither provide an over abundance or scarcity of resources.
-Agents all learn via a success-biased strategy.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch1/ABM_Output2.png")
-```
-
 ## Expected Products
@@
 This paper will show logically how we expect that protected areas function in the context of community level cooperative conservation.
-An example of how this model may inform implementation of protected areas for various resource mobilities or starting resource degradation is displayed in figures \@ref(fig:abmSim1) & \@ref(fig:abmSim2).
+I also expect that this research will provide actionable insights for how designing protected areas while considering inter-group dynamics may allow for optimal resource provisioning under a much more broad set of circumstances than considering solitary groups alone.
 
@@
 
-```{r abmSim1, echo=FALSE, fig.cap="Simulated agent-based model output of conservation adoption under single group dynamics.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch1/abmSim1.png")
-```
-
-```{r abmSim2, echo=FALSE, fig.cap="Simulated agent-based model output of conservation adoption under multiple group dynamics.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch1/abmSim2.png")
-```
 \newpage
@@
 With few studies that measure transmission fidelity in a naturalistic setting, where outcomes have real economic and health impacts - we have an opportunity to explore how environmental constraints and social learning interact when outcomes matter.
-This project seeks to further investigate how and why behaviors are modified during the transmission process and how modification (both enrounious and intentional) affect performance targets.
+This project seeks to further investigate how and why behaviors are modified during the transmission process and how modification (both erroneous and intentional) affect performance targets.
 
@@
 Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "75%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch3/MapLines.png")
+knitr::include_graphics("Figures/Ch3/InsetMap.png")
 ```

### 2020-11-24 Abm_MultipleReplicationsPlots.R

@@ -7,16 +7,16 @@ abm<-function(#Specified parameters
   
   StartPercCarryingCapacity = 0.65, #amount of resources available in the landscape at the start in proportion to CC
   
-  PercProtected=0.4, #percent of the total resource that's in a protected area
+  PercProtected=0.8, #percent of the total resource that's in a protected area
   
   CoopPercStart=0.9, #percent of individuals who start by following the rules at t0
   
   LearningStrategy = "Success Bias", #options are Success Bias & Conformist
   
   TimeSteps=30,
-  ResourceRegenerationPerTimeStep=1.5,
+  ResourceRegenerationPerTimeStep=1.3,
   harvestMax=17,
-  ProbOfMobility=0.2){
+  ProbOfMobility=0.6){
   
   ## Create output dfs
   
@@ -59,7 +59,7 @@ abm<-function(#Specified parameters
                        PayoffProtectedLastTime = rep(NA,Individuals),
                        PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
     
-    ProtectPerDefect<-ifelse(DefNumStart ==0,0,as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
+    ProtectPerDefect<-ifelse(DefNumStart ==0,as.integer(StartResourceProtected),as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
     WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
     
     for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
@@ -214,7 +214,7 @@ abm<-function(#Specified parameters
     par(mfrow=c(2,2))
     
     #mean payoff
-    plot(rowMeans(outputmeanPayoff), type = 'l', ylab = "Value", xlab = "Timestep", ylim=c(0,harvestMax), lwd = 3, main = "Mean Payoff")
+    plot(rowMeans(outputmeanPayoff), type = 'l', ylab = "Harvest Payoff", xlab = "Timestep", ylim=c(0,harvestMax), lwd = 3, main = "Mean Payoff")
     
     for (j in 1:Runs) {  
       
[…]

### 2020-11-24 MultipleReplicationsNoPlots.R

@@ -249,20 +249,31 @@ FillData%<>%mutate(PercProtected=round(PercProtected,digits=1))%>%mutate(Resourc
 
 
 for (i in 1:nrow(FillData)){
-  abmnp(Runs=3,Individuals = 100,TotalCarryingCapacity=10000,
-      StartPercCarryingCapacity = 0.65,
+  abmnp(Runs=5,Individuals = 100,TotalCarryingCapacity=10000,
+      StartPercCarryingCapacity = 1.00,
       PercProtected = FillData[i,]$PercProtected, 
       ProbOfMobility = FillData[i,]$ResourceMobility,CoopPercStart=1.0, 
       LearningStrategy = "Success Bias",
       TimeSteps=30,
       ResourceRegenerationPerTimeStep=1.3,
       harvestMax=25)
-  FillData[i,3]<-apply(FullOutput$meanPayoff[30,1:3],1,mean)  
-  
+  FillData[i,3]<-apply(FullOutput$meanPayoff[30,1:5],1,mean)  
+  print(paste0(i/nrow(FillData)*100,"%"))
 }
 
 ggplot(data=FillData, mapping=aes(x=PercProtected, y=ResourceMobility, fill= meanPayoff)) + 
-  geom_tile()+scale_fill_viridis()
+  geom_tile()+scale_fill_viridis()+theme_minimal()+ylab("Resource Mobility")+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        plot.title = element_text( size=18, color="black",face="bold"),
+        axis.title.x = element_text( size=18),
+        axis.title.y = element_text( size=18),
+        axis.text=(element_text(color="black", size=14)),
+        legend.title = element_text(colour="black", size=18),
+        legend.text = element_text( size = 14))+
+  scale_x_continuous(labels = scales::percent, expand = c(0, 0))+
+  scale_y_continuous( labels = scales::percent,expand = c(0, 0))+
+  labs(fill = "Average Harvest\nPayoff")+
+  xlab("Percent Protected")
 
 
 
@@ -273,15 +284,16 @@ FillData2%<>%mutate(PercProtected=round(PercProtected,digits=1))%>%mutate(Resour
 
 
 for (i in 1:nrow(FillData2)){
-  abmnp(Runs=3,Individuals = 100,TotalCarryingCapacity=10000,
-        StartPercCarryingCapacity = 0.65,
[…]

### 2020-11-24 ProtectedAreasABM.R

@@ -216,7 +216,7 @@ abm<-function(#Specified parameters
   
   p2<-ggplot(data=output,aes(x=timeStep))+
     geom_line(aes(y=meanPayoff),size=3,color="#993404")+
-    theme_classic()+ylab("Resource Units")+
+    theme_classic()+ylab("Resource Units")+ylim(0,harvestMax+1)+
     ggtitle("Mean Individual Payoff")
   
   p3<-ggplot(data=output,aes(x=timeStep))+

### 2020-11-24 app.R

@@ -106,7 +106,7 @@ server <- function(input, output) {
                            PayoffProtectedLastTime = rep(NA,Individuals),
                            PayoffWorkingLastTime= rep(NA,Individuals))   #dataframe to be filled with initial payoffs
         
-        ProtectPerDefect<-as.integer(StartResourceProtected/DefNumStart) #protected area resources per each individual harvesting there
+        ProtectPerDefect<-ifelse(DefNumStart ==0,as.integer(StartResourceProtected),as.integer(StartResourceProtected/DefNumStart)) #protected area resources per each individual harvesting there
         WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
         
         for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
@@ -205,6 +205,7 @@ server <- function(input, output) {
             
             
             
+            
             PercTimeWorking<- 1-PercTimeProtected   #time foraging in the working landscape
             
             
@@ -215,12 +216,16 @@ server <- function(input, output) {
             
             
             #New resource pools to pull from
-            #make sure they dont regenerate past their carrying capacity
+            #Resources over the CC move with probability 'ProbOfMobility'
             
             NewProtectedResourcesTotal<-round(output$NumResourcesProtect[t-1] * ResourceRegenerationPerTimeStep,digits=0)
-            NewProtectedResourcesTotal<-ifelse(NewProtectedResourcesTotal<=  TotalCCResourceProtected,NewProtectedResourcesTotal, TotalCCResourceProtected)
+            ProtectedResourcesOverCC<-round(NewProtectedResourcesTotal-TotalCCResourceProtected,digits=0)
+            ProtectedResourcesOverCC<- ifelse(ProtectedResourcesOverCC<=  0,0, ProtectedResourcesOverCC)
+            
             NewWorkingResourcesTotal<-round(output$NumResourcesWorking[t-1] * ResourceRegenerationPerTimeStep,digits=0)
-            NewWorkingResourcesTotal<-ifelse(NewWorkingResourcesTotal<=TotalCCResourceWorking,NewWorkingResourcesTotal,TotalCCResourceWorking)
+            WorkingResourcesOverCC<-round(NewWorkingResourcesTotal-TotalCCResourceWorking,digits=0)
[…]

### 2020-12-10 DissertationProposal/DissertationProposal.Rmd
Test edits to intro and moon map

@@
 # Introduction
-## Phenomenon
 Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources.
-Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources [@network2019global].
-Today, on average, we degenerate these resources at a rate 56% faster than regeneration [@network2020global].
+Until approximately 50 years ago, the rate of overall ecological regeneration exceeded the rate at which we consume and degrade our environment [@network2019global].
+Today, on average, we degenerate our renewable resources at a rate 56% faster than regeneration [@network2020global].
 This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining [@beltratti_sustainable_1993].
@@
 
-The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity [@haines2010links; @redford1995human].
-As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target [@parrish_are_2003].
-Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) [@iucn2020protectedplanet].
-These area-based conservation initiatives are from hereon referred to as protected areas.
+Slowing the rate of loss and reduction of earth’s natural areas is decisively, the first step to stemming the global decline in natural resources and ecosystem function [@unep2018ngs; @haines2010links].
+Achieving this will require both the adoption of new, efficient technologies and the preservation of currently intact areas [@secretariat2009review].
+Toward this aim, each year we design and distribute technologies that use fewer or alternative resources such as improved efficiency cookstoves and solar panels [@ryan2012spreading].
+In conjunction, we section off an ever growing proportion of earth’s marine and terrestrial areas to protect key resources (hotspots), the flow of resources across landscapes (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) [@iucn2020protectedplanet].
 
@@
 
-Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local communities [@adams_biodiversity_2004; @leverington_global_2010].
-While high-profile protected area successes are salient in the conservation narrative, protected areas also commonly fail to reach their intended biodiversity targets.
-The ubiquity of these failures is somewhat speculative as unsuccessful protected areas go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004].
-Regardless of the high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].
+Both of these efforts benefit from a great deal of top-down support.
+Governments demarcate parks and wilderness areas, Western nonprofit organizations drop thousands of cookstoves across Equatorial Africa, and the global carbon market funds much of these efforts in the name of climate change mitigation [@redmond2015global; @wara2007global].
+While they may be well supported in the onset, many conservation initiatives are intended only to seed a specific conservation paradigm or technology, relying on subsequent widespread community adoption for long-term benefits [@mills2019conservation; @sunderlin2012redd].
+From hereon, we refer to geographic areas managed for conservation by communities themselves as *community conservation areas*, regardless of whether the *origin* of an area is community-based or top-down.
 
-Community norms around protected area adherence are characterized as individually costly, yet beneficial at the group level.
-Substantial variation exists in the degree to which communities adhere to protected area regulations; some communities develop strong and lasting conservation norms, while in others these norms degrade over time or may fail to emerge at all [@rogers_diffusion_2010; @simon_win-win_2012].
-Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017].
-This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present [@ostrom_governing_1990; @richerson2001evolution].
-What is less understood however, is the emergence and initial adoption of conservation behaviors.
+Community conservation areas can be a powerful tool for sustaining ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local communities [@adams_biodiversity_2004; @leverington_global_2010].
+The success of these areas is largely contingent on local buy-in and adherence to area regulations — conservation doesn’t work if no one does it [@catalano_learning_2019].
+Substantial variation exists in the degree to which communities adhere to conservation area regulations; some communities develop strong and lasting conservation norms, while in others these norms degrade over time or may fail to emerge at all [@simon_win-win_2012].
+The degree to which conservation areas fail to catch on in communities is likely much greater than we recognize; unsuccessful conservation projects go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004].
 
-Research on the diffusion of conservation behaviors suffers from a lack of generalizability.
-Case studies tend to be highly context dependent and offer little mechanistic insight into the emergence and spread of these behaviors.
-Outside of conservation science, the field of evolutionary anthropology has devoted considerable effort to understanding how and why individuals adopt costly cooperative behaviors.
-Of particular relevance to understanding group adherence to protected area regulations, the theory of cultural group selection offers a mechanistic explanation for the emergence and spread of individually costly group norms.
+Similar to community-based conservation areas, new efficient technologies have the potential to mitigate natural resource degradation while benefiting human populations, but rely greatly on community buy-in.
+Adoption of these technologies may diffuse and become ubiquitous within a population [@bollinger2012peer].
+This is, however, not always the case.
+The reasons for if and when an individual adopts a new technology, especially in the contexts of conservation, are often unclear [@smith2014early].
+Even in cases where a technology does spread within a population, the specific use behavior and use outcome may become so greatly modified along the transmission chain that the population level result is unrecognizable from the original intent [@lewis_transmission_2012; @mcguigan_over-imitation_2011].
+Theoretical insights and controlled empirical research give suggestions for how intended technological adoption outcomes may be more likely to be retained during transmission from individual to individual or group to group.
+So far, these insights are yet to be validated empirically in a conservation context where technology adoption has real impacts for individuals, communities, and the environment, and those impacts may affect future adoptions [@king2007use; @muthukrishna_when_2016].
 
-## Primer on cultural group selection
-The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations [@boyd_transmission_2010].
-Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not.
-As an individual’s behavior is flexible, individuals may migrate from less to more successful groups or directly copy the behaviors they observe in visibly successful cooperative groups [@henrich_cultural_2004].
-This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection.
-In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall [@mcelreath_mathematical_2008; @okasha_multilevel_2004; @price1972extension].
+The efficacy of both community conservation areas and improved technologies depends on widespread behavioral adoption at both the individual and community level.
+Cultural group selection theory poses mechanisms for the spread and persistence of behaviors within and between groups [@boyd_transmission_2010].
+Under this theory, groups develop characteristics or institutions such as a resource management agreement or community conservation norms and those characteristics can spread between groups through one of three mechanisms.
+Groups may outcompete or kill other groups with different institutions, individuals in one group may emigrate to another group, or individuals in less successful groups may copy behaviors they observe in more successful groups [@henrich_cultural_2004; @richerson_cultural_2016].
+This change in the frequency of behavioral traits in groups is analogous to the spread and perpetuation of beneficial alleles between mixing groups in genetic natural selection.
+In these contexts, selective pressures still act on individuals and push them to engage in non-cooperative behaviors like stealing resources from a community conservation area, but they also experience selective pressure at the group level which may lead to cooperative behavior even at some individual cost [@mcelreath_mathematical_2008; @okasha_multilevel_2004; @price1972extension].
 
-Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors [@boyd_rapid_2011; @lehmann_cultural_2008].
-Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically.
-Conservation behaviors are fundamentally different from generic cooperative behaviors as they operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates.
-As such, cultural group selection theory must be extended to include these dynamics in order to be effectively applied to conservation.
-
-## Applying cultural group selection to the adoption of conservation behaviors
-Protected areas are an ideal microcosm for studying and applying cultural group selection.
-Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals [@adams_biodiversity_2004].
-While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice [@naughton-treves_role_2005].
-Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful.
-Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection.
-Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.
-
-## Fidelity of behavioral transmission
-The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission.
-In situations where behaviors do emerge and spread within populations, behaviors are transmitted along a chain of individuals, where they may be modified considerably to the point where the behavioral outcome is unrecognizable [@lewis_transmission_2012; @mcguigan_over-imitation_2011].
-This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves.
-Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator [@king2007use; @muthukrishna_when_2016].
-These insights are crucial for practitioners aiming to implement long-lasting, effective conservation projects, but must first be tested empirically in a conservation context where an individual’s access to a behavioral demonstration is realistic and outcomes of adopted behaviors have real impacts for individuals, communities, and the environment.
+Researchers have posited that cultural group selection theory should be applicable to understanding the spread and persistence of conservation behaviors such as adherence to community conservation areas or adoption of new efficient technologies [@kline2018designing].
+Cultural group selection theory offers mechanistic insights for the emergence and spread of behaviors within and between groups broadly, but its applicability to conservation specifically is unknown.
+Conservation has many aspects which make it unique compared to behavioral adoption in the abstract.
+In conservation, an individual’s behavior directly affects their environment, which may in turn become altered in a way which shapes future individual decision making.
+Feedbacks between individual behavior and resource availability are further complicated by resource characteristics such as carrying capacity, mobility, and regeneration rate.
+These specifics make conservation a fundamentally different system from general behavioral adoption and cooperation, thus we must extend cultural group selection theory considering conservation explicitly.
+The aim of this research is to expand the theories of behavioral adoption via cultural group selection to explicitly consider factors unique to conservation, and to empirically examine the mechanisms of conservation behavioral adoption within and between groups.
+This research will offer insights for strategically seeding both community conservation areas and new technologies so that they may offer the greatest benefit to communities and the preservation of renewable resources.
 
 ## Dissertation objectives
-The overarching question this research seeks to answer is: What are the underlying social mechanisms which cause conservation behaviors to emerge and spread efficiently within populations?
+The overarching question this research seeks to answer is: What are the underlying social mechanisms which cause conservation behaviors to emerge, spread, and persist within populations?
 
-* Objective 1: Create a dynamic, agent-based model which extends theories of cultural group selection into the contexts of conservation.
-Formally test the intuition that between group imitation can lead to the spread of community wide protected area adoption whereas solitary groups may fail to cooperate under identical environmental constraints such as resource degradation, regeneration, and mobility.
+* Objective 1: Create an agent based model to theoretically test whether cultural group selection via success copying or direct competition results in greater conservation adoption and better environmental outcomes, particularly in environmental contexts where solitary groups fail to cooperate.
 
-* Objective 2: Calculate theoretical adoption probabilities of protected areas known as Community Forest Management Agreements (CoFMAs) for each ward on Pemba Island, Tanzania.
-Probabilities are based on the salience of group cultural identities and the costs and benefits of CoFMAs to individuals and groups.
-Compare theoretical predictions to the presence or absence of CoFMAs in reality.
+* Objective 2: Empirically test whether imitating the successes of adjacent groups or competing against them for resources is driving the spread of community conservation areas on Pemba Island, Tanzania.
 
-* Objective 3: Quantitatively examine how social and physical access to a behavioral demonstration during behavioral transmission affects both the copying fidelity as well as the performance of copied behavior.
-This project will make use of the ongoing spread of fuel efficient cookstove technology on Pemba Island, Tanzania.
+* Objective 3: Examine if the efficiency of an improved cookstove design has been modified as it has spread through three communities on Pemba Island, Tanzania.
+Quantitatively assess the relationship between transmission fidelity and the social dynamics between demonstrating/copying individuals.
 
 ## Study System
-All empirical research for this dissertation will take place on Pemba Island, Tanzania.
+All empirical research for this dissertation will take place on Pemba Island, Tanzania; Pemba is one of two main islands in the Zanzibar archipelago (Figure \@ref(fig:moon)).
 In 2014, Pemba Island had a reported population of 406,808 [@RGZ2014Zanzibar].
-This population is divided into wards of 2,000 to 6,000 individuals.
-The majority of Pemban wards are described as forest-dependent.
-The Department of Forests and Non-Renewable Natural Resources allows wards to enter into Community Forest Management Agreements (CoFMAs), which among other things, establish forest protected areas and exclude outside individuals [@benjaminsen2014between].
-As of 2018, 18 Pemban wards had adopted CoFMAs [@andrews_cultural_2018].
+This population is divided into 121 wards of 2,000 to 6,000 individuals.
+The majority of Pemban wards are described as forest-dependent, with 90% of households relying exclusively on forest products (fuelwood and charcoal) to meet their daily cooking needs.
+The yearly deforestation rate across all of Zanzibar was 1.1% in 2014 and is worsening as the population increases by 3.2% each year [@RGZ2014Zanzibar].
+On Pemba, the fuelwood offtake around the cities is approximately ten times that of sustainable harvest [@ely2000participatory].
+Unlike many areas with high deforestation, on Pemba, 86% of all forest loss is attributed to the local communities directly, rather than outside entities (e.g.
+foregin corporations) [@terracarbonreport].
+
+All of the above factors contributed to Zanzibar being selected as one of eight pilot sites in Tanzania for the REDD+ program in 2009 [@burgess2010getting].
+As part of this program on Pemba, Zanzibar’s Department of Forests and Non-Renewable Natural Resources worked with 18 wards to establish Community Forest Management Agreements (CoFMAs) [@andrews_cultural_2018].
+Crucially, these agreements establish both a *use area* and a *conservation area* inside the forest in that ward and are meant to exclude outside individuals from harvesting inside the forest as well (Figure \@ref(fig:moon)) [@benjaminsen2014between].
+
+```{r moon, echo=FALSE, fig.cap="Pemba Island, Tanzania.
+Showing wards with and without CoFMAs.
+For each CoFMA, the size (ha) is proportional to the size of the circle and the percent area designated as Use and Conservation is indicated by the blue and green colors respectively.", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/MoonMapEditedNoLabs.png")
+```
+
+A secondary component of the REDD+ program on Pemba was the distribution of fuel efficient cookstoves in three wards [@terracarbonreport].
+These cookstoves have been shown to reduce fuelwood use by up to 50% [@lung_large-scale_2019].
+The specific type of stove that was distributed can be built with materials readily available on Pemba.
+The intent of this aspect of the REDD+ project was to seed this design so that it may be widely copied throughout the communities and reduce deforestation beyond that of the original cookstoves.
+
+The final key component of the REDD+ project as it relates to this research was the promise of payments to communities for their conservation efforts.
+This aspect of the program never came to fruition due to issues with the carbon validation process [@andrewsdoes].
+Given that this aspect was unrealized, the CoFMAs have no outside budget and are now entirely community run, although still designated by the Department of Forests and Non-Renewable Natural Resources.
+Since the establishment of the original 18 CoFMAs, ten more wards have elected to enter the process of establishing CoFMAs, even without the promise of community payments from REDD+ [@andrewsdoes].
+Furthermore, the seeded cookstove design has continued to spread across communities.
+While continuing to spread, both the CoFMAs and cookstoves have had variable success in reducing deforestation and fuelwood use respectively.
+This variation in both spread and efficacy of both of these initiatives within and between communities serves as the basis for the empirical research in this dissertation.
 
@@
 At every time step agents will attempt to harvest resources from the working landscape up until the specified maximum harvest amount.
-The outcome of this harvest attempt for each agent is the result of one random draw from a Binomial distribution where the number of trials is equal to the number of resources available on the working landscape divided by the number of individuals (i.e.
+The outcome of this harvest attempt for each agent is the result of one random draw from a binomial distribution where the number of trials is equal to the number of resources available on the working landscape divided by the number of individuals (i.e.
 that individual’s share of the available resources) and the probability of success is equal to the percent of an agent’s harvest time spent in the working landscape; in this scenario we are forcing obligatory cooperation so that percentage is always 100%.
@@
 If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e.
-not be available for agents in the following time step) or move to the other area (protected or working) with some specified probability also applied to a Binomial distribution.
+not be available for agents in the following time step) or move to the other area (protected or working) with some specified probability also applied to a binomial distribution.
 Resources which move to the adjacent area but exceed the carrying capacity will also die.
@@
 Note that the sustainable harvest value is now dependent on the percent of the resources in the working landscape vs the protected area.
-We can run the model for every combination of protected area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (fig \@ref(fig:abm2)).
+We can run the model for every combination of protected area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (figure \@ref(fig:abm2)).
 In the figure below, we can see that up until the maximum sustainable harvest value for this scenario (16), agents will always reach the set maximum harvest value until the resources are no longer made available to them (i.e.
@@
 
-The probability of resource mobility is operationalized in the model by using random draws from a Binomial distribution.
+The probability of resource mobility is operationalized in the model by using random draws from a binomial distribution.
 For every resource unit regenerated over the carrying capacity of the protected area, we take one random binomial draw with the probability of success equal to the probability of resource mobility.
@@
 During model initialization, we can set the percentage of agents who we will allow to spend a proportion of their harvest effort in the protected area (defecting).
-The proportion of effort an individual agent spends in the protected area is drawn from the beta distribution below (fig \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.
+The proportion of effort an individual agent spends in the protected area is drawn from the beta distribution below (figure \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.
 
@@
 The learning strategy of the overall population is currently set to either success-biased or conformist-biased learning for all agents, but in future versions of the model, it will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents.
-The entire model process is visualized conceptually in the figure below and a table of all the model starting conditions, key variables, and outcomes can be found in Table \@ref(tab:tabAbmVars)).
+The entire model process is visualized conceptually in the figure below and a table of all the model starting conditions, key variables, and outcomes can be found in Table \@ref(tab:tabAbmVars).
 
@@
 
-With all model parameters kept constant as in the output shown in (fig \@ref(fig:abm3)), we can run the model with just 5% of the agents assigned as having the possibility to defect at initialization and all agents modifying harvest strategies via success-biased learning.
-The average final time step outcomes from five runs of this model can be seen in (fig \@ref(fig:abm4)).
+With all model parameters kept constant as in the output shown in (figure \@ref(fig:abm3)), we can run the model with just 5% of the agents assigned as having the possibility to defect at initialization and all agents modifying harvest strategies via success-biased learning.
+The average final time step outcomes from five runs of this model can be seen in (figure \@ref(fig:abm4)).
 
@@
 Allowing agents to modify harvest strategies and harvest from the protected area greatly reduces the overall harvest outcomes.
-To better understand the underlying dynamics, we set the starting model parameters to the ideal conditions for maximum payoff as identified in (fig \@ref(fig:abm3)) and observe the model behavior at each time step.
+To better understand the underlying dynamics, we set the starting model parameters to the ideal conditions for maximum payoff as identified in (figure \@ref(fig:abm3)) and observe the model behavior at each time step.
 
@@
                                       "Protected area percent of total carrying capacity")))
-kbl(abmVars,format="latex", caption = "Description of all model variables and outcomes",escape=F) %>% 
+kbl(abmVars,format="latex", caption = "Description of all model variables and outcomes",escape=F,booktabs = T) %>% 
   kable_classic_2(full_width=F)%>% row_spec(0,bold=TRUE)%>%
@@
     column_spec(2, width = "8em")%>%
-  column_spec(3, width = "20em")
+  column_spec(3, width = "20em")%>%kableExtra::kable_styling(latex_options = "striped")
 
@@
 ### Methodological Specifics
-I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)).
+I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (figure \@ref(fig:mapp)).
 With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.

### 2020-12-11 DissertationProposal/DissertationProposal.Rmd
abm changes to text and fig added

@@
 
-# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Area-Based Conservation. 
+# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Community Conservation Areas. 
 ## Background
-Cultural group selection theorizes that individuals will adopt cooperative behavior if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985].
-Under this framework, the presence of multiple competing groups drives community cooperative norms to emerge under a much broader set of conditions compared to solitary groups [@richerson_cultural_2016].
-Protected area conservation is a cooperative phenomenon where individuals sacrifice some degree of their own harvest potential in order to benefit the group as a whole.
-Therefore, insights from cultural group selection may help conservation scientists better understand and predict community adherence to protected area regulations [@brooks_applying_2018].
+For community conservation areas to tangibly reduce the loss of earth’s natural resources, they must be adopted at large spatial scales and remain intact in perpetuity [@turner2007global].
+Predicting if and how conservation actions will spread and evolve across a landscape has proven to be a major challenge for conservation scientists and practitioners [@child2010conceptual].
+Theory around the long-term adoption of conservation norms is sparse [@mahajan2020theory].
+Because of this gap, most researchers default to the diffusion of innovations literature to understand and draw conclusions around conservation adoption [@mascia2018conservation; @rogers_diffusion_2010].
+This body of literature has proven valuable for demonstrating *how* conservation paradigms spread, but is relatively mute for understanding the underlying social mechanisms of *why* conservation paradigmes spread and last [@pannell2006understanding; @mills2019conservation].
 
-While cultural group selection theory offers mechanistic insights for the emergence of cooperative norms in groups broadly, its applicability to conservation specifically is unknown.
-Conservation has many aspects which make it unique compared to cooperation in the abstract.
-In conservation, an individual’s behavior directly affects their environment, which may in turn become altered in a way which shapes future individual decision making.
-Feedbacks between individual behavior and resource availability are further complicated by resource characteristics such as mobility and regeneration rate.
-These specifics make conservation a fundamentally different system from general cooperation, thus we must extend cultural group selection theory considering conservation explicitly.
+Cultural group selection offers theories on the mechanisms behind the spread and perpetuation of cooperative behaviors such as adhering to community conservation area regulations.
+Existing cultural group selection theory however, is highly abstract, relying on simple, analytical models of nonspecific behaviors [@boyd_rapid_2011; @lehmann_cultural_2008].
+Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically [although see @waring_coevolution_2017; & -@waring2015multilevel].
+This is problematic as conservation behaviors have behavior-environment feedbacks which are fundamentally different from generic cooperative behaviors.
+They operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates.
+As such, cultural group selection theory must be extended to include these dynamics in order to be effectively applied to conservation.
 
 This project aims to logically test the applicability of cultural group selection to conservation.
-Specifically, this research will address how behavior-environment feedbacks and resource mobility and regeneration rates affect the emergence and spread of protected area adoption.
-In this chapter I will develop a dynamic, agent-based simulation of individuals harvesting resources from a landscape under a variety of environmental constraints.
-I will then use this simulation to test under what circumstances adherence to protected area regulations emerges as a group norm within solitary and multiple, competing groups.
-This model will generate qualitative predictions about system outcomes given a very simple set of assumptions.
+Specifically, this research will address how behavior-environment feedbacks and resource mobility and regeneration rates affect the emergence and spread of community conservation areas under three different social contexts.
+Cultural group selection theorizes that individuals will adopt  cooperative behaviors that they observe in individuals in outside, successful groups.
+Under this framework, the presence of multiple groups drives cooperative community norms to emerge under a much broader set of conditions compared to solitary groups [@richerson_cultural_2016].
+Building on this, Andrews and Borgerhoff Mulder [-@andrews_cultural_2018] argue that multiple groups should drive the adoption of cooperative norms to an even greater degree when those groups are competing against one another for common resources and the benefits of cooperation are observable.
+The three social contexts this project will test are therefore: solitary groups, multiple separate groups, and multiple competing groups.
+Insights from this theoretical research will give qualitative predictions of how social and environmental factors innate to a landscape may facilitate or hinder the emergence, spread, and persistence of community conservation areas [@brooks_applying_2018].
 
 ### Question:
-Under what environmental contexts might the mechanisms of cultural group selection influence individual decision making to generate lasting group norms around the adherence to protected area regulations?
+Can the presence of multiple, observable groups drive widespread adherence to community conservation areas in environmental contexts where such norms do not persist for solitary groups?
+Secondly, how do separate vs shared group resource pools affect the adoption outcome?
 
 ### Predictions:
-* P1: Groups will adopt protected area regulations in both the solitary and multiple competing groups scenarios under specific environmental contexts.
+* P1: In the presence of multiple observable groups, the environmental contexts in which groups adhere to community conservation areas will be much more broad compared to solitary groups.
 
-* P2: In the presence of multiple competing groups, the environmental contexts which generate group adoption of protected area regulations will be much more broad compared to solitary groups.
+* P2: When multiple groups share a common resource pool, group adherence to community conservation areas will emerge and persist under a broader set of environmental contexts compared to when groups may only observe one another.
 
@@
 
-An agent-based modeling framework is ideal for demonstrating the theory of cultural group selection regarding protected area norms, because it allows such group level properties to emerge out of simple decision rules for each individual [@railsback_agent-based_2012; @wilensky_introduction_2015].
-This design mirrors our real-world understanding of the important system components.
+Agent-based models allow for system level properties to emerge from very simple constraints and rule sets for individual actors [@railsback_agent-based_2012].
+In this project we are interested in understanding how group cooperative conservation norms can emerge and persist from individual community members responding to social and environmental information.
+This modeling design is ideal for testing these theories as it will formalize the logic and intuition under which we can and can not expect cultural group selection to be applicable to conservation adoption.
+Given the emergent properties of agent-based modeling, this research may also elucidate unseen errors in our logic or may lead to the development of future research questions around the adoption of community conservation areas [@wilensky_introduction_2015].
+The specific research question posed above will build on an existing agent-based model described and validated below.
 
@@
 We can write a simple analytical model to better denote these dynamics and determine interesting and appropriate starting conditions for more complex permutations of the agent-based model, as well as validate our intuitions as to how the model should behave.
+
 Given that the resources regenerate multiplicatively, any starting conditions which cause the resource pool to diminish after the first time step will further deplete the ever-shrinking resource pool with each additional time step.
@@
 
-We can extend this model to explore how the sink/source dynamics between protected areas and adjacent working landscapes may provision more renewable resources than working landscapes alone.
+We can extend this model to explore how the sink/source dynamics between conservation areas and adjacent working landscapes may provision more renewable resources than working landscapes alone.
 In this formation, we still have one homogenous group of agents who harvest regenerating resources from a landscape.
-This landscape now has some specified percentage sectioned off and inaccessible to the agents — the protected area.
+This landscape now has some specified percentage sectioned off and inaccessible to the agents — the conservation area.
 At every time step agents will attempt to harvest resources from the working landscape up until the specified maximum harvest amount.
@@
 
-After the agents harvest resources from the landscape, the resources in the working and protected areas will reproduce by the specified regeneration factor.
+After the agents harvest resources from the landscape, the resources in the working and conservation areas will reproduce by the specified regeneration factor.
 If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e.
-not be available for agents in the following time step) or move to the other area (protected or working) with some specified probability also applied to a binomial distribution.
+not be available for agents in the following time step) or move to the other area (conservation or working) with some specified probability also applied to a binomial distribution.
 Resources which move to the adjacent area but exceed the carrying capacity will also die.
-If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the protected area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).
+If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the conservation area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).
 
@@
 
-Note that the sustainable harvest value is now dependent on the percent of the resources in the working landscape vs the protected area.
-We can run the model for every combination of protected area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (figure \@ref(fig:abm2)).
+Note that the sustainable harvest value is now dependent on the percent of the resources in the working landscape vs the conservation area.
+We can run the model for every combination of conservation area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (figure \@ref(fig:abm2)).
 In the figure below, we can see that up until the maximum sustainable harvest value for this scenario (16), agents will always reach the set maximum harvest value until the resources are no longer made available to them (i.e.
-all protected).
-Above the maximum sustainable harvest value, the size of the protected area becomes greatly important for determining average harvest payoff to agents.
-In this scenario where the resources regenerate at a relatively high rate (20%) and the probability of mobility is 100%, the sink/source dynamics allow agents to reach the greatest average harvest value when the protected area is large and therefore provisioning many resources to the ever depleting working landscape at every time step.
+all conservation).
+Above the maximum sustainable harvest value, the size of the conservation area becomes greatly important for determining average harvest payoff to agents.
+In this scenario where the resources regenerate at a relatively high rate (20%) and the probability of mobility is 100%, the sink/source dynamics allow agents to reach the greatest average harvest value when the conservation area is large and therefore provisioning many resources to the ever depleting working landscape at every time step.
 
-```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) for many combinationa of maximum harvest values and protected area sizes.
+```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) for many combinationa of maximum harvest values and conservation area sizes.
 Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
@@
 
-Up to this point, we have assumed that excess resources produced in the protected area will move to the working landscape, up until a specified carrying capacity, with 100% probability.
+Up to this point, we have assumed that excess resources produced in the conservation area will move to the working landscape, up until a specified carrying capacity, with 100% probability.
 Key insights from this model are gained from experimenting with different probabilities of resource mobility.
-To frame this in the context of real world protected areas, we can assume that insights from model runs with high probabilities of resource mobility may be applied to wide ranging organisms such as large aquatic animals, while model runs with low probabilities of resource mobility may be applied to relatively low vagility organisms such short or large-seeded trees [@thomson2011seed].
+To frame this in the context of real world conservation areas, we can assume that insights from model runs with high probabilities of resource mobility may be applied to wide ranging organisms such as large aquatic animals, while model runs with low probabilities of resource mobility may be applied to relatively low vagility organisms such short or large-seeded trees [@thomson2011seed].
 
 The probability of resource mobility is operationalized in the model by using random draws from a binomial distribution.
-For every resource unit regenerated over the carrying capacity of the protected area, we take one random binomial draw with the probability of success equal to the probability of resource mobility.
-With all other model specifications kept equal and a maximum harvest value set slightly above the maximum sustainable harvest value for a working landscape alone, we can explore the interaction between protected area size and resource mobility given obligatory cooperation.
+For every resource unit regenerated over the carrying capacity of the conservation area, we take one random binomial draw with the probability of success equal to the probability of resource mobility.
+With all other model specifications kept equal and a maximum harvest value set slightly above the maximum sustainable harvest value for a working landscape alone, we can explore the interaction between conservation area size and resource mobility given obligatory cooperation.
 
 ```{r abm3, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model.
-Outcomes are the result of many combinations of resource mobilities and protected area sizes.
+Outcomes are the result of many combinations of resource mobilities and conservation area sizes.
 Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
@@
 
-To apply this model to conservation adoption and different mechanisms of social learning, we can allow agents to alter their harvest strategies beyond obligatory cooperation around protected area regulations.
-During model initialization, we can set the percentage of agents who we will allow to spend a proportion of their harvest effort in the protected area (defecting).
-The proportion of effort an individual agent spends in the protected area is drawn from the beta distribution below (figure \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.
+To apply this model to conservation adoption and different mechanisms of social learning, we can allow agents to alter their harvest strategies beyond obligatory cooperation around conservation area regulations.
+During model initialization, we can set the percentage of agents who we will allow to spend a proportion of their harvest effort in the conservation area (defecting).
+The proportion of effort an individual agent spends in the conservation area is drawn from the beta distribution below (figure \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.
 
 ```{r betastart, out.width = "75%",echo=FALSE, fig.align='center', fig.pos= "H", fig.cap="One thousand random draws from a beta distribution with the shape 1 and 2 parameters set at 1 and 2 respectively.
-This distribution is used to assign the harvest effort spent in the protected area for defecting agents during model initialization."}
+This distribution is used to assign the harvest effort spent in the conservation area for defecting agents during model initialization."}
 hist(rbeta(1000,1,2),main=NULL)
@@
 
-After harvesting from the working landscape, agents who defect to some degree and did not already reach the maximum harvest value, spend the remainder of their harvest effort in the protected area.
-Following the same logic as harvest from the working landscape, the number of resources agents receive from the protected area is determined by one draw from the binomial distribution where the number of trials is equal to the number of resources in the protected area divided by the number of agents who defect in that time step and the probability of success is equal to the percentage of effort that agent spends in the protected area.
+After harvesting from the working landscape, agents who defect to some degree and did not already reach the maximum harvest value, spend the remainder of their harvest effort in the conservation area.
+Following the same logic as harvest from the working landscape, the number of resources agents receive from the conservation area is determined by one draw from the binomial distribution where the number of trials is equal to the number of resources in the conservation area divided by the number of agents who defect in that time step and the probability of success is equal to the percentage of effort that agent spends in the conservation area.
 
@@
 ```{r abm4, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model.
-Outcomes are the result of many combinations of resource mobilities and protected area sizes.
+Outcomes are the result of many combinations of resource mobilities and conservation area sizes.
 Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
@@
 ```
-Allowing agents to modify harvest strategies and harvest from the protected area greatly reduces the overall harvest outcomes.
+Allowing agents to modify harvest strategies and harvest from the conservation area greatly reduces the overall harvest outcomes.
 To better understand the underlying dynamics, we set the starting model parameters to the ideal conditions for maximum payoff as identified in (figure \@ref(fig:abm3)) and observe the model behavior at each time step.
@@
 Starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
-Eighty percent of resources are inside the protected area and the probability of resource mobility is 100 percent.
+Eighty percent of resources are inside the conservation area and the probability of resource mobility is 100 percent.
 The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25.
@@
 
-Model dynamics show that over time, agents begin to harvest from the protected area as they copy more successful individuals.
-This behavior depletes the resource pool in the protected area and lowers the number of resources which regenerate and agents can harvest in the subsequent time step.
-Note that full defection is defined as agents spending an equal proportion of their effort in the protected area as the proportion protected (i.e.
+Model dynamics show that over time, agents begin to harvest from the conservation area as they copy more successful individuals.
+This behavior depletes the resource pool in the conservation area and lowers the number of resources which regenerate and agents can harvest in the subsequent time step.
+Note that full defection is defined as agents spending an equal proportion of their effort in the conservation area as the proportion conservation (i.e.
 they treat the entire landscape equally).
@@
 The results displayed above are for one homogenous group of agents harvesting resources from a common pool of resources.
-This model is useful for exploring the sink/source utility of protected areas to ensure long-term resource provisioning, especially in the contexts of social learning.
+This model is useful for exploring the sink/source utility of conservation areas to ensure long-term resource provisioning, especially in the contexts of social learning.
 To further explore this model, I have created an interactive application which performs one run of the model and can be found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).
@@
 abmVars <- data.frame(Parameters = c("Individuals","Total Carrying Capacity", "Percent Carrying Capacity", 
-                                    "Percent Protected", "Starting Cooperators", "Time Steps", "Max Harvest",
-                                     "Resource Regeneration",  "Resource Mobility","Groups","Learning Strategy", 
-                                     "Percent Cooperation", "Average Payoff", "Protected Area Degradation",
+                                    "Percent conservation", "Starting Cooperators", "Time Steps", "Max Harvest",
+                                     "Resource Regeneration",  "Resource Mobility","Group Scenario","Percent Group Copiers","Learning Strategy", 
+                                     "Percent Cooperation", "Average Payoff", "conservation Area Degradation",
                                      "Working Landscape Degradation"), 
@@
                                "Starting Condition","Starting Condition","Starting Condition",
-                               "Starting Condition","Key Independent Variable","Key Independent Variable",
+                               "Starting Condition","Key Independent Variable","Key Independent Variable","Key Independent Variable",
                                "Key Independent Variable","Key Independent Variable",
@@
                                       "Percent of carrying capacity at initialization",
-                                      "Percent of landscape in protected area",
+                                      "Percent of landscape in conservation area",
                                       "Starting percent of agents fully cooperating",
@@
                                       "Maximum number of resources agents will harvest during one time step",
-                                      "The percent of available resources which will regenerate at each time step in both protected and working landscapes",
-                                      "The probability that any available resource will move from the  protected to working landscape or vice versa",
-                                      "Solitary or multiple competing groups",
+                                      "The percent of available resources which will regenerate at each time step in both conservation and working landscapes",
+                                      "The probability that any available resource will move from the  conservation to working landscape or vice versa",
+                                      "Solitary, separate, or competing groups","Percent of agents copying the successful group strategy",
                                       "Assigns a percent of the population to success-biased imitation and the remaining population to conformist-biased imitation",
@@
                                       "Working landscape percent of total carrying capacity",
-                                      "Protected area percent of total carrying capacity")))
+                                      "conservation area percent of total carrying capacity")))
 kbl(abmVars,format="latex", caption = "Description of all model variables and outcomes",escape=F,booktabs = T) %>% 
@@
 
-In the multiple interacting groups scenario, two groups with separate resource pools will each follow the processes described above.
+Incorporating multiple, observable groups into the above model will allow us to test the theories proposed by cultural group selection in a conservation context.
+The agent-based model described above shows that community conservation areas have the potential to provision a greater number of harvestable resources than working landscapes alone.
+Further, the success of those areas depends on the degree to which individuals adhere to the community conservation area regulations.
+Cultural group selection suggests that given this scenario, the presence of multiple groups may cause individuals to adhere to community conservation areas under a broader set of environmental conditions than for solitary groups.
+We propose two different versions of this multiple group scenario in which we will test this prediction.
+
+In the seperate groups scenario, two groups with separate resource pools will each follow the processes described above.
 However, at the end of each time step the groups will compare the average harvest rates.
@@
 
-Overall, the goal of this model is to test how environmental conditions affect cultural group selection mechanisms.
+In the competing groups scenario, the model will follow the exact same scheduling process as in the separate groups scenario, but a specified portion of the working landscape(s) will be shared among the groups.
+Mobile resources leaving the conservation areas will emigrate to the working landscape as in the single group model.
+If excess resources exist in the working landscape, they will move to whichever conservation area is below its carrying capacity.
+If both conservation areas are below carrying capacity, resources will have an equal chance of moving to either.
+Lastly, the carrying capacity of the working landscape associated with each group will include the shared area, but not the other group’s working landscape.
+This will ensure that resources do not always reach an equilibrium between the groups and the groups may have variable success.
+See figure \@ref(fig:conceptDiagram2) for a conceptual description of the two multi-group scenarios.
+
+```{r conceptDiagram2, echo=FALSE, fig.cap="Concept diagram of the two different multi-group dynamics that will be tested in the proposed agent-based model.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch1/ABMConceptDiagram2.png")
+```
+
+Overall, the goal of this model is to test how the mechanisms of cultural group selection operate under different social and environmental conditions.
 Cultural group selection may lead to community adoption of sustainable norms, where solitary groups fail to adopt cooperative behaviors.
@@
 The insights gained from this model will be published in a peer reviewed, academic journal such as Environmental Modeling and Software.
-This paper will show logically how we expect that protected areas function in the context of community level cooperative conservation.
-I also expect that this research will provide actionable insights for how designing protected areas while considering inter-group dynamics may allow for optimal resource provisioning under a much more broad set of circumstances than considering solitary groups alone.
+This paper will show logically how we expect that community conservation areas function in the context of community level cooperative conservation.
+I also expect that this research will provide actionable insights for how designing conservation areas while considering inter-group dynamics may allow for optimal resource provisioning under a much more broad set of circumstances than considering solitary groups alone.
 

### 2020-12-11 DissertationProposal/DissertationProposal.Rmd
ditching the abstract

@@
 
-\newpage
-# Summary
-Area-based conservation is critical for stemming global biodiversity loss.
-As such, each year billions of dollars are spent to protect key areas.
-It is common, however, for protected areas to fail to reach their stated conservation goals.
-The most chronic cause of conservation failure is a lack of local support and adoption of mandates.
-Conversely, historic protected area successes are also attributed to widespread adoption of conservation behaviors.
-Research examining how and why protected areas are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes.
-This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin local adoption of conservation practices in and around protected areas.
-Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.
-
-1. Develop a dynamic, theoretical model to generate predictions regarding if and how group-level adoption of protected areas can emerge and persist via competition between groups.
-
-2. Empirically test theoretical predictions of protected area adoption as a function of group competition and perceived costs and benefits to local community members.
-
-3. Using the ongoing spread of an improved efficiency cookstove design, examine how the nature of interpersonal relationships may modify conservation behaviors as they spread within groups.
-
-Globally, we have expended a tremendous amount of resources to identify conservation goals which will deliver the greatest benefits to biodiversity and ecosystem function.
-For these benefits to come to fruition however, conservation initiatives must be adopted at a rate which exceeds ecological degradation.
-This research will bring together cultural evolutionary science and ecology to identify the key mechanisms of conservation adoption.
 \newpage

### 2020-12-11 DissertationProposal/DissertationProposal.Rmd
deforestation Map

@@
 The reasons for if and when an individual adopts a new technology, especially in the contexts of conservation, are often unclear [@smith2014early].
-Even in cases where a technology does spread within a population, the specific use behavior and use outcome may become so greatly modified along the transmission chain that the population level result is unrecognizable from the original intent [@lewis_transmission_2012; @mcguigan_over-imitation_2011].
-Theoretical insights and controlled empirical research give suggestions for how intended technological adoption outcomes may be more likely to be retained during transmission from individual to individual or group to group.
-So far, these insights are yet to be validated empirically in a conservation context where technology adoption has real impacts for individuals, communities, and the environment, and those impacts may affect future adoptions [@king2007use; @muthukrishna_when_2016].
+Even in cases where a technology does spread within a population, the way individuals use the technology may change dramatically from person to person so that the population level result is unrecognizable from the original intent [@lewis_transmission_2012; @mcguigan_over-imitation_2011].
+Theoretical insights and controlled empirical research give suggestions for how intended technological adoption outcomes may be more likely to be retained during transmission from individual to individual or group to group [@king2007use; @muthukrishna_when_2016].
+So far, these insights are yet to be validated empirically in a conservation context where technology adoption has real impacts for individuals, communities, and the environment, and those impacts may affect future adoptions.
 
@@
 
-* Objective 1: Create an agent based model to theoretically test whether cultural group selection via success copying or direct competition results in greater conservation adoption and better environmental outcomes, particularly in environmental contexts where solitary groups fail to cooperate.
+* Objective 1: Create an agent-based model to theoretically test whether cultural group selection via success copying or direct competition results in greater conservation adoption and better environmental outcomes, particularly in environmental contexts where solitary groups fail to cooperate.
 
@@
 
+```{r ForestCoverMap, echo=FALSE, fig.cap="Change in annual forest cover loss in original wards with CoFMAs.
+Change values are the difference between average annual forest cover loss from 2001 - 2010 and the average annual forest cover loss from 2010 - 2018.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch2/DeforestationMap.png")
+```
+
 This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly.
@@
 
-```{r mapp, echo=FALSE, fig.cap="Satellite map of Pemba Island, Tanzania.
-Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "75%", fig.align='center', dpi=72, fig.pos= "H"}
+```{r mapp, echo=FALSE, fig.cap="Pemba Island, Tanzania.
+Three marked locations are the wards which will be surveyed for fuel-efficient cookstove copying fidelity and outcomes.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
 knitr::include_graphics("Figures/Ch3/InsetMap.png")

### 2020-12-11 DeforestationMap.R

@@ -1,0 +1,55 @@
+library(sf)
+library(tidyverse)
+library(magrittr)
+library(rmapshaper)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(ggspatial)
+
+
+
+sizes<-read_csv("~/Pemba_Project/CoFMASizes.csv")
+sizes%<>%group_by(CoFMA)%>%mutate(Total=sum(AlternativeUse,ConservationArea,UtilizationArea))%>%
+  mutate(UseArea=sum(AlternativeUse,UtilizationArea))%>%
+  mutate(UseAreaPerc=UseArea/Total)%>%mutate(ConsAreaPerc=ConservationArea/Total)
+
+sizes%<>%mutate(ChangeYearlyRateForestLoss_FromCofma=-1*(CoverChange_PercPerYear_2010_2018-CoverChange_PercPerYear_2001_2010))
+
+gadm3_3 <- readRDS("~/Pemba_Project/MapData/gadm36_TZA_3_sf.rds")
+#ggplot() + geom_sf(data = gadm3_3) + theme_bw()
+Pemba <- st_crop(gadm3_3, xmin = 39.5, xmax = 39.9,
+                 ymin = -5.6, ymax = -4.75)
+
+#ggplot() + geom_sf(data = Pemba) + theme_bw()
+Pemba[Pemba$NAME_3 == "Mgogoni" & Pemba$NAME_2 =="Chake",]$NAME_3<-"Mgogoni_2" #duplicate "mgogni" ward Does NOT have Cofma
+
+CofmaWards<-c("Changaweni", "Fundo", "Gando", "Kambini","Kangani","Kifundi","Kisiwa Panza",
+              "Mgelema","Mgogoni","Michenzani","Mjimbini","Mjini Wingwi","Msuka Magharibi","Mtambwe Kaskazini",
+              "Mtambwe Kusini","Shumba Mjini","Tondooni","Tumbe Magharibi")
+
+
+Pemba$CoFMA<-ifelse(Pemba$NAME_3 %in% CofmaWards, "CoFMA","No CoFMA")
+sizes$NAME_3<-sizes$CoFMA
+outcome<-sizes%>%ungroup()%>%select(NAME_3,ChangeYearlyRateForestLoss_FromCofma)
+Pemba<-merge(Pemba,outcome,by="NAME_3",all.x = TRUE)
+
+
+
+
+ggplot() + geom_sf(data = Pemba, aes(fill=ChangeYearlyRateForestLoss_FromCofma))+
[…]

### 2020-12-15 DissertationProposal/DissertationProposal.Rmd
changed ch2 intro. Next is methods.

@@
 Resources which move to the adjacent area but exceed the carrying capacity will also die.
-If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the conservation area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).
+If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the conservation area (${C_{mob}}$) up until a specified carrying capacity ($W_{cc}$).
 
-$$P_{mob} = (R*P_{r})-P_{r}$$
+$$C_{mob} = (R*C_{r})-C_{r}$$
 
-$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)+min\{P_{mob},W_{cc}-W_{r}\}}{N} $$
+$$H_{sust} \leq \frac{(1-R)*(W_{r}-H_{max}*N)+min\{C_{mob},W_{cc}-W_{r}\}}{N} $$
 
@@
 
-```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) for many combinationa of maximum harvest values and conservation area sizes.
+```{r abm2, out.width = "75%",echo=FALSE, fig.cap="Average harvest values accross five model runs for all agents in the final time step (30) for many combinationa of maximum harvest values and conservation area sizes.
 Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
@@
 ## Background
-Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection [@molleman_cultural_2013].
-One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use [@waring_coevolution_2017].
-Under this prediction, selection acts both on individuals and on groups of individuals.
-When the ratio of group benefit to individual cost of cooperation outweighs the individual fitness consequences of cooperation, those behaviors will be favored by selection.
-Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups [@boyd_transmission_2010; @henrich_cultural_2004].
+Human social structures developed and still exist in nested groups, ranging from a nuclear family unit to a national or even multi-national identity.
+Group behavioral norms can have profound impacts on the wellbeing of group members and their environment, far beyond that of a single individual.
+Cultural group selection theorizes, and to a small extent has empirically demonstrated, the mechanisms which establish and perpetuate group-level norms [@henrich_cultural_2004; @bell_culture_2009].
+Most relevant to community conservation is the development of norms which are beneficial for the group, but cause an individual to incur some personal cost; we refer to these as costly cooperative behaviors [@molleman_cultural_2013].
+Central to theories around the adoption of costly cooperative behaviors is the insight that selection acts on both individuals and groups.
+Individuals may engage in personally costly behavior if it is beneficial for their group, which they also depend on for their wellbeing [@handley_human_2020].
+The degree to which an individual depends on each of their nested groups is directly proportional to the personal costs they will incur for that group [@okasha_multilevel_2004; @price1972extension].
+To the extent that conservation is a collaborative dilemma, these insights from cultural group selection may be paramount to implementing effective community based conservation [@waring2015multilevel; @kline2018designing].
 
-While these predictions are theoretically sound, they must be tested empirically [@kline2018designing].
-Two such studies have begun to test the basis of cultural group selection in the natural world.
-The first demonstrated that culture offers greater fodder for group selection than do genetics.
-Researchers estimated the total variance in both cultural and genetic traits within and between groups of individuals.
-The proportion of variance in an allele or trait found between groups is referred to as genetic $F_{ST}$ or cultural $F_{ST}$ respectively.
-Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits [@bell_culture_2009].
+Conservation is a particularly wicked collection action problem because the group level costs and benefits operate at a much longer time scale than those of the individual.
+The group level benefits of conservation are often much less certain than the immediate individual costs of forgoing resources.
+Humans tend to steeply discount the benefits of their actions as those benefits move further away in time and the benefits become more uncertain [@dasgupta2005uncertainty].
+This hyperbolic discounting makes conservation norms very unlikely to emerge and persist under our current understanding of cultural group selection — the individual costs are high and group benefits are low.
+The role of institutions in human society is to reduce these barriers and to encourage collaborative behavior [@north1990institutions; @ostrom1990governing].
+One particular institution aimed at promoting costly cooperative behaviors around conservation is REDD+ [@angelsen2017redd; @angelsen2008moving].
 
-A second empirical study tested the prediction that large scale human cooperation is in fact a product of competition between cultural groups.
-Researchers created pair-wise estimates of cultural $F_{ST}$ between 12 ethnic groups.
-These estimates of cultural differentiation were combined with a survey of individual willingness to cooperate with members of the other groups.
-Results of this study showed a strong association between cultural similarity and willingness to cooperate.
-This research indicates that the presence of variable cultural groups influences within-group cooperative behavior [@handley_human_2020].
+REDD+ in Zanzibar was designed to give performance based payments to communities who successfully reduce deforestation [@burgess2010getting].
+In theory, these immediate, promised payments should alter the ratio of individual costs to group benefits so that costly cooperation is favored under cultural group selection.
+While this project was implemented in only a select few communities (Figure \@ref(fig:moon)), this setup should also promote the spread of community conservation norms to adjacent communities through success-biased imitation [@boyd_transmission_2010].
+
+The REDD+ program on Pemba helped establish CoFMAs in 18 wards on the island, some of which have been successful in slowing and even stopping deforestation (Figure \@ref(fig:ForestCoverMap).
+The payments for the conservation efforts however were never given to communities.
+While this is ethically troublesome, it is interesting in light of the fact that ten additional wards on the island have subsequently elected to establish CoFMAs.
+This leaves open the question of why community conservation areas are spreading on the island while they are not expected to do so under cultural group selection.
+One explanation posited by Andrews and Borgerhoff Mulder [-@andrews_cultural_2018] actually hinges on a major critique of REDD+; leakage [@andam2008measuring].
 
@@
 
-This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly.
-Insights from cultural group selection may provide solutions to pressing environmental challenges of collective action.
-If validated experimentally, principles of cultural group selection may allow conservation scientists to better predict the viability of conservation initiatives.
+Leakage [as defined by @andam2008measuring] is a form of land use spillover where an intervention, such as the CoFMAs established on Pemba by REDD+, has a negative, counteracting effect on the targeted outcome variable in another place.
+On Pemba, we have anecdotal testimonies that individuals in wards with CoFMAs have simply shifted their harvest pressure onto adjacent wards.
+While dubious for the efficacy of REDD+ writ large, on Pemba, leakage may actually be driving the spread of community based conservation observed in the ten new communities which elected to create CoFMAs post REDD+ [@andrewsdoes].
+Just as the performance based payments from REDD+ were designed to alter the individual/community cost/benefit matrix so that selection favors costly collaboration via immediate community payments, we hypothesize that leakage from wards with CoFMAs may also act to favor costly cooperative behavior as individuals must act to defend their access to resources they would not have otherwise.
 
-Protected areas are an ideal study system for testing the mechanisms of cultural group selection in real world conservation.
-When a community adopts a protected area, like the Community Forest Management Agreements observed on Pemba Island, Tanzania, individuals agree to forego a portion of their personal harvest potential for the benefit of the community.
-Under cultural group selection, we only expect this behavior to emerge when the proportion of dissimilar to similar cultural traits between a group and its neighbors outweighs the benefit/cost ratio of engaging in cooperative conservation.
-Plainly put, communities should be more likely to adopt CoFMAs if they are culturally dissimilar to their surrounding communities.
-This project will test this theory by evaluating the perceived individual costs and group benefits of CoFMAs, within and between group cultural similarities (cultural $F_{st}$), and actual adoption of CoFMAs in forest dependent wards on Pemba Island [@kline2018designing].
+If the spread of CoFMAs on Pemba is indeed a counterintuitive result of leakage, it may have considerable implications for REDD+ overall and community conservation areas broadly.
+To answer this, we first need to confirm that leakage is indeed occurring on Pemba as a result of individuals adhering to CoFMAs.
+Secondly, we need to evaluate the motivations of individuals who elected to establish CoFMAs in the ten wards who subsequently elected to do so.
+An alternative hypothesis is that individuals are copying the behaviors of more successful groups and are ambivalent towards or even unaware of leakage.
+The overall goal of this project is to quantify the degree to which leakage is driving the cascade of community conservation area adoption on Pemba.
 
 ### Question: 
-Do Pemban wards with stronger cultural differentiation from surrounding wards adopt CoFMAs at a higher rate than wards who are more culturally similar to their neighbors?
+Is leakage occurring as a result of individuals adhering to CoFMA regulations? If leakage is occurring, is it driving the spread of CoFMAs on Pemba?
 
 ### Predictions
-* P1: Cultural differentiation of Pemban wards will predict CoFMA adoption only when the perceived individual costs of CoFMAs are relatively low.
+* P1: Regardless of the demarcated size of the conservation area in each CoFMA, the leakage pressure on adjacent wards will be associated with the rate of deforestation, i.e.
+the extent to which individuals adhere to conservation area regulations.
 
-* P2: High levels of forest dependence and perceived costs of CoFMAs will prevent CoFMA adoption, even when groups are considerably culturally dissimilar to adjacent groups.
+* P2: In wards that have subsequently applied for CoFMAS, the primary motivator will be to combat leakage from adjacent wards, more so than copying their success or intrinsic motivations.
 

### 2020-12-16 DissertationProposal/DissertationProposal.Rmd
done for the day. Forever?

@@
 --- 
-title: Cultural Evolution of Conservation Behavior at the Intersection of Working
-  Landscapes and Protected Areas
+title: Cultural Evolution of Conservation Behaviors Within and Between Communities
 author: "Matt Clark"
@@
 
-# Chapter 2: Effects of inter-group resource competition and selection on community adoption of costly conservation policy.
+# Chapter 2: Effects of Inter-Group Resource Competition and Selection on Community Adoption of Costly Conservation Policy.
 
 ## Background
-Human social structures developed and still exist in nested groups, ranging from a nuclear family unit to a national or even multi-national identity.
+Human social structures are characterized by nested groups, ranging from a nuclear family unit to a national or even multi-national identity.
 Group behavioral norms can have profound impacts on the wellbeing of group members and their environment, far beyond that of a single individual.
@@
 In theory, these immediate, promised payments should alter the ratio of individual costs to group benefits so that costly cooperation is favored under cultural group selection.
-While this project was implemented in only a select few communities (Figure \@ref(fig:moon)), this setup should also promote the spread of community conservation norms to adjacent communities through success-biased imitation [@boyd_transmission_2010].
+While this project was implemented in only a select few wards (Figure \@ref(fig:moon)), this setup should also promote the spread of community conservation norms to adjacent wards through success-biased imitation [@boyd_transmission_2010].
 
-The REDD+ program on Pemba helped establish CoFMAs in 18 wards on the island, some of which have been successful in slowing and even stopping deforestation (Figure \@ref(fig:ForestCoverMap).
+The REDD+ program on Pemba helped establish CoFMAs in 18 wards on the island, some of which have been successful in slowing and even stopping deforestation (Figure \@ref(fig:ForestCoverMap)).
 The payments for the conservation efforts however were never given to communities.
@@
 
-```{r ForestCoverMap, echo=FALSE, fig.cap="Change in annual forest cover loss in original wards with CoFMAs.
-Change values are the difference between average annual forest cover loss from 2001 - 2010 and the average annual forest cover loss from 2010 - 2018.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch2/DeforestationMap.png")
-```
-
 Leakage [as defined by @andam2008measuring] is a form of land use spillover where an intervention, such as the CoFMAs established on Pemba by REDD+, has a negative, counteracting effect on the targeted outcome variable in another place.
 On Pemba, we have anecdotal testimonies that individuals in wards with CoFMAs have simply shifted their harvest pressure onto adjacent wards.
-While dubious for the efficacy of REDD+ writ large, on Pemba, leakage may actually be driving the spread of community based conservation observed in the ten new communities which elected to create CoFMAs post REDD+ [@andrewsdoes].
+While dubious for the efficacy of REDD+ writ large, on Pemba, leakage may actually be driving the spread of community based conservation observed in the ten new communities which elected to create CoFMAs post-REDD+ [@andrewsdoes].
 Just as the performance based payments from REDD+ were designed to alter the individual/community cost/benefit matrix so that selection favors costly collaboration via immediate community payments, we hypothesize that leakage from wards with CoFMAs may also act to favor costly cooperative behavior as individuals must act to defend their access to resources they would not have otherwise.
@@
 ### Predictions
-* P1: Regardless of the demarcated size of the conservation area in each CoFMA, the leakage pressure on adjacent wards will be associated with the rate of deforestation, i.e.
-the extent to which individuals adhere to conservation area regulations.
+* P1: Reduced deforestation rates inside CoFMA conservation areas will be correlated with increased deforestation in adjacent wards.
 
-* P2: In wards that have subsequently applied for CoFMAS, the primary motivator will be to combat leakage from adjacent wards, more so than copying their success or intrinsic motivations.
+* P2: In wards that have subsequently applied for CoFMAS, individuals who voted in favor of implementing the CoFMA will report leakage as their primary motivation.
 
 ## Research Design
-Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level.
-These group level selection pressures are what we hypothesize to drive the adoption of CoFMAs.
-Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are blurred.
-For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.
+This project has two main components, both of which benefit from preliminary work done by our collaborators already working in Pemba, and both of which will require additional on-the-ground data collection.
+The first component is to assess if individuals adhering to CoFMA regulations is associated with accelerated forest loss in nearby wards (leakage).
+The second component is to assess people’s motivations for electing to create CoFMAs in their own wards and see if those have changed over time.
 
-I will conduct an in-person survey of randomly sampled forest users in each Pemban ward.
-This survey will elicit information on approximately 50 cultural norms ranging from crime, cultural markers, punishment, family dynamics, etc.
-Cultural $F_{ST}$ values will be calculated for each trait given the following formula:
+*Assessing leakage*
 
-$$F_{STx} = \frac{\frac{n_{i}}{n_{i}+n_{j}} (p_{i}-\bar{p})^2 + \frac{n_{j}}{n_{i}+n_{j}} (p_{j}-\bar{p})^2}{\bar{p}(1-\bar{p})}$$
+Using a combination of Landsat 5, Landsat 7, and Landsat 8 images our collaborators have created composite images of average forest cover across all of Pemba for the years 2001, 2010, and 2018 [@cohen2004landsat].
+Data produced from these images classifies forest vs non-forest with approximately 95% accuracy across all three time periods [@andrewsdoes].
+As shown in (Figure \@ref(fig:ForestCoverMap)), we have shapefiles for the individual wards on Pemba [accessed from GADM -@areas2020gadm] and can compare rates of forest loss for each ward across years.
+While this data is useful, it does not allow us to differentiate between forest loss inside or outside the demarcated conservation and use areas for each CoFMA.
+These data may then be skewed by misinterpreting tree planting for woodlots as reforestation, or forest loss inside designated use areas as deforestation.
 
-Where $i$ and $j$ are the two populations being compared, $n_{i}$ and $n_{j}$ represent the number of survey respondents for each group respectively.
-For trait $x$, $\bar{p}$ is the total frequency of trait $x$ in both populations, $p_{i}$ and $p_{j}$ are the trait frequency in each population.
-The equation can be interpreted as the variance in the relative frequency of trait $x$ between populations divided by the total variance in the trait across both populations.
-These values will be averaged across all traits to produce the total cultural $F_{ST}$ values between each pair of wards.
+```{r ForestCoverMap, echo=FALSE, fig.cap="Change in annual forest cover loss in original wards with CoFMAs.
+Change values are the difference between average annual forest cover loss from 2001 - 2010 and the average annual forest cover loss from 2010 - 2018.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
+knitr::include_graphics("Figures/Ch2/DeforestationMap.png")
+```
 
-In addition to measuring the between and within group cultural variance, the in-person survey will also elicit information regarding individual’s perceived costs/benefits of CoFMAs, as well as group costs/benefits of CoFMAs.
-We can then use the cultural $F_{ST}$ values, paired with perceived costs and benefits to assess whether or not group cooperation will be favored by selection using the Rogers model of group selection by selective migration, from hereon referred to as the Rogers Inequality.
+To better assess if individuals are adhering to CoFMA regulations, we need to sort out the originally demarcated use and conservation areas in each CoFMA.
+To our knowledge, these data do not currently exist or are not available.
+As a first step in this research, we will attempt to access this information through the Department of Forestry and Non-Renewable Natural Resources libraries on Pemba.
+Where/if this data is not available, we will drive or walk the boundary of each area to create a shapefile to use for analysis.
 
-The Rogers Inequality model was initially designed to measure genetic evolution of groups via selective migration, but is equally applicable to, and has been adapted for cultural evolution [@bowles2006group].
-This model presents an inequality under which individually costly, yet group beneficial traits are favored by selection.
-The emergence of group cooperative norms will be favored by selection when the ratio of between-group variance ($F_{ST}$) to within-group variance ($1-F_{ST}$) exceeds the ratio of individual costs of a behavior ($C_{w}$) to group level population benefit ($B_{p}$).
-Thus, if the following inequality is true, then individually costly, group beneficial behavior is favored by selection [@kline2018designing].
+We will compare the pre-CoFMA and post-CoFMA deforestation rates inside the conservation area for each ward to get a better measure of the degree to which individuals actually adhered to CoFMA regulations.
+To then estimate leakage, we will compare these values to the changes in deforestation rates between the two time periods for all ten of the wards that subsequently elected to create CoFMAs.
+If leakage did occur, we expect to see an association between the decline of deforestation inside CoFMA conservation areas and an increase in deforestation in nearby wards.
 
-$$\frac{c_{w}}{b_{p}} < \frac{F_{ST}}{(1-F_{ST})}$$
+*Assessing motivations*
 
-While the primary output of this model is a true/false estimate of whether the emergence of a costly cooperative norm is favored by selection, we can also use the ratio presented in the inequality as a measure of selection strength.
-For example, we will expect Community Forest Management Agreements to be more strongly selected for when the right-hand side of the inequality is far greater than the left-hand side compared to when the sides are close to equal.
+The ten wards which have subsequently elected to create CoFMAs did so in 2018 and 2019.
+We know from our collaborators that on the order of magnitude of approximately 500 individuals in each ward voted on the proposition.
+Our intention is to randomly sample individuals in each ward to gain a representative sample of each of these voting populations.
+By surveying these individuals, we hope to gain insights on two main questions.
+What was the reason they voted in favor of adopting a CoFMA if they did?
+And have their motivations changed?
+
+To understand why individuals voted to adopt CoFMAs, we will ask about their perceptions of outside individuals stealing resources from their ward during that time period.
+We will ask a variety of questions about the individuals they believe were responsible and if they were aware of CoFMAs in adjacent wards.
+We will also ask a series of questions in regard to the long-term benefits of forest conservation in the absence of threats from outside groups.
+These questions will be targeted to let us tease apart the alternative explanations for why individuals voted to adopt CoFMAs.
+
+To further triangulate on the reason(s) why individuals adopt community conservation areas, we will rely on the local, anecdotal insight that in the absence of REDD+ payments, some of the original CoFMAs are no longer being adhered to by individuals harvesting from the forest.
+If this is the case *and* leakage is causing the cascade of community conservation areas on Pemba, we expect that leakage pressure will have decreased in some CoFMA adjacent wards and individuals will no longer support subsequent creation of CoFMAs.
+In the survey, we will also ask about individual’s current willingness to adopt CoFMAs and their current perceptions of tree poaching and conservation benefits.
+
+*Drawing conclusions*
+
+The situation on Pemba in regard to REDD+ and community conservation areas is a bit of a natural experiment.
+There are 18 original wards with CoFMAs, each of those with a different proportion conservation-designated forest.
+Subsequently, each ward has shown different rates in forest cover loss and external wards have elected to create CoFMAs themselves.
+We hope to tie these disparate sources of information together to create a full picture of the social-ecological system.
+
+We plan to analyze these results in a regression context where we will estimate the probability that an individual supports costly cooperation as a function of leakage.
+We plan to control for traits unique to individuals such as wealth and relative forest dependence.
+A final question we intend to answer is whether the hypothesized cascade of community conservation areas as a result of leakage is a permanent feature of the social-ecological landscape or just a temporary blip.
+The degree of drop off in support for CoFMAs between 2018/2019 and this research will serve as a good indicator for how the system may develop in the future.
 
@@
 This project benefits from preliminary work done by our collaborators at the Max Planck Institute for Evolutionary Anthropology (MPI).
-MPI researchers conducted a preliminary qualitative survey and collation of oral histories with the assistance of Department of Forests and Non-Renewable Natural Resources staff to confirm anecdotally that competition between wards has led to CoFMA adoption at the ward level.
+MPI researchers conducted a preliminary qualitative survey and collation of oral histories with the assistance of Department of Forests and Non-Renewable Natural Resources staff to confirm anecdotally that leakage is occurring and has led to CoFMA adoption at the ward level.
 These researchers have also visited wards across the island and recorded which wards have adopted CoFMAs and which have not.
-These adoption characteristics will be compared against the adoption characteristics recorded in my work as a form of data validation.
+This is in combination with the work mentioned above that our MPI collaborators have done on measuring deforestation rates on the island.
 
 ## Expected Outputs
-The primary outputs of data collection and analysis in this project will produce binary predictions of protected area adoption for each study ward, as well as the calculated strength of selection for protected area adoption for each study ward.
-I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of CoFMAs.
-Figure \@ref(fig:selPress) shows an example of what the results of this study might look like.
+This research is designed to contribute to our understanding of how and why community conservation areas spread and persist, and the adoption of costly cooperative norms broadly.
+Community buy-in and social-ecological outcomes of conservation areas vary significantly in practice [@naughton-treves_role_2005].
+Agrawal and Redford [-@agrawal2006poverty] argue that this is because conservation areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause conservation to be widely adopted and subsequently successful.
+This research will help close the gap in our understanding of conservation adoption and help us better allocate limited resources for conservation.
 The findings of this study will be submitted for publication in a peer reviewed, academic journal.
+Lastly, the findings of this research will help to empirically validate the insights from the agent-based model presented in the previous chapter.
+As we build validity in this model, we can use it to ask different and more in-depth questions about the nature of conservation adoption.
+The compatibility between this empirical study system and the agent-based model may allow us to generate and answer novel questions synergistically in a way neither project could do alone.
 
-```{r selPress, echo=FALSE, fig.cap="Simulated data showing protected area adoption vs calculated selection pressure for group adoption of protected areas.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
-knitr::include_graphics("Figures/Ch3/SelPressureAdopt.png")
-```
 \newpage
 
-# Chapter 3: Social influences on transmission fidelity in household level diffusion of beneficial technology.
+# Chapter 3: Social Influences on Transmission Fidelity in Household Level Diffusion of Beneficial Technology.
 
@@
 This step is critical as conservation is a fundamentally different system from where these theories were originally developed.
-The second step will empirically validate these theoretical insights by comparing calculated cultural group selection pressures to real-world adoption of Community Forest Management Agreements on Pemba Island, Tanzania.
+The second step will empirically validate these theoretical insights by determining why community adoption of conservation areas is spreading on Pemba Island, Tanzania.
 Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities to examine how behavioral efficiency may be altered by social contexts (e.g.

### 2021-02-23 HCRI_Grant/ExecutiveSummary.Rmd
HCRI Grant stuff

---
title: Resilient Landscapes and Livelihoods Through Better Conservation Planning
author: "Matt Clark"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: false
  pdf_document:
    toc: no
    number_sections: false
  bookdown::html_document2:
    toc: no
    number_sections: false
fontsize: 12pt
link-citations: yes
csl: AmJBot.csl
subtitle: Proposal for Boise State Hazard and Climate Resilience Institute Research Grant
---

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')
```

# Executive Summary
From 2001–2019, the world lost 9.7% of total tree cover, largely due to deforestation. These losses directly reduce ecosystem and human resilience to climate change. Despite decades of effort into international and community-based conservation, preventing deforestation is still a critical problem. Conservation projects that work in one place are abandoned in others, or worse, have adverse impacts on local communities. Much of the literature around deforestation has focused either on the relevant ecological processes or social processes in isolation. This project will explicitly study the feedbacks between social and ecological processes in the context of conservation interventions. Thus, our project aims to understand how features of the natural system interact with features of the social system to create conservation outcomes.

Our research integrates both empirical data collection and theoretical land use simulations of our study system, Pemba Island, Tanzania. Because of the variation in natural features, as well as the fact that Pemba has been host to a variety of both international and community-based conservation interventions, the island serves as an ideal place to assess the factors driving conservation success. To ensure that our research produces findings that help local communities in Pemba combat deforestation, we are collaborating with Community Forests Pemba, a nonprofit organization focused on climate resilience through sustainable agroforestry.

Results of this project will significantly improve our ability to predict conservation outcomes for both the environment and communities. To communicate simulation outcomes, we will produce an interactive dashboard displaying alternative conservation scenarios. We will host a workshop with our community partner to demonstrate how project results can inform ongoing conservation efforts. This work advances the HCRI Natural Environment pillar by providing a roadmap for conservation planning on Pemba and in complex social-ecological landscapes broadly. This work also contributes to the Economy & Society pillar by explicitly demonstrating the local economic viability of intervention strategies.

\newpage

### 2021-02-23 HCRI_Grant/Proposal.Rmd
HCRI Grant stuff

---
title: Resilient Landscapes and Livelihoods Through Better Conservation Planning
author: "Matt Clark"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: false
  pdf_document:
    toc: no
    number_sections: false
  bookdown::html_document2:
    toc: no
    number_sections: false
fontsize: 12pt
link-citations: yes
csl: pnas.csl
subtitle: Proposal for Boise State Hazard and Climate Resilience Institute Research Grant
bibliography: [Refs.bib]
header-includes:
- \usepackage{pdfpages}
- \usepackage[font={small}]{caption}
---

```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)
```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')
```
\newpage

# Executive Summary
From 2001–2019, the world lost 9.7% of total tree cover, largely due to deforestation. These losses directly reduce ecosystem and human resilience to climate change. Despite decades of effort into international and community-based conservation, preventing deforestation is still a critical problem. Conservation projects that work in one place are abandoned in others, or worse, have adverse impacts on local communities. Much of the literature around deforestation has focused either on the relevant ecological processes or social processes in isolation. This project will explicitly study the feedbacks between social and ecological processes in the context of conservation interventions. Thus, our project aims to understand how features of the natural system interact with features of the social system to create conservation outcomes.

Our research integrates both empirical data collection and theoretical land use simulations of our study system, Pemba Island, Tanzania. Because of the variation in natural features, as well as the fact that Pemba has been host to a variety of both international and community-based conservation interventions, the island serves as an ideal place to assess the factors driving conservation success. To ensure that our research produces findings that help local communities in Pemba combat deforestation, we are collaborating with Community Forests Pemba, a nonprofit organization focused on climate resilience through sustainable agroforestry.

Results of this project will significantly improve our ability to predict conservation outcomes for both the environment and communities. To communicate simulation outcomes, we will produce an interactive dashboard displaying alternative conservation scenarios. We will host a workshop with our community partner to demonstrate how project results can inform ongoing conservation efforts. This work advances the HCRI Natural Environment pillar by providing a roadmap for conservation planning on Pemba and in complex social-ecological landscapes broadly. This work also contributes to the Economy & Society pillar by explicitly demonstrating the local economic viability of intervention strategies.

\newpage

# Project Team

\newpage
# Project Narrative
### Background and Rationale
Extreme climatic events such as the hurricane that devastated Puerto Rico in 2017, the deadly European heatwave in 2019, and the recent Texas snowpocalypse are becoming more common due to anthropogenic-caused climate change [@easterling2000climate; @beniston2004extreme].  Intact natural areas buffer the effects of these extreme climatic events on both ecosystems and human communities [@lloret2012extreme; @alongi2008mangrove]. Conserving a broad and diverse array of natural areas is unequivocally our best insurance policy against the worsening impacts of climate change [@isbell2015biodiversity; @loreau2001biodiversity; @oliver2015biodiversity].

Today, on average, we degenerate our natural areas at a rate 56% faster than they regenerate [@network2020global]. In response, $24 billion is allocated to conserving natural areas annually [@waldron2020protecting; @waldron2017reductions]. Thirty six biodiversity hotspots have been identified globally as priority areas for these conservation dollars [@mittermeier2011global; @larsen2012conserving]. These biodiversity hotspots greatly overlap with the world’s poorest communities [@fisher2007poverty]. Because of this overlap and also due to the price and availability of land, many conservation areas are adjacent to, or even contain low income, often indigenous populations [@naughton2005role].

One area in particular, the Coastal Forests of Eastern Africa biodiversity hotspot (Figure \@ref(fig:Mapp)), has seen considerable foregin and community-based investment in conservation [@burgess2017two]. Despite these efforts, deforestation and biodiversity loss in the area continue to worsen [@burgess2017two; @cuadros2018empty]. Pemba Island, our study system for this project is one of two main islands in the Zanzibar archipelago, located in the Coastal Forests of Eastern Africa biodiversity hotspot.

```{r Mapp, echo=FALSE, fig.cap="Pemba Island, Tanzania. Black lines show geographic boundaries for 121 distinct wards on the island. Subfigure shows East Africa with the Coastal Forests of Eastern Africa biodiversity hotspot highlighted in red.", out.width = "40%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("InsetSimple.png")
```

Deforestation is a perennial issue on Pemba, as the yearly deforestation rate is nearly 2% and 90% of households rely exclusively on forest products (fuelwood and charcoal) to meet their daily cooking needs [@RGZ2014Zanzibar; @ely2000participatory]. As such, Pemba has been subject to a stream of foregin conservation initiatives since the mid 20th century. Meanwhile, despite the initial enthusiasm for conservation, deforestation remains apace in Pemba. Unfortunately, anecdotal evidence suggests that resentment is mounting amongst residents towards the train of conservation projects that have failed to deliver any kind of sustained economic compensation or ecological benefits.

This phenomenon is not rare. Conservation projects, while well intentioned, often fail to meet both ecological and community development goals [@christie_marine_2004; @watson_performance_2014]. In areas like Pemba, where the environment and livelihoods are closely linked, economic and ecological processes may have complex interactions and feedbacks, which make conservation outcomes difficult to predict.

Community Forests Pemba, our local collaborator, is a Pemban nonprofit organization established in 2009 with a mission to conserve and restore Pembas forests while supporting livelihood development on the island. The organization specifically seeks to build a resilient Pemba, ecologically and economically, with respect to climate change. Community Forests Pemba also works in 35 vulnerable communities across the island to disseminate information on sustainable agroforestry and gender equity to remote households. Our collaborative effort, if funded by this proposal, will seek to build an integrated understanding of why conservation and development initiatives have continued to fall short on Pemba, and how we can do better in the future.

### Aims and Objectives
Forest conservation is a complex mix of individual behaviors, informal group-level institutions, and formal programs or interventions. These social features exist on top of underlying variation in the natural environment. We will construct and parameterize an agent-based land use simulation to produce insights about how all of these moving parts interact and provide a roadmap for implementing successful conservation projects on-the-ground. Our model will generate predictions for ecological and economic viability of two common conservation interventions: protected areas and payments for ecosystem services Table \@ref(tab:tabinterventions).

We will produce two specific deliverables to immediately inform conservation efforts on-the-ground in Pemba.

1. We will create and host an interactive web tool to let individuals run our agent-based simulations to test alternative scenarios and better predict conservation outcomes. This web tool will accompany the scholarly publication of this work to better communicate findings to a broad audience.

2. In the winter following the work outlined here, we will hold a community information and training session with Community Forests Pemba to communicate the findings and implications of this project and to give a live demonstration of the web tool described above. This session will be designed to immediately inform ongoing conservation efforts.

### Research Methods
*Data Collection*

Empirical data collected for this project will inform the decision rules built into our agent-based land use simulations. Our objective is to understand how individuals make decisions in response to varying ecological and economic conditions. To achieve this, our Community Forests Pemba will identify community members who have been subject to a variety of different conservation interventions and rely on each of the four forest types on the island: high tropical rainforest, mangroves, coral rag forest, and agroforestry scrub matrix. Only community members who actively rely on harvesting fuelwood for household needs will be interviewed. Respondents will be compensated for their time (see budget).

With the help of a local translator arranged by Community Forests Pemba, Matt Clark and Dr. Jeff Andrews will conduct structured interviews with each of these community members. Interviews will consist of a series of scenarios where the respondent is asked how they would modify their wood harvesting strategy based on the environmental and economic conditions of interest.

*Modeling*

Dr. Vicken Hillis and Dr. Monique Borgerhoff-Mulder will oversee the agent-based modeling portion of this project. All model coding will be done by Matt Clark. Our simulations will vary features of the natural environment including: dispersal distance, market substitutability, regrowth rate, and starting degradation. We will then impose three different conservation intervention scenarios and features within them, all listed in Table \@ref(tab:tabinterventions).

Within this backdrop, individuals make harvesting decisions about where (protected or not) and how much fuelwood to harvest. Decision-making and learning rules are motivated by insights from cultural evolutionary theory. In particular, we’re interested in how intervention types affect individual’s access to resources given their environment, how individuals respond to these changes, and how community response further impacts their environment.

```{r tabinterventions, echo=F, fig.align='center', warning=FALSE}
intVars <- data.frame(Intervention = c("No intervention","Protected areas", "Payments for Ecosystem Services"), 
                      Scheme = c("Individuals harvest renewable resources at a rate which matches their household need, regardless of the location on the landscape or harvest rate compared to resource regeneration rate.",
                                 "Some portion of the landscape is designated for conservation and resources inside are socially unacceptable to harvest. Resource units produced inside conservation area may spillover to use area under sink/source dynamics. Some agents may defect and harvest from the conservation area.",
                                 "Agents are incentivised to forgo harvesting resources in return for economic payments. Payments may offset resource collection when agents previously sold harvested resources, but may fail to change incentive structures when agents directly depend on resources harvested."), 
                      Features = c("N/A",
                                      "Protected area size",
                                      "Market value for harvested fuelwood, Payment amount"))

kbl(intVars,format="latex", caption = "Three conservation intervention scenarios to be tested with the proposed agent-based model.",escape=F,booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = c("striped","hold_position"), font_size = 9)%>%
  #kable_classic_2(full_width=T)%>% 
row_spec(0,bold=TRUE)%>%
    column_spec(1, width = "8em")%>%
    column_spec(2, width = "25em")%>%
  column_spec(3, width = "8em")

```

### Key Implications
*Intellectual Merit*

Predicting if and how conservation actions will evolve has proven to be a major challenge for conservation scientists and practitioners [@child2010conceptual]. Conservation social science largely relies on case studies to draw correlations between social phenomena and conservation outcomes. Widely applicable theory around the long-term adoption of conservation norms is therefore sparse [@mahajan2020theory]. Recently however, theories from cultural evolution, a field of anthropology, have been identified as a potentially useful tool to better understand the mechanisms which drive the evolution of conservation initiatives [@brooks_applying_2018].

Cultural evolution oﬀers theories on the mechanisms behind the evolution of cooperative behaviors such as adhering to community conservation. Existing cultural group selection theory however, is highly abstract, relying on simple, analytical models of nonspeciﬁc behaviors [@lehmann_evolution_2006; @boyd_rapid_2011]. Current theory therefore does not oﬀer predictions about the evolution of conservation norms speciﬁcally [although see @waring2015multilevel; & -@waring_coevolution_2017]. This is problematic as conservation behaviors have behavior-environment feedbacks which are fundamentally different from generic cooperative behaviors. They operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates. Specific resources may also be substitutable for other goods on the market or communities may rely on them directly. As such, cultural group selection theory must be extended to include these dynamics in order to be eﬀectively applied to conservation.

*Broader Impacts*

The broader impacts of this research will advance the overall mission of HCRI to build resilient communities by providing methodology and tools to better plan conservation projects for nature and communities as an integrated system, both on Pemba and elsewhere. The project as described also closely aligns with both the Natural Environment and Economy & Society HRCI pillars in the following ways:

* *Natural Environment* - The work here will provide immediately actionable insights for conservation on Pemba Island and Zanzibar overall. Our community partner, among others, will be provided with the tools to make use of these insights as part of ongoing conservation and community development efforts. More generally, conservation is in need of unifying theory about which kinds of interventions succeed or fail as a result of underlying ecological and social conditions. This research will serve as a foundational building block for conservation science to consider complete social-ecological landscapes moving forward.

* *Economy & Society* - Fuelwood and charcoal are the primary source of energy to 40% of the world’s rural poor, producing more renewable energy than solar, wind, and hydroelectric sources combined [@muller2018state]. Conservation may be the key to ensuring continued access to vital resources for communities in an uncertain climatic future, but it may also backfire and disenfranchise communities to much needed resources. Similar to the pillar above, the work described here will be immediately relevant to better designing conservation projects in our study system, but the implications of the scholarly work are notably far-reaching. Our simulations will explicitly consider the access to resources of local communities both as an outcome and as a driver of conservation efficacy.

### Dissemination Plan
Scholarly results from this project that advance conservation science broadly will be published in a peer-reviewed academic journal. Currently we are targeting *Environmental Modeling and Software*. All model code will be posted on a public GitHub repository and linked to the publication. Before publication we will contribute to open, reproducible science by making our pre-print manuscripts available online using the preprint science repository arXiv.org.

Pemba specific results and the web tool which will accompany this project will be presented at a workshop with Community Forests Pemba described earlier in this proposal. This workshop is currently planned as an online, remote workshop given budget restrictions and the covid-19 pandemic. If possible, we would like to host this workshop in-person if conditions change.

### Timeline 
```{r timeline2,echo=FALSE, cache=TRUE, message = FALSE, warning=FALSE,fig.cap="Timeline of proposed research activities and team members involved. Top (green) row shows activities, bottom five (grey) rows show which team members will be involved in each activity.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H" }

knitr::include_graphics("TimelineBetter.png")
```

```{r timeline, eval=FALSE,echo=FALSE, cache=TRUE, message = FALSE, warning=FALSE} 
library(vistime)
library(ggplot2)

data <- read.csv(text="event,group,start,end,color
                       Data Collection,Phase,2021-07-01,2021-08-15,#ccece6
                       Model Building,Phase,2021-08-15,2021-12-15,#99d8c9
                       Workshop,Phase,2021-12-15,2021-12-31,#66c2a4
                       Matt Clark, Matt Clark,2021-07-01,2021-12-31,#bdbdbd
                       Community Forests Pemba, Community Forests Pemba,2021-07-01,2021-08-15,#bdbdbd
                         , Community Forests Pemba,2021-12-15,2021-12-31,#bdbdbd
                       Jeff Andrews, Jeff Andrews,2021-07-01,2021-12-31,#bdbdbd
                       Monique Borgerhoff-Mulder, Monique Borgerhoff-Mulder,2021-08-15,2021-12-15,#bdbdbd
                       Vicken Hillis, Vicken Hillis ,2021-08-15,2021-12-15,#bdbdbd"

                       )

#vistime(data) 
#p<-gg_vistime(data, title = "Table 2: Timeline of proposed research activities and team members involved")
p <- gg_vistime(data, optimize_y = T, title = "Table 2: Timeline of proposed research activities (green) and\nteam members involved (grey).")

p +theme_bw()+scale_x_datetime(date_breaks = "1 month")+ ggplot2::theme(
  plot.title = element_text(hjust = 0, size=15, color="black"),
  axis.text.x = element_text(size = 8, color = "black"),
  axis.text.y = element_blank(),
  axis.ticks.y = element_blank(),
  panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

```

### Future Directions
The Human-Environment Systems research center at Boise State has built a partnership with researchers at the Max Planck Institute for Evolutionary Anthropology already working on conservation research on Pemba Island, Tanzania. Although we consider our partnership to be strong, Boise State personnel have not yet visited the field site. If funded, this project will begin what we anticipate will be a long-lasting and fruitful research program. Thus, this proposal will not only fund necessary conservation research, but will also fund an international research partnership between both institutions.

Future research will seek to collect quantitative data in our system to validate the outcomes of our land use simulations. These data will include biodiversity and land cover data from Pemban forests, as well as community metrics on exposure to conservation interventions and willingness to participate in future interventions.

\newpage

# Matt Clark CV
\includepdf[pages={-}]{CURRICULUMVITAE.pdf}

\newpage

# Community Partner Letter of Support

\newpage

# References

### 2021-02-23 HCRI_Grant/Map_DiversityHotspot.R

@@ -1,0 +1,65 @@
+library(sf)
+library(tidyverse)
+library(magrittr)
+library(rmapshaper)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(ggspatial)
+library(grid)
+
+gadm3_3 <- readRDS("~/Pemba_Project/MapData/gadm36_TZA_3_sf.rds")
+Pemba <- st_crop(gadm3_3, xmin = 39.5, xmax = 39.9,
+                 ymin = -5.6, ymax = -4.75)
+Pemba[Pemba$NAME_3 == "Mgogoni" & Pemba$NAME_2 =="Chake",]$NAME_3<-"Mgogoni_2" #duplicate "mgogni" ward Does NOT have Cofma
+
+CofmaWards<-c("Changaweni", "Fundo", "Gando", "Kambini","Kangani","Kifundi","Kisiwa Panza",
+              "Mgelema","Mgogoni","Michenzani","Mjimbini","Mjini Wingwi","Msuka Magharibi","Mtambwe Kaskazini",
+              "Mtambwe Kusini","Shumba Mjini","Tondooni","Tumbe Magharibi")
+
+
+Pemba$CoFMA<-ifelse(Pemba$NAME_3 %in% CofmaWards, "CoFMA","No CoFMA")
+CWs<-Pemba%>%filter( CoFMA =="CoFMA")%>%
+  mutate(lon=map_dbl(geometry, ~st_centroid(.x)[[1]]), # add centroid values for labels
+         lat=map_dbl(geometry, ~st_centroid(.x)[[2]]))
+
+cols <- c("No CoFMA" = "#a6cee3", "CoFMA" = "#33a02c")
+
+africamap <- ne_countries(scale = 'medium', type = 'map_units',
+                          returnclass = 'sf',continent="africa")
+
+
+EastAfrica <- st_crop(africamap, xmin = 32, xmax = 60,
+                      ymin = -26, ymax = 5)
+
+
+hotspots<-read_sf("hotspots_2016_1.shp")
+CFEA<-hotspots%>%filter(NAME=="Coastal Forests of Eastern Africa" )
+ggplot()+geom_sf(data=CFEA)
+
+EAfr<-ggplot() + geom_sf(data = EastAfrica, fill="#d9d9d9")+
+    geom_sf(data = CFEA, fill="red")+
+    coord_sf(expand = FALSE)+ theme_bw()+
[…]

### 2021-02-23 HCRI_Grant/timeline.R

@@ -1,0 +1,47 @@
+library(vistime)
+library(ggplot2)
+
+data <- read.csv(text="event,group,start,end,color
+                       Data Collection,Phase,2021-07-01,2021-08-15,#c7e9c0
+                       Model Building,Phase,2021-08-15,2021-10-15,#a1d99b
+                       Dashboard Building, Phase, 2021-10-15,2021-12-15, #74c476
+                       Workshop,Phase,2021-12-15,2021-12-31,#41ab5d
+                       Matt Clark, Matt Clark,2021-07-01,2021-12-31,#bdbdbd
+                       Community Forests Pemba, Community Forests Pemba,2021-07-01,2021-08-15,#bdbdbd
+                         , Community Forests Pemba,2021-12-15,2021-12-31,#bdbdbd
+                       Jeff Andrews, Jeff Andrews,2021-07-01,2021-10-15,#bdbdbd
+                         , Jeff Andrews,2021-12-15,2021-12-31,#bdbdbd
+                       Monique Borgerhoff-Mulder, Monique Borgerhoff-Mulder,2021-08-15,2021-10-15,#bdbdbd
+                       Vicken Hillis, Vicken Hillis ,2021-08-15,2021-10-15,#bdbdbd"
+                 
+                       )
+
+
+
+#vistime(data) 
+#p<-gg_vistime(data, title = "Table 2: Timeline of proposed research activities and team members involved")
+p <- gg_vistime(data, optimize_y = T)
+
+
+p +theme_bw()+scale_x_datetime(date_breaks = "1 month")+ ggplot2::theme(
+  plot.title = element_text(hjust = 0, size=12, color="black"),
+  axis.text.x = element_text(size = 15, color = "black"),
+  axis.text.y = element_blank(),
+  axis.ticks.y = element_blank(),
+  panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+
+
+
+
+x<-lubridate::ymd(unique(data$start))
+x<-as.POSIXct(x,format='%Y-%M-%D')
+x<-format(as.POSIXct(x,format='%m/%d/%Y %H:%M:%S'),format='%m/%d/%Y')
+p +theme_bw()+scale_x_datetime(breaks =x )+ ggplot2::theme(
+  plot.title = element_text(hjust = 0, size=12, color="black"),
+  axis.text.x = element_text(size = 15, color = "black"),
[…]

### 2021-02-26 HCRI_Grant/Proposal.Rmd
updates with cfp letter

@@
 - \usepackage[font={small}]{caption}
+- \usepackage{wrapfig}
 ---
@@
 ```
+
 \newpage
@@
 We will host a workshop with our community partner to demonstrate how project results can inform ongoing conservation efforts.
-This work advances the HCRI Natural Environment pillar by providing a roadmap for conservation planning on Pemba and in complex social-ecological landscapes broadly.
+This work advances the HCRI Natural Environment pillar by providing a roadmap for conservation planning in Pemba and in complex social-ecological landscapes broadly.
 This work also contributes to the Economy & Society pillar by explicitly demonstrating the local economic viability of intervention strategies.
@@
 ### Background and Rationale
-Extreme climatic events such as the hurricane that devastated Puerto Rico in 2017, the deadly European heatwave in 2019, and the recent Texas snowpocalypse are becoming more common due to anthropogenic-caused climate change [@easterling2000climate; @beniston2004extreme].
-Intact natural areas buffer the effects of these extreme climatic events on both ecosystems and human communities [@lloret2012extreme; @alongi2008mangrove].
-Conserving a broad and diverse array of natural areas is unequivocally our best insurance policy against the worsening impacts of climate change [@isbell2015biodiversity; @loreau2001biodiversity; @oliver2015biodiversity].
+Diverse and intact natural landscapes are unequivocally our best insurance policy against the worsening impacts of climate change [@isbell2015biodiversity; @loreau2001biodiversity; @oliver2015biodiversity; @lloret2012extreme; @alongi2008mangrove].
+In response, $24 billion is spent on conservation interventions annually [@waldron2020protecting; @waldron2017reductions].
+As the world’s most biodiverse areas greatly overlap with the world’s poorest communities, many conservation interventions require community development and buy-in from vulnerable, often indenous populations [@fisher2007poverty; @mittermeier2011global; @larsen2012conserving; @naughton2005role].
 
-Today, on average, we degenerate our natural areas at a rate 56% faster than they regenerate [@network2020global].
-In response, $24 billion is allocated to conserving natural areas annually [@waldron2020protecting; @waldron2017reductions].
-Thirty six biodiversity hotspots have been identified globally as priority areas for these conservation dollars [@mittermeier2011global; @larsen2012conserving].
-These biodiversity hotspots greatly overlap with the world’s poorest communities [@fisher2007poverty].
-Because of this overlap and also due to the price and availability of land, many conservation areas are adjacent to, or even contain low income, often indigenous populations [@naughton2005role].
+\begin{wrapfigure}{r}{0.5\textwidth}
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{InsetSimple.png}
+  \end{center}
+  \caption{Pemba Island, Tanzania.
+Black lines show geographic boundaries for 121 distinct wards on the island.
+Subfigure shows East Africa with the Coastal Forests of Eastern Africa biodiversity hotspot highlighted in red.}
+\end{wrapfigure}
 
-One area in particular, the Coastal Forests of Eastern Africa biodiversity hotspot (Figure \@ref(fig:Mapp)), has seen considerable foregin and community-based investment in conservation [@burgess2017two].
-Despite these efforts, deforestation and biodiversity loss in the area continue to worsen [@burgess2017two; @cuadros2018empty].
-Pemba Island, our study system for this project is one of two main islands in the Zanzibar archipelago, located in the Coastal Forests of Eastern Africa biodiversity hotspot.
+Conservation interventions, while well intentioned, frequently fail to meet both ecological and community development goals [@christie_marine_2004; @watson_performance_2014].
+When environments and livelihoods are closely linked, economic and ecological processes may have complex interactions and feedbacks that make conservation outcomes difficult to predict.
+One area in particular, the Coastal Forests of Eastern Africa biodiversity hotspot, has seen considerable foreign and community-based investment in conservation, yet deforestation and biodiversity loss in the area continue to worsen [ @burgess2017two; @cuadros2018empty].
 
-```{r Mapp, echo=FALSE, fig.cap="Pemba Island, Tanzania.
+This study focuses on Pemba Island, Tanzania, located in the Coastal Forests of Eastern Africa biodiversity hotspot.
+Deforestation is a perennial issue in Pemba, as the yearly deforestation rate is nearly 2%, and 90% of households rely exclusively on forest products (fuelwood and charcoal) to meet their daily cooking needs [@RGZ2014Zanzibar; @ely2000participatory].
+As such, Pemba has been subject to a stream of foreign conservation initiatives since the mid 20th century, yet deforestation remains apace.
+Unfortunately, anecdotal evidence suggests that resentment is mounting amongst residents towards the train of conservation projects that have failed to deliver any sustained economic compensation or ecological benefits.
+Our collaborative effort will build land use simulations that integrate ecology and economics to demonstrate why conservation and development initiatives have continued to fall short in Pemba, and how we can do better in the future.
+
+This project is in collaboration with the Max Planck Institute for Evolutionary Anthropology, and Community Forests Pemba, a Pemban nonprofit organization established in 2009 with a mission to conserve and restore Pemba’s forests while supporting livelihood development on the island.
+The organization specifically seeks to build a resilient Pemba, ecologically and economically, with respect to climate change.
+This effort is mutually beneficial as Community Forests Pemba facilitate data collection and provide on-the-ground logistical support, while resulting land use simulations will inform future conservation efforts by the organization
+
+If funded, the project will establish a formal collaboration between the Hazards and Climate Resilience Center, the Max Planck Institute for Evolutionary Anthropology, Community Forests Pemba, and Boise State Human-Environment Systems.
+This funding will specifically support the dissertation work of one PhD student at Boise State, Matt Clark, and will yield an interactive web dashboard displaying all land use simulations described below.
+
+```{r Mapp, echo=FALSE,eval=FALSE, fig.cap="Pemba Island, Tanzania.
 Black lines show geographic boundaries for 121 distinct wards on the island.
-Subfigure shows East Africa with the Coastal Forests of Eastern Africa biodiversity hotspot highlighted in red.", out.width = "40%", fig.align='center', dpi=72, fig.pos= "H"}
+Subfigure shows East Africa with the Coastal Forests of Eastern Africa biodiversity hotspot highlighted in red.", out.width = "40%" ,fig.align='center', dpi=72, fig.pos= "H"}
 knitr::include_graphics("InsetSimple.png")
@@
 
-Deforestation is a perennial issue on Pemba, as the yearly deforestation rate is nearly 2% and 90% of households rely exclusively on forest products (fuelwood and charcoal) to meet their daily cooking needs [@RGZ2014Zanzibar; @ely2000participatory].
-As such, Pemba has been subject to a stream of foregin conservation initiatives since the mid 20th century.
-Meanwhile, despite the initial enthusiasm for conservation, deforestation remains apace in Pemba.
-Unfortunately, anecdotal evidence suggests that resentment is mounting amongst residents towards the train of conservation projects that have failed to deliver any kind of sustained economic compensation or ecological benefits.
-
-This phenomenon is not rare.
-Conservation projects, while well intentioned, often fail to meet both ecological and community development goals [@christie_marine_2004; @watson_performance_2014].
-In areas like Pemba, where the environment and livelihoods are closely linked, economic and ecological processes may have complex interactions and feedbacks, which make conservation outcomes difficult to predict.
-
-Community Forests Pemba, our local collaborator, is a Pemban nonprofit organization established in 2009 with a mission to conserve and restore Pembas forests while supporting livelihood development on the island.
-The organization specifically seeks to build a resilient Pemba, ecologically and economically, with respect to climate change.
-Community Forests Pemba also works in 35 vulnerable communities across the island to disseminate information on sustainable agroforestry and gender equity to remote households.
-Our collaborative effort, if funded by this proposal, will seek to build an integrated understanding of why conservation and development initiatives have continued to fall short on Pemba, and how we can do better in the future.
-
 ### Aims and Objectives
-Forest conservation is a complex mix of individual behaviors, informal group-level institutions, and formal programs or interventions.
-These social features exist on top of underlying variation in the natural environment.
-We will construct and parameterize an agent-based land use simulation to produce insights about how all of these moving parts interact and provide a roadmap for implementing successful conservation projects on-the-ground.
-Our model will generate predictions for ecological and economic viability of two common conservation interventions: protected areas and payments for ecosystem services Table \@ref(tab:tabinterventions).
+Our project will generate predictions for ecological and economic viability of two common conservation interventions: protected areas and payments for ecosystem services (Table \@ref(tab:tabinterventions), and provide a roadmap for implementing successful conservation projects on-the-ground.
 
@@
 1.
-We will create and host an interactive web tool to let individuals run our agent-based simulations to test alternative scenarios and better predict conservation outcomes.
+We will create and host an interactive web dashboard to let individuals run our agent-based simulations to visualize alternative scenarios and better predict conservation outcomes.
 This web tool will accompany the scholarly publication of this work to better communicate findings to a broad audience.
@@
 2.
-In the winter following the work outlined here, we will hold a community information and training session with Community Forests Pemba to communicate the findings and implications of this project and to give a live demonstration of the web tool described above.
+We will hold a community information and training session with Community Forests Pemba to communicate the findings and implications of this project and to give a live demonstration of the web dashboard described above.
 This session will be designed to immediately inform ongoing conservation efforts.
@@
 
+All simulations will be displayed visually on an interactive dashboard created using the Flexdashboard and R Shiny functionality in the R statistical programming language [@R; @flexD; @shiny].
+The dashboard will allow users to test all three intervention scenarios and their variations against all described social and ecological starting conditions.
+
 ```{r tabinterventions, echo=F, fig.align='center', warning=FALSE}
-intVars <- data.frame(Intervention = c("No intervention","Protected areas", "Payments for Ecosystem Services"), 
+intVars <- data.frame(Intervention = c("No intervention","Protected areas", "Payments for ecosystem services"), 
                       Scheme = c("Individuals harvest renewable resources at a rate which matches their household need, regardless of the location on the landscape or harvest rate compared to resource regeneration rate.",
@@
 
-The broader impacts of this research will advance the overall mission of HCRI to build resilient communities by providing methodology and tools to better plan conservation projects for nature and communities as an integrated system, both on Pemba and elsewhere.
+The broader impacts of this research will advance the overall mission of HCRI to build resilient communities by providing methodology and tools to better plan conservation projects for nature and communities as an integrated system, both in Pemba and elsewhere.
 The project as described also closely aligns with both the Natural Environment and Economy & Society HRCI pillars in the following ways:
 
-* *Natural Environment* - The work here will provide immediately actionable insights for conservation on Pemba Island and Zanzibar overall.
+* *Natural Environment* - The work here will provide immediately actionable insights for conservation in Pemba Island and Zanzibar overall.
 Our community partner, among others, will be provided with the tools to make use of these insights as part of ongoing conservation and community development efforts.
@@
 ### Timeline 
-```{r timeline2,echo=FALSE, cache=TRUE, message = FALSE, warning=FALSE,fig.cap="Timeline of proposed research activities and team members involved.
-Top (green) row shows activities, bottom five (grey) rows show which team members will be involved in each activity.", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H" }
+```{r timeline2,echo=FALSE, cache=TRUE, message = FALSE, warning=FALSE,fig.cap="Timeline of proposed research activities and team members involved.", out.width = "80%", fig.align='center', dpi=72, fig.pos= "H" }
 
-knitr::include_graphics("TimelineBetter.png")
+knitr::include_graphics("Timeline.png")
 ```
@@
 ### Future Directions
-The Human-Environment Systems research center at Boise State has built a partnership with researchers at the Max Planck Institute for Evolutionary Anthropology already working on conservation research on Pemba Island, Tanzania.
+The Human-Environment Systems research center at Boise State has built a partnership with researchers at the Max Planck Institute for Evolutionary Anthropology already working on conservation research in Pemba Island, Tanzania.
 Although we consider our partnership to be strong, Boise State personnel have not yet visited the field site.
@@
 
-# Matt Clark CV
 \includepdf[pages={-}]{CURRICULUMVITAE.pdf}
@@
 
-# Community Partner Letter of Support
+\includepdf[pages={-}]{CFP_Letter.pdf}
 

### 2021-03-01 HCRI_Grant/Proposal.Rmd
Proposal submitted

@@
 - \usepackage{pdfpages}
-- \usepackage[font={small}]{caption}
+- \usepackage[font={footnotesize}]{caption}
 - \usepackage{wrapfig}
+
 ---
@@
 # Executive Summary
-From 2001–2019, the world lost 9.7% of total tree cover, largely due to deforestation.
-These losses directly reduce ecosystem and human resilience to climate change.
-Despite decades of effort into international and community-based conservation, preventing deforestation is still a critical problem.
-Conservation projects that work in one place are abandoned in others, or worse, have adverse impacts on local communities.
-Much of the literature around deforestation has focused either on the relevant ecological processes or social processes in isolation.
-This project will explicitly study the feedbacks between social and ecological processes in the context of conservation interventions.
-Thus, our project aims to understand how features of the natural system interact with features of the social system to create conservation outcomes.
+Biodiversity buffers ecosystems and human communities against the worsening impacts of climate change.
+Despite decades of effort into conservation, global biodiversity loss is still a critical problem.
+Interventions that work in one place are abandoned in others, or worse, harm local communities.
+This project will combine empirical data collection with simulation modeling to study how feedbacks between the natural and social systems impact conservation.
 
-Our research integrates both empirical data collection and theoretical land use simulations of our study system, Pemba Island, Tanzania.
-Because of the variation in natural features, as well as the fact that Pemba has been host to a variety of both international and community-based conservation interventions, the island serves as an ideal place to assess the factors driving conservation success.
-To ensure that our research produces findings that help local communities in Pemba combat deforestation, we are collaborating with Community Forests Pemba, a nonprofit organization focused on climate resilience through sustainable agroforestry.
+This work will advance the HCRI Natural Environment and Economy & Society pillars by providing a roadmap for conservation planning in complex social-ecological landscapes while considering the local economic viability of intervention strategies.
+The aim of this project is to produce land use simulations demonstrating the viability of two conservation paradigms in our study system, Pemba Island, Tanzania.
+Pemba Island serves as an ideal place to study the factors driving conservation success due to the biological variation across the island and its history of both international and community-based conservation interventions.
+Simulations will be informed using interviews with local community members and insights from conservation social science.
 
-Results of this project will significantly improve our ability to predict conservation outcomes for both the environment and communities.
-To communicate simulation outcomes, we will produce an interactive dashboard displaying alternative conservation scenarios.
-We will host a workshop with our community partner to demonstrate how project results can inform ongoing conservation efforts.
-This work advances the HCRI Natural Environment pillar by providing a roadmap for conservation planning in Pemba and in complex social-ecological landscapes broadly.
-This work also contributes to the Economy & Society pillar by explicitly demonstrating the local economic viability of intervention strategies.
+To apply project insights locally, we are collaborating with Community Forests Pemba, a nonprofit organization focused on climate resilience through sustainable agroforestry.
+We will produce an interactive dashboard which will visualize how alternative conservation scenarios will impact the environment and local communities.
+Following completion of this dashboard, we will host a workshop with Community Forests Pemba to demonstrate how simulation results can directly inform future conservation planning.
+
+This scholarly work will build upon ongoing research in Pemba by the Max Planck Institute for Evolutionary Anthropology.
+Thus, funding this project will support international and cross-disciplinary collaboration, produce an interactive, public-facing dashboard, improve the conservation capacity of Community Forests Pemba, and support the published dissertation work of one Ecology, Evolution, & Behavior PhD student at Boise State University.
 
@@
 # Project Team
+**Matt Clark, doctoral student - Boise State University, Human-Environment Systems**
+
+This project is a major component of Matt’s dissertation work.
+As such, he will be involved with all stages of the project.
+Matt will travel to Pemba and work with Mbarouk M.
+Omar and Dr.
+Jeff Andrews to interview community members involved in conservation projects.
+Matt will code the land use model and build the web dashboard displaying simulation results.
+Matt, with the help of Mbarouk M.
+Omar, will co-lead the workshop presenting the project outcomes.
+
+\
+
+**Vicken Hillis PhD, assistant professor - Boise State University, Human-Environment Systems**
+
+Dr.
+Vicken Hillis will supervise Matt Clark at all stages of this project.
+Dr.
+Hillis will oversee the development of the land use model and the publication of the scholarly work.
+Dr.
+Hillis is also actively involved with securing additional funding for this project.
+\
+
+**Jeff Andrews PhD, postdoctoral researcher - Max Planck Institute for Evolutionary Anthropology**
+
+Dr.
+Jeff Andrews has worked in Pemba since 2017 and will help coordinate on-the-ground logistics for this project.
+Dr.
+Andrews will also help with data collection and serve as a co-author on all published work.
+
+\
+
+**Monique Borgerhoff-Mulder PhD,  professor - University of California Davis, Anthropology**
+
+Dr.
+Monique Borgerhoff-Mulder has worked in international conservation research for multiple decades.
+Dr.
+Borgerhoff-Mulder will advise the production of all scholarly work and ensure that it advances conservation social science.
+Dr.
+Borgerhoff-Mulder is also involved with securing additional financial support for this project, and leads our team research visa with Tanzania’s Department of Forestry and Nonrenewable Natural Resources.
+
+\
+
+**Mbarouk Mussa Omar, executive director - Community Forests Pemba**
+
+Mbarouk Mussa Omar will give logistical support to this project in Pemba.
+This will include arranging interviews with community members, assigning an employee of Community Forests Pemba to work as a translator, and coordinating help with language study for Matt Clark.
+Mbarouk Mussa Omar will co-lead the workshop where Matt Clark will present the results of this project to Community Forests Pemba and will make use of the insights for real-world application of our research findings.
 
@@
 In response, $24 billion is spent on conservation interventions annually [@waldron2020protecting; @waldron2017reductions].
-As the world’s most biodiverse areas greatly overlap with the world’s poorest communities, many conservation interventions require community development and buy-in from vulnerable, often indenous populations [@fisher2007poverty; @mittermeier2011global; @larsen2012conserving; @naughton2005role].
+These interventions are fundamentally about shaping people’s behavior through institutional incentives.
+Interventions commonly achieve this by directly paying local communities to forgo resources (payments for ecosystem services), or by legally protecting a portion of the landscape (protected areas) [@jack2008designing; @naughton2005role].
 
-\begin{wrapfigure}{r}{0.5\textwidth}
+\begin{wrapfigure}{r}{0.45\textwidth}
   \begin{center}
-    \includegraphics[width=0.48\textwidth]{InsetSimple.png}
+    \includegraphics[width=0.45\textwidth]{InsetSimpleCropped.png}
   \end{center}
+  \captionsetup{margin=0.25cm}
   \caption{Pemba Island, Tanzania.
-Black lines show geographic boundaries for 121 distinct wards on the island.
+Black lines show geographic boundaries for 121 distinct communities on the island.
 Subfigure shows East Africa with the Coastal Forests of Eastern Africa biodiversity hotspot highlighted in red.}
@@
 Conservation interventions, while well intentioned, frequently fail to meet both ecological and community development goals [@christie_marine_2004; @watson_performance_2014].
-When environments and livelihoods are closely linked, economic and ecological processes may have complex interactions and feedbacks that make conservation outcomes difficult to predict.
-One area in particular, the Coastal Forests of Eastern Africa biodiversity hotspot, has seen considerable foreign and community-based investment in conservation, yet deforestation and biodiversity loss in the area continue to worsen [ @burgess2017two; @cuadros2018empty].
+Economic and ecological processes are inherently interwoven, creating complex interactions that make conservation outcomes difficult to predict.
+One area in particular, the Coastal Forests of Eastern Africa biodiversity hotspot, has seen considerable foreign and community-based investment in conservation, yet deforestation and biodiversity loss in the area continue to worsen [@burgess2017two; @cuadros2018empty].
 
-This study focuses on Pemba Island, Tanzania, located in the Coastal Forests of Eastern Africa biodiversity hotspot.
+This study focuses on Pemba Island, Tanzania, located in the Coastal Forests of Eastern Africa biodiversity hotspot (Figure 1).
 Deforestation is a perennial issue in Pemba, as the yearly deforestation rate is nearly 2%, and 90% of households rely exclusively on forest products (fuelwood and charcoal) to meet their daily cooking needs [@RGZ2014Zanzibar; @ely2000participatory].
@@
 The organization specifically seeks to build a resilient Pemba, ecologically and economically, with respect to climate change.
-This effort is mutually beneficial as Community Forests Pemba facilitate data collection and provide on-the-ground logistical support, while resulting land use simulations will inform future conservation efforts by the organization
+This effort is mutually beneficial as Community Forests Pemba will facilitate data collection and provide on-the-ground logistical support, while resulting land use simulations will inform future conservation efforts by the organization.
 
-If funded, the project will establish a formal collaboration between the Hazards and Climate Resilience Center, the Max Planck Institute for Evolutionary Anthropology, Community Forests Pemba, and Boise State Human-Environment Systems.
-This funding will specifically support the dissertation work of one PhD student at Boise State, Matt Clark, and will yield an interactive web dashboard displaying all land use simulations described below.
+If funded, the project will establish a formal collaboration between the Hazards and Climate Resilience Institute, the Max Planck Institute for Evolutionary Anthropology, Community Forests Pemba, and Boise State Human-Environment Systems.
+This funding will specifically support the published dissertation work of one Ecology, Evolution, & Behavior PhD student at Boise State, Matt Clark, and will yield an interactive web dashboard displaying all land use simulations described below.
 
@@
 ### Aims and Objectives
-Our project will generate predictions for ecological and economic viability of two common conservation interventions: protected areas and payments for ecosystem services (Table \@ref(tab:tabinterventions), and provide a roadmap for implementing successful conservation projects on-the-ground.
+Our project will generate predictions for ecological and economic viability of two common conservation interventions: protected areas and payments for ecosystem services (Table \@ref(tab:tabinterventions2)).
 
-We will produce two specific deliverables to immediately inform conservation efforts on-the-ground in Pemba.
+We will produce two specific deliverables to immediately inform conservation efforts in Pemba.
 
 1.
-We will create and host an interactive web dashboard to let individuals run our agent-based simulations to visualize alternative scenarios and better predict conservation outcomes.
+We will create and host an interactive web dashboard to let individuals run our simulations and visualize alternative scenarios to better predict conservation outcomes.
 This web tool will accompany the scholarly publication of this work to better communicate findings to a broad audience.
@@
 
-Empirical data collected for this project will inform the decision rules built into our agent-based land use simulations.
-Our objective is to understand how individuals make decisions in response to varying ecological and economic conditions.
-To achieve this, our Community Forests Pemba will identify community members who have been subject to a variety of different conservation interventions and rely on each of the four forest types on the island: high tropical rainforest, mangroves, coral rag forest, and agroforestry scrub matrix.
+All empirical data collection for this project will be in the form of qualitative interviews with local community members.
+Community Forests Pemba will identify community members who have been subject to a variety of different conservation interventions and rely on each of the four forest types on the island: high tropical rainforest, mangroves, coral rag forest, and agroforestry scrub matrix.
 Only community members who actively rely on harvesting fuelwood for household needs will be interviewed.
@@
 Jeff Andrews will conduct structured interviews with each of these community members.
-Interviews will consist of a series of scenarios where the respondent is asked how they would modify their wood harvesting strategy based on the environmental and economic conditions of interest.
+Interviews will consist of a series of scenarios where the respondent is asked how they would modify their fuelwood harvesting strategy based on the environmental and economic conditions of interest.
+These data will inform how individuals make fuelwood harvesting decisions in our land use simulations in response to changes in the social-ecological landscape.
 
@@
 Vicken Hillis and Dr.
-Monique Borgerhoff-Mulder will oversee the agent-based modeling portion of this project.
+Monique Borgerhoff-Mulder will oversee the modeling portion of this project.
 All model coding will be done by Matt Clark.
-Our simulations will vary features of the natural environment including: dispersal distance, market substitutability, regrowth rate, and starting degradation.
-We will then impose three different conservation intervention scenarios and features within them, all listed in Table \@ref(tab:tabinterventions).
+Simulations, in the form of an agent-based model, will vary features of the natural and social environment, impose three different conservation intervention scenarios on the landscape, and observe how they impact the environment and community access to resources.
+All variable features are listed in Table \@ref(tab:tabinterventions2).
 
-Within this backdrop, individuals make harvesting decisions about where (protected or not) and how much fuelwood to harvest.
-Decision-making and learning rules are motivated by insights from cultural evolutionary theory.
-In particular, we’re interested in how intervention types affect individual’s access to resources given their environment, how individuals respond to these changes, and how community response further impacts their environment.
+Within this backdrop, simulated individuals will make harvesting decisions about where (protected or not) and how much fuelwood to harvest.
+Individuals will learn and adjust harvest strategies in response to their environment.
+Overall, we’re interested in how environmental conditions determine the economic impact of interventions, how individuals respond to these impacts, and how community response further shapes their environment.
 
@@
 
-```{r tabinterventions, echo=F, fig.align='center', warning=FALSE}
+```{r tabinterventions, echo=F, fig.align='center', warning=FALSE, eval=FALSE}
 intVars <- data.frame(Intervention = c("No intervention","Protected areas", "Payments for ecosystem services"), 
-                      Scheme = c("Individuals harvest renewable resources at a rate which matches their household need, regardless of the location on the landscape or harvest rate compared to resource regeneration rate.",
+                      Scheme = c("Baseline comparison.
+Individuals harvest renewable resources at a rate which matches their household need, regardless of the location on the landscape or harvest rate compared to resource regeneration rate.",
                                  "Some portion of the landscape is designated for conservation and resources inside are socially unacceptable to harvest.
-Resource units produced inside conservation area may spillover to use area under sink/source dynamics.
-Some agents may defect and harvest from the conservation area.",
-                                 "Agents are incentivised to forgo harvesting resources in return for economic payments.
+Resources produced inside the conservation area may spillover to the adjacent landscape.
+Some individuals may steal resources from the conservation area.",
+                                 "Individuals are incentivised to forgo harvesting resources in return for economic payments.
 Payments may offset resource collection when agents previously sold harvested resources, but may fail to change incentive structures when agents directly depend on resources harvested."),
@@
 
+```{r tabinterventions2, echo=F, fig.align='center', warning=FALSE}
+intVars2 <- data.frame(Intervention = c("No intervention","Protected areas", "Payments for ecosystem services"), 
+                      Scheme = c("Baseline comparison.
+Individuals harvest renewable resources at a rate which matches their household need, regardless of the location on the landscape or harvest rate compared to resource regeneration rate.",
+                                 "Some portion of the landscape is designated for conservation and resources inside are socially unacceptable to harvest.
+Resources produced inside the conservation area may spillover to adjacent landscapes.
+Some individuals may steal resources from the conservation area.",
+                                 "Individuals are incentivised to forgo harvesting resources in return for economic payments.
+Payments may offset resource collection when individuals previously sold harvested resources, but may fail to change incentive structures when individuals directly depend on resources harvested."),
+                      IFeatures = c("N/A",
+                                      "Protected area size",
+                                      "Market value for fuelwood, Payment amount"),
+                     EFeatures = c("Regrowth rate, Starting degradation", "Seed dispersal distance, Regrowth rate, Starting degradation", "Market substitutability, Regrowth rate, Starting degradation") )
+
+colnames(intVars2)[3]<-"Intervention features"
+colnames(intVars2)[4]<-"Environmental features"
+
+kbl(intVars2,format="latex", caption = "Three conservation intervention scenarios to be tested with the proposed land use model.",escape=F,booktabs = T) %>% 
+  kableExtra::kable_styling(latex_options = c("striped","hold_position"), font_size = 9)%>%
+  #kable_classic_2(full_width=T)%>% 
+row_spec(0,bold=TRUE)%>%
+    column_spec(1, width = "8em")%>%
+    column_spec(2, width = "23em")%>%
+  column_spec(3, width = "8em")%>%
+  column_spec(4, width = "10em")
+
+```
+
+\
+
 ### Key Implications
@@
 
-Predicting if and how conservation actions will evolve has proven to be a major challenge for conservation scientists and practitioners [@child2010conceptual].
-Conservation social science largely relies on case studies to draw correlations between social phenomena and conservation outcomes.
-Widely applicable theory around the long-term adoption of conservation norms is therefore sparse [@mahajan2020theory].
-Recently however, theories from cultural evolution, a field of anthropology, have been identified as a potentially useful tool to better understand the mechanisms which drive the evolution of conservation initiatives [@brooks_applying_2018].
-
-Cultural evolution oﬀers theories on the mechanisms behind the evolution of cooperative behaviors such as adhering to community conservation.
-Existing cultural group selection theory however, is highly abstract, relying on simple, analytical models of nonspeciﬁc behaviors [@lehmann_evolution_2006; @boyd_rapid_2011].
-Current theory therefore does not oﬀer predictions about the evolution of conservation norms speciﬁcally [although see @waring2015multilevel; & -@waring_coevolution_2017].
+Predicting if and how conservation initiatives will evolve has proven to be a major challenge for conservation scientists and practitioners [@child2010conceptual].
+Widely applicable theory around the long-term social viability of conservation projects is sparse [@mahajan2020theory].
+Conservation is a complex mix of individual behaviors, informal group-level institutions, and formal interventions.
+These social features exist on top of underlying variation in the natural environment.
+Theories from cultural evolution, a field of anthropology, examine the evolution of behaviors generally and may be applicable to conservation [@brooks_applying_2018].
+Existing cultural evolution theory is highly abstract, offering insights only about nonspeciﬁc cooperative behaviors [@lehmann_evolution_2006; @boyd_rapid_2011].
 This is problematic as conservation behaviors have behavior-environment feedbacks which are fundamentally different from generic cooperative behaviors.
@@
 Specific resources may also be substitutable for other goods on the market or communities may rely on them directly.
-As such, cultural group selection theory must be extended to include these dynamics in order to be eﬀectively applied to conservation.
+As such, cultural evolutionary theory must be extended to include these dynamics in order to be eﬀectively applied to conservation.
 
@@
 The broader impacts of this research will advance the overall mission of HCRI to build resilient communities by providing methodology and tools to better plan conservation projects for nature and communities as an integrated system, both in Pemba and elsewhere.
-The project as described also closely aligns with both the Natural Environment and Economy & Society HRCI pillars in the following ways:
+The project also closely aligns with both the Natural Environment and Economy & Society HRCI pillars in the following ways:
 
@@
 * *Economy & Society* - Fuelwood and charcoal are the primary source of energy to 40% of the world’s rural poor, producing more renewable energy than solar, wind, and hydroelectric sources combined [@muller2018state].
-Conservation may be the key to ensuring continued access to vital resources for communities in an uncertain climatic future, but it may also backfire and disenfranchise communities to much needed resources.
-Similar to the pillar above, the work described here will be immediately relevant to better designing conservation projects in our study system, but the implications of the scholarly work are notably far-reaching.
-Our simulations will explicitly consider the access to resources of local communities both as an outcome and as a driver of conservation efficacy.
+Conservation may be the key to ensuring continued access to vital resources for communities in an uncertain climatic future, but it may also backfire and disenfranchise communities of much needed resources, especially as the world’s most biodiverse areas greatly overlap with the world’s poorest communities [@fisher2007poverty; @mittermeier2011global; @larsen2012conserving; @naughton2005role].
+Similar to the pillar above, the work described here will be immediately relevant to better designing conservation projects in our study system, but the implications of the scholarly work are far-reaching.
+Our simulations will explicitly consider the access to resources by local communities both as an outcome and as a driver of conservation efficacy.
 
@@
 
-Pemba specific results and the web tool which will accompany this project will be presented at a workshop with Community Forests Pemba described earlier in this proposal.
-This workshop is currently planned as an online, remote workshop given budget restrictions and the covid-19 pandemic.
+Pemba specific results and the interactive dashboard which will accompany this project will be presented at a workshop with Community Forests Pemba described earlier in this proposal.
+This workshop is currently planned as an online, remote workshop given budget restrictions and the COVID-19 pandemic.
 If possible, we would like to host this workshop in-person if conditions change.
@@
 ### Future Directions
-The Human-Environment Systems research center at Boise State has built a partnership with researchers at the Max Planck Institute for Evolutionary Anthropology already working on conservation research in Pemba Island, Tanzania.
-Although we consider our partnership to be strong, Boise State personnel have not yet visited the field site.
-If funded, this project will begin what we anticipate will be a long-lasting and fruitful research program.
-Thus, this proposal will not only fund necessary conservation research, but will also fund an international research partnership between both institutions.
-
+We view the work proposed here as the first step toward a long-term, cross disciplinary collaboration between HCRI, the Max Planck Institute for Evolutionary Anthropology, Community Forests Pemba, and Human-Environment Systems.
+We hope to establish a research program in Pemba that will create lasting impacts for community members and produce significant scholarly contribution.
 Future research will seek to collect quantitative data in our system to validate the outcomes of our land use simulations.
 These data will include biodiversity and land cover data from Pemban forests, as well as community metrics on exposure to conservation interventions and willingness to participate in future interventions.
+This study system is well positioned to investigate questions from conservation social science and community resilience to climate change.
+All research will be co-produced with Community Forests Pemba to ensure broader impacts and real-world applicability.
 
 \newpage
+# Budget 
+```{r tabbudget, echo=F, fig.align='center', warning=FALSE}
+budgVars <- data.frame(Item = c("Airfare to Zanzibar","Airfare to Pemba","Lodging", "Field interpreter wage","Swahili tutoring", "Interview compensation", "Land transportation","TOTAL"), 
+                      Amount = c("2,500","200",
+                                 "650","500","300", "200", "300", "4,650"), 
+                      Details = c("Air travel round trip for Matt Clark from Boise to Abeid Amani Karume International Airport as of 2/15/2021","Air travel round trip Abeid Amani Karume International Airport to Pemba Airport as of 2/15/2021",
+                                      "Lodging for Matt Clark at Hill View Guest House, Wete Town, Pemba Island for 50 days @ 13/day. ",
+                                      "Payment for interpreter to assist with community interviews.
+Interpreter will be hired from Community Forests Pemba staff.", "Language tutoring for Matt Clark.
+Tutor will be coordinated by Community Forests Pemba.", "Compensation for community members who participate in qualitative interviews.", "Transportation in Pemba including gasoline for motorbike and various taxi travel."," " ))
 
-\includepdf[pages={-}]{CURRICULUMVITAE.pdf}
+colnames(budgVars)[2]<-"Amount (USD)"
+
+kbl(budgVars,format="latex", caption = "Project budget and justification.",escape=F,booktabs = T) %>% 
+  kableExtra::kable_styling(latex_options = c("striped","hold_position"), font_size = 12)%>%
+  #kable_classic_2(full_width=T)%>% 
+row_spec(0,bold=TRUE)%>%
+    column_spec(1, width = "10em")%>%
+    column_spec(2, width = "4em")%>%
+  column_spec(3, width = "27em")
+
+```
+
+\includepdf[pages={-}]{AbbreviatedCURRICULUMVITAE.pdf}
 

### 2021-04-12 AnthroposphereDraft.Rmd
cleaning up

--- 
title: Anthroposphere
author: "Matt Clark"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
fontsize: 12pt
link-citations: yes
csl: HCRI_Grant/pnas.csl
subtitle: 
bibliography: [DissertationProposal/References/Dissertationproposal.bib,DissertationProposal/References/NonZoteroRefs.bib,HCRI_Grant/Refs.bib,Anthroposphere.bib, Anthroposphere2.bib]
---
Extreme climatic events like the hurricane that devastated Puerto Rico in 2017 and the recent Texas snowpocalypse garner much of the public attention when it comes to the risks of climate change. While it is true that these events are concerning and becoming more frequent, the ever-worsening press of continuous impacts such as sea level rise, rangeland conifer encroachment, and coastal erosion are much more disastrous for human livelihoods [@easterling2000climate; @beniston2004extreme]. The livelihoods which are most impacted by these changes are those of the world's poorest and most vulnerable communities — the populations least responsible for greenhouse gas emissions [@patz2007climate; @voelkel2018assessing; @green2009disproportionate]. While the global economy is slowly moving away from fossil fuels and toward renewable energy sources, the climatic shifts we are seeing today are projected to worsen in frequency and severity for the foreseeable future [@IPCC2007climate].

Our best insurance policy against the worsening impacts of climate change is unequivocally to conserve a broad and diverse array of natural areas [@isbell2015biodiversity; @loreau2001biodiversity; @oliver2015biodiversity; @lloret2012extreme]. Mangrove forests for example, buffer communities against storms and floods, hillside vegetation reduces the risk of landslides, and most importantly, ecosystems suck up atmospheric carbon [@spiker2003national; @alongi2008mangrove].

Luckily for us, much like the way a retiree might live off the interest of a 401(k) account without impacting the principal, these systems regenerate themselves even while we skim off the top. Actually, until approximately 50 years ago, the overall rate at which our natural resources were growing back exceeded the rate at which we were consuming and degrading them [@network2019global]. Today however, on average we degenerate our natural areas at a rate 56% faster than they regenerate — we are tapping into the principal [@network2020global].

With this recognition, conservation has emerged as a major focus for political, private, and individual agendas, with $24 billion invested in conservation worldwide annually [@waldron2020protecting]. Both the number and size of geographic areas designated for conservation is skyrocketing each year [@IUCN_Protect]. The crown jewel of these conservation efforts are the world’s 36 biodiversity hotspots. Biodiversity hotspots are areas that have thousands of species found nowhere else on earth and, by definition, have lost at least 70%  of their native vegetation. The deluge of conservation dollars into these areas, which make up just 2.4% of earth’s land surface, is juxtaposed against the fact that they largely overlap with the world’s poorest, most rural communities. Thus, the day-to-day needs of local communities and Western conservation projects inevitably collide.

Conservation is a game of spillovers. Marine protected areas for example are intended to give safe harbor to fish populations where they can multiply, leave, and increase yields for fishing boats far outside their boundaries [@marshall2019underestimating]. On land, parks have up front costs to establish, but quickly bolster booming recreation and tourism economies [@koontz2017visitor]. Social and ecological systems are inherently interwoven, especially in biodiversity hotspots where many people rely directly on their environment to meet their daily needs. It may come as no surprise that these spillovers, or consequences of conservation interventions outside the initial geographic and temporal scope of the project, are not always beneficial for local people or the environment.

One particularly wicked spillover in biodiversity hotspots is leakage. Leakage occurs when interventions limit harvest or land development inside a specific geographic area, but that harvest pressure is only displaced onto adjacent landscapes, resulting in no net reduction of deforestation. Leakage is a perennial problem in land conservation. It has attenuated efforts to slow the spread of oil palm in Borneo, limit agricultural expansion in the United States, establish an eco-tourism economy in Costa Rica, reverse deforestation in Cambodia and Vietnam, and curb the high profile land conversion in the Peruvian and Brazilian Amazon [@heilmayr_deforestation_2020; @jacobson_temporal_2014; @robalino_heterogeneous_2017; @magliocca_direct_2020; @meyfroidt_forest_2009; @oliveira_land-use_2007; @alix-garcia_forest_2017].

It is well documented that leakage and other spillovers are increasing in frequency and magnitude along with conservation interventions worldwide [@pfaff_spillovers_2017]. While these increases inherently make leakage difficult to predict and document, climatic instability is also making it more difficult to predict the long term and large scale outcomes of conservation interventions. The last half-century of human disturbance has caused regime shifts in ecosystems globally and will continue to do so. This underlying base rate of environmental change, combined with the interconnectedness of ecological and social systems causes conservation interventions to be fraught with unintended consequences.

Communities living in biodiversity hotspots have a two-fold lived experience of climate change. As previously mentioned, these communities are among the least responsible for greenhouse gas emissions, but bear the brunt of its consequences such as hurricanes and agricultural salinization. An underappreciated impact of climate change on these communities however, is the ever increasingly forceful onslaught of Western conservation interventions. Conservation happens where land is cheap, and land is cheap near the world’s most vulnerable communities. Thus, corporate net-zero promises and the privatization of a carbon credit market are drastically altering the livelihoods of some of the world’s poorest people with limited knowledge of the large scale or long term social, or even environmental, impacts.

“Unfortunately, anecdotal evidence suggests that resentment is mounting amongst residents towards the train of conservation projects that have failed to deliver any sustained economic compensation or ecological beneﬁts.” -Dr. Jeff Andrews, anthropologist studying conservation impacts in Pemba Island, Tanzania

Conservation does not have to be a shot in the dark. Research has repeatedly demonstrated that local communities can better predict the outcomes of conservation interventions than can foreign practitioners [@halbrendt2014differences; @agrawal1999enchantment]. Community-based conservation initiatives are also highly effective in catalyzing the behavioral changes required for most interventions to be effective.

Instead of implementing conservation interventions directly, the conservation dollars spent on biodiversity hotspots would go much further to support and monitor community-based conservation projects. In fact, expanding monitoring is among the greatest contributions that Western conservation funding can make to protecting the world’s natural landscapes. Conservation interventions have effects on the landscape and on communities far outside of their initial geographic and temporal scope. Monitoring these effects is the first step to understanding and addressing the leakage which plagues current conservation efforts.

Under our current trajectory, global biodiversity hotspots are poised to become a graveyard of trialed and abandoned conservation interventions. As with all impacts of climate change, it is the world’s most vulnerable communities who are paying the price. Implementing meaningful change in complex social-ecological landscapes is always going to be difficult, but we have ways to make it better. We need to greatly expand the scope with which we monitor the outcomes of these projects and let communities themselves take the wheel.

\newpage
#Work Cited

### 2021-04-12 DissertationProposal/DissertationProposal.Rmd
cleaning up

@@
 Researchers have posited that cultural group selection theory should be applicable to understanding the spread and persistence of conservation behaviors such as adherence to community conservation areas or adoption of new efficient technologies [@kline2018designing].
-Cultural group selection theory offers mechanistic insights for the emergence and spread of behaviors within and between groups broadly, but its applicability to conservation specifically is unknown.
+Cultural group selection theory offers mechanistic insights for the emergence and spread of behaviors within and between groups broadly, but its applicability to conservation is unknown.
 Conservation has many aspects which make it unique compared to behavioral adoption in the abstract.
-In conservation, an individual’s behavior directly affects their environment, which may in turn become altered in a way which shapes future individual decision making.
+In conservation, an individual’s behavior affects their environment, which may in turn become altered in a way which shapes future individual decision making.
 Feedbacks between individual behavior and resource availability are further complicated by resource characteristics such as carrying capacity, mobility, and regeneration rate.
-These specifics make conservation a fundamentally different system from general behavioral adoption and cooperation, thus we must extend cultural group selection theory considering conservation explicitly.
+These specifics make conservation a fundamentally different system from general behavioral adoption and cooperation, thus we must extend cultural group selection theory considering conservation.
 The aim of this research is to expand the theories of behavioral adoption via cultural group selection to explicitly consider factors unique to conservation, and to empirically examine the mechanisms of conservation behavioral adoption within and between groups.
-This research will offer insights for strategically seeding both community conservation areas and new technologies so that they may offer the greatest benefit to communities and the preservation of renewable resources.
+This research will offer insights for implementing both community conservation areas and new technologies so that they may offer the greatest benefit to communities and the preservation of renewable resources.
 
@@
 Showing wards with and without CoFMAs.
-For each CoFMA, the size (ha) is proportional to the size of the circle and the percent area designated as Use and Conservation is indicated by the blue and green colors respectively.", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
+For each CoFMA, the size (ha) is proportional to the size of the circle and the percent area designated as \\textit{use} and \\textit{conservation} is indicated by the blue and green colors respectively.", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
 knitr::include_graphics("Figures/MoonMapEditedNoLabs.png")
@@
 
-# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Community Conservation Areas. 
+# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Community Conservation Areas 
 ## Background
@@
 
-```{r FirstABMb, out.width = "100%",echo=FALSE, fig.cap="Resource quantity and mean harvest vales over time for one run of the agent-based model at a maximum harvest valuue one above the calculated value for sustainable harvest.
+```{r FirstABMb, out.width = "100%",echo=FALSE, fig.cap="Resource quantity and mean harvest vales over time for one run of the agent-based model at a maximum harvest value one above the calculated value for sustainable harvest.
 Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources.
@@
 This behavior depletes the resource pool in the conservation area and lowers the number of resources which regenerate and agents can harvest in the subsequent time step.
-Note that full defection is defined as agents spending an equal proportion of their effort in the conservation area as the proportion conservation (i.e.
+Note that full defection is when agents spending an equal proportion of their effort in the conservation area as the proportion of resources conserved (i.e.
 they treat the entire landscape equally).
@@
 
-In the competing groups scenario, the model will follow the exact same scheduling process as in the separate groups scenario, but a specified portion of the working landscape(s) will be shared among the groups.
+In the competing groups scenario, the model will follow the same scheduling process as in the separate groups scenario, but a specified portion of the working landscape(s) will be shared among the groups.
 Mobile resources leaving the conservation areas will emigrate to the working landscape as in the single group model.
@@
 
-# Chapter 2: Effects of Inter-Group Resource Competition and Selection on Community Adoption of Costly Conservation Policy.
+# Chapter 2: Effects of Inter-Group Resource Competition and Selection on Community Adoption of Costly Conservation Policy
 
@@
 Cultural group selection theorizes, and to a small extent has empirically demonstrated, the mechanisms which establish and perpetuate group-level norms [@henrich_cultural_2004; @bell_culture_2009].
-Most relevant to community conservation is the development of norms which are beneficial for the group, but cause an individual to incur some personal cost; we refer to these as costly cooperative behaviors [@molleman_cultural_2013].
+Most relevant to community conservation is the development of norms which are beneficial for the group, but cause individuals to incur some personal cost; we refer to these as costly cooperative behaviors [@molleman_cultural_2013].
 Central to theories around the adoption of costly cooperative behaviors is the insight that selection acts on both individuals and groups.
@@
 The degree to which an individual depends on each of their nested groups is directly proportional to the personal costs they will incur for that group [@okasha_multilevel_2004; @price1972extension].
-To the extent that conservation is a collaborative dilemma, these insights from cultural group selection may be paramount to implementing effective community based conservation [@waring2015multilevel; @kline2018designing].
+To the extent that conservation is a collaborative dilemma, these insights from cultural group selection may be paramount to implementing effective community-based conservation [@waring2015multilevel; @kline2018designing].
 
@@
 The group level benefits of conservation are often much less certain than the immediate individual costs of forgoing resources.
-Humans tend to steeply discount the benefits of their actions as those benefits move further away in time and the benefits become more uncertain [@dasgupta2005uncertainty].
+Humans tend to steeply discount the benefits of their actions as those benefits move further away in time and the benefits become less certain [@dasgupta2005uncertainty].
 This hyperbolic discounting makes conservation norms very unlikely to emerge and persist under our current understanding of cultural group selection — the individual costs are high and group benefits are low.
@@
 
-Leakage [as defined by @andam2008measuring] is a form of land use spillover where an intervention, such as the CoFMAs established on Pemba by REDD+, has a negative, counteracting effect on the targeted outcome variable in another place.
+Leakage [as defined by @meyfroidt2020focus] is a form of land use spillover where an intervention, such as the CoFMAs established on Pemba by REDD+, has a negative, counteracting effect on the targeted outcome variable in another place.
 On Pemba, we have anecdotal testimonies that individuals in wards with CoFMAs have simply shifted their harvest pressure onto adjacent wards.
-While dubious for the efficacy of REDD+ writ large, on Pemba, leakage may actually be driving the spread of community based conservation observed in the ten new communities which elected to create CoFMAs post-REDD+ [@andrewsdoes].
+While dubious for the efficacy of REDD+ writ large, on Pemba, leakage may actually be driving the spread of community-based conservation observed in the ten new communities which elected to create CoFMAs post-REDD+ [@andrewsdoes].
 Just as the performance based payments from REDD+ were designed to alter the individual/community cost/benefit matrix so that selection favors costly collaboration via immediate community payments, we hypothesize that leakage from wards with CoFMAs may also act to favor costly cooperative behavior as individuals must act to defend their access to resources they would not have otherwise.
@@
 
-To better assess if individuals are adhering to CoFMA regulations, we need to sort out the originally demarcated use and conservation areas in each CoFMA.
+To better assess if individuals are adhering to CoFMA regulations, we need to map the originally demarcated use and conservation areas in each CoFMA.
 To our knowledge, these data do not currently exist or are not available.
@@
 
-# Chapter 3: Social Influences on Transmission Fidelity in Household Level Diffusion of Beneficial Technology.
+# Chapter 3: Social Influences on Transmission Fidelity in Household Level Diffusion of Beneficial Technology
 
@@
 Alternatively, some behaviors require only very low-fidelity replication to achieve identical outcomes [@enquist_one_2010].
-In fact, it is likely that if a behavior is under strong environmental constraints then individual learning should help correct errors in the transmission process.
+In fact, it is likely that if a behavior is under strong environmental constraints then individual learning should help correct errors in the transmission process; therefore, functional traits which directly affect an individual’s wellbeing should be more resilient to modification during transmission [@andersson2013fidelity].
 
@@
 Geographic distance will be measured according to the shortest walking distance between the two stoves.
-Communication frequency will be measured in terms of weekly communication and taken as a measure of social connectedness between the original stove builder and the copying individual.
+Communication frequency will be measured in terms of weekly communication of any type and taken as a measure of social connectedness between the original stove builder and the copying individual.
+As recall data tends to be unreliable, I will use current social connectedness as opposed to the time period the stove was built.
 I will also use this survey to collect data on the decision making processes individuals exhibit when deciding to build a fuel-efficient cookstove.
+Lastly, a variety of external factors may contribute to the outcome of cookstove construction.
+I will collect information to control for household wealth, access to materials, and intended cookstove use.
 
@@
 I will compare standardized parameter estimates, to determine which factors (geographic distance or social relatedness) most strongly contribute to copying fidelity in this study system.
-Similar analyses will be used to determine which factors, if any, are strongly associated with stove fuel efficiency.
+Similar analyses will be used to determine if these same factors, if any, are strongly associated with stove fuel efficiency.
 

### 2021-04-12 LeakageLitReview.Rmd
cleaning up

---
title: "Leakage Literature Review"
author: "Matt Clark"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
fontsize: 12pt
link-citations: yes
csl: DissertationProposal/AMJBot.csl
subtitle: 
bibliography: [Anthroposphere2.bib]
---

**Review Papers (no order)**

Overview of leakage. Calls for better monitoring of leakage and conservation outcomes both spatially and temporally [@ewers_estimates_2008].

Gives definition of leakage compared to other spillovers such as “indirect land use change.” General review of spillover processes [@meyfroidt_focus_2020].

Little is known about the predictability of leakage and blockage. Study looked at 3,398 protected areas globally to assess rates of both. 11.8% show leakage while 54.8% show blockage. GDP is the strongest predictor of whether or not blockage occurs [@fuller_first_2019].

Conservation via protected areas is increasing worldwide in terms of both the number and size of areas. Spillovers of all kinds are increasing with them [@pfaff_spillovers_2017]

REDD+ takes advantage of market globalization and the fact that land conservation is cheaper in some places compared to others to leverage the greatest conservation potential with the conservation funding available. Leakage plagues REDD+ projects globally and is therefore a major problem for global efforts to reduce greenhouse gasses [@atmadja_review_2012]

Leakage is shifting from an issue of local land use displacement to global displacement resulting from conservation interventions. Theory suggests that this might lead to spatial matching of land use and productivity potential. In practice we see high income to low income transitions where low income countries have worse agricultural technology and require more area to produce the same amount of resources, resulting in a net loss of forest cover [@meyfroidt_globalization_2013].

Leakage is commonly viewed as a failing of governance, but this is not necessarily true. In Vietnam, evidence suggests that deforestation leakage info adjacent Cambodia was planned. Leakage is a feature of policy/governance and often constitutes a choice rather than a failing [@bastos_lima_leakage_2019].

Protected areas, payments for ecosystem services, and technicological (mostly agricultural) advances are intended to mitigate habitat loss. Leakage and selection biases threaten the efficacy of protected areas and payments for ecosystem services. When multiple substitutable forest products exist such as various forms of timber or palm oil/ rapeseed, increased productivity (technological advance) in one may actually lead to a net increase in deforestation as demand for the cheaper product decreases demand for the more expensive product and that dedicated land may not be converted [@lim_perverse_2017].

**Case Studies (no order)**

An assessment of protected areas in East Africa’s evergreen forests showed that leakage increased commensurate with the efficacy of an area's protections within its boundaries. Social factors were also important in predicting deforestation rates [@pfeifer_protected_2012].

In the world’s tropical and subtropical forests, an analysis of 120 protected areas showed that 55 experienced leakage. In 78.2% of those 55 protected areas, the forest conserved by the protected area was not enough to offset the deforestation in the buffer region [@ford_deforestation_2020].

Authors present a method for better monitoring leakage through global trade [@godar_towards_2015].

The Soybean Moratorium and zero-deforestation beef agreements halved deforestation in the Brazilian Amazon, but resulted in a nearly seven times increase in deforestation in the neighboring biome from legally converted soybean fields and rangeland [@dou_spillover_2018].

Global analysis of leakage via global trade shows that approximately a third of forest gains in countries with net forest increases are offset by those countries increasing demand for forest commodities from other countries [@pendrill_deforestation_2019].

Land prices in Brazil predict future deforestation from leakage following land-use policies aimed at forest conservation. Conservation policies cause adjacent lands to become more valuable and more likely to get deforested [@miranda_land_2019]

From 2000-2016, Cambodia lost approximately 22% of all forest cover. Leakage resulting from conservation focused land concessions may have contributed to up to 10% of this loss [@magliocca_direct_2020].

Theoretical model of leakage processes shows that opportunistic conservation is more likely to result in leakage than when conservation is pursued strategically. Model is validated empirically with data from Kalimantan Indonesia (Borneo) [@renwick_reserves_2015].

From 2009-2016, researchers looked at the effects of oil palm certification programs in Indonesian Borneo (Kalimantan). Comparisons against counterfactuals show that certification programs increased forest cover compared to business as usual, but the outcomes were weakened by leakage [@heilmayr_deforestation_2020].

Vietnam has experienced net forest growth in recent years. Up to 40% of this forest regrowth however, is offset by leakage. Woodlots mitigated much of the potential leakage effects [@meyfroidt_forest_2009].

Parks created for eco-tourism in Costa Rica show mixed outcomes of leakage. Areas close to roads were much more likely to exhibit leakage than those which are distant from roads. Suggests preexisting deforestation pressure plays an important role [@robalino_heterogeneous_2017].

In a temporal spillover, land set aside as part of the Conservation Reserve Program in the United States was 25% more likely to be converted to agriculture after the program ended than were other comparable lands [@jacobson_temporal_2014].

Positive conservation effects resulting from zero-deforestation cattle agreements in the Brazilian Amazon are “washed out” by leakage [@alix-garcia_forest_2017].

Land conservation in the Peruvian Amazon is largely working, but is somewhat mitigated by leakage [@oliveira_land-use_2007].

Empirically grounded in the $6 billion spent on land acquisition by the Nature Conservancy on land acquisition in the last 50 years, authors create an economic model showing that buying up land for conservation can increase the value and deforestation pressure on adjacent lands [@armsworth_land_2006].

### 2021-04-12 MultipleReplicationsNoPlots.R

@@ -249,13 +249,13 @@ FillData%<>%mutate(PercProtected=round(PercProtected,digits=1))%>%mutate(Resourc
 
 
 for (i in 1:nrow(FillData)){
-  abmnp(Runs=3,Individuals = 100,TotalCarryingCapacity=10000,
+  abmnp(Runs=5,Individuals = 100,TotalCarryingCapacity=10000,
       StartPercCarryingCapacity = 1.00,
       PercProtected = FillData[i,]$PercProtected, 
       ProbOfMobility = FillData[i,]$ResourceMobility,CoopPercStart=1.0, 
       LearningStrategy = "Success Bias",
       TimeSteps=30,
-      ResourceRegenerationPerTimeStep=1.05,
+      ResourceRegenerationPerTimeStep=1.20,
       harvestMax=25)
   FillData[i,3]<-apply(FullOutput$meanPayoff[30,1:3],1,mean)  
   print(paste0(i/nrow(FillData)*100,"%"))
@@ -273,7 +273,7 @@ ggplot(data=FillData, mapping=aes(x=PercProtected, y=ResourceMobility, fill= mea
   scale_x_continuous(labels = scales::percent, expand = c(0, 0))+
   scale_y_continuous( labels = scales::percent,expand = c(0, 0))+
   labs(fill = "Average Harvest\nPayoff")+
-  xlab("Percent Protected")
+  xlab("Percent in Conservation Area")
 
 
 
@@ -309,7 +309,7 @@ ggplot(data=FillData2, mapping=aes(x=PercProtected, y=ResourceMobility, fill= me
   scale_x_continuous(labels = scales::percent, expand = c(0, 0))+
   scale_y_continuous( labels = scales::percent,expand = c(0, 0))+
   labs(fill = "Average Harvest\nPayoff")+
-  xlab("Percent Protected")
+  xlab("Percent in Conservation Area")
 
 
 FillData<-data.frame(PercProtected=seq(0.1,1,0.1),HarvestMax=seq(12,30,2))
@@ -344,7 +344,7 @@ ggplot(data=FillData, mapping=aes(x=PercProtected, y=HarvestMax, fill= meanPayof
    scale_x_continuous(labels = scales::percent, expand = c(0, 0))+
   scale_y_continuous( breaks=c(12,16,20,24,28),expand = c(0, 0))+
   labs(fill = "Average Harvest\nPayoff")+
-  xlab("Percent Protected")
[…]

### 2021-10-06 DOI_Review/Manuscript.rmd
Lots of things mostly manuscript though

---
title: Utility of Diffusion Models for Predicting and Diagnosing the Spread of Conservation Behaviors
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: Boise State University- Human Environment Systems
  name: Vicken Hillis
- affiliation: Max Planck Institute for Evolutionary Anthropology
  name: Jeffrey Andrews

date: "`r format(Sys.time(), '%B %d, %Y')`"
keywords: Community-based conservation, diffusion of innovations, cultural evolution, complex systems
bibliography: [./DOI_Conservation_Review.bib]
abstract: Conservation researchers are increasingly identifying diffusion theory as useful for understanding the spread of community-based conservation initiatives. This work has nevertheless been largely ambiguous as to the specific opportunities and shortcomings for diffusion theory in conservation research, and how it might be practically applied. Here, we employ agent-based simulations to confirm that process-based diffusion models reasonably predict the long-term patterns in adoption of conservation behaviors even under tightly coupled behavior-environment feedbacks. We caution however, that while diffusion theory may predict the pattern with which a conservation intervention will spread through a population, it gives little insight as to the underlying transmission mechanism. We show that the transmission mechanism employed by a population can alter the final proportion of individuals that adopt an intervention, and argue that conservation researchers and practitioners can leverage theory from social learning and cultural evolution to better design and diagnose projects. Lastly, we emphasize that developing and fitting process models in statistical frameworks traditionally reserved for regression alone, may pose a productive new avenue for conservation science, particularly in our understanding of conservation as a complex system.

fontsize: 12pt
link-citations: yes
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
header-includes:
- \usepackage{pdfpages}
- \usepackage[font={footnotesize}]{caption}
- \usepackage{bm}
- \usepackage{amsmath}

---

```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)
```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)

## Caching code
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')

setwd("C:/Users/Matthewclark989/Documents/Pemba_Project/DOI_Review")
```

> "Diffusion study...is something like the use of radioactive tracers in studying the process of plant   growth: it helps illuminate the process."
>
> `r tufte::quote_footer('--- Everett M. Rogers, Diffusion of Innovations, 2003')`

It is becoming increasingly apparent that humanity will not reach the global conservation goals we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002]. The sweeping behavioral changes required to meet these targets will thus have to originate from and be maintained by the communities living in and around threatened natural areas themselves [@amel_beyond_2017; @catalano_learning_2019]. Community-based conservation projects aim to do just this by incentivizing rural, and often very poor, communities to implement these types of changes with little to no external oversight [@western_natural_1994]. While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote long-term behavioral changes which spread from person to person and even community to community [@brooks_assessing_2013; @horwich_community_2007].

While some debate still exists and exact methods vary, it is largely accepted that community-based conservation interventions are a viable method for seeding self-spreading, -sustaining, and ethical protections on natural resources [TURNER 2017; BEREKES 2007; ADAMS2004]. A major question in implementing these projects is if a specific intervention will indeed catch-on in a population and if so, how quickly [Radeloff 2013]? Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [Mangubhai 2011; Visconti 2019 ].

# Diffusion Theory
Inducing and/or speeding up the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time. Diffusion theory recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [ROGERS 2010]. When an individual is introduced to a new behavior, they often have a high degree of uncertainty as to the risks and benefits of adopting and will wait to adopt until that uncertainty is reduced to a tolerable level. The primary way individuals reduce uncertainty about the outcome of adopting a particular behavior is through second-hand observation of early adopters. Therefore, the core of diffusion theory is the notion that individuals imitate those around them when making decisions about adopting a novel behavior or not, rather than weighing the pros and cons independently.

This assumption about human decision-making has been validated repeatedly in cultures around the world; individuals do not act as cost-benefit analysts, in which they adapt their behaviors based on the environment alone [HENRICH et al 2001]. Our dependence on social learning to make decisions about behavioral adoption has profound and specific implications for patterns of behavioral adoption at the population level [SMALDINO 2018; another cite??]. We observe these patterns through diffusion curves which show the proportion of a population that has adopted a behavior as a function of time [HOPPITT 2013; ROGERS 2010; RYAN & GROSS 1950].

Behavioral spread through imitation produces a sigmoidal “S” shaped diffusion curve, which is observed in the vast majority of empirical studies on behavioral adoption [HENRICH 2001]. This “S” shaped curve is produced regardless of the social learning mechanism employed by a population and can even be observed when a portion of the population relies on individual learning [HENRICH 2001]. Behavioral adoption thus exhibits equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [BARRETT 2019].

# Diffusion Theory in Conservation

Individuals tend to place more weight on imitation when making decisions in which the outcomes can not be well predicted independently [PERREAULT 2012]. This phenomenon is particularly salient under variable, rapidly changing environments [BOYD & RICHERSON 1988; Richerson & Boyd 2008]. It follows then, that the adoption of novel conservation behaviors by a community is likely to be achieved through social imitation (i.e. diffusion), especially given worsening climatic instability and regime changes resulting from widespread biodiversity loss.

Indeed, the sigmoidal pattern of behavioral adoption characteristic to diffusion has been observed in over 80% of case studies on community-based conservation interventions worldwide [MASCIA & MILLS 2018; MILLS et al 2019]. Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the adoption of community based conservation interventions is considerably attenuated [PIERTI et al 2009]. With the recognition that the patterns of uptake for community-based conservation practices are directly linked to behavioral imitation, researchers have begun suggesting diffusion theory as a lens for understanding and implementing individual interventions [ABERNETHY et al 2014; MBARU & BARNES 2017; NIEMIC et al 2021].

Meta-level studies have recently come to the same conclusion. For example, in a review of 22 conservation interventions referenced above, the majority of  initiatives were adopted by less than 30% of potential adopters, leading the authors to point to diffusion theory as an avenue for increasing the proportion of adopters [MILLS et al 2019]. And a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices further identified diffusion theory as useful for guiding the development of conservation interventions [MAHAJAN et al 2020].

These studies are correct in their conclusion that diffusion theory can be productively applied to the spread of community-based conservation interventions, but ambiguous in exactly how a practitioner or researcher might practically apply this insight. The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors, and what the limitations of this theory are for complex adaptive systems like community-based conservation.

# Quantitative applications

The sigmoidal pattern of behavioral spread over time mirrors that of disease transmission [SMALDINO & JONES 2020]. This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral uptake [CAVALLI-SFORZA and FELDMAN 1981; STRANG 1991]. Specifically, the canonical susceptible-infectious-recovered (SIR) model and SIRS variation where individuals are added back to the pool of susceptibles after recovery (eq 1), have been demonstrated to be well suited to fitting the diffusion process [FARZANEH et al 2019;SMALDINO & JONES 2020].

\begin{align*}
&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} \\
&\frac{dR}{dt} = \gamma* \text{I}_{t} - \chi * \text{R}_{t}
\end{align*}

When the natural history characteristics of some pathogen is known, the SI, SIR, SIS, &  SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [HUPPERT & KATRIEL 2013]. For conservation interventions however, there are no natural history features of a specific intervention, such as the rate or duration of adoption, which can be known ahead of time. We can however, use Bayesian inference to probabilistically estimate these parameters from data on behavioral uptake over time [Grinsztajn 2020; CARPENTER 2018]. Below, we use a minimal agent-based model of conservation adoption contingent on behavior-environment feedbacks to simulate data on the long-term adoption dynamics of a community-based conservation intervention. Then, using just the first 30% of the data, we estimate the rate of intervention spread, rate of dropout, and time after dropout before agents are available to adopt again. We use these estimated parameters to predict the overall patterns of adoption over all time steps and compare our observed to predicted values.

# Agent-based simulation
We developed a minimal agent-based simulation which we use to generate data on conservation adoption within a population. We use these data to test the performance of the SIRS diffusion model in predicting patterns of conservation behavior adoption under both unbiased and success-biased social learning. Success-biased social learning is contingent on behavior-environment feedbacks. Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [FERRARO 2011]. We establish a degraded landscape/seascape (from herein referred to as landscape) where all simulated individuals are dependent on harvesting some renewable, mobile resource. In the first timestep, we introduce a conservation intervention which delineates 20% of the landscape as a conservation area and individuals are rewarded for abstaining from harvesting there. In each subsequent timestep, agents either choose to enroll in the conservation program, receive their reward, and continue to harvest just from the remaining 80% of the unprotected landscape, or they choose to harvest from the entire landscape and forgo the reward. After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior [BALDINI 2013, BARRETT 2019]. The protected and unprotected resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.

![Conceptual diagram showing process at each timestep in the agnt-based simulation. This simulation was used to produce data on the adoption of conservation behaviors as a function of social learning and environmental outcomes.](./ConceptDiagram.png)

A visual description of this minimal model is shown in figure 1, and the full ODD+ model description can be found in the supplemental material (S1). Interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at this URL [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/). As this simulation is not the emphasis of this study, we show just two runs below, one in which. The starting conditions of these runs are identical except for the social learning mechanism employed by the population. As the unbiased learning mechanism is independent of the resource levels, the adoption curve follows the traditional “S” pattern and nears fixation in the long-term. Under environmentally mediated success-biased learning, the adoption pattern spikes initially, but drops off dramatically as the resources in the unprotected landscape are reduced following increased harvest pressure there as a result of the conservation intervention. The full output of this later simulation is shown below in figure 2. Additional simulation runs can be found in the supplemental material (S2) or by using the interactive web application linked above.

![One run of the agent based simulation used to produce data for analysis with the process -based SIRS model of behavioral diffusion. Starting conditions for this simulation were: 100 individuals, total landscape, carrying capacity of 10000 resources, 20% starting resource level, individual payment amount of 9 resources for enrollment in the conservation intervention, 2% initial enrollment, success-biased learning strategy at a strength of 1.1, 30% probability of resource mobility when carrying capacity reached, a maximum harvest value of 15, and 30% resource regeneration rate per time step.](./SiccessBiasedFullRun.png)

# Predicting adoption of conservation behaviors

With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described above to the first 30% of observations. We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time:

$$ \text{Y}_{t} \sim Binomial(\text{N}, \text{I}_{t}) $$
\begin{align*}
&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} \\
&\frac{dR}{dt} = \gamma* \text{I}_{t} - \chi * \text{R}_{t}
\end{align*}

$$\beta,\gamma,\chi \sim Normal(0,0.5)$$

For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an rhat of 1 for all estimated parameters (fig 3). As the parameters $\beta$, $\gamma$, and $\chi$ are directly interpretable as the rates of adoption, dropout, and willingness to retry a previously abandoned intervention, accurate estimation of these values alone, especially in the early stages of an intervention, is a practical and straightforward application of diffusion theory to community-based conservation interventions.

![Markov chains for the three parameters of interest from the SIRS model fit to simulated data. The agent-based simulation used to produce the data employed environmentally mediated success-biased learning. A figure showing the Markov chains for the unbiased learning simulation can be found in the supplemental material (S3).](./SuccessPars.png)

We can also use these parameter values, estimated from the first 30% of the time series, to predict the long-term patterns of adoption for community-based conservation initiatives. In figures 4 & 5, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step. Grey and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data respectively. These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning. R and STAN code for these models can be found in the supplemental material (S4)

![SIRS model fit of agent-based simulation data under unbiased social learning shown as translucent black lines representing 1000 draws from the posterior. Simulated data are overlaid with grey points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./DirectCopying.png)

![SIRS model fit of agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior. Simulated data are overlaid with grey points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./SuccessBiasedCopying.png)

Figures 4 & 5 also demonstrate that different learning mechanisms, while both producing a sigmoidal adoption pattern, can result in different overall rates in adoption of an intervention, even under identical starting conditions. In the simulation of a population employing unbiased learning, the population fixates at nearly 90% adoption of the intervention behavior. For success-biased learning on the other hand, resource scarcity in the unprotected area caused the population to adopt the intervention at approximately 60% in the long-term.

# Shortcomings for application of diffusion theory to community-based conservation

As demonstrated above, the standard SIRS diffusion model can reliably capture the long-term patterns of behavioral adoption for community-based conservation interventions, even when adoption is contingent on feedbacks with ecological processes. This model, while accurately capturing the pattern of spread, is unable to provide insight as to the mechanism of spread or causality in the system. As shown above and in existing literature, the social transmission mechanism, or mix of mechanisms, employed by a population can dramatically affect the overall rates of adoption for a given behavior [SMALDINO et al 2017; JEFF&MONIQUE2018; WARING2021]. Therefore, while we agree that diffusion theory can be productively applied to community-based conservation, we point to theories of social learning and cultural evolution as a better alternative for diagnosing and planning interventions.

Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [CHENG et al 2020]. Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes [TURNER et al 2021]. Specifically, in modeling social-ecological systems, either using analytical approaches or simulation, processes should be based on realistic human cognitions which allow actors to modify their behaviors in response to new information [MEYFROIDT 2013]. Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [SCHILL et al 2019].

We argue that when prediction is the goal, analytical or statistical approaches like the SIRS model presented above are always preferable to simulation based approaches [EPSTEIN 2008; TREDENNINCK et al 2021]. The practice of applying statistical frameworks of parameter estimation to analytical process models, and in particular to ordinary differential equations, is still relatively new and has been embraced mostly by ecology [HEFLEY 2017]. Extending the SIRS diffusion model to include a term which represents changes in the environment over time and the effect of resource levels on adoption rates may more accurately describe the system of adoption in community-based conservation interventions. An example of such an extension is presented below:

\begin{align*}
&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} - \phi * \text{E}_{t} \\
&\frac{dR}{dt} = \gamma * \text{I}_{t} - \chi * \text{R}_{t} + \phi * \text{E}_{t} \\
&\frac{dE}{dt} =  - \rho * (\text{S}_{t} + \text{R}_{t}) + \theta*\text{E}_{t}
\end{align*}

While process models may be preferable for predicting adoption outcomes of conservation interventions, they might not always be possible. In complex systems, these models can quickly require many parameters in order to accurately account for all moving parts and thus are prone to issues of identifiability in parameter estimation  [LAUBMEIER 2020]. Further, complex social-ecological systems have properties which are inherently stochastic and may be impossible to model analytically. For example, coastal fisheries have been shown to approach mathematical chaos in the number and proportions of species present [WILSON et al 1994].

In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for tipping points and emergence, two key features of the complex systems which can not be accounted for with process models such the SIRS or SIRS extension models presented above [REYERS et al 2018; RAILSBACK & GRIMM; WILENSKY & RAND]. Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically  [LIU et al 2007].

# Conclusions

Overall, we argue that fitting process models of behavioral diffusion to data on the spread of conservation behaviors is a practical way for researchers and practitioners to apply diffusion theory to community-based conservation. We show that these models can be powerful for estimating the natural history characteristics of a particular intervention, such as the rate of spread, and for predicting the long-term patterns of adoption. We also introduce cultural evolutionary research as a better set of theories for diagnosing the spread, or lack thereof, in such interventions and for planning future interventions. Lastly, we describe conservation as a complex adaptive system and propose simulation modeling as an alternative to process-based models when they may be analytically intractable or ill suited for system dynamics.

\newpage

# References

### 2021-10-06 DOI_Review/ApplySTANtoABMdat.R

@@ -151,21 +151,22 @@ params_monitor = c("y_hat", "y0", "params", "fake_I") #fake_I from generated qua
 mod = stan("~/Pemba_Project/DOI_Review/SIRS.stan",
            data = stan_d,
            pars = params_monitor,
-           chains = 1,
+           chains = 4,
            warmup = 500,
            iter = 1500)
 
 
 # You should do some MCMC diagnostics, including:
 traceplot(mod, pars="lp__")
+names(mod)[94:96]<-c("Beta","Gamma","Chi")
 
-traceplot(mod, pars=c("params", "y0"))
+traceplot(mod, pars=c("params"))+mytheme+theme(strip.text.x = element_text(size = 15))
 #summary(mod)$summary[,"Rhat"]
 
 # These all check out for my model, so I'll move on.
 
 # Extract the posterior samples to a structured list:
-posts <- extract(mod)
+posts <- rstan::extract(mod)
 #hist(posts$params[,1])
 #hist(posts$params[,2])
 
@@ -208,9 +209,11 @@ mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_b
                 legend.text = element_text( size = 14))
 
 
+#cor(mod_median[-(1:sample_days)],(SucBiasdat[(sample_days+1):sample_n,]$Enrolled)/sample_n)
+
 ggplot(df_sample, aes(x=sample_time, y=sample_prop)) +
   #samples
-  geom_line(data=draws,mapping = aes(x = as.integer(mod_time), y=value, group = draw), alpha = 0.2, size=0.1) +
+  geom_line(data=draws,mapping = aes(x = as.integer(mod_time), y=value, group = draw), alpha = 0.05, size=0.1) +
 
   #points
   geom_point(aes(fill="Training"),col="black" ,shape = 21, size = 5,stroke=2) +

### 2021-10-06 DOI_Review/Binary_Adoption_ABM.R

@@ -173,7 +173,7 @@ abm<-function(#Specified parameters
   if(LearningStrategy=="Conformist Bias"){
     
     spreadProb<-0.25 #0.25
-    recProb<-0.1 #0.01
+    recProb<-0.03 #0.01
     
     NewEnroll<-rep(NA,Individuals)
   
@@ -287,7 +287,7 @@ abm<-function(#Specified parameters
     theme_classic()+ylab("Carrying Capacity")+xlab("")+
     scale_y_continuous(labels = scales::percent,limits = c(0,1.0) )+
     scale_colour_manual(name = '', 
-                        values =c('Protected'='#b2df8a','Working'='#1f78b4'), labels = c('Protected Area','Working Area'))+
+                        values =c('Protected'='#b2df8a','Working'='#1f78b4'), labels = c('Conservation Area','Unprotected Area'))+
     ggtitle("Resource Integrity")+mytheme +
     theme(legend.position="bottom")
   
@@ -311,74 +311,47 @@ abm<-function(#Specified parameters
 
 set.seed(1) 
 
-deqdatPOS_Conform<-abm(#Specified parameters
-  Individuals=100, #number of total resource users in a population
-  TotalCarryingCapacity=100000, #total available resource units
-  StartPercCarryingCapacity = 0.25, #amount of resources available in the landscape at the start in proportion to CC
-  PercProtected=0.2, #percent of the total resource that's in a protected area
-  EnrollPercStart=0.01, #percent of individuals who start by following the rules at t0
-  LearningStrategy = "Conformist Bias", #options are Success Bias & Conformist...not implementing this...for now
-  BiasStrength = 1.05,
-  TimeSteps=100,
-  ResourceRegenerationPerTimeStep=1.15,
-  harvestMax=35,
-  ProbOfMobility=0.9)
+#deqdatPOS_Conform<-abm(#Specified parameters
+ # Individuals=100, #number of total resource users in a population
+#  TotalCarryingCapacity=100000, #total available resource units
+ # StartPercCarryingCapacity = 0.25, #amount of resources available in the landscape at the start in proportion to CC
+  #PercProtected=0.2, #percent of the total resource that's in a protected area
+  #EnrollPercStart=0.01, #percent of individuals who start by following the rules at t0
+  #LearningStrategy = "Conformist Bias", #options are Success Bias & Conformist...not implementing this...for now
[…]

### 2021-10-06 DOI_Review/DOI_ReviewApp.R

@@ -1,0 +1,433 @@
+library(shiny)
+library(ggplot2)
+library(shinythemes)
+library(DT)
+
+# Define UI for application that draws a histogram
+
+ui<-shinyUI(navbarPage("Protected Areas ABM",
+                       theme = shinytheme("flatly"),
+                       tabPanel("AbmOutput",
+                                # plot the map
+                                fluidRow(column(12,
+                                                h1("AbmOutput"),
+                                                p("Instructions Go Here"),
+                                                br(),
+                                                strong("Important information goes here"))),
+                                
+                                hr(),
+                                fluidRow(sidebarPanel(width = 3,
+                                                      
+                                                      helpText("Help Text"),
+                                                      sliderInput("individuals",
+                                                                  "Number of individuals:",
+                                                                  min = 10,
+                                                                  max = 1000,
+                                                                  value = 100),
+                                                      selectInput("LearningStrategy","Learning Strategy",
+                                                                  c("Success Bias" = "Success Bias",
+                                                                    "Conformist Bias"= "Conformist Bias")),
+                                                      sliderInput("totCareCapac","Total Carrying Capacity Landscape",
+                                                                  min = 100,max=100000,value=10000),
+                                                      sliderInput("StartPercCarryingCapacity","Starting Percent Carrying Capacity",
+                                                                  min=0.01,max=1.0,value=0.2),
+                                                      sliderInput("timeSteps",
+                                                                  "Time Steps",
+                                                                  min = 10,
+                                                                  max = 1000,
+                                                                  value = 100),
+                                                      sliderInput("percProtect",
+                                                                  "Percent of Resources Protected",
+                                                                  min = 0.00,
[…]

### 2021-10-06 DOI_Review/Manuscript.rmd
refs ~done~

@@
 
-While some debate still exists and exact methods vary, it is largely accepted that community-based conservation interventions are a viable method for seeding self-spreading, -sustaining, and ethical protections on natural resources [TURNER 2017; BEREKES 2007; ADAMS2004].
-A major question in implementing these projects is if a specific intervention will indeed catch-on in a population and if so, how quickly [Radeloff 2013]?
-Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [Mangubhai 2011; Visconti 2019 ].
+While some debate still exists and exact methods vary, it is largely accepted that community-based conservation interventions are a viable method for seeding self-spreading, -sustaining, and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004].
+A major question in implementing these projects is if a specific intervention will indeed catch-on in a population and if so, how quickly [@radeloff_hot_2013]?
+Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [@mangubhai_not_2011; @visconti_protected_2019].
 
@@
 Inducing and/or speeding up the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time.
-Diffusion theory recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [ROGERS 2010].
+Diffusion theory recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010].
 When an individual is introduced to a new behavior, they often have a high degree of uncertainty as to the risks and benefits of adopting and will wait to adopt until that uncertainty is reduced to a tolerable level.
@@
 
-This assumption about human decision-making has been validated repeatedly in cultures around the world; individuals do not act as cost-benefit analysts, in which they adapt their behaviors based on the environment alone [HENRICH et al 2001].
-Our dependence on social learning to make decisions about behavioral adoption has profound and specific implications for patterns of behavioral adoption at the population level [SMALDINO 2018; another cite??].
-We observe these patterns through diffusion curves which show the proportion of a population that has adopted a behavior as a function of time [HOPPITT 2013; ROGERS 2010; RYAN & GROSS 1950].
+This assumption about human decision-making has been validated repeatedly in cultures around the world; individuals do not act as cost-benefit analysts, in which they adapt their behaviors based on the environment alone [@henrich_search_2001].
+Our dependence on social learning to make decisions about behavioral adoption has profound and specific implications for patterns of behavioral adoption at the population level [@smaldino_adoption_2017].
+We observe these patterns through diffusion curves which show the proportion of a population that has adopted a behavior as a function of time [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
 
-Behavioral spread through imitation produces a sigmoidal “S” shaped diffusion curve, which is observed in the vast majority of empirical studies on behavioral adoption [HENRICH 2001].
-This “S” shaped curve is produced regardless of the social learning mechanism employed by a population and can even be observed when a portion of the population relies on individual learning [HENRICH 2001].
-Behavioral adoption thus exhibits equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [BARRETT 2019].
+Behavioral spread through imitation produces a sigmoidal “S” shaped diffusion curve, which is observed in the vast majority of empirical studies on behavioral adoption [@henrich_cultural_2001].
+This “S” shaped curve is produced regardless of the social learning mechanism employed by a population and can even be observed when a portion of the population relies on individual learning [@henrich_cultural_2001].
+Behavioral adoption thus exhibits equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [@barrett_equifinality_2019].
 
@@
 
-Individuals tend to place more weight on imitation when making decisions in which the outcomes can not be well predicted independently [PERREAULT 2012].
-This phenomenon is particularly salient under variable, rapidly changing environments [BOYD & RICHERSON 1988; Richerson & Boyd 2008].
+Individuals tend to place more weight on imitation when making decisions in which the outcomes can not be well predicted independently [@perreault_bayesian_2012].
+This phenomenon is particularly salient under variable, rapidly changing environments [@boyd_culture_1988; @richerson_not_2008].
 It follows then, that the adoption of novel conservation behaviors by a community is likely to be achieved through social imitation (i.e.
@@
 
-Indeed, the sigmoidal pattern of behavioral adoption characteristic to diffusion has been observed in over 80% of case studies on community-based conservation interventions worldwide [MASCIA & MILLS 2018; MILLS et al 2019].
-Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the adoption of community based conservation interventions is considerably attenuated [PIERTI et al 2009].
-With the recognition that the patterns of uptake for community-based conservation practices are directly linked to behavioral imitation, researchers have begun suggesting diffusion theory as a lens for understanding and implementing individual interventions [ABERNETHY et al 2014; MBARU & BARNES 2017; NIEMIC et al 2021].
+Indeed, the sigmoidal pattern of behavioral adoption characteristic to diffusion has been observed in over 80% of case studies on community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
+Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the adoption of community based conservation interventions is considerably attenuated [@pietri_information_2009-1].
+With the recognition that the patterns of uptake for community-based conservation practices are directly linked to behavioral imitation, researchers have begun suggesting diffusion theory as a lens for understanding and implementing individual interventions [@abernethy_two_2014; @mbaru_key_2017; @niemiec_efficacy-based_2021].
 
 Meta-level studies have recently come to the same conclusion.
-For example, in a review of 22 conservation interventions referenced above, the majority of  initiatives were adopted by less than 30% of potential adopters, leading the authors to point to diffusion theory as an avenue for increasing the proportion of adopters [MILLS et al 2019].
-And a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices further identified diffusion theory as useful for guiding the development of conservation interventions [MAHAJAN et al 2020].
+For example, in a review of 22 conservation interventions referenced above, the majority of  initiatives were adopted by less than 30% of potential adopters, leading the authors to point to diffusion theory as an avenue for increasing the proportion of adopters [@mills_how_2019].
+And a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices further identified diffusion theory as useful for guiding the development of conservation interventions [@mahajan_theorybased_2020].
 
@@
 
-The sigmoidal pattern of behavioral spread over time mirrors that of disease transmission [SMALDINO & JONES 2020].
-This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral uptake [CAVALLI-SFORZA and FELDMAN 1981; STRANG 1991].
-Specifically, the canonical susceptible-infectious-recovered (SIR) model and SIRS variation where individuals are added back to the pool of susceptibles after recovery (eq 1), have been demonstrated to be well suited to fitting the diffusion process [FARZANEH et al 2019;SMALDINO & JONES 2020].
+The sigmoidal pattern of behavioral spread over time mirrors that of disease transmission [@smaldino_coupled_2020].
+This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral uptake [@cavalli-sforza_cultural_1981; @strang_adding_1991].
+Specifically, the canonical susceptible-infectious-recovered (SIR) model and SIRS variation where individuals are added back to the pool of susceptibles after recovery (eq 1), have been demonstrated to be well suited to fitting the diffusion process [@famil-rohani_development_2019; @smaldino_coupled_2020].
+In this instantiation, $\beta$ represents the rate of infection in the population, $\gamma$ represents the rate of recovery, and $\chi$ is the rate that recovered individuals are added back into the pool of susceptibles.
 
@@
 
-When the natural history characteristics of some pathogen is known, the SI, SIR, SIS, &  SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [HUPPERT & KATRIEL 2013].
+When the natural history characteristics of some pathogen is known, the SI, SIR, SIS, &  SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [@huppert_mathematical_2013].
 For conservation interventions however, there are no natural history features of a specific intervention, such as the rate or duration of adoption, which can be known ahead of time.
-We can however, use Bayesian inference to probabilistically estimate these parameters from data on behavioral uptake over time [Grinsztajn 2020; CARPENTER 2018].
+We can however, use Bayesian inference to probabilistically estimate these parameters from data on behavioral uptake over time [@margossian_review_2019; @margossian_differential_2017;  @carpenter_predator-prey_2018].
 Below, we use a minimal agent-based model of conservation adoption contingent on behavior-environment feedbacks to simulate data on the long-term adoption dynamics of a community-based conservation intervention.
@@
 Success-biased social learning is contingent on behavior-environment feedbacks.
-Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [FERRARO 2011].
+Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [@ferraro_future_2011].
 We establish a degraded landscape/seascape (from herein referred to as landscape) where all simulated individuals are dependent on harvesting some renewable, mobile resource.
@@
 In each subsequent timestep, agents either choose to enroll in the conservation program, receive their reward, and continue to harvest just from the remaining 80% of the unprotected landscape, or they choose to harvest from the entire landscape and forgo the reward.
-After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior [BALDINI 2013, BARRETT 2019].
+After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior [@baldini_success-biased_2012; @baldini_two_2013, @barrett_equifinality_2019].
 The protected and unprotected resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
@@
 With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described above to the first 30% of observations.
-We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time:
+We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time.
+Here, "infected" individuals are those that are enrolled in the conservation intervention.
 
@@
 This model, while accurately capturing the pattern of spread, is unable to provide insight as to the mechanism of spread or causality in the system.
-As shown above and in existing literature, the social transmission mechanism, or mix of mechanisms, employed by a population can dramatically affect the overall rates of adoption for a given behavior [SMALDINO et al 2017; JEFF&MONIQUE2018; WARING2021].
+As shown above and in existing literature, the social transmission mechanism, or mix of mechanisms, employed by a population can dramatically affect the overall rates of adoption for a given behavior [@smaldino_adoption_2017; @andrews_cultural_2018; @tam_measuring_2021].
 Therefore, while we agree that diffusion theory can be productively applied to community-based conservation, we point to theories of social learning and cultural evolution as a better alternative for diagnosing and planning interventions.
 
-Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [CHENG et al 2020].
-Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes [TURNER et al 2021].
-Specifically, in modeling social-ecological systems, either using analytical approaches or simulation, processes should be based on realistic human cognitions which allow actors to modify their behaviors in response to new information [MEYFROIDT 2013].
-Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [SCHILL et al 2019].
+Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [@cheng_strengthen_2020].
+Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes [@turner_land-useland-cover_2021].
+Specifically, in modeling social-ecological systems, either using analytical approaches or simulation, processes should be based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013].
+Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [@schill_more_2019].
 
-We argue that when prediction is the goal, analytical or statistical approaches like the SIRS model presented above are always preferable to simulation based approaches [EPSTEIN 2008; TREDENNINCK et al 2021].
-The practice of applying statistical frameworks of parameter estimation to analytical process models, and in particular to ordinary differential equations, is still relatively new and has been embraced mostly by ecology [HEFLEY 2017].
+We argue that when prediction is the goal, analytical or statistical approaches like the SIRS model presented above are always preferable to simulation based approaches [@epstein_why_2008; @tredennick_practical_2021].
+The practice of applying statistical frameworks of parameter estimation to analytical process models, and in particular to ordinary differential equations, is still relatively new and has been embraced mostly by ecology [@hefley_when_2017].
 Extending the SIRS diffusion model to include a term which represents changes in the environment over time and the effect of resource levels on adoption rates may more accurately describe the system of adoption in community-based conservation interventions.
-An example of such an extension is presented below:
+An example of such an extension is presented below where we include another equation for the change in a resource pool over time $\frac{dE}{dt}$.
+We also must include three additional parameters: $\phi$.
+$\rho$, and $\theta$, which represent the effect of resource level on enrollment (infection), the effect of harvest on the resource level, and the regeneration rate of the resource respectively.
 
@@
 While process models may be preferable for predicting adoption outcomes of conservation interventions, they might not always be possible.
-In complex systems, these models can quickly require many parameters in order to accurately account for all moving parts and thus are prone to issues of identifiability in parameter estimation  [LAUBMEIER 2020].
+In complex systems, these models can quickly require many parameters in order to accurately account for all moving parts and thus are prone to issues of identifiability in parameter estimation [@laubmeier_ecological_2020].
 Further, complex social-ecological systems have properties which are inherently stochastic and may be impossible to model analytically.
-For example, coastal fisheries have been shown to approach mathematical chaos in the number and proportions of species present [WILSON et al 1994].
+For example, coastal fisheries have been shown to approach mathematical chaos in the number and proportions of species present [@wilson_chaos_1994].
 
-In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for tipping points and emergence, two key features of the complex systems which can not be accounted for with process models such the SIRS or SIRS extension models presented above [REYERS et al 2018; RAILSBACK & GRIMM; WILENSKY & RAND].
-Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically  [LIU et al 2007].
+In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for tipping points and emergence, two key features of the complex systems which can not be accounted for with process models such the SIRS or SIRS extension models presented above [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
+Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically  [@liu_complexity_2007].
 

### 2021-10-08 DOI_Review/Manuscript.rmd
Text edits mostly

@@
 bibliography: [./DOI_Conservation_Review.bib]
-abstract: Conservation researchers are increasingly identifying diffusion theory as useful for understanding the spread of community-based conservation initiatives.
-This work has nevertheless been largely ambiguous as to the specific opportunities and shortcomings for diffusion theory in conservation research, and how it might be practically applied.
+abstract: A common goal of conservation researchers and practitioners is to design interventions that benefit both human and ecological well-being and as a result are adopted robustly by local communities.
+Diffusion theory provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding the spread of these community-based conservation initiatives.
+This work has nevertheless been largely ambiguous as to the practical applications of diffusion theory to community-based conservation and has not addressed the potential shortcomings for its application to complex, coupled human-natural systems.
 Here, we employ agent-based simulations to confirm that process-based diffusion models reasonably predict the long-term patterns in adoption of conservation behaviors even under tightly coupled behavior-environment feedbacks.
@@
 We show that the transmission mechanism employed by a population can alter the final proportion of individuals that adopt an intervention, and argue that conservation researchers and practitioners can leverage theory from social learning and cultural evolution to better design and diagnose projects.
-Lastly, we emphasize that developing and fitting process models in statistical frameworks traditionally reserved for regression alone, may pose a productive new avenue for conservation science, particularly in our understanding of conservation as a complex system.
+Lastly, we emphasize that developing and fitting process models in statistical frameworks traditionally reserved for regression alone, may pose a productive new avenue for conservation science, particularly in concert with simulation techniques and insights from complex systems.
 
@@
 Specifically, the canonical susceptible-infectious-recovered (SIR) model and SIRS variation where individuals are added back to the pool of susceptibles after recovery (eq 1), have been demonstrated to be well suited to fitting the diffusion process [@famil-rohani_development_2019; @smaldino_coupled_2020].
+These compartmental differential equations assign every member of a population to a category (S, I, or R), depending on the proportion of individuals in the other categories and parameter values for the natural history characteristics of a given disease [@tolles_modeling_2020].
 In this instantiation, $\beta$ represents the rate of infection in the population, $\gamma$ represents the rate of recovery, and $\chi$ is the rate that recovered individuals are added back into the pool of susceptibles.
@@
 \begin{align*}
-&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
-&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} \\
-&\frac{dR}{dt} = \gamma* \text{I}_{t} - \chi * \text{R}_{t}
+&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \\
+&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
+&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t}
 \end{align*}
@@
 We can however, use Bayesian inference to probabilistically estimate these parameters from data on behavioral uptake over time [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018].
-Below, we use a minimal agent-based model of conservation adoption contingent on behavior-environment feedbacks to simulate data on the long-term adoption dynamics of a community-based conservation intervention.
-Then, using just the first 30% of the data, we estimate the rate of intervention spread, rate of dropout, and time after dropout before agents are available to adopt again.
+
+Below, we develop a minimal agent-based model to highlight our core dynamics of interest.
+In this abstraction, we simulate conservation adoption contingent on behavior-environment feedbacks to produce data on the long-term adoption dynamics of a community-based conservation intervention.
+Then, using just the first 30% of the data, we estimate the rate of intervention spread, rate of dropout, and readoption.
 We use these estimated parameters to predict the overall patterns of adoption over all time steps and compare our observed to predicted values.
@@
 We developed a minimal agent-based simulation which we use to generate data on conservation adoption within a population.
-We use these data to test the performance of the SIRS diffusion model in predicting patterns of conservation behavior adoption under both unbiased and success-biased social learning.
-Success-biased social learning is contingent on behavior-environment feedbacks.
+We use these data to test the performance of the SIRS diffusion model in predicting patterns of conservation behavior adoption under two social learning mechanisms.
+The first is unbiased copying, where agents have the chance to copy an alternative behavior with some set probability independent of the benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
+The second mechanism is success-biased learning, under which, agents disproportionately copy the behaviors of more successful agents [@baldini_success-biased_2012].
 Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [@ferraro_future_2011].
@@
 In each subsequent timestep, agents either choose to enroll in the conservation program, receive their reward, and continue to harvest just from the remaining 80% of the unprotected landscape, or they choose to harvest from the entire landscape and forgo the reward.
-After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior [@baldini_success-biased_2012; @baldini_two_2013, @barrett_equifinality_2019].
+After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior [@baldini_two_2013].
 The protected and unprotected resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
+As behavioral adoption under success-biased learning is dependent on harvest success, the successfulness of each behavior is determined largely by the number of available resources in each area, and the number of resources in each area is a result of behavioral adoption, this learning strategy is contingent on behavior-environment feedbacks, unlike unbiased adoption which occurs independent of harvest success.
 
@@
 \begin{align*}
-&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
-&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} - \phi * \text{E}_{t} \\
-&\frac{dR}{dt} = \gamma * \text{I}_{t} - \chi * \text{R}_{t} + \phi * \text{E}_{t} \\
-&\frac{dE}{dt} =  - \rho * (\text{S}_{t} + \text{R}_{t}) + \theta*\text{E}_{t}
+&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \\
+&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} - \phi \cdot \text{E}_{t} \\
+&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} + \phi \cdot \text{E}_{t} \\
+&\frac{dE}{dt} =  - \rho \cdot (\text{S}_{t} + \text{R}_{t}) + \theta \cdot \text{E}_{t}
 \end{align*}

### 2021-10-08 DOI_Review/DOI_ReviewApp.R

@@ -26,7 +26,7 @@ ui<-shinyUI(navbarPage("Protected Areas ABM",
                                                                   value = 100),
                                                       selectInput("LearningStrategy","Learning Strategy",
                                                                   c("Success Bias" = "Success Bias",
-                                                                    "Conformist Bias"= "Conformist Bias")),
+                                                                    "Unbiased"= "Unbiased")),
                                                       sliderInput("totCareCapac","Total Carrying Capacity Landscape",
                                                                   min = 100,max=100000,value=10000),
                                                       sliderInput("StartPercCarryingCapacity","Starting Percent Carrying Capacity",
@@ -234,7 +234,7 @@ server <- function(input, output) {
       
       
       #start conformist learning
-      if(LearningStrategy=="Conformist Bias"){
+      if(LearningStrategy=="Unbiased"){
         
         spreadProb<-0.25 #0.25
         recProb<-0.1 #0.01

### 2021-10-16 DOI_Review/Manuscript.rmd
Tons of updates on DOI review and HCRI proj.

--- 2021-10-06 DOI_Review/Manuscript.rmd
@@
   name: Vicken Hillis
-- affiliation: Max Planck Institute for Evolutionary Anthropology
-  name: Jeffrey Andrews
 
@@
 bibliography: [./DOI_Conservation_Review.bib]
-abstract: Conservation researchers are increasingly identifying diffusion theory as useful for understanding the spread of community-based conservation initiatives.
-This work has nevertheless been largely ambiguous as to the specific opportunities and shortcomings for diffusion theory in conservation research, and how it might be practically applied.
-Here, we employ agent-based simulations to confirm that process-based diffusion models reasonably predict the long-term patterns in adoption of conservation behaviors even under tightly coupled behavior-environment feedbacks.
-We caution however, that while diffusion theory may predict the pattern with which a conservation intervention will spread through a population, it gives little insight as to the underlying transmission mechanism.
-We show that the transmission mechanism employed by a population can alter the final proportion of individuals that adopt an intervention, and argue that conservation researchers and practitioners can leverage theory from social learning and cultural evolution to better design and diagnose projects.
-Lastly, we emphasize that developing and fitting process models in statistical frameworks traditionally reserved for regression alone, may pose a productive new avenue for conservation science, particularly in our understanding of conservation as a complex system.
+abstract: A common goal of conservation practitioners is to design interventions that benefit both people and the environment, and as a result are adopted robustly by local communities.
+Diffusion theory provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
+Current prescriptions have nevertheless been largely ambiguous as to the practical applications of diffusion theory to conservation work, and in particular, in coupled human-natural systems.
+Here, we demonstrate the capacity for a process-based diffusion model to reasonably predict the long-term rates of conservation adoption under tightly coupled behavior-environment feedbacks.
+Data on conservation adoption are generated through an agent-based simulation which highlights our dynamics of interest.
+We caution however, that while diffusion models may predict the pattern with which a conservation intervention will spread through a population, they give little diagnostic insight as to why.
+We show that the social learning mechanism employed by a population can alter the final proportion of individuals that adopt an intervention, and argue that conservation researchers and practitioners can leverage theory from social learning and cultural evolution to better design and diagnose projects.
+Lastly, we emphasize that developing and fitting process models in statistical frameworks traditionally reserved for linear regression alone, may pose a productive new avenue for conservation science, particularly in concert with simulation techniques and insights from complex systems.
 
@@
 >
-> `r tufte::quote_footer('--- Everett M. Rogers, Diffusion of Innovations, 2003')`
+> `r tufte::quote_footer('--- Everett M. Rogers, Diffusion of Innovations, 2010')`
 
@@
 While some debate still exists and exact methods vary, it is largely accepted that community-based conservation interventions are a viable method for seeding self-spreading, -sustaining, and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004].
-A major question in implementing these projects is if a specific intervention will indeed catch-on in a population and if so, how quickly [@radeloff_hot_2013]?
+A major question in implementing these projects is if a specific intervention will indeed catch on in a population and if so, how quickly [@radeloff_hot_2013]?
 Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [@mangubhai_not_2011; @visconti_protected_2019].
@@
 # Diffusion Theory
-Inducing and/or speeding up the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time.
+Promoting the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time.
 Diffusion theory recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010].
@@
 This assumption about human decision-making has been validated repeatedly in cultures around the world; individuals do not act as cost-benefit analysts, in which they adapt their behaviors based on the environment alone [@henrich_search_2001].
-Our dependence on social learning to make decisions about behavioral adoption has profound and specific implications for patterns of behavioral adoption at the population level [@smaldino_adoption_2017].
+Our dependence on social learning to make decisions has profound and specific implications for patterns of behavioral adoption at the population level [@smaldino_adoption_2017].
 We observe these patterns through diffusion curves which show the proportion of a population that has adopted a behavior as a function of time [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
@@
 Behavioral adoption thus exhibits equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [@barrett_equifinality_2019].
+Critically, this makes it virtually impossible to infer the learning mechanism(s) used by a population from the pattern of adoption over time alone.
 
@@
 
-Individuals tend to place more weight on imitation when making decisions in which the outcomes can not be well predicted independently [@perreault_bayesian_2012].
-This phenomenon is particularly salient under variable, rapidly changing environments [@boyd_culture_1988; @richerson_not_2008].
+Individuals tend to place more weight on imitation when making decisions in which the outcomes can not be easily predicted independently [@perreault_bayesian_2012].
+This has been explicitly demonstrated under variable, rapidly changing environments [@boyd_culture_1988; @richerson_not_2008].
 It follows then, that the adoption of novel conservation behaviors by a community is likely to be achieved through social imitation (i.e.
@@
 Indeed, the sigmoidal pattern of behavioral adoption characteristic to diffusion has been observed in over 80% of case studies on community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
-Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the adoption of community based conservation interventions is considerably attenuated [@pietri_information_2009-1].
+Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the adoption of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
 With the recognition that the patterns of uptake for community-based conservation practices are directly linked to behavioral imitation, researchers have begun suggesting diffusion theory as a lens for understanding and implementing individual interventions [@abernethy_two_2014; @mbaru_key_2017; @niemiec_efficacy-based_2021].
@@
 
-These studies are correct in their conclusion that diffusion theory can be productively applied to the spread of community-based conservation interventions, but ambiguous in exactly how a practitioner or researcher might practically apply this insight.
-The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors, and what the limitations of this theory are for complex adaptive systems like community-based conservation.
+These studies are compelling in their conclusion that diffusion theory can be productively applied to the spread of community-based conservation interventions, but ambiguous in exactly how someone might practically apply this insight.
+We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and/or what they can do to promote its spread.
+The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors given realistic data, and the limitations of this theory as a diagnostic tool and for application to complex, coupled human-natural systems.
 
@@
 The sigmoidal pattern of behavioral spread over time mirrors that of disease transmission [@smaldino_coupled_2020].
-This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral uptake [@cavalli-sforza_cultural_1981; @strang_adding_1991].
+This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral spread [@cavalli-sforza_cultural_1981; @strang_adding_1991].
 Specifically, the canonical susceptible-infectious-recovered (SIR) model and SIRS variation where individuals are added back to the pool of susceptibles after recovery (eq 1), have been demonstrated to be well suited to fitting the diffusion process [@famil-rohani_development_2019; @smaldino_coupled_2020].
-In this instantiation, $\beta$ represents the rate of infection in the population, $\gamma$ represents the rate of recovery, and $\chi$ is the rate that recovered individuals are added back into the pool of susceptibles.
+These compartmental differential equations assign every member of a population to a category (S, I, or R), depending on the proportion of individuals in the other categories and parameter values for the natural history characteristics of a given disease [@tolles_modeling_2020].
+In this instantiation, the natural history parameter $\beta$ represents the rate of infection in the population, $\gamma$ represents the rate of recovery, and $\chi$ is the rate that recovered individuals are added back into the pool of susceptibles.
 
-\begin{align*}
-&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
-&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} \\
-&\frac{dR}{dt} = \gamma* \text{I}_{t} - \chi * \text{R}_{t}
-\end{align*}
+\begin{align}
+&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
+&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
+&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} \nonumber
+\end{align}
 
@@
 We can however, use Bayesian inference to probabilistically estimate these parameters from data on behavioral uptake over time [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018].
-Below, we use a minimal agent-based model of conservation adoption contingent on behavior-environment feedbacks to simulate data on the long-term adoption dynamics of a community-based conservation intervention.
-Then, using just the first 30% of the data, we estimate the rate of intervention spread, rate of dropout, and time after dropout before agents are available to adopt again.
-We use these estimated parameters to predict the overall patterns of adoption over all time steps and compare our observed to predicted values.
+
+Below, we develop a minimal agent-based model to highlight our core dynamics of interest.
+In this abstraction, we simulate the spread in adoption of a conservation intervention from just a few initial individuals, to a stable proportion of the population.
+We use this simulation to produce realistic data on long-term adoption dynamics under tightly coupled behavior-environment feedbacks unique to community-based conservation interventions.
+Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using the SIRS model shown above.
+We use these estimated parameters to predict the long-term patterns of adoption for the remaining 70% of time steps and compare our predicted values to those observed in the simulation.
 
@@
 We developed a minimal agent-based simulation which we use to generate data on conservation adoption within a population.
-We use these data to test the performance of the SIRS diffusion model in predicting patterns of conservation behavior adoption under both unbiased and success-biased social learning.
-Success-biased social learning is contingent on behavior-environment feedbacks.
+We use these data to test the performance of the SIRS diffusion model in predicting patterns of conservation behavior adoption under two social learning mechanisms.
+The first is unbiased copying, where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
+The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of more successful agents [@baldini_two_2013].
+This mechanism is formally demonstrated in equation 1, for two behaviors: $A$ & $B$, where the probability that an individual adopts behavior $A$ is a function of the mean pay-off, $\bar{P}$, of both behaviors at each time step and the strength of the bias, $\beta$ [@barrett_equifinality_2019].
+
+\begin{equation}
+ \text{Pr}(A|A,B)_{t+1} = \frac{\text{exp}\beta\bar{P}_{A,t}}{\text{exp}\beta\bar{P}_{A,t} + \text{exp}\beta\bar{P}_{B,t}}  
+\end{equation}
+
 Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [@ferraro_future_2011].
-We establish a degraded landscape/seascape (from herein referred to as landscape) where all simulated individuals are dependent on harvesting some renewable, mobile resource.
-In the first timestep, we introduce a conservation intervention which delineates 20% of the landscape as a conservation area and individuals are rewarded for abstaining from harvesting there.
-In each subsequent timestep, agents either choose to enroll in the conservation program, receive their reward, and continue to harvest just from the remaining 80% of the unprotected landscape, or they choose to harvest from the entire landscape and forgo the reward.
-After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior [@baldini_success-biased_2012; @baldini_two_2013, @barrett_equifinality_2019].
-The protected and unprotected resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
+We establish a degraded landscape/seascape (hereafter referred to as landscape) where all simulated individuals are dependent on harvesting some renewable, mobile resource.
+In the first time step, we introduce a conservation intervention which delineates 20% of the landscape as a conservation area and is initially adopted by just 2% of the population.
+In each subsequent time step, agents enrolled in the conservation program receive a reward, and continue to harvest just from the remaining 80% of the unconserved landscape.
+Agents not enrolled in the intervention harvest from the entire landscape and forgo the reward.
+After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior in the next time step [@baldini_success-biased_2012].
+The conservation and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
+As behavioral adoption under success-biased learning is dependent on harvest success, the success of each behavior is determined largely by the amount of available resources in each area, and the amount of resources in each area is a result of behavioral adoption, this learning strategy is contingent on behavior-environment feedbacks, unlike unbiased adoption which occurs independent of harvest success.
 
@@
 A visual description of this minimal model is shown in figure 1, and the full ODD+ model description can be found in the supplemental material (S1).
-Interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at this URL [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
-As this simulation is not the emphasis of this study, we show just two runs below, one in which.
+Interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
+As this simulation is not the emphasis of this study, we show just two runs below.
 The starting conditions of these runs are identical except for the social learning mechanism employed by the population.
 As the unbiased learning mechanism is independent of the resource levels, the adoption curve follows the traditional “S” pattern and nears fixation in the long-term.
-Under environmentally mediated success-biased learning, the adoption pattern spikes initially, but drops off dramatically as the resources in the unprotected landscape are reduced following increased harvest pressure there as a result of the conservation intervention.
-The full output of this later simulation is shown below in figure 2.
-Additional simulation runs can be found in the supplemental material (S2) or by using the interactive web application linked above.
+Under environmentally mediated success-biased learning, the adoption pattern spikes initially, but drops off dramatically as the resources in the unconserved landscape are reduced following increased harvest pressure there as a result of the conservation intervention.
+The full output of this latter simulation is shown below in figure 2.
 
@@
 $$ \text{Y}_{t} \sim Binomial(\text{N}, \text{I}_{t}) $$
-\begin{align*}
-&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
-&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} \\
-&\frac{dR}{dt} = \gamma* \text{I}_{t} - \chi * \text{R}_{t}
-\end{align*}
+\begin{align}
+&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
+&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
+&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t}\nonumber
+\end{align}
 
@@
 
-For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an rhat of 1 for all estimated parameters (fig 3).
+For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an $\hat{R}$ of 1 for all estimated parameters (fig 3).
 As the parameters $\beta$, $\gamma$, and $\chi$ are directly interpretable as the rates of adoption, dropout, and willingness to retry a previously abandoned intervention, accurate estimation of these values alone, especially in the early stages of an intervention, is a practical and straightforward application of diffusion theory to community-based conservation interventions.
+For example, a practitioner might fit this model to real-world participation rates during the early weeks of an intervention in order to estimate whether they should focus more on increasing rates of uptake or retention going forward.
 
@@
 
-We can also use these parameter values, estimated from the first 30% of the time series, to predict the long-term patterns of adoption for community-based conservation initiatives.
+We can also use these parameter values, estimated from the first 30% of the time series, to predict the long-term patterns of adoption for a community-based conservation initiative.
 In figures 4 & 5, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step.
@@
 These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning.
-R and STAN code for these models can be found in the supplemental material (S4)
+R and STAN code for these models can be found in the supplemental material (S4).
 
@@
 Figures 4 & 5 also demonstrate that different learning mechanisms, while both producing a sigmoidal adoption pattern, can result in different overall rates in adoption of an intervention, even under identical starting conditions.
-In the simulation of a population employing unbiased learning, the population fixates at nearly 90% adoption of the intervention behavior.
-For success-biased learning on the other hand, resource scarcity in the unprotected area caused the population to adopt the intervention at approximately 60% in the long-term.
+In the simulation of a population employing unbiased learning, the population equilibrates at nearly 90% adoption of the intervention behavior.
+For success-biased learning on the other hand, resource scarcity in the unconserved area caused the population to adopt the intervention at approximately 60% in the long-term.
 
@@
 Therefore, while we agree that diffusion theory can be productively applied to community-based conservation, we point to theories of social learning and cultural evolution as a better alternative for diagnosing and planning interventions.
+For example, Tam et al.
+[-@tam_measuring_2021] and Kline et al.
+[-@kline_designing_2018] give practical examples of how to measure social learning strategies in the context of conservation, and tools (game/survey instrument, code, etc.) with which to collect and analyze the data.
 
 Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [@cheng_strengthen_2020].
-Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes [@turner_land-useland-cover_2021].
-Specifically, in modeling social-ecological systems, either using analytical approaches or simulation, processes should be based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013].
+Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013; @turner_land-useland-cover_2021].
 Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [@schill_more_2019].
 
-We argue that when prediction is the goal, analytical or statistical approaches like the SIRS model presented above are always preferable to simulation based approaches [@epstein_why_2008; @tredennick_practical_2021].
-The practice of applying statistical frameworks of parameter estimation to analytical process models, and in particular to ordinary differential equations, is still relatively new and has been embraced mostly by ecology [@hefley_when_2017].
-Extending the SIRS diffusion model to include a term which represents changes in the environment over time and the effect of resource levels on adoption rates may more accurately describe the system of adoption in community-based conservation interventions.
-An example of such an extension is presented below where we include another equation for the change in a resource pool over time $\frac{dE}{dt}$.
-We also must include three additional parameters: $\phi$.
-$\rho$, and $\theta$, which represent the effect of resource level on enrollment (infection), the effect of harvest on the resource level, and the regeneration rate of the resource respectively.
+The practice of applying statistical frameworks for parameter estimation to analytical process models, and in particular to ordinary differential equations, is still relatively new and has been embraced mostly in ecology [@hefley_when_2017; @laubmeier_ecological_2020].
+This emerging methodology allows for the estimation of interesting and directly interpretable process parameters, not possible with linear regression.
+We believe this is an exciting new avenue for conservation science as the direct interpretability and applicability of process parameters lend themselves well to tight integration with applied research as is common in the field.
+Future research may be interested in extending the SIRS diffusion model to include a term which represents change in the environment over time and the effect of resource levels on adoption rates to more accurately reflect the social-ecological coupling.
+An example of such an extension is presented below, where we include another equation for the change in a resource pool over time $\frac{dE}{dt}$.
+We also must include three additional parameters: $\phi$, $\rho$, and $\theta$, which represent the effect of resource level on enrollment (infection), the effect of harvest on the resource level, and the regeneration rate of the resource respectively.
 
-\begin{align*}
-&\frac{dS}{dt} = - \beta * \text{S}_{t} * \text{I}_{t} + \chi * \text{R}_{t} \\
-&\frac{dI}{dt} = \beta * \text{S}_{t} * \text{I}_{t} - \gamma * \text{I}_{t} - \phi * \text{E}_{t} \\
-&\frac{dR}{dt} = \gamma * \text{I}_{t} - \chi * \text{R}_{t} + \phi * \text{E}_{t} \\
-&\frac{dE}{dt} =  - \rho * (\text{S}_{t} + \text{R}_{t}) + \theta*\text{E}_{t}
-\end{align*}
+\begin{align}
+&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
+&\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} - \phi \cdot \text{E}_{t} \\
+&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} + \phi \cdot \text{E}_{t} \nonumber\\
+&\frac{dE}{dt} = - \rho \cdot (\text{S}_{t} + \text{R}_{t}) + \theta \cdot \text{E}_{t}\nonumber
+\end{align}
 
-While process models may be preferable for predicting adoption outcomes of conservation interventions, they might not always be possible.
-In complex systems, these models can quickly require many parameters in order to accurately account for all moving parts and thus are prone to issues of identifiability in parameter estimation [@laubmeier_ecological_2020].
-Further, complex social-ecological systems have properties which are inherently stochastic and may be impossible to model analytically.
-For example, coastal fisheries have been shown to approach mathematical chaos in the number and proportions of species present [@wilson_chaos_1994].
+While process models may be preferable to simulation and linear regression approaches for predicting adoption outcomes of conservation interventions, they might not always be possible.
+In complex systems, these models can quickly require many parameters in order to accurately account for all moving parts and thus are prone to issues of identifiability in parameter estimation and may even become mathematically intractable [@laubmeier_ecological_2020].
+Further, as has been demonstrated in coastal fisheries, these systems can exhibit mathematical chaos, where our inability to measure the parameters of interest at a fine enough resolution results in seemingly stochastic, unpredictable outcomes [@wilson_chaos_1994].
 
-In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for tipping points and emergence, two key features of the complex systems which can not be accounted for with process models such the SIRS or SIRS extension models presented above [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
 Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically [@liu_complexity_2007].
+In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems which are difficult or even impossible to account for with process models such the SIRS or SIRS extension models presented above [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
 

### 2021-10-16 DOI_Review/Binary_Adoption_ABM.R

@@ -287,7 +287,7 @@ abm<-function(#Specified parameters
     theme_classic()+ylab("Carrying Capacity")+xlab("")+
     scale_y_continuous(labels = scales::percent,limits = c(0,1.0) )+
     scale_colour_manual(name = '', 
-                        values =c('Protected'='#b2df8a','Working'='#1f78b4'), labels = c('Conservation Area','Unprotected Area'))+
+                        values =c('Protected'='#b2df8a','Working'='#1f78b4'), labels = c('Conservation Area','Unconserved Area'))+
     ggtitle("Resource Integrity")+mytheme +
     theme(legend.position="bottom")
   

### 2021-10-16 HCRI_Grant/ProjectFiles/CreateFireProbabilityLayer.R

@@ -1,19 +1,33 @@
+library(raster)
+Pem2018LC<-raster::raster("Pemba2018_LandCover.tif")
 stackRS<-raster::stack("PembaFiresAndPredictors.tif")
-names(stackRS)<-c("roads_proximity","slope","soil_cat","fires2019")
+#Make function to reproject LC to the stack DF
+preprocess_raster<-function(raster,method){
+  if(method=="ngb"){ x<-raster::projectRaster(raster,crs=crs(stackRS),method="ngb")}
+  if(method=="bilinear"){ x<-raster::projectRaster(raster,crs=crs(stackRS))}
+  return(resample(x,stackRS,method=method))
+}
+Pem2018LC<-preprocess_raster(Pem2018LC,method="ngb")
+Pem2018LC<-mask(Pem2018LC,stackRS$PembaFiresAndPredictors.1)
+stackRS<-stack(stackRS,Pem2018LC)
+names(stackRS)<-c("roads_proximity","slope","soil_cat","fires2019","LandCover2018")
 StackDF <- raster::as.data.frame(stackRS,xy=TRUE)
 StackDF<-na.omit(StackDF)
 StackDF$soil_cat<-as.character(StackDF$soil_cat)
+StackDF$LandCover2018<-as.character(StackDF$LandCover2018)
+
 #ggplot(v,aes(x=roads_proximity,y=layer))+geom_jitter(alpha=0.05)+geom_smooth(method="gam")
 
 #stdize function as per Gelman reccomendation
 stdize<-function(x){
   (x-mean(x))/(2*sd(x))}
-StackDF<-StackDF%>%mutate(slope_std=stdize(slope),road_proximity_std=stdize(roads_proximity))
+StackDF<-StackDF%>%filter(LandCover2018%in%c("5","1"))%>% #Only include burnable pixels
+  mutate(slope_std=stdize(slope),road_proximity_std=stdize(roads_proximity))
 
 ########THIS NEEDS TO BE EDITED SO WE ONLY PREDICT FIRES ON BURNABLE LAND! #####################
 # Filter out all non-burnable land before running the regression!!!!!!!!!!!!!
 
-fit<-glm(fires2019~slope_std+road_proximity_std+soil_cat,family="binomial",data=StackDF)
+fit<-glm(fires2019~slope_std+road_proximity_std+soil_cat+LandCover2018,family="binomial",data=StackDF)
[…]

### 2021-10-16 HCRI_Grant/ProjectFiles/RunABMonPemData.R

@@ -12,6 +12,21 @@ rLndCvr2018<-rProbBurn
 rLndCvr2018[]<-sample(classes,size=ncell(rLndCvr2018),replace=T,prob=c(0.1,0.1,0.7,0.1))
 rLndCvr2018[is.na(rProbBurn)]<-NA
 
+#These will be used to assign working/fallow ag randomly
+FallowRechargeTime<-3
+AgLimit<-2
+propstrt<-AgLimit/FallowRechargeTime
+
+#Import 2018 Lc Data
+Pem2018LC<-raster::raster("~/Pemba_Project/HCRI_Grant/ProjectFiles/Pem_2018_LC.tif")
+
+rLndCvr2018<-Pem2018LC
+rLndCvr2018[which(rLndCvr2018[]==2)]<-sample(c(100,2),size=length(which(rLndCvr2018[]==2)),
+                                             replace=T,prob=c(propstrt,1-propstrt))
+rLndCvr2018[]<-as.factor(ifelse(rLndCvr2018[] %in% c(1,5),"Burnable",
+                             ifelse(rLndCvr2018[] == 2,"Agriculture",
+                                    ifelse(rLndCvr2018[] == 100,"Bare","Unburnable")))) #bare means fallow here
+
 
 #stack all starting layers together
 rstack<-stack(stackRS$fires2019,rProbBurn,rLndCvr2018 )
@@ -31,12 +46,11 @@ rstack <- mask(r2, PembaSUB)
 ############### model copied from simple abm
 ag2018<-which(rstack$LndCvr2018[]==1) #ag from 2018 land cover layer
 burned2019<-which(rstack$Burn2019[]==1) #get cell numbers that burned in 2019
-bare2018<-which(rstack$LndCvr2018[]==2) #fallow/bare from 2018 land cover layer
+fallow2018<-which(rstack$LndCvr2018[]==2) #fallow/bare from 2018 land cover layer
 burnable2018<-which(rstack$LndCvr2018[]==3) #burnable from 2018 landcover layer
 unburnable2018<-which(rstack$LndCvr2018[]==4) #unburnable from 2018 land cover layer
 
-FallowRechargeTime<-3
-AgLimit<-2
+
 
 #Make the starting ag layer
 rstack$ag2019<-NA #Make all NA to start
@@ -45,7 +59,7 @@ rstack$ag2019[burned2019]<-1
 
 #Make starting fallow layer
[…]

### 2021-11-20 DOI_Review/Manuscript.rmd
Text edits through quantitative applications

@@
 Diffusion theory provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
-Current prescriptions have nevertheless been largely ambiguous as to the practical applications of diffusion theory to conservation work, and in particular, in coupled human-natural systems.
-Here, we demonstrate the capacity for a process-based diffusion model to reasonably predict the long-term rates of conservation adoption under tightly coupled behavior-environment feedbacks.
+Here, we elaborate on the need for better understanding the spread of conservation behaviors, and review diffusion theory and its past application to these problems.
+We follow this with a practical demonstration of how the mathematical models which fit the diffusion process can be used to diagnose and predict the uptake of conservation behaviors even under tightly coupled behavior-environment feedbacks.
 Data on conservation adoption are generated through an agent-based simulation which highlights our dynamics of interest.
-We caution however, that while diffusion models may predict the pattern with which a conservation intervention will spread through a population, they give little diagnostic insight as to why.
-We show that the social learning mechanism employed by a population can alter the final proportion of individuals that adopt an intervention, and argue that conservation researchers and practitioners can leverage theory from social learning and cultural evolution to better design and diagnose projects.
-Lastly, we emphasize that developing and fitting process models in statistical frameworks traditionally reserved for linear regression alone, may pose a productive new avenue for conservation science, particularly in concert with simulation techniques and insights from complex systems.
+Throughout this review, situate diffusion theory in the broader literature on cultural evolution and use these insights to refine our claims about how diffusion can be productively applied to conservation.
+Specifically, we show that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention.
+Overally, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion in these systems.
 
@@
 The primary way individuals reduce uncertainty about the outcome of adopting a particular behavior is through second-hand observation of early adopters.
-Therefore, the core of diffusion theory is the notion that individuals imitate those around them when making decisions about adopting a novel behavior or not, rather than weighing the pros and cons independently.
+Therefore, at the core of diffusion theory is the observation that individuals imitate those around them when making decisions about adopting a novel behavior or not, rather than weighing the pros and cons independently [@henrich_search_2001].
 
-This assumption about human decision-making has been validated repeatedly in cultures around the world; individuals do not act as cost-benefit analysts, in which they adapt their behaviors based on the environment alone [@henrich_search_2001].
-Our dependence on social learning to make decisions has profound and specific implications for patterns of behavioral adoption at the population level [@smaldino_adoption_2017].
-We observe these patterns through diffusion curves which show the proportion of a population that has adopted a behavior as a function of time [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
+When modifying their beliefs or behaviors, people do not imitate any other individual at random [@bandura_social_1977].
+Instead, they preferentially adopt traits by employing a variety of learning biases such as success-biased learning where individuals who are perceived as more successful are more likely to be copied than unsuccessful individuals [@boyd_culture_1988; @laland_social_2004].
+As a result, the cumulative culture of a population evolves over time as various beliefs and behaviors are unequally copied and their frequency within a population changes.
+While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information which is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016].
+This information affects individuals’ behaviors, which in-turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991].
+Thus, cultural evolution provides an intuitive way to scale from individual learning and decision making to population level patterns and even population level behavior-environment feedbacks.
 
-Behavioral spread through imitation produces a sigmoidal “S” shaped diffusion curve, which is observed in the vast majority of empirical studies on behavioral adoption [@henrich_cultural_2001].
-This “S” shaped curve is produced regardless of the social learning mechanism employed by a population and can even be observed when a portion of the population relies on individual learning [@henrich_cultural_2001].
-Behavioral adoption thus exhibits equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [@barrett_equifinality_2019].
-Critically, this makes it virtually impossible to infer the learning mechanism(s) used by a population from the pattern of adoption over time alone.
+Diffusion of novel beliefs and behaviors is a specific instance of cultural evolution, where the trait being transmitted is new to the population and outcompetes some previous cultural variant, such as in the classic example of hybrid corn seed adoption in Iowa farmers [@mesoudi_cultural_2016; @ryan_acceptance_1950].
+We observe these patterns through diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
+The exact shape of these diffusion curves is highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010 ].
 
-# Diffusion Theory in Conservation
+In the vast majority of empirical studies, the diffusion of a novel trait produces a sigmoidal “S” shape, where it initially spreads slowly, followed by a period of rapid adoption, then again spreads slowly until it reaches fixation [@rogers_diffusion_2010 ].
+This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001].
+These two curve shapes however are not completely discrete, as the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].
 
-Individuals tend to place more weight on imitation when making decisions in which the outcomes can not be easily predicted independently [@perreault_bayesian_2012].
-This has been explicitly demonstrated under variable, rapidly changing environments [@boyd_culture_1988; @richerson_not_2008].
-It follows then, that the adoption of novel conservation behaviors by a community is likely to be achieved through social imitation (i.e.
-diffusion), especially given worsening climatic instability and regime changes resulting from widespread biodiversity loss.
+![Canonical "S" and "R" diffusion curves.
+"R" curves are generally indicative of a higher proportion of individuals employing independent learning.
+"S" curves are generally indicative of a higher proportion of individuals employing social learning.
+](./SvsR2.png)
 
-Indeed, the sigmoidal pattern of behavioral adoption characteristic to diffusion has been observed in over 80% of case studies on community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
-Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the adoption of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
-With the recognition that the patterns of uptake for community-based conservation practices are directly linked to behavioral imitation, researchers have begun suggesting diffusion theory as a lens for understanding and implementing individual interventions [@abernethy_two_2014; @mbaru_key_2017; @niemiec_efficacy-based_2021].
+# Diffusion Theory in Community-Based Conservation
 
-Meta-level studies have recently come to the same conclusion.
-For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters, leading the authors to point to diffusion theory as an avenue for increasing the proportion of adopters [@mills_how_2019].
-And a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices further identified diffusion theory as useful for guiding the development of conservation interventions [@mahajan_theorybased_2020].
+Just as the sigmoidal pattern indicative of social learning is seen in the majority of cases on behavioral adoption broadly, scholars have also noticed this pattern in over 80% of case studies on the uptake of community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
+High-level studies have begun identifying diffusion theory as potentially useful for diagnosing why some community-based conservation interventions spread and persist, while others fizzle out.
+For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters.
+The authors point to a sigmoidal diffusion pattern to explain the slow start of many initiatives and suggest that practitioners may actively spread information on interventions (as in Abernethy et al.
+[-@abernethy_two_2014]), or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004].
+Further, a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices identifies qualitative insights, such as intervention observability, from diffusion theory as useful in guiding the development of conservation interventions [@mahajan_theorybased_2020].
 
-These studies are compelling in their conclusion that diffusion theory can be productively applied to the spread of community-based conservation interventions, but ambiguous in exactly how someone might practically apply this insight.
+As these high-level studies suggest, qualitative insights from diffusion theory have been productively applied to understand the rate and success of spread of conservation behaviors in single case studies around the world.
+In Kenya for example, Mbaru and Barnes [-@mbaru_key_2017] show that identifying socially well-connected individuals can facilitate the rapid diffusion of information on conservation objectives.
+Similarly, in Fort Collins, CO, researchers demonstrate that both the speed and extent of diffusion of information on native plant gardening likely increase with after motivating communication in the population [@niemiec_efficacy-based_2021].
+Romero-de-Diego et al.
+[-@romero-de-diego_drivers_2021] use expert interviews to identify that observability of benefits was a key driver of the rapid scaling-up of wildlife management areas in Mexico.
+Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
+This finding mirrors those of studies which look at more general cultural adoption across island chains, further suggesting that conservation behaviors may spread similarly to technological or ideological innovations [@kline_population_2010; @beheim_inheritance_2011].
+
+These studies are compelling in their conclusion that diffusion theory can be productively applied to the spread of community-based conservation, but have been critiqued as qualitative suggestions may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020].
+Below, we make a quantitative argument for using diffusion theory to understand the spread of community-based conservation interventions.
 We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and/or what they can do to promote its spread.
-The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors given realistic data, and the limitations of this theory as a diagnostic tool and for application to complex, coupled human-natural systems.
+The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors given realistic data, and diagnose when in the lifetime of an intervention individuals may be dropping out.
 

### 2021-11-23 DOI_Review/Manuscript.rmd
catching up for PR

@@
 Diffusion theory provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
-Here, we elaborate on the need for better understanding the spread of conservation behaviors, and review diffusion theory and its past application to these problems.
+Here, we elaborate on the need for better understanding the spread of conservation behaviors and review diffusion theory and its past application to these problems.
 We follow this with a practical demonstration of how the mathematical models which fit the diffusion process can be used to diagnose and predict the uptake of conservation behaviors even under tightly coupled behavior-environment feedbacks.
@@
 Specifically, we show that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention.
-Overally, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion in these systems.
+Overally, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion of community-based conservation initiatives.
 
@@
 
-When modifying their beliefs or behaviors, people do not imitate any other individual at random [@bandura_social_1977].
+When modifying their beliefs or behaviors, people do not imitate any other individuals at random [@bandura_social_1977].
 Instead, they preferentially adopt traits by employing a variety of learning biases such as success-biased learning where individuals who are perceived as more successful are more likely to be copied than unsuccessful individuals [@boyd_culture_1988; @laland_social_2004].
@@
 We observe these patterns through diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
-The exact shape of these diffusion curves is highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010 ].
+The exact shape of these diffusion curves is highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010].
 
-In the vast majority of empirical studies, the diffusion of a novel trait produces a sigmoidal “S” shape, where it initially spreads slowly, followed by a period of rapid adoption, then again spreads slowly until it reaches fixation [@rogers_diffusion_2010 ].
+In the vast majority of empirical studies, the diffusion of a novel trait produces a sigmoidal “S” shape, where it initially spreads slowly, followed by a period of rapid adoption, then again spreads slowly until it reaches fixation [@rogers_diffusion_2010].
 This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001].
@@
 # Diffusion Theory in Community-Based Conservation
-
 Just as the sigmoidal pattern indicative of social learning is seen in the majority of cases on behavioral adoption broadly, scholars have also noticed this pattern in over 80% of case studies on the uptake of community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
@@
 Below, we make a quantitative argument for using diffusion theory to understand the spread of community-based conservation interventions.
-We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and/or what they can do to promote its spread.
-The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors given realistic data, and diagnose when in the lifetime of an intervention individuals may be dropping out.
+We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and what they can do to promote its spread.
+The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors given realistic data and diagnose when in the lifetime of an intervention individuals may be dropping out.
 
 # Quantitative applications
-
-The sigmoidal pattern of behavioral spread over time mirrors that of disease transmission [@smaldino_coupled_2020].
+In this section we explain how we can fit the pattern of diffusion mathematically to derive inference about the rate and degree to which an intervention will be adopted by a population.
+The “S” shaped diffusion pattern of behavioral spread over time mirrors that of community disease transmission [@smaldino_coupled_2020].
 This parallel implies that the well-established mathematics used to describe patterns of disease transmission can be applied to fitting models of behavioral spread [@cavalli-sforza_cultural_1981; @strang_adding_1991].
@@
 
-\begin{align}
-&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
+\begin{align*}
+&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \\
 &\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
-&\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} \nonumber
-\end{align}
+&\frac{dR}{dt} = \gamma\cdot \text{I}_{t} - \chi \cdot \text{R}_{t}
+\end{align*}
 
-When the natural history characteristics of some pathogen is known, the SI, SIR, SIS, & SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [@huppert_mathematical_2013].
+When the natural history characteristics of some pathogen are known, the SI, SIR, SIS, & SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [@huppert_mathematical_2013].
 For conservation interventions however, there are no natural history features of a specific intervention, such as the rate or duration of adoption, which can be known ahead of time.
-We can however, use Bayesian inference to probabilistically estimate these parameters from data on behavioral uptake over time [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018].
+Instead, the qualitative insights from diffusion theory referenced above aim to increase the rate of spread and reduce the rate of dropout so that the intervention persists in the population and may even spread to new communities.
+While we can not know exactly what the rates of spread and dropout will be ahead of time, we can use Bayesian inference to probabilistically estimate them once the project has been implemented [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018].
+Knowing these parameters may allow us to predict the final rate of uptake from early in the project, identify whether adoption or dropout are the primary issue, and apply the qualitative insights from diffusion theory more targetedly.
 
-Below, we develop a minimal agent-based model to highlight our core dynamics of interest.
-In this abstraction, we simulate the spread in adoption of a conservation intervention from just a few initial individuals, to a stable proportion of the population.
-We use this simulation to produce realistic data on long-term adoption dynamics under tightly coupled behavior-environment feedbacks unique to community-based conservation interventions.
-Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using the SIRS model shown above.
+In some contexts, the uptake of community-based conservation is subject to behavior-environment feedbacks, a fundamental difference from behavioral adoption in the abstract [@liu_complexity_2007].
+In the case of marine protected areas for example, success in conserving fish populations may actually disincentivize individuals to continue supporting conservation programs [@christie_marine_2004].
+A key question then is whether the diffusion models explained above will still reliably fit the spread of conservation behaviors when those behaviors alter the environment, and changes in the environment alter the payoff structure of various behavioral choices.
+To answer this question, we use an agent-based simulation to generate data on the spread of a conservation behavior via social learning, contingent on feedbacks with the environment.
+Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using a Bayesian instantiation of the SIRS model shown above.
 We use these estimated parameters to predict the long-term patterns of adoption for the remaining 70% of time steps and compare our predicted values to those observed in the simulation.
+While the purpose of this example is to test whether the SIRS model can capture our core dynamics of interest, it also serves as an explicit demonstration of how a practitioner or other researchers can directly apply the quantitative process described above.
+We first briefly describe the simulation below.
 
@@
 We developed a minimal agent-based simulation which we use to generate data on conservation adoption within a population.
-We use these data to test the performance of the SIRS diffusion model in predicting patterns of conservation behavior adoption under two social learning mechanisms.
-The first is unbiased copying, where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
+We use these data to test the performance of the SIRS diffusion model in fitting patterns of conservation behavior adoption under two social learning mechanisms.
+The first is unbiased, frequency-based, copying where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
 The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of more successful agents [@baldini_two_2013].
@@
 
-\begin{equation}
- \text{Pr}(A|A,B)_{t+1} = \frac{\text{exp}\beta\bar{P}_{A,t}}{\text{exp}\beta\bar{P}_{A,t} + \text{exp}\beta\bar{P}_{B,t}}  
-\end{equation}
+$$ \text{Pr}(A|A,B)_{t+1} = \frac{\text{exp}\beta\bar{P}_{A,t}}{\text{exp}\beta\bar{P}_{A,t} + \text{exp}\beta\bar{P}_{B,t}} $$
 
-Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [@ferraro_future_2011].
+Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [@ferraro_future_2011; @rakotonarivo_role_2021].
 We establish a degraded landscape/seascape (hereafter referred to as landscape) where all simulated individuals are dependent on harvesting some renewable, mobile resource.
@@
 Agents not enrolled in the intervention harvest from the entire landscape and forgo the reward.
-After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior in the next time step [@baldini_success-biased_2012].
-The conservation and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
+After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior of their partner in the next time step [@baldini_success-biased_2012].
+The conserved and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
 As behavioral adoption under success-biased learning is dependent on harvest success, the success of each behavior is determined largely by the amount of available resources in each area, and the amount of resources in each area is a result of behavioral adoption, this learning strategy is contingent on behavior-environment feedbacks, unlike unbiased adoption which occurs independent of harvest success.
@@
 
-A visual description of this minimal model is shown in figure 1, and the full ODD+ model description can be found in the supplemental material (S1).
+A visual description of this minimal model is shown in figure 2, and the full ODD+ model description can be found in the supplemental material (S1).
 Interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
@@
 Under environmentally mediated success-biased learning, the adoption pattern spikes initially, but drops off dramatically as the resources in the unconserved landscape are reduced following increased harvest pressure there as a result of the conservation intervention.
-The full output of this latter simulation is shown below in figure 2.
+The full output of this latter simulation is shown below in figure 3.
 

### 2021-11-23 HCRI_Grant/ProjectFiles/ProcessLcEstimates.R

@@ -1,5 +1,5 @@
 library(raster)
-Pem2018LC<-raster::raster("~/Pemba_Project/HCRI_Grant/ProjectFiles/UnProcessed_2018_LC_Estimates_20m.tif")
+Pem2018LC<-raster::raster("~/Pemba_Project/HCRI_Grant/ProjectFiles/UnProcessed_Pem_2018_LC_20m_V2.tif")
 fireRS<-raster::raster("PredFire2019.tif")
 stackRS<-raster::stack("PembaFiresAndPredictors.tif")
 preprocess_raster<-function(raster,method){
@@ -10,7 +10,7 @@ preprocess_raster<-function(raster,method){
 Pem2018LC<-preprocess_raster(Pem2018LC,method="ngb")
 Pem2018LC<-raster::mask(Pem2018LC,stackRS$PembaFiresAndPredictors.1)
 plot(Pem2018LC)
-raster::writeRaster(Pem2018LC,"~/Pemba_Project/HCRI_Grant/ProjectFiles/Pem_2018_LC_20m.tif",overwrite=T)
+raster::writeRaster(Pem2018LC,"~/Pemba_Project/HCRI_Grant/ProjectFiles/Pem_2018_LC_20m_V2.tif",overwrite=T)
 
 #Plot it
 r <- as.factor(Pem2018LC)

### 2021-12-02 DOI_Review/Manuscript.rmd
text edits

@@
 ---
-title: Utility of Diffusion Models for Predicting and Diagnosing the Spread of Conservation Behaviors
+title: Utility of Diffusion of Innovations for Diagnosing the Spread of Conservation Behaviors
 author:
@@
 abstract: A common goal of conservation practitioners is to design interventions that benefit both people and the environment, and as a result are adopted robustly by local communities.
-Diffusion theory provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
-Here, we elaborate on the need for better understanding the spread of conservation behaviors and review diffusion theory and its past application to these problems.
+Diffusion of innovations theory (DOI) provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
+Here, we elaborate on the need for better understanding conservation institutions as a product of individual behavioral adoption and review DOI and its past application to these problems.
 We follow this with a practical demonstration of how the mathematical models which fit the diffusion process can be used to diagnose and predict the uptake of conservation behaviors even under tightly coupled behavior-environment feedbacks.
 Data on conservation adoption are generated through an agent-based simulation which highlights our dynamics of interest.
-Throughout this review, situate diffusion theory in the broader literature on cultural evolution and use these insights to refine our claims about how diffusion can be productively applied to conservation.
+Throughout this review, we situate DOI in the broader literature on cultural evolution and use these insights to refine our claims about how diffusion can be productively applied to conservation.
 Specifically, we show that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention.
-Overally, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion of community-based conservation initiatives.
+Overall, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion of community-based conservation initiatives.
 
@@
 It is becoming increasingly apparent that humanity will not reach the global conservation goals we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002].
-The sweeping behavioral changes required to meet these targets will thus have to originate from and be maintained by the communities living in and around threatened natural areas themselves [@amel_beyond_2017; @catalano_learning_2019].
-Community-based conservation projects aim to do just this by incentivizing rural, and often very poor, communities to implement these types of changes with little to no external oversight [@western_natural_1994].
-While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote long-term behavioral changes which spread from person to person and even community to community [@brooks_assessing_2013; @horwich_community_2007].
+We are thus turning to local communities to self-impose and maintain sustainable limits on resource extraction and even to restore degraded natural areas themselves [@amel_beyond_2017; @catalano_learning_2019].
+Community-based conservation projects aim to incentivize these behaviors in rural, and often very poor, communities with little to no external oversight [@western_natural_1994].
+While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote behavioral changes which are adopted robustly by communities who harvest natural resources to meet their daily needs [@brooks_assessing_2013; @horwich_community_2007].
+While some debate still exists, it is largely accepted that empowering communities to self-run conservation institutions and manage environments for multiple interests is a viable method for seeding self-sustaining and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004; @agrawal_enchantment_1999].
 
-While some debate still exists and exact methods vary, it is largely accepted that community-based conservation interventions are a viable method for seeding self-spreading, -sustaining, and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004].
-A major question in implementing these projects is if a specific intervention will indeed catch on in a population and if so, how quickly [@radeloff_hot_2013]?
+Broadly defined, institutions, like those introduced by community-based conservation interventions, are socially constructed and transmitted prescriptions on behavior, which generally aim to incentivize collective action [@lott_review_1992; @currie_evolution_2016; @richerson_tribal_2009].
+In natural resource management, the desired outcome of any particular institution varies on a case-by-case basis but may include goals like increasing the livelihood benefits produced from a resource or improving the biodiversity of a particular area [@agrawal_studying_2014].
+Regardless of the exact social or ecological outcome of interest, “success” of a conservation institution has become synonymous with the degree to which the particular behavioral prescription is adopted by a population and whether or not that behavior is sustained over time [@agrawal_common_2001].
+
+Decades of research have identified key contextual factors that support the durability of resource safeguarding behaviors in a population, particularly with regard to common-pool resources [@agrawal_common_2001; @baland_halting_1996; @ostrom_governing_1990].
+These factors are largely narrative descriptions of the qualities of, and interactions between, resource types, political structures, and resource users which are most likely to lead to stable and sustainable resource management institutions [@ostrom_general_2009].
+Although at least 40 unique factors have been identified as contributing to the success of community-based conservation institutions, well defined social and ecological boundaries, and well connected social systems stand out as major important themes [@ostrom_governing_1990; @ostrom_general_2009; @agrawal_common_2001].
+Research on common property additionally recommends that institutions be matched in scale to the underlying ecological processes of the resources they aim to manage [@mckean_common_2000].
+
+Many of the factors that support the long-term viability of conservation institutions are intertwined such that the applicability of one may depend on the level of another and some factors may be inextricably correlated [@agrawal_group_2001; @agrawal_sustainable_2003].
+Further, which specific factors are most important is highly specific to a given community or location, suggesting that there is not possibly one unifying panacea for successful conservation institutions [@brooks_how_2012; @geist_proximate_2002; @agrawal_communities_2001].
+Thus, even with decades of productive research on sustainable resource management, complete quantitative syntheses and insights have remained elusive.
+
+We argue here that considering the stability of institutions within a population from the perspective of individuals, rather than entire communities may provide valuable insight for implementing community-based conservation projects, both qualitatively and quantitatively.
+Sustainable resource management institutions are observed at the community level, but in reality they are an emergent property of individual level decision making — communities do not adopt behaviors, individuals do [@currie_cultural_2021].
+The normative conditions imposed on individuals by their ecological and social environments affect the likelihood that they will endogenously engage in cooperative resource management practices or adopt them when exogenously introduced by an intervention.
+A major question in implementing these projects is if a specific intervention will indeed be adopted by members of a population and if so, how quickly [@radeloff_hot_2013]?
 Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [@mangubhai_not_2011; @visconti_protected_2019].
+By examining sustainable resource management institutions as the emergent outcome of individual behavioral changes, we can leverage insights from other areas of social science to further guide theory on why some institutions emerge and others fail to be adopted long-term [@berkes_rethinking_2004].
 
-# Diffusion Theory
+In this article, we examine one social science theory, diffusion of innovations (DOI), and its previous and potential applications to community-based conservation institutions [@rogers_diffusion_2010].
+Our primary aim is to give a practical demonstration of how a focus on individual behavioral adoption can expand our inference about the success or failure of conservation interventions beyond narrative descriptions.
+We do this by first simulating data on conservation adoption in a resource-dependent population and then statistically fitting these data to a model of diffusion.
+Adoption of resource safeguarding behavior is fundamentally different from behavioral adoption in the abstract in that individual adoptions affect the natural environment and alter the payoff structure under which individuals continue to adopt in future time periods.
+Our simulation highlights this feedback of interest in order to answer the key question of whether DOI is applicable to community-based conservation specifically.
+Throughout this article, we situate DOI theory in the broader literature on cultural evolution and argue that this body of work may provide more actionable insight than DOI in isolation, particularly by considering underlying mechanisms of behavioral transmission.
+
+# Diffusion of Innovations
 Promoting the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time.
-Diffusion theory recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010].
+DOI recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010].
 When an individual is introduced to a new behavior, they often have a high degree of uncertainty as to the risks and benefits of adopting and will wait to adopt until that uncertainty is reduced to a tolerable level.
 The primary way individuals reduce uncertainty about the outcome of adopting a particular behavior is through second-hand observation of early adopters.
-Therefore, at the core of diffusion theory is the observation that individuals imitate those around them when making decisions about adopting a novel behavior or not, rather than weighing the pros and cons independently [@henrich_search_2001].
+Therefore, at the core of DOI is the observation that individuals imitate those around them when making decisions about adopting a novel behavior or not, rather than weighing the pros and cons independently [@henrich_search_2001].
 
@@
 While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information which is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016].
-This information affects individuals’ behaviors, which in-turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991].
-Thus, cultural evolution provides an intuitive way to scale from individual learning and decision making to population level patterns and even population level behavior-environment feedbacks.
+This information affects individuals’ behaviors, which in-turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003].
+Thus, cultural evolution provides an intuitive way to scale from individual decision-making to population level patterns and even population level behavior-environment feedbacks.
 
@@
 This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001].
-These two curve shapes however are not completely discrete, as the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].
+These two curve shapes however are not completely distinct, as the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].
 
 ![Canonical "S" and "R" diffusion curves.
-"R" curves are generally indicative of a higher proportion of individuals employing independent learning.
+"R" curves are generally indicative of a higher proportion of individuals employing individual learning.
 "S" curves are generally indicative of a higher proportion of individuals employing social learning.
@@
 
-# Diffusion Theory in Community-Based Conservation
+# Diffusion of Innovations in Community-Based Conservation
 Just as the sigmoidal pattern indicative of social learning is seen in the majority of cases on behavioral adoption broadly, scholars have also noticed this pattern in over 80% of case studies on the uptake of community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
-High-level studies have begun identifying diffusion theory as potentially useful for diagnosing why some community-based conservation interventions spread and persist, while others fizzle out.
-For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters.
-The authors point to a sigmoidal diffusion pattern to explain the slow start of many initiatives and suggest that practitioners may actively spread information on interventions (as in Abernethy et al.
+Fittingly, high-level studies have begun identifying DOI as potentially useful for diagnosing why some community-based conservation interventions spread and persist, while others fizzle out.
+For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters at the sigmoidal equilibrium.
+The authors take insights from DOI to suggest that practitioners may actively spread information on interventions (as in Abernethy et al.
 [-@abernethy_two_2014]), or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004].
-Further, a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices identifies qualitative insights, such as intervention observability, from diffusion theory as useful in guiding the development of conservation interventions [@mahajan_theorybased_2020].
+Further, a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices identifies qualitative insights, such as intervention observability, from DOI as useful in guiding the development of conservation interventions [@mahajan_theorybased_2020].
 
-As these high-level studies suggest, qualitative insights from diffusion theory have been productively applied to understand the rate and success of spread of conservation behaviors in single case studies around the world.
+As these high-level studies suggest, qualitative insights from DOI have been productively applied to understand the rate and success of spread of conservation behaviors in single case studies around the world.
 In Kenya for example, Mbaru and Barnes [-@mbaru_key_2017] show that identifying socially well-connected individuals can facilitate the rapid diffusion of information on conservation objectives.
-Similarly, in Fort Collins, CO, researchers demonstrate that both the speed and extent of diffusion of information on native plant gardening likely increase with after motivating communication in the population [@niemiec_efficacy-based_2021].
+Similarly, in Fort Collins, CO, researchers demonstrate that both the speed and extent of diffusion of information on native plant gardening likely increase when community members are offered incentives to increase communication [@niemiec_efficacy-based_2021].
 Romero-de-Diego et al.
-[-@romero-de-diego_drivers_2021] use expert interviews to identify that observability of benefits was a key driver of the rapid scaling-up of wildlife management areas in Mexico.
-Conversely, in contexts where communication and thus imitation are limited, such as across island chains, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
+[-@romero-de-diego_drivers_2021] use expert interviews to identify that observability of benefits was a key driver of the rapid adoption of a voluntary wildlife management program among landholders in Mexico.
+Conversely, in contexts where observability and thus imitation are limited, such as across island chains, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
 This finding mirrors those of studies which look at more general cultural adoption across island chains, further suggesting that conservation behaviors may spread similarly to technological or ideological innovations [@kline_population_2010; @beheim_inheritance_2011].
 
-These studies are compelling in their conclusion that diffusion theory can be productively applied to the spread of community-based conservation, but have been critiqued as qualitative suggestions may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020].
-Below, we make a quantitative argument for using diffusion theory to understand the spread of community-based conservation interventions.
+These studies are compelling in their conclusion that DOI can be productively applied to the spread of community-based conservation, but have been critiqued as general, narrative suggestions from DOI may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020].
+Below, we make a quantitative argument for using DOI to understand the spread of community-based conservation interventions.
 We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and what they can do to promote its spread.
-The aim of this review is to explicitly demonstrate how diffusion theory can be used to generate predictions about the patterns of uptake of conservation behaviors given realistic data and diagnose when in the lifetime of an intervention individuals may be dropping out.
+The aim of this review is to explicitly demonstrate how DOI can be used to generate predictions about the patterns of uptake of conservation behaviors given behavior-environment feedbacks and diagnose how an intervention may be best improved.
 
 # Quantitative applications
-In this section we explain how we can fit the pattern of diffusion mathematically to derive inference about the rate and degree to which an intervention will be adopted by a population.
+In this section we explain how we can fit the pattern of diffusion mathematically to estimate the rates of adoption and dropout, as well as to predict the speed and degree to which an intervention will be adopted by a population.
 The “S” shaped diffusion pattern of behavioral spread over time mirrors that of community disease transmission [@smaldino_coupled_2020].
@@
 For conservation interventions however, there are no natural history features of a specific intervention, such as the rate or duration of adoption, which can be known ahead of time.
-Instead, the qualitative insights from diffusion theory referenced above aim to increase the rate of spread and reduce the rate of dropout so that the intervention persists in the population and may even spread to new communities.
+Instead, the qualitative insights from DOI and the study of sustainable resource management institutions referenced above aim to increase the rate of spread and reduce the rate of dropout so that the intervention persists in the population and may even spread to new communities.
 While we can not know exactly what the rates of spread and dropout will be ahead of time, we can use Bayesian inference to probabilistically estimate them once the project has been implemented [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018].
-Knowing these parameters may allow us to predict the final rate of uptake from early in the project, identify whether adoption or dropout are the primary issue, and apply the qualitative insights from diffusion theory more targetedly.
+Knowing these parameters may allow us to predict the final rate of uptake from early in the project, identify whether adoption or dropout are limiting the overall rate of uptake, and apply the qualitative insights from DOI more targetedly.
 
-In some contexts, the uptake of community-based conservation is subject to behavior-environment feedbacks, a fundamental difference from behavioral adoption in the abstract [@liu_complexity_2007].
+In some contexts, the uptake of community-based conservation is subject to behavior-environment feedbacks, a fundamental difference from behavioral adoption generally [@liu_complexity_2007].
 In the case of marine protected areas for example, success in conserving fish populations may actually disincentivize individuals to continue supporting conservation programs [@christie_marine_2004].
@@
 # Agent-based simulation
-We developed a minimal agent-based simulation which we use to generate data on conservation adoption within a population.
-We use these data to test the performance of the SIRS diffusion model in fitting patterns of conservation behavior adoption under two social learning mechanisms.
-The first is unbiased, frequency-based, copying where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
-The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of more successful agents [@baldini_two_2013].
-This mechanism is formally demonstrated in equation 1, for two behaviors: $A$ & $B$, where the probability that an individual adopts behavior $A$ is a function of the mean pay-off, $\bar{P}$, of both behaviors at each time step and the strength of the bias, $\beta$ [@barrett_equifinality_2019].
+We seek to test whether the SIRS model shown above can produce reliable insights about community uptake of a conservation behavior, especially given behavior-environment feedbacks common in natural resource management institutions.
+To obtain data which explicitly incorporate this dynamic, we develop a minimal agent-based model which we use to simulate data on the spread of a resource safeguarding behavior within a population.
+For comparison, we simulate adoption data both independent of, and contingent on feedbacks with the environment.
+These two dynamics are operationalized using two different social learning mechanisms for behavioral transmission.
+The first is unbiased, frequency-based, copying where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the harvest benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
+We include a simulation using this mechanism as a baseline for how behavior-environment feedbacks inherent to our second mechanism affect the overall pattern of adoption.
+The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of other agents with the greatest harvesting success [@baldini_two_2013].
+This mechanism is formally demonstrated in equation 1, for two behaviors: $A$ & $B$, where the probability that an individual adopts behavior $A$ is a function of the mean harvest, $\bar{P}$, from behaviors at each time step and the strength of the bias, $\beta$ [@barrett_equifinality_2019].
+Under biased social learning, the strength of the learning biases (e.g.
+$\beta$) control how disproportionately agents copy others based on a given condition (e.g.
+success); a $\beta$ = 1 in this formulation would thus mean that the probability that one agent copies another is directly proportional to the ratio of their harvests.
 
@@
 
-Our simulated system emulates the common performance payments or payments for ecosystem services approach to conservation [@ferraro_future_2011; @rakotonarivo_role_2021].
-We establish a degraded landscape/seascape (hereafter referred to as landscape) where all simulated individuals are dependent on harvesting some renewable, mobile resource.
-In the first time step, we introduce a conservation intervention which delineates 20% of the landscape as a conservation area and is initially adopted by just 2% of the population.
-In each subsequent time step, agents enrolled in the conservation program receive a reward, and continue to harvest just from the remaining 80% of the unconserved landscape.
-Agents not enrolled in the intervention harvest from the entire landscape and forgo the reward.
-After each harvest period, agents are randomly paired, and employ either unbiased or success-biased learning to decide whether or not to adopt the conservation behavior of their partner in the next time step [@baldini_success-biased_2012].
+Our simulated system is conceptualized as a fishery where agents are directly compensated for forgoing resource extraction from a particular area if they choose [@rakotonarivo_role_2021].
+We establish a fishery, loosely based on a global average stock integrity, at 20% of its potential carrying capacity [@fao_state_2020].
+In the first time step, we introduce an intervention which delineates 20% of the fishery as a conservation area and just 2% of agents initially change their behavior to forgo harvesting there.
+Fish reproduce proportionally to the population in each time period, and move freely between the conservation area and unconserved area.
+In each subsequent time step, agents who adopted the resource safeguarding behavior (enrolled) receive a reward, and continue to harvest just from the remaining 80% of the unconserved fishery.
+Agents not enrolled in the intervention harvest from the entire fishery and forgo the reward.
+After each harvest period, agents are randomly paired, and employ either frequency-based or success-biased learning to decide whether or not to adopt the behavior of their partner in the next time step [@baldini_success-biased_2012].
 The conserved and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
-As behavioral adoption under success-biased learning is dependent on harvest success, the success of each behavior is determined largely by the amount of available resources in each area, and the amount of resources in each area is a result of behavioral adoption, this learning strategy is contingent on behavior-environment feedbacks, unlike unbiased adoption which occurs independent of harvest success.
+A visual description of this minimal model is shown in figure 2 andInterested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
 
@@
 
-A visual description of this minimal model is shown in figure 2, and the full ODD+ model description can be found in the supplemental material (S1).
-Interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
-As this simulation is not the emphasis of this study, we show just two runs below.
-The starting conditions of these runs are identical except for the social learning mechanism employed by the population.
-As the unbiased learning mechanism is independent of the resource levels, the adoption curve follows the traditional “S” pattern and nears fixation in the long-term.
-Under environmentally mediated success-biased learning, the adoption pattern spikes initially, but drops off dramatically as the resources in the unconserved landscape are reduced following increased harvest pressure there as a result of the conservation intervention.
-The full output of this latter simulation is shown below in figure 3.
+We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, frequency-based and success biased.
+Aside from the learning mechanism employed by the population, all model conditions are identical for these two scenarios.
+Frequency-based transmission is used as a reference as it reliably produces the standard sigmoidal “S” curve of behavioral adoption and does not incorporate any feedbacks between adoption and the environment.
+Under the success-biased learning scenario, agents compare their harvests to others in each time step and disproportionately copy more successful agents.
+In this instantiation then, adoption is dependent on harvest success and harvest success is determined largely by the amount of available resources in each area.
+As the amount of available resources in an area is determined by the number of agents who harvest there, this learning mechanism incorporates behavior-environment feedbacks typical of natural resource management institutions.
+Because agents compare their own harvests to those of others in each time step under the success-biased learning scenario, rather than exclusively imitating other individuals, this scenario actually represents a mix of individual and social learning.
+The adoption pattern observed in this simulation results in a mix between an “R” and “S” type curve, reinforcing the interpretation of this scenario as a mix of individual and social learning [@henrich_cultural_2001].
+Figure3 shows the full output of the later simulation where adoption spikes initially, but drops off dramatically as the resources in the unconserved area are reduced following increased harvest pressure there as a result of the conservation intervention.
 
@@
 # Predicting adoption of conservation behaviors
-
 With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described above to the first 30% of observations.
@@
 
-For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an $\hat{R}$ of 1 for all estimated parameters (fig 3).
-As the parameters $\beta$, $\gamma$, and $\chi$ are directly interpretable as the rates of adoption, dropout, and willingness to retry a previously abandoned intervention, accurate estimation of these values alone, especially in the early stages of an intervention, is a practical and straightforward application of diffusion theory to community-based conservation interventions.
+For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an $\hat{R}$ of 1 for all estimated parameters (fig 4).
+As the parameters $\beta$, $\gamma$, and $\chi$ are directly interpretable as the rates of adoption, dropout, and willingness to retry a previously abandoned intervention, accurate estimation of these values, especially in the early stages of an intervention, provides practical and straightforward intuition to diagnose areas for improvement in community-based conservation interventions.
 For example, a practitioner might fit this model to real-world participation rates during the early weeks of an intervention in order to estimate whether they should focus more on increasing rates of uptake or retention going forward.
+In addition, this insight may help practitioners better decide which factors from the institutional design literature referenced above may be most impactful for their specific system.
 
@@
 We can also use these parameter values, estimated from the first 30% of the time series, to predict the long-term patterns of adoption for a community-based conservation initiative.
-In figures 4 & 5, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step.
+In figures 5 & 6, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step.
 Grey and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data respectively.
 These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning.
-R and STAN code for these models can be found in the supplemental material (S4).
+R and STAN code for these models can be found in the supplemental material (S3).
 
@@
 
-Figures 4 & 5 also demonstrate that different learning mechanisms, while both producing a sigmoidal adoption pattern, can result in different overall rates in adoption of an intervention, even under identical starting conditions.
+Figures 5 & 6 also demonstrate that different scenarios and respective learning mechanisms, while both producing fixation in the adoption rate, can result in different overall rates in adoption of an intervention, even under identical starting conditions.
 In the simulation of a population employing unbiased learning, the population equilibrates at nearly 90% adoption of the intervention behavior.
 For success-biased learning on the other hand, resource scarcity in the unconserved area caused the population to adopt the intervention at approximately 60% in the long-term.
+Lastly, the mix of social and independent learning employed by agents in the success-biased learning scenario produces a somewhat “R” shaped curve as explained in the section above.
+These results show that the SIRS model of diffusion can reliably fit the pattern of behavioral adoption even when the pattern does not follow a strict sigmoidal “S” shape.
 
 # Shortcomings for application of diffusion theory to community-based conservation
-
 As demonstrated above, the standard SIRS diffusion model can reliably capture the long-term patterns of behavioral adoption for community-based conservation interventions, even when adoption is contingent on feedbacks with ecological processes.

### 2021-12-19 DOI_Review/Manuscript.rmd
File path change 4 new comp

@@
 abstract: A common goal of conservation practitioners is to design interventions that benefit both people and the environment, and as a result are adopted robustly by local communities.
-Diffusion of innovations theory (DOI) provides generalized heuristics and models for how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
+Diffusion of innovations theory (DOI) provides generalized heuristics and models that represent how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
 Here, we elaborate on the need for better understanding conservation institutions as a product of individual behavioral adoption and review DOI and its past application to these problems.
@@
 
-setwd("C:/Users/Matthewclark989/Documents/Pemba_Project/DOI_Review")
+setwd("C:/Users/Matt/Documents/Pemba_Project/DOI_Review")
 ```
@@
 It is becoming increasingly apparent that humanity will not reach the global conservation goals we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002].
-We are thus turning to local communities to self-impose and maintain sustainable limits on resource extraction and even to restore degraded natural areas themselves [@amel_beyond_2017; @catalano_learning_2019].
+We are thus turning to local communities to self-impose and maintain sustainable limits on resource extraction and even to restore degraded natural areas [@amel_beyond_2017; @catalano_learning_2019].
 Community-based conservation projects aim to incentivize these behaviors in rural, and often very poor, communities with little to no external oversight [@western_natural_1994].
 While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote behavioral changes which are adopted robustly by communities who harvest natural resources to meet their daily needs [@brooks_assessing_2013; @horwich_community_2007].
-While some debate still exists, it is largely accepted that empowering communities to self-run conservation institutions and manage environments for multiple interests is a viable method for seeding self-sustaining and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004; @agrawal_enchantment_1999].
+While some debate still exists, it is largely accepted that empowering communities to self-run conservation institutions and manage environments for multiple interests is a viable method for seeding self-sustaining and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004; @agrawal_enchantment_1999; @yates_purpose_2019].
 
@@
 In natural resource management, the desired outcome of any particular institution varies on a case-by-case basis but may include goals like increasing the livelihood benefits produced from a resource or improving the biodiversity of a particular area [@agrawal_studying_2014].
-Regardless of the exact social or ecological outcome of interest, “success” of a conservation institution has become synonymous with the degree to which the particular behavioral prescription is adopted by a population and whether or not that behavior is sustained over time [@agrawal_common_2001].
+Regardless of the exact social or ecological outcome of interest, “success” of a conservation institution has become synonymous with the degree to which the particular behavioral prescription is adopted by a population and whether or not that behavior is sustained over time [@agrawal_common_2001; @yates_purpose_2019].
 
@@
 Sustainable resource management institutions are observed at the community level, but in reality they are an emergent property of individual level decision making — communities do not adopt behaviors, individuals do [@currie_cultural_2021].
-The normative conditions imposed on individuals by their ecological and social environments affect the likelihood that they will endogenously engage in cooperative resource management practices or adopt them when exogenously introduced by an intervention.
+The normative conditions imposed on individuals by their ecological and social environments (i.e.
+the factors referenced above) affect the likelihood that they will endogenously engage in cooperative resource management practices or adopt them when exogenously introduced by an intervention.
+
 A major question in implementing these projects is if a specific intervention will indeed be adopted by members of a population and if so, how quickly [@radeloff_hot_2013]?
@@
 
-In this article, we examine one social science theory, diffusion of innovations (DOI), and its previous and potential applications to community-based conservation institutions [@rogers_diffusion_2010].
-Our primary aim is to give a practical demonstration of how a focus on individual behavioral adoption can expand our inference about the success or failure of conservation interventions beyond narrative descriptions.
-We do this by first simulating data on conservation adoption in a resource-dependent population and then statistically fitting these data to a model of diffusion.
+In this article, we examine one social science theory specifically focused on individual behavior adoption, diffusion of innovations (DOI), and its previous and potential applications to community-based conservation institutions [@rogers_diffusion_2010].
+Our primary aim is to give a practical demonstration of how insights from DOI can be applied to expand our inference about the spread and persistence of conservation interventions beyond narrative descriptions.
+We do this by first simulating data on conservation adoption in a resource-dependent population and then statistically fitting a model of diffusion to these data.
 Adoption of resource safeguarding behavior is fundamentally different from behavioral adoption in the abstract in that individual adoptions affect the natural environment and alter the payoff structure under which individuals continue to adopt in future time periods.
-Our simulation highlights this feedback of interest in order to answer the key question of whether DOI is applicable to community-based conservation specifically.
+Our simulation highlights this feedback of interest in order to answer the key question of whether DOI is applicable to community-based conservation specifically, given this unique adoption dynamic.
 Throughout this article, we situate DOI theory in the broader literature on cultural evolution and argue that this body of work may provide more actionable insight than DOI in isolation, particularly by considering underlying mechanisms of behavioral transmission.
 
-# Diffusion of Innovations
+# Diffusion of innovations
 Promoting the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time.
@@
 The primary way individuals reduce uncertainty about the outcome of adopting a particular behavior is through second-hand observation of early adopters.
-Therefore, at the core of DOI is the observation that individuals imitate those around them when making decisions about adopting a novel behavior or not, rather than weighing the pros and cons independently [@henrich_search_2001].
+Therefore, at the core of DOI is the observation that individuals imitate those around them when making decisions about adopting a novel behavior, rather than weighing the pros and cons independently [@henrich_search_2001].
 
-When modifying their beliefs or behaviors, people do not imitate any other individuals at random [@bandura_social_1977].
-Instead, they preferentially adopt traits by employing a variety of learning biases such as success-biased learning where individuals who are perceived as more successful are more likely to be copied than unsuccessful individuals [@boyd_culture_1988; @laland_social_2004].
+When modifying their beliefs or behaviors, people do not imitate other individuals at random [@bandura_social_1977].
+Instead, they preferentially adopt traits by employing a variety of learning biases, such as success-biased learning, where individuals who are perceived as more successful are disproportionately likely to be copied [@boyd_culture_1988; @laland_social_2004].
 As a result, the cumulative culture of a population evolves over time as various beliefs and behaviors are unequally copied and their frequency within a population changes.
 While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information which is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016].
-This information affects individuals’ behaviors, which in-turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003].
-Thus, cultural evolution provides an intuitive way to scale from individual decision-making to population level patterns and even population level behavior-environment feedbacks.
+This information affects individuals’ behaviors, which in turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003].
+Thus, cultural evolution provides an intuitive way to integrate  individual decision-making, population level patterns, and even population level behavior-environment feedbacks.
 
 Diffusion of novel beliefs and behaviors is a specific instance of cultural evolution, where the trait being transmitted is new to the population and outcompetes some previous cultural variant, such as in the classic example of hybrid corn seed adoption in Iowa farmers [@mesoudi_cultural_2016; @ryan_acceptance_1950].
-We observe these patterns through diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
+We observe these patterns using diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
 The exact shape of these diffusion curves is highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010].
@@
 This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001].
-These two curve shapes however are not completely distinct, as the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].
+These two curve shapes however are not completely distinct.
+As the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].
 
@@
 
-# Diffusion of Innovations in Community-Based Conservation
+# Diffusion of innovations in community-based conservation
 Just as the sigmoidal pattern indicative of social learning is seen in the majority of cases on behavioral adoption broadly, scholars have also noticed this pattern in over 80% of case studies on the uptake of community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
@@
 As these high-level studies suggest, qualitative insights from DOI have been productively applied to understand the rate and success of spread of conservation behaviors in single case studies around the world.
-In Kenya for example, Mbaru and Barnes [-@mbaru_key_2017] show that identifying socially well-connected individuals can facilitate the rapid diffusion of information on conservation objectives.
-Similarly, in Fort Collins, CO, researchers demonstrate that both the speed and extent of diffusion of information on native plant gardening likely increase when community members are offered incentives to increase communication [@niemiec_efficacy-based_2021].
+In Kenya for example, Mbaru and Barnes [-@mbaru_key_2017] showed that identifying socially well-connected individuals can facilitate the rapid diffusion of information on conservation objectives.
+Similarly, in Fort Collins, CO, researchers demonstrated that both the speed and extent of diffusion of information on native plant gardening likely increase when community members are offered incentives to increase communication [@niemiec_efficacy-based_2021].
 Romero-de-Diego et al.
-[-@romero-de-diego_drivers_2021] use expert interviews to identify that observability of benefits was a key driver of the rapid adoption of a voluntary wildlife management program among landholders in Mexico.
+[-@romero-de-diego_drivers_2021] used expert interviews to identify that observability of benefits was a key driver of the rapid adoption of a voluntary wildlife management program among landholders in Mexico.
 Conversely, in contexts where observability and thus imitation are limited, such as across island chains, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
@@
 
-These studies are compelling in their conclusion that DOI can be productively applied to the spread of community-based conservation, but have been critiqued as general, narrative suggestions from DOI may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020].
+These studies are compelling in their conclusion that DOI can be productively applied to the spread of community-based conservation, but have been critiqued as general, narrative suggestions from DOI that may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020].
 Below, we make a quantitative argument for using DOI to understand the spread of community-based conservation interventions.
 We envision a practitioner who is monitoring the weekly, monthly, or even yearly uptake rates of an intervention and is hoping to predict if and when it will “take off” in the population, what the final proportion of adopters will be, and what they can do to promote its spread.
-The aim of this review is to explicitly demonstrate how DOI can be used to generate predictions about the patterns of uptake of conservation behaviors given behavior-environment feedbacks and diagnose how an intervention may be best improved.
+In this review, we therefore explicitly demonstrate how DOI can be used to generate predictions about the patterns of uptake of conservation behaviors given behavior-environment feedbacks and diagnose how an intervention may be best improved.
 
 # Quantitative applications
-In this section we explain how we can fit the pattern of diffusion mathematically to estimate the rates of adoption and dropout, as well as to predict the speed and degree to which an intervention will be adopted by a population.
+In this section we explain how we can fit an observed pattern of diffusion mathematically to estimate the rates of adoption and dropout, as well as to predict the speed and degree to which an intervention will be adopted by a population.
 The “S” shaped diffusion pattern of behavioral spread over time mirrors that of community disease transmission [@smaldino_coupled_2020].
@@
 
-\begin{align*}
-&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \\
+\begin{align}
+&\frac{dS}{dt} = - \beta \cdot \text{S}_{t} \cdot \text{I}_{t} + \chi \cdot \text{R}_{t} \nonumber\\
 &\frac{dI}{dt} = \beta \cdot \text{S}_{t} \cdot \text{I}_{t} - \gamma \cdot \text{I}_{t} \\
-&\frac{dR}{dt} = \gamma\cdot \text{I}_{t} - \chi \cdot \text{R}_{t}
-\end{align*}
+&\frac{dR}{dt} = \gamma\cdot \text{I}_{t} - \chi \cdot \text{R}_{t}\nonumber
+\end{align}
 
@@
 While we can not know exactly what the rates of spread and dropout will be ahead of time, we can use Bayesian inference to probabilistically estimate them once the project has been implemented [@margossian_review_2019; @margossian_differential_2017; @carpenter_predator-prey_2018].
-Knowing these parameters may allow us to predict the final rate of uptake from early in the project, identify whether adoption or dropout are limiting the overall rate of uptake, and apply the qualitative insights from DOI more targetedly.
+Knowing these parameters may allow us to predict the final rate of uptake from early in the project, identify whether adoption or dropout are limiting the overall rate of uptake, and apply the qualitative insights from DOI and natural resource management institutions in a more targeted fashion.
 
-In some contexts, the uptake of community-based conservation is subject to behavior-environment feedbacks, a fundamental difference from behavioral adoption generally [@liu_complexity_2007].
+The uptake of community-based conservation is subject to behavior-environment feedbacks, a fundamental difference from behavioral adoption generally [@liu_complexity_2007].
 In the case of marine protected areas for example, success in conserving fish populations may actually disincentivize individuals to continue supporting conservation programs [@christie_marine_2004].
@@
 To answer this question, we use an agent-based simulation to generate data on the spread of a conservation behavior via social learning, contingent on feedbacks with the environment.
-Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using a Bayesian instantiation of the SIRS model shown above.
+Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using a Bayesian instantiation of the SIRS model shown in equation 1.
 We use these estimated parameters to predict the long-term patterns of adoption for the remaining 70% of time steps and compare our predicted values to those observed in the simulation.
@@
 # Agent-based simulation
-We seek to test whether the SIRS model shown above can produce reliable insights about community uptake of a conservation behavior, especially given behavior-environment feedbacks common in natural resource management institutions.
-To obtain data which explicitly incorporate this dynamic, we develop a minimal agent-based model which we use to simulate data on the spread of a resource safeguarding behavior within a population.
+We seek to test whether the SIRS model shown in equation 1 can produce reliable insights about community uptake of a conservation behavior, especially given behavior-environment feedbacks common in natural resource management institutions.
+To answer this question, we first need complete time series data in-hand which are a product of our dynamics of interest.
+To obtain these data, we develop a minimal agent-based model which we use to simulate data on the spread of a resource safeguarding behavior within a population.
 For comparison, we simulate adoption data both independent of, and contingent on feedbacks with the environment.
-These two dynamics are operationalized using two different social learning mechanisms for behavioral transmission.
-The first is unbiased, frequency-based, copying where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the harvest benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
+These two dynamics are operationalized using two different social learning mechanisms for behavioral transmission [@kendal_social_2018].
+The first is frequency-dependent copying where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the harvest benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
 We include a simulation using this mechanism as a baseline for how behavior-environment feedbacks inherent to our second mechanism affect the overall pattern of adoption.
 The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of other agents with the greatest harvesting success [@baldini_two_2013].
-This mechanism is formally demonstrated in equation 1, for two behaviors: $A$ & $B$, where the probability that an individual adopts behavior $A$ is a function of the mean harvest, $\bar{P}$, from behaviors at each time step and the strength of the bias, $\beta$ [@barrett_equifinality_2019].
+This mechanism is formally demonstrated in equation 2, for two behaviors: $A$ & $B$, where the probability that an individual adopts behavior $A$ is a function of the mean harvest, $\bar{P}$, from each behavior at each time step and the strength of the bias, $\beta$ [@barrett_equifinality_2019].
 Under biased social learning, the strength of the learning biases (e.g.
@@
 
-$$ \text{Pr}(A|A,B)_{t+1} = \frac{\text{exp}\beta\bar{P}_{A,t}}{\text{exp}\beta\bar{P}_{A,t} + \text{exp}\beta\bar{P}_{B,t}} $$
+\begin{align}
+\text{Pr}(A|A,B)_{t+1} = \frac{\text{exp}\beta\bar{P}_{A,t}}{\text{exp}\beta\bar{P}_{A,t} + \text{exp}\beta\bar{P}_{B,t}} 
+\end{align}
 
@@
 Agents not enrolled in the intervention harvest from the entire fishery and forgo the reward.
-After each harvest period, agents are randomly paired, and employ either frequency-based or success-biased learning to decide whether or not to adopt the behavior of their partner in the next time step [@baldini_success-biased_2012].
+After each harvest period, agents are randomly paired, and employ either frequency-dependent or success-biased learning to decide whether or not to adopt the behavior of their partner in the next time step [@baldini_success-biased_2012].
 The conserved and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
-A visual description of this minimal model is shown in figure 2 andInterested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
+A visual description of this minimal model is shown in figure 2 and interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
 
-![Conceptual diagram showing process at each timestep in the agnt-based simulation.
+![Conceptual diagram showing process at each time step in the agnt-based simulation.
 This simulation was used to produce data on the adoption of conservation behaviors as a function of social learning and environmental outcomes.](./ConceptDiagram.png)
 
-We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, frequency-based and success biased.
+We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, frequency-dependent and success biased.
 Aside from the learning mechanism employed by the population, all model conditions are identical for these two scenarios.
-Frequency-based transmission is used as a reference as it reliably produces the standard sigmoidal “S” curve of behavioral adoption and does not incorporate any feedbacks between adoption and the environment.
+Frequency-dependent transmission is used as a reference as it reliably produces the standard sigmoidal “S” curve of behavioral adoption and does not incorporate any feedbacks between adoption and the environment.
 Under the success-biased learning scenario, agents compare their harvests to others in each time step and disproportionately copy more successful agents.
@@
 The adoption pattern observed in this simulation results in a mix between an “R” and “S” type curve, reinforcing the interpretation of this scenario as a mix of individual and social learning [@henrich_cultural_2001].
-Figure3 shows the full output of the later simulation where adoption spikes initially, but drops off dramatically as the resources in the unconserved area are reduced following increased harvest pressure there as a result of the conservation intervention.
+Figure 3 shows the full output of the later simulation where adoption (enrollment) spikes initially, but drops off dramatically as the resources in the unconserved area are reduced following increased harvest pressure there as a result of the conservation intervention.
 
-![One run of the agent based simulation used to produce data for analysis with the process -based SIRS model of behavioral diffusion.
-Starting conditions for this simulation were: 100 individuals, total landscape, carrying capacity of 10000 resources, 20% starting resource level, individual payment amount of 9 resources for enrollment in the conservation intervention, 2% initial enrollment, success-biased learning strategy at a strength of 1.1, 30% probability of resource mobility when carrying capacity reached, a maximum harvest value of 15, and 30% resource regeneration rate per time step.](./SiccessBiasedFullRun.png)
+![One run of the agent-based simulation used to produce data for analysis with the SIRS model of behavioral adoption under the success-biased learning scenario.
+The top figure shows the number of resources in both the conservation and unconserved areas over time, as a percent of the carrying capacity in those areas.
+The middle figure shows the average number of resources harvested (payoff) by each agent at each time step.
+The bottom figure shows the number of agents who have adopted the conservation behavior (enrolled) at every time step.](./SiccessBiasedFullRun.png)
 
 # Predicting adoption of conservation behaviors
-With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described above to the first 30% of observations.
-We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time.
+With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described in equation 1 to the first 30% of observations.
+We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation (3) below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time.
 Here, "infected" individuals are those that are enrolled in the conservation intervention.
@@
 
-![Markov chains for the three parameters of interest from the SIRS model fit to simulated data.
-The agent-based simulation used to produce the data employed environmentally mediated success-biased learning.
+![Markov chains for each of the three parameters of interest from the SIRS model fit to simulated data.
+Beta is interpretable as the rate of adoption of the behavior (enrollment).
+Gamma is an estimate of the rate of dropout.
+Chi represents the rate at which individuals who previously dropped out are added back into the susceptible category.
+The agent-based simulation used to produce the data employed environmentally mediated, success-biased learning.
 A figure showing the Markov chains for the unbiased learning simulation can be found in the supplemental material (S3).](./SuccessPars.png)
@@
 In figures 5 & 6, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step.
-Grey and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data respectively.
+Gray and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data respectively.
 These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning.
@@
 ![SIRS model fit of agent-based simulation data under unbiased social learning shown as translucent black lines representing 1000 draws from the posterior.
-Simulated data are overlaid with grey points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./DirectCopying.png)
+Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./DirectCopying.png)
 
 ![SIRS model fit of agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior.
-Simulated data are overlaid with grey points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./SuccessBiasedCopying.png)
+Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./SuccessBiasedCopying.png)
 
 Figures 5 & 6 also demonstrate that different scenarios and respective learning mechanisms, while both producing fixation in the adoption rate, can result in different overall rates in adoption of an intervention, even under identical starting conditions.
-In the simulation of a population employing unbiased learning, the population equilibrates at nearly 90% adoption of the intervention behavior.
+In the simulation of a population employing frequency-dependent learning, the population equilibrates at nearly 90% adoption of the intervention behavior.
 For success-biased learning on the other hand, resource scarcity in the unconserved area caused the population to adopt the intervention at approximately 60% in the long-term.
@@
 
-# Shortcomings for application of diffusion theory to community-based conservation
-As demonstrated above, the standard SIRS diffusion model can reliably capture the long-term patterns of behavioral adoption for community-based conservation interventions, even when adoption is contingent on feedbacks with ecological processes.
-This model, while accurately capturing the pattern of spread, is unable to provide insight as to the mechanism of spread or causality in the system.
-As shown above and in existing literature, the social transmission mechanism, or mix of mechanisms, employed by a population can dramatically affect the overall rates of adoption for a given behavior [@smaldino_adoption_2017; @andrews_cultural_2018; @tam_measuring_2021].
-Therefore, while we agree that diffusion theory can be productively applied to community-based conservation, we point to theories of social learning and cultural evolution as a better alternative for diagnosing and planning interventions.
+# Shortcomings for application of diffusion of innovations to community-based conservation
+As demonstrated in figures 5 & 6, the standard SIRS diffusion model can reliably capture the long-term patterns of behavioral adoption for community-based conservation interventions, even when adoption is contingent on feedbacks with ecological processes.
+Fitting this model to observed data on the uptake of a behavior can provide us with estimates of the rates of adoption, dropout, and readoption for a specific intervention and allow us to predict the final stable rate of adoption in the population from early in the project.
+These insights, while useful, are somewhat limited in their ability to prescribe improvements for an intervention.
+As highlighted by the quote in the beginning of this article, diffusion study can explain what is happening in a system, but gives little intuition as to the mechanisms behind why it is happening [@rogers_diffusion_2010].
+
+As first described by Henrich [-@henrich_cultural_2001] and demonstrated in figures 5 & 6, we see clear changes in the population patterns of adoption when individuals employ varying levels of social or independent learning.
+But when comparing many biased social learning mechanisms, the pattern of adoption may exhibit equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [@barrett_equifinality_2019].
+Critically, this makes it virtually impossible to infer the specific biased social learning mechanism(s) used by a population from the pattern of adoption over time alone.
+This information is however important for researchers and practitioners, as the social transmission mechanism, or mix of mechanisms, employed by a population can dramatically affect the overall rates of adoption for a given behavior [@smaldino_adoption_2017; @andrews_cultural_2018; @tam_measuring_2021].
+Plainly, DOI can tell us if an intervention is being robustly adopted or not, but we need to understand the actual transmission mechanism to effectively act on this information.
+Thankfully, this information is possible to collect.
 For example, Tam et al.
@@
 
+With information on the patterns and mechanisms of uptake of community-based conservation behaviors we can more selectively apply the many narrative insights from DOI and the study of natural resource management institutions.
+In populations primarily employing prestige-biased learning (where prestigious individuals are disproportionately copied) for instance, the identification of and collaboration with community change agents may be particularly effective [@jimenez_prestige-biased_2019; @henrich_evolution_2001; @dietz_drama_2002].
+Conversely, the observability of both resources and the benefits of conservation are likely of increased importance when the population primarily employs content-biased learning (where some characteristic of the trait itself makes it more desirable than a previous trait)  [@ostrom_governing_1990; @rogers_diffusion_2010; @boyd_culture_1988-1].
+
+# Future work
 Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [@cheng_strengthen_2020].
-Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013; @turner_land-useland-cover_2021].
+Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013; @turner_land-useland-cover_2021; @le_polain_de_waroux_role_2021].
 Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [@schill_more_2019].
 
-The practice of applying statistical frameworks for parameter estimation to analytical process models, and in particular to ordinary differential equations, is still relatively new and has been embraced mostly in ecology [@hefley_when_2017; @laubmeier_ecological_2020].
+Mechanistic, or process models allow for us to formally describe these complex interactions while testing the logical consequences of our assumptions and the relation of our hypothesized processes to observed data [@epstein_why_2008].
+The practice of applying statistical frameworks for parameter estimation to analytical process models, and in particular to ordinary differential equations, is however still relatively new and has been embraced mostly in ecology [@hefley_when_2017; @laubmeier_ecological_2020; @morales_mechanistic_2021].
 This emerging methodology allows for the estimation of interesting and directly interpretable process parameters, not possible with linear regression.
@@
 &\frac{dR}{dt} = \gamma \cdot \text{I}_{t} - \chi \cdot \text{R}_{t} + \phi \cdot \text{E}_{t} \nonumber\\
-&\frac{dE}{dt} = - \rho \cdot (\text{S}_{t} + \text{R}_{t}) + \theta \cdot \text{E}_{t}\nonumber
+&\frac{dE}{dt} = - \rho \cdot (\text{S}_{t} + \text{R}_{t}) + \theta\cdot\text{E}_{t}\nonumber
 \end{align}
@@
 Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically [@liu_complexity_2007].
-In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems which are difficult or even impossible to account for with process models such the SIRS or SIRS extension models presented above [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
+In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems which are difficult or even impossible to account for with process models such the SIRS  or SIRS extension model presented in equation 4 [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
+Agent-based simulations allow modelers to impose a realism in the governance or institutional dynamics that are largely intractable in closed-form mathematical equations [@bravo_agents_2011; joffre_combining_2015].
+As the number of purely subsistence communities shrinks each year worldwide, and multiple levels of government are increasingly involved in natural resource management, incorporating these features into our models is paramount for reliable inference [@mckean_common_2000; @armitage_adaptive_2009].
 
 # Conclusions
-
-Overall, we argue that fitting process models of behavioral diffusion to data on the spread of conservation behaviors is a practical way for researchers and practitioners to apply diffusion theory to community-based conservation.
-We show that these models can be powerful for estimating the natural history characteristics of a particular intervention, such as the rate of spread, and for predicting the long-term patterns of adoption.
-We also introduce cultural evolutionary research as a better set of theories for diagnosing the spread, or lack thereof, in such interventions and for planning future interventions.
-Lastly, we describe conservation as a complex adaptive system and propose simulation modeling as an alternative to process-based models when they may be analytically intractable or ill suited for system dynamics.
+Overall, we argue that fitting process models of behavioral diffusion to data on the spread of conservation behaviors is a practical way for researchers and practitioners to apply DOI to community-based conservation.
+We test the utility of the SIRS model of diffusion on simulated data which highlight the feedbacks between behavioral adoption and environmental change commonly seen in community-based conservation interventions.
+We show that this model can be a powerful tool for estimating the natural history characteristics of a particular intervention, such as the rate of spread, and for predicting the long-term patterns of adoption, even under reciprocal  behavior-environment interactions.
+We also introduce the broader field of cultural evolution as a more comprehensive set of theories for understanding the underlying mechanisms of behavioral adoption than DOI alone.
+We hope that this review provides a model for how practitioners might practically apply insights from DOI to community-based conservation interventions, while also clarifying the position of DOI in the current literature on the adoption of sustainable resource management practices.
 

### 2021-12-19 DOI_Review/Manuscript.rmd
Kelly edits & added Jeff

@@
   name: Matt Clark
-- affiliation: Boise State University- Human Environment Systems
+- affiliation: Boise State University - Human Environment Systems
   name: Vicken Hillis
+- affiliation: Max Planck Institute for Evolutionary Anthropology 
+  name: Jeffrey Andrews
 
@@
 Decades of research have identified key contextual factors that support the durability of resource safeguarding behaviors in a population, particularly with regard to common-pool resources [@agrawal_common_2001; @baland_halting_1996; @ostrom_governing_1990].
-These factors are largely narrative descriptions of the qualities of, and interactions between, resource types, political structures, and resource users which are most likely to lead to stable and sustainable resource management institutions [@ostrom_general_2009].
-Although at least 40 unique factors have been identified as contributing to the success of community-based conservation institutions, well defined social and ecological boundaries, and well connected social systems stand out as major important themes [@ostrom_governing_1990; @ostrom_general_2009; @agrawal_common_2001].
+These factors are largely narrative descriptions of the qualities of, and interactions between, resource types, governance structures, and resource users which are most likely to lead to stable and sustainable resource management institutions [@ostrom_general_2009].
+Although at least 40 unique factors have been identified as contributing to the success of community-based conservation institutions, well defined social and ecological boundaries, and well connected social systems stand out as major important themes [@ostrom_governing_1990; @ostrom_general_2009; @agrawal_common_2001; @cox_review_2010].
 Research on common property additionally recommends that institutions be matched in scale to the underlying ecological processes of the resources they aim to manage [@mckean_common_2000].
@@
 We observe these patterns using diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
-The exact shape of these diffusion curves is highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010].
+The exact shape of these diffusion curves can be highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010].
 
@@
 In figures 5 & 6, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step.
-Gray and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data respectively.
+Gray and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data, respectively.
 These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning.
@@
 An example of such an extension is presented below, where we include another equation for the change in a resource pool over time $\frac{dE}{dt}$.
-We also must include three additional parameters: $\phi$, $\rho$, and $\theta$, which represent the effect of resource level on enrollment (infection), the effect of harvest on the resource level, and the regeneration rate of the resource respectively.
+We also must include three additional parameters: $\phi$, $\rho$, and $\theta$, which represent the effect of resource level on enrollment (infection), the effect of harvest on the resource level, and the regeneration rate of the resource, respectively.
 
@@
 Agent-based simulations allow modelers to impose a realism in the governance or institutional dynamics that are largely intractable in closed-form mathematical equations [@bravo_agents_2011; joffre_combining_2015].
-As the number of purely subsistence communities shrinks each year worldwide, and multiple levels of government are increasingly involved in natural resource management, incorporating these features into our models is paramount for reliable inference [@mckean_common_2000; @armitage_adaptive_2009].
+As the number of purely subsistence communities shrinks each year worldwide, and multiple levels of institutions are increasingly involved in natural resource management, incorporating these features into our models is paramount for reliable inference [@mckean_common_2000; @armitage_adaptive_2009].
 

### 2022-01-27 DOI_Review/Manuscript.rmd
text edits

@@
 bibliography: [./DOI_Conservation_Review.bib]
-abstract: A common goal of conservation practitioners is to design interventions that benefit both people and the environment, and as a result are adopted robustly by local communities.
-Diffusion of innovations theory (DOI) provides generalized heuristics and models that represent how learned behaviors spread within a population and has been increasingly identified by researchers as useful for understanding conservation adoption.
-Here, we elaborate on the need for better understanding conservation institutions as a product of individual behavioral adoption and review DOI and its past application to these problems.
-We follow this with a practical demonstration of how the mathematical models which fit the diffusion process can be used to diagnose and predict the uptake of conservation behaviors even under tightly coupled behavior-environment feedbacks.
-Data on conservation adoption are generated through an agent-based simulation which highlights our dynamics of interest.
-Throughout this review, we situate DOI in the broader literature on cultural evolution and use these insights to refine our claims about how diffusion can be productively applied to conservation.
+abstract: The study of community-based conservation is plagued by a large number of important variables, nonlinear dynamics, and interactions and feedbacks among predictors.
+This complexity has made the quantification of causal mechanisms and between-site comparative analyses notoriously difficult.
+Here, we argue that analyzing the emergence and persistence of community-based conservation institutions at the level of the individual, rather than the entire community, can yield important quantitative insights.
+We highlight the utility of Diffusion of Innovations theory (DOI) and the broader field of cultural evolution.
+In particular, the differential models of disease transmission, on which DOI is founded, may provide a productive avenue for quantifying the spread and persistence of community-based conservation institutions.
+We first simulate data on community adoption of a conservation institution, contingent on feedbacks between individual behavior and environmental processes.
+We then demonstrate that fitting these data to a probabilistic differential model of diffusion can produce reliable estimates of the rates of adoption, dropout, and long-term uptake from early on in the lifetime of an institution.
 Specifically, we show that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention.
-Overall, we provide a quantitative approach for modeling the spread of conservation behaviors and argue for a more complete incorporation of cultural evolutionary theory when considering diffusion of community-based conservation initiatives.
+Overall, we explore a new quantitative approach for modeling the spread of conservation behaviors using probabilistic differential equations and argue for the incorporation of cultural evolutionary theory into the field.
 
@@
 
-It is becoming increasingly apparent that humanity will not reach the global conservation goals we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002].
-We are thus turning to local communities to self-impose and maintain sustainable limits on resource extraction and even to restore degraded natural areas [@amel_beyond_2017; @catalano_learning_2019].
+It is becoming increasingly apparent that humanity will not reach the global conservation targets we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002].
+Conservationists are thus turning to local communities to self-impose and maintain sustainable limits on resource extraction and even to restore degraded natural areas [@amel_beyond_2017; @catalano_learning_2019].
 Community-based conservation projects aim to incentivize these behaviors in rural, and often very poor, communities with little to no external oversight [@western_natural_1994].
-While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote behavioral changes which are adopted robustly by communities who harvest natural resources to meet their daily needs [@brooks_assessing_2013; @horwich_community_2007].
-While some debate still exists, it is largely accepted that empowering communities to self-run conservation institutions and manage environments for multiple interests is a viable method for seeding self-sustaining and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004; @agrawal_enchantment_1999; @yates_purpose_2019].
+While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote behavioral changes which are adopted robustly by resource dependent communities [@brooks_assessing_2013; @horwich_community_2007].
+While some debate still exists, it is largely accepted that empowering communities to self-run conservation institutions is a viable method for seeding self-sustaining and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004; @agrawal_enchantment_1999; @yates_purpose_2019].
 
@@
 
-Decades of research have identified key contextual factors that support the durability of resource safeguarding behaviors in a population, particularly with regard to common-pool resources [@agrawal_common_2001; @baland_halting_1996; @ostrom_governing_1990].
+Decades of research have identified key contextual factors that support the durability of resource safeguarding behaviors in populations, particularly with regard to common-pool resources [@agrawal_common_2001; @baland_halting_1996; @ostrom_governing_1990].
 These factors are largely narrative descriptions of the qualities of, and interactions between, resource types, governance structures, and resource users which are most likely to lead to stable and sustainable resource management institutions [@ostrom_general_2009].
-Although at least 40 unique factors have been identified as contributing to the success of community-based conservation institutions, well defined social and ecological boundaries, and well connected social systems stand out as major important themes [@ostrom_governing_1990; @ostrom_general_2009; @agrawal_common_2001; @cox_review_2010].
+Scholars have identified at least 40 unique factors as contributing to the success of community-based conservation institutions and some overarching patterns have begun to emerge [@ostrom_governing_1990; @ostrom_general_2009; @agrawal_common_2001].
+Well-defined social and ecological boundaries, and well-connected social systems particularly stand out as important themes [@cox_review_2010].
 Research on common property additionally recommends that institutions be matched in scale to the underlying ecological processes of the resources they aim to manage [@mckean_common_2000].
@@
 Many of the factors that support the long-term viability of conservation institutions are intertwined such that the applicability of one may depend on the level of another and some factors may be inextricably correlated [@agrawal_group_2001; @agrawal_sustainable_2003].
-Further, which specific factors are most important is highly specific to a given community or location, suggesting that there is not possibly one unifying panacea for successful conservation institutions [@brooks_how_2012; @geist_proximate_2002; @agrawal_communities_2001].
-Thus, even with decades of productive research on sustainable resource management, complete quantitative syntheses and insights have remained elusive.
+Predictors and outcomes of interest commonly exhibit threshold effects, time lags, feedbacks,  and a variety of other nonlinear dynamics, making causal inference through standard regression practices difficult [@folke_socialecological_2007; @levin_social-ecological_2013].
+And further, which individual factors are most important is highly specific to a given community or location, suggesting that there is not possibly one unifying panacea for successful conservation institutions [@brooks_how_2012; @geist_proximate_2002; @agrawal_communities_2001].
+This complexity of community-based resource management systems, as well as logistical constraints in measuring the sheer number of important variables has made multi-site comparative analyses elusive [@sesmad_social-ecological_2014; @biggs_routledge_2021].
+Thus, even with decades of productive research on sustainable resource management, all major syntheses remain descriptive in nature, lacking a robust methodology to quantify causal insights [@frey_synthesis_2017; @mcginnis_social-ecological_2014; @gutierrez_leadership_2011].
 
@@
 Predicting the final proportion of a population that will adopt an intervention is critical, as many interventions require buy-in from nearly all community members to be effective [@mangubhai_not_2011; @visconti_protected_2019].
-By examining sustainable resource management institutions as the emergent outcome of individual behavioral changes, we can leverage insights from other areas of social science to further guide theory on why some institutions emerge and others fail to be adopted long-term [@berkes_rethinking_2004].
+By examining sustainable resource management institutions as the emergent outcome of individual behavioral changes, we can leverage insights from other areas of social science to further guide theory on why some institutions emerge and persist, and others fail to be adopted long-term [@berkes_rethinking_2004].
 
@@
 Promoting the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time.
-DOI recognizes that this process is primarily governed by the level of uncertainty an individual has about a novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010].
+DOI recognizes that this process is primarily governed by the level of uncertainty an individual has  about the utility of the novel behavior and their personal threshold for the amount of certainty required to make a decision [@rogers_diffusion_2010].
 When an individual is introduced to a new behavior, they often have a high degree of uncertainty as to the risks and benefits of adopting and will wait to adopt until that uncertainty is reduced to a tolerable level.
@@
 While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information which is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016].
-This information affects individuals’ behaviors, which in turn may affect their environments, and further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003].
+This information affects individuals’ behaviors, which in turn may affect their environments, and feedback to further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003].
+Observed variation among groups, such as differences in resource management institutions, are produced by such social and environmental histories, then reinforced through conformist-biased learning where the dominant behavioral trait is disproportionately copied [@henrich_evolution_1998].
 Thus, cultural evolution provides an intuitive way to integrate individual decision-making, population level patterns, and even population level behavior-environment feedbacks.
@@
 When the natural history characteristics of some pathogen are known, the SI, SIR, SIS, & SIRS suite of models reliably give accurate, deterministic predictions about the number of individuals in a population who will become infected and for how long a disease will persist within a population [@huppert_mathematical_2013].
-For conservation interventions however, there are no natural history features of a specific intervention, such as the rate or duration of adoption, which can be known ahead of time.
+For a specific conservation interventions however, there is no way to precisely know the natural history features, such as the rate or duration of adoption, ahead of time.
 Instead, the qualitative insights from DOI and the study of sustainable resource management institutions referenced above aim to increase the rate of spread and reduce the rate of dropout so that the intervention persists in the population and may even spread to new communities.
@@
 A key question then is whether the diffusion models explained above will still reliably fit the spread of conservation behaviors when those behaviors alter the environment, and changes in the environment alter the payoff structure of various behavioral choices.
+
 To answer this question, we use an agent-based simulation to generate data on the spread of a conservation behavior via social learning, contingent on feedbacks with the environment.

### 2022-01-27 ParticipatoryMapping/LeakagePM.R
Gridded basemaps. Figures out how to set cell size

@@ -1,0 +1,28 @@
+library(sf)
+library(tidyverse)
+PembaBound<-read_sf("~/Pemba_Project/PembaShapeFile.shp")
+
+PembaBound<-PembaBound%>%mutate(StudyCom=ifelse(NAME_3 =="Mjimbini","Yes","No" ))
+
+cols<- c("No" = "#d9d9d9", "Yes" = "#525252")
+
+ggplot()+
+  geom_sf(data=PembaBound,aes(fill=StudyCom),color="#bdbdbd")+
+  theme_bw()+
+  geom_sf_text(data=PembaBound,aes(label=NAME_3),size=1)+
+  scale_colour_manual(
+    values = cols,
+    aesthetics = ( "fill")
+  )+
+  theme(axis.text.x=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks=element_blank(),
+        axis.title.x=element_blank(),
+        axis.title.y=element_blank(),
+        legend.position="none",
+        panel.grid.major=element_blank(),
+        panel.grid.minor=element_blank())+
+ggspatial::annotation_north_arrow(location = "bl", which_north = "true",height=unit(1,"cm"), 
+                                  width=unit(1,"cm"),
+                                  pad_x = unit(0.0, "in"), pad_y = unit(0, "in"), 
+                                  style = north_arrow_fancy_orienteering)

### 2022-01-27 ParticipatoryMapping/MakeGriddedBasemaps.R

@@ -1,0 +1,129 @@
+library(ggmap)
+library(sf)
+library(tidyverse)
+register_google(key="AIzaSyCh-KGA6T9qaPF-wjfnBV8hb7zzGYMlGTw")
+
+Pemba <- read_sf("~/Pemba_Project/PembaShapeFile.shp")
+
+####Make fishnet
+initial<-st_transform(Pemba, 3857) %>% dplyr::select(NAME_3, geometry)
+initial$index_target <- 1:nrow(initial)
+target <- st_geometry(initial)
+grid <- st_make_grid(target,
+                     cellsize=1000, #In meters
+                    
+                     crs = st_crs(initial),
+                     what = "polygons",
+                     square = TRUE
+)
+
+# To sf
+grid <- st_sf(index = 1:length(lengths(grid)), grid) # Add index
+cent_grid <- st_centroid(grid)
+cent_merge <- st_join(cent_grid, initial["index_target"], left = F)
+grid_new <- inner_join(grid, st_drop_geometry(cent_merge))
+
+# Fishnet
+Fishgeom <- aggregate(grid_new,
+                      by = list(grid_new$index_target),
+                      FUN = min,
+                      do_union = FALSE
+)
+
+# Lets add the df
+Fishnet <- left_join(
+  Fishgeom %>% dplyr::select(index_target),
+  st_drop_geometry(initial)
+) %>%
+  dplyr::select(-index_target)
+
+plot(st_geometry(Fishnet), main = "Fishnet")
+
+##########Plotting function for merging ggmap and sf
+
+# Transform nc to EPSG 3857 (Pseudo-Mercator, what Google uses)
+Pemba_3857 <- st_transform(Pemba, 3857)
+
+map <- get_googlemap("pemba island", scale=2)
+
+
+# Define a function to fix the bbox to be in EPSG:3857
+ggmap_bbox <- function(map) {
+  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
+  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
+  # and set the names to what sf::st_bbox expects:
+  map_bbox <- setNames(unlist(attr(map, "bb")), 
+                       c("ymin", "xmin", "ymax", "xmax"))
+  
+  # Coonvert the bbox to an sf polygon, transform it to 3857, 
[…]

### 2022-03-12 DOI_Review/Manuscript.rmd
updating

@@
 We then demonstrate that fitting these data to a probabilistic differential model of diffusion can produce reliable estimates of the rates of adoption, dropout, and long-term uptake from early on in the lifetime of an institution.
-Specifically, we show that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention.
-Overall, we explore a new quantitative approach for modeling the spread of conservation behaviors using probabilistic differential equations and argue for the incorporation of cultural evolutionary theory into the field.
+We emphasize that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention.
+Overall, we explore a new quantitative approach for modeling the spread of conservation behaviors using probabilistic differential equations and argue for more complete incorporation of cultural evolutionary theory into the field.
 
@@
 ```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
+#
 library(knitr)
@@
 This complexity of community-based resource management systems, as well as logistical constraints in measuring the sheer number of important variables has made multi-site comparative analyses elusive [@sesmad_social-ecological_2014; @biggs_routledge_2021].
-Thus, even with decades of productive research on sustainable resource management, all major syntheses remain descriptive in nature, lacking a robust methodology to quantify causal insights [@frey_synthesis_2017; @mcginnis_social-ecological_2014; @gutierrez_leadership_2011].
+Thus, even with decades of productive research on sustainable resource management, all major syntheses remain descriptive in nature, lacking a robust methodology to quantify causal insights [@cumming_advancing_2020; @frey_synthesis_2017; @mcginnis_social-ecological_2014; @gutierrez_leadership_2011].
 
@@
 The normative conditions imposed on individuals by their ecological and social environments (i.e.
-the factors referenced above) affect the likelihood that they will endogenously engage in cooperative resource management practices or adopt them when exogenously introduced by an intervention.
+the factors referenced above) affect the likelihood that they will endogenously engage in cooperative resource management practices or adopt them when exogenously introduced by an intervention [@wilson_generalizing_2013; @andrews_cultural_2018].
 
@@
 The primary way individuals reduce uncertainty about the outcome of adopting a particular behavior is through second-hand observation of early adopters.
-Therefore, at the core of DOI is the observation that individuals imitate those around them when making decisions about adopting a novel behavior, rather than weighing the pros and cons independently [@henrich_search_2001].
+Therefore, at the core of DOI is the observation that individuals imitate those around them when making decisions about adopting a novel behavior, rather than weighing the pros and cons independently [@rogers_diffusion_2010; @henrich_search_2001].
 
@@
 This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001].
+The curvilinear “R” shape produced by independent, or “environmental” learning is a product of probability theory, in that late adopters represent the far tail of the distribution, having not discovered (and thus adopted) the novel behavior even after many trials [@henrich_cultural_2001].
 These two curve shapes however are not completely distinct.
@@
 For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters at the sigmoidal equilibrium.
-The authors take insights from DOI to suggest that practitioners may actively spread information on interventions (as in Abernethy et al.
-[-@abernethy_two_2014]), or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004].
+The authors take insights from DOI to suggest that practitioners may actively spread information on interventions [as in @abernethy_two_2014], or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004].
 Further, a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices identifies qualitative insights, such as intervention observability, from DOI as useful in guiding the development of conservation interventions [@mahajan_theorybased_2020].
@@
 
-We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, frequency-dependent and success biased.
+We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, frequency-dependent and success-biased.
 Aside from the learning mechanism employed by the population, all model conditions are identical for these two scenarios.
@@
 As the amount of available resources in an area is determined by the number of agents who harvest there, this learning mechanism incorporates behavior-environment feedbacks typical of natural resource management institutions.
-Because agents compare their own harvests to those of others in each time step under the success-biased learning scenario, rather than exclusively imitating other individuals, this scenario actually represents a mix of individual and social learning.
+Having independently sampled from one or both environments, agents then compare their harvests to those of others in each time step.
+In this way, the success-biased learning scenario actually represents a mix of individual and social learning, rather than a scenario where individuals make decisions purely from social information.
 The adoption pattern observed in this simulation results in a mix between an “R” and “S” type curve, reinforcing the interpretation of this scenario as a mix of individual and social learning [@henrich_cultural_2001].
@@
 
-![SIRS model fit of agent-based simulation data under unbiased social learning shown as translucent black lines representing 1000 draws from the posterior.
+![SIRS model fit of agent-based simulation data under frequency-dependent social learning shown as translucent black lines representing 1000 draws from the posterior.
 Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./DirectCopying.png)
@@
 
+Lastly, while we have shown here that DOI can inform community-based conservation efforts in silico, the real-world may constrain such productive application.
+In a sharp response to Mascia and Mills [-@mascia_when_2018], authors Lund and Blewstein [-@lund_when_2018] question the narrative of “conservation interventions as innovations that can be freely adopted.” They argue that top-down support and manipulation are common in these contexts and may hinder the applicability of “demand-driven” conceptualizations of how conservation practices spread.
+These criticisms are paralleled by issues raised by Rogers himself when identifying the pro-innovation and individual blame biases of DOI, which portray lack of adoption as a flaw of the individual, opposed to the innovation [@rogers_diffusion_2010].
+Given the problematic history of colonial-type conservation, we urge readers to seriously consider who is providing the narrative when assessing how and why conservation institutions diffuse within populations [@agrawal_enchantment_1999-1].
+
 # Future work
@@
 In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems which are difficult or even impossible to account for with process models such the SIRS  or SIRS extension model presented in equation 4 [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
-Agent-based simulations allow modelers to impose a realism in the governance or institutional dynamics that are largely intractable in closed-form mathematical equations [@bravo_agents_2011; joffre_combining_2015].
+Agent-based simulations allow modelers to impose a realism in the governance or institutional dynamics that are largely intractable in closed-form mathematical equations [@bravo_agents_2011; @joffre_combining_2015].
 As the number of purely subsistence communities shrinks each year worldwide, and multiple levels of institutions are increasingly involved in natural resource management, incorporating these features into our models is paramount for reliable inference [@mckean_common_2000; @armitage_adaptive_2009].

### 2022-03-12 DOI_Review/ApplySTANtoABMdat.R

@@ -191,7 +191,7 @@ mod_time = stan_d$fake_ts
 df_sample = data.frame(sample_prop, sample_time)
 df_fit = data.frame(mod_median, mod_low, mod_high, mod_time)
 
-draws<-as_tibble(posts$fake_I[,,2])%>%add_column(draw=1:1000)
+draws<-as_tibble(posts$fake_I[,,2])%>%add_column(draw=1:4000)
 names(draws)[1:100]<-1:100
 
 draws <-  pivot_longer(draws, c(1:100) , names_to = "mod_time")

### 2022-07-21 DOI_Review/Manuscript.rmd
updated post-reviewers' comments

@@
 ---
-title: Utility of Diffusion of Innovations for Diagnosing the Spread of Conservation Behaviors
+title: A Quantitative Application of Diffusion of Innovations for Modeling the Spread
+  of Conservation Behaviors
 author:
@@
   name: Jeffrey Andrews
-
 date: "`r format(Sys.time(), '%B %d, %Y')`"
-keywords: Community-based conservation, diffusion of innovations, cultural evolution, complex systems
-bibliography: [./DOI_Conservation_Review.bib]
-abstract: The study of community-based conservation is plagued by a large number of important variables, nonlinear dynamics, and interactions and feedbacks among predictors.
-This complexity has made the quantification of causal mechanisms and between-site comparative analyses notoriously difficult.
-Here, we argue that analyzing the emergence and persistence of community-based conservation institutions at the level of the individual, rather than the entire community, can yield important quantitative insights.
-We highlight the utility of Diffusion of Innovations theory (DOI) and the broader field of cultural evolution.
-In particular, the differential models of disease transmission, on which DOI is founded, may provide a productive avenue for quantifying the spread and persistence of community-based conservation institutions.
-We first simulate data on community adoption of a conservation institution, contingent on feedbacks between individual behavior and environmental processes.
-We then demonstrate that fitting these data to a probabilistic differential model of diffusion can produce reliable estimates of the rates of adoption, dropout, and long-term uptake from early on in the lifetime of an institution.
-We emphasize that the social learning mechanism individuals employ can alter the population pattern of adoption and the final proportion of individuals that adopt an intervention.
-Overall, we explore a new quantitative approach for modeling the spread of conservation behaviors using probabilistic differential equations and argue for more complete incorporation of cultural evolutionary theory into the field.
-
+output:
+  pdf_document:
+    keep_tex: yes
+    fig_caption: yes
+    latex_engine: pdflatex
+    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
+  word_document: default
+bibliography: ./DOI_Conservation_Review.bib
+abstract: The study of community-based conservation is challenged by a large number
+  of important variables and nonlinear dynamics. This complexity has made quantitative
+  and comparative analyses notoriously difficult. Here, we argue that analyzing the
+  emergence and persistence of community-based conservation institutions as an emergent
+  phenomenon of individual decision-making can yield important quantitative insights.
+  We first review diffusion of innovations theory (DOI) and the broader field of cultural
+  evolution. We then simulate data on community adoption of a conservation institution,
+  contingent on feedbacks between individual behavior and environmental processes.
+  We demonstrate that fitting these data to differential models of disease transmission,
+  on which DOI is founded, can produce reliable estimates of the rates of adoption,
+  dropout, and long-term uptake of an institution. Overall, we explore a new quantitative
+  approach for modeling the spread of conservation behaviors using probabilistic differential
+  equations and argue for further incorporation of cultural evolutionary theory into
+  the field.
 fontsize: 12pt
 link-citations: yes
-output: 
-  pdf_document:
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
+keywords: Community-based conservation, diffusion of innovations, cultural evolution,
+  complex systems
 header-includes:
@@
 - \usepackage{amsmath}
-
 ---
@@
 
+# 1 - Introduction
+
 > "Diffusion study...is something like the use of radioactive tracers in studying the process of plant   growth: it helps illuminate the process."
@@
 
-It is becoming increasingly apparent that humanity will not reach the global conservation targets we have set for ourselves through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002].
+## 1.1 - Motivation
+It is becoming increasingly apparent that humanity will not reach global conservation targets through top-down policies alone [@dudley_essential_2018; @mair_achieving_2021; @tittensor_mid-term_2014; @brockington_fortress_2002].
 Conservationists are thus turning to local communities to self-impose and maintain sustainable limits on resource extraction and even to restore degraded natural areas [@amel_beyond_2017; @catalano_learning_2019].
 Community-based conservation projects aim to incentivize these behaviors in rural, and often very poor, communities with little to no external oversight [@western_natural_1994].
-While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote behavioral changes which are adopted robustly by resource dependent communities [@brooks_assessing_2013; @horwich_community_2007].
+While the details of each intervention vary considerably, the general strategy is to closely link community development and conservation goals to promote behavioral changes that are adopted robustly by resource dependent communities [@brooks_assessing_2013; @horwich_community_2007].
 While some debate still exists, it is largely accepted that empowering communities to self-run conservation institutions is a viable method for seeding self-sustaining and ethical protections on natural resources [@turner_global_2007; @berkes_community-based_2007; @adams_biodiversity_2004; @agrawal_enchantment_1999; @yates_purpose_2019].
 
-Broadly defined, institutions, like those introduced by community-based conservation interventions, are socially constructed and transmitted prescriptions on behavior, which generally aim to incentivize collective action [@lott_review_1992; @currie_evolution_2016; @richerson_tribal_2009].
+Broadly defined, institutions, like those introduced by community-based conservation interventions, are socially constructed and transmitted prescriptions on behavior that often aim to incentivize collective action [@lott_review_1992; @currie_evolution_2016; @richerson_tribal_2009].
 In natural resource management, the desired outcome of any particular institution varies on a case-by-case basis but may include goals like increasing the livelihood benefits produced from a resource or improving the biodiversity of a particular area [@agrawal_studying_2014].
@@
 Decades of research have identified key contextual factors that support the durability of resource safeguarding behaviors in populations, particularly with regard to common-pool resources [@agrawal_common_2001; @baland_halting_1996; @ostrom_governing_1990].
-These factors are largely narrative descriptions of the qualities of, and interactions between, resource types, governance structures, and resource users which are most likely to lead to stable and sustainable resource management institutions [@ostrom_general_2009].
+These factors are largely narrative descriptions of the qualities of, and interactions between, resource types, governance structures, and resource users that are most likely to lead to stable and sustainable resource management institutions [@ostrom_general_2009].
 Scholars have identified at least 40 unique factors as contributing to the success of community-based conservation institutions and some overarching patterns have begun to emerge [@ostrom_governing_1990; @ostrom_general_2009; @agrawal_common_2001].
-Well-defined social and ecological boundaries, and well-connected social systems particularly stand out as important themes [@cox_review_2010].
+Continual monitoring, graduate sanctions for offenses, and well-defined social boundaries particularly stand out as important themes [@cox_review_2010; @baggio_explaining_2016].
 Research on common property additionally recommends that institutions be matched in scale to the underlying ecological processes of the resources they aim to manage [@mckean_common_2000].
@@
 Predictors and outcomes of interest commonly exhibit threshold effects, time lags, feedbacks,  and a variety of other nonlinear dynamics, making causal inference through standard regression practices difficult [@folke_socialecological_2007; @levin_social-ecological_2013].
-And further, which individual factors are most important is highly specific to a given community or location, suggesting that there is not possibly one unifying panacea for successful conservation institutions [@brooks_how_2012; @geist_proximate_2002; @agrawal_communities_2001].
+And further, which individual factors are most important is highly specific to a given community or location, suggesting that there is not possibly one unifying panacea for common-pool resource management issues [@brooks_how_2012; @geist_proximate_2002; @agrawal_communities_2001].
 This complexity of community-based resource management systems, as well as logistical constraints in measuring the sheer number of important variables has made multi-site comparative analyses elusive [@sesmad_social-ecological_2014; @biggs_routledge_2021].
@@
 
-In this article, we examine one social science theory specifically focused on individual behavior adoption, diffusion of innovations (DOI), and its previous and potential applications to community-based conservation institutions [@rogers_diffusion_2010].
+In this article, we examine one social science theory specifically focused on individual behavior adoption, diffusion of innovations (DOI), and its previous and potential applications to community-based conservation institutions [@rogers_diffusion_2010; @mascia_when_2018].
 Our primary aim is to give a practical demonstration of how insights from DOI can be applied to expand our inference about the spread and persistence of conservation interventions beyond narrative descriptions.
 We do this by first simulating data on conservation adoption in a resource-dependent population and then statistically fitting a model of diffusion to these data.
-Adoption of resource safeguarding behavior is fundamentally different from behavioral adoption in the abstract in that individual adoptions affect the natural environment and alter the payoff structure under which individuals continue to adopt in future time periods.
+Adoption of resource safeguarding behaviors is fundamentally different from adoption of more standard behaviors such as clothing choice.
+By design, conservation behaviors alter the natural environment and adjust the arena in which individuals make future choices, such as to maintain the behavior.
 Our simulation highlights this feedback of interest in order to answer the key question of whether DOI is applicable to community-based conservation specifically, given this unique adoption dynamic.
@@
 
-# Diffusion of innovations
+## 1.2 - Diffusion of innovations
 Promoting the widespread adoption of some novel (to the individual) behavior is not an issue unique to community-based conservation; producers of cigarettes and snack foods, as well as public health professionals, have been grappling with these questions for quite some time.
@@
 Instead, they preferentially adopt traits by employing a variety of learning biases, such as success-biased learning, where individuals who are perceived as more successful are disproportionately likely to be copied [@boyd_culture_1988; @laland_social_2004].
-As a result, the cumulative culture of a population evolves over time as various beliefs and behaviors are unequally copied and their frequency within a population changes.
-While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information which is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016].
-This information affects individuals’ behaviors, which in turn may affect their environments, and feedback to further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003].
+As a result, the cumulative culture of a population evolves over time as various beliefs and behaviors are preferentially copied and their frequency within a population changes.
+While a full review of cultural evolution is beyond the scope of this paper, the general intuition is that populations carry a variable pool of information that is shaped by events in the lives of individuals such as social observation or natural selection [@mesoudi_cultural_2016].
+This information affects individuals’ behaviors, that in turn may affect their environments, and feedback to further alter which traits are adopted in the future [@durham_coevolution_1991; @henrich_evolution_2003].
 Observed variation among groups, such as differences in resource management institutions, are produced by such social and environmental histories, then reinforced through conformist-biased learning where the dominant behavioral trait is disproportionately copied [@henrich_evolution_1998].
@@
 Diffusion of novel beliefs and behaviors is a specific instance of cultural evolution, where the trait being transmitted is new to the population and outcompetes some previous cultural variant, such as in the classic example of hybrid corn seed adoption in Iowa farmers [@mesoudi_cultural_2016; @ryan_acceptance_1950].
-We observe these patterns using diffusion curves which show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
+We observe these patterns using diffusion curves that show the proportion of a population that has adopted a belief or behavior on the Y-axis, as a function of time on the X-axis [@hoppitt_effect_2010; @rogers_diffusion_2010; @ryan_acceptance_1950].
 The exact shape of these diffusion curves can be highly dependent on the learning mechanism employed by individuals in the population, as well as contextual factors like the population density or ease of trialing the cultural variant [@smaldino_adoption_2017; @henrich_cultural_2001; @rogers_diffusion_2010].
 
-In the vast majority of empirical studies, the diffusion of a novel trait produces a sigmoidal “S” shape, where it initially spreads slowly, followed by a period of rapid adoption, then again spreads slowly until it reaches fixation [@rogers_diffusion_2010].
+In the vast majority of empirical studies, the diffusion of a novel trait produces a sigmoidal “S” shape, where it initially spreads slowly, followed by a period of rapid adoption, then again spreads slowly until it reaches an equilibrium [@rogers_diffusion_2010].
 This “S” shaped curve is indicative of biased social learning, in contrast to the relatively rare “R” shaped curves observed when individuals learn through independent trial and error (fig 1) [@henrich_cultural_2001].
-The curvilinear “R” shape produced by independent, or “environmental” learning is a product of probability theory, in that late adopters represent the far tail of the distribution, having not discovered (and thus adopted) the novel behavior even after many trials [@henrich_cultural_2001].
+The curvilinear “R” shape produced by individual, or environmental learning is a result of the fact that late adopters represent the tail of the distribution, having not discovered (and thus adopted) the novel behavior even after many trials [@henrich_cultural_2001].
 These two curve shapes however are not completely distinct.
-As the combination of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].
+As the proportion of learning strategies employed by individuals in a population shifts more toward individual or social learning, we expect to see a shift toward an “R” shape or “S” shape respectively [@henrich_cultural_2001].
 
@@
 
-# Diffusion of innovations in community-based conservation
+## 1.3 - Diffusion of innovations in community-based conservation
 Just as the sigmoidal pattern indicative of social learning is seen in the majority of cases on behavioral adoption broadly, scholars have also noticed this pattern in over 80% of case studies on the uptake of community-based conservation interventions worldwide [@mascia_when_2018; @mills_how_2019].
@@
 For example, in a review of 22 conservation interventions referenced above, the majority of initiatives were adopted by less than 30% of potential adopters at the sigmoidal equilibrium.
-The authors take insights from DOI to suggest that practitioners may actively spread information on interventions [as in @abernethy_two_2014], or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004].
+The authors use insights from DOI to suggest that practitioners may actively spread information on interventions (as in Abernethy et al.
+[-@abernethy_two_2014]), or facilitate communication to increase adoption rates [@mills_how_2019; @greenhalgh_diffusion_2004].
 Further, a recent framework for understanding the establishment, persistence, and spread of community‐based conservation practices identifies qualitative insights, such as intervention observability, from DOI as useful in guiding the development of conservation interventions [@mahajan_theorybased_2020].
@@
 [-@romero-de-diego_drivers_2021] used expert interviews to identify that observability of benefits was a key driver of the rapid adoption of a voluntary wildlife management program among landholders in Mexico.
-Conversely, in contexts where observability and thus imitation are limited, such as across island chains, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
-This finding mirrors those of studies which look at more general cultural adoption across island chains, further suggesting that conservation behaviors may spread similarly to technological or ideological innovations [@kline_population_2010; @beheim_inheritance_2011].
+Conversely, in contexts where observability and thus imitation are limited, such as across neighboring islands, the rate and degree of spread of community-based conservation interventions is considerably attenuated [@pietri_information_2009-1].
+This finding mirrors those of studies which look at more general cultural adoption between islands, further suggesting that conservation behaviors may spread similarly to technological or ideological innovations [@kline_population_2010; @beheim_inheritance_2011].
 
-These studies are compelling in their conclusion that DOI can be productively applied to the spread of community-based conservation, but have been critiqued as general, narrative suggestions from DOI that may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020].
+These studies are compelling in their conclusion that DOI can be productively applied to the spread of community-based conservation, but have been critiqued as general, narrative suggestions from DOI that may miss important contextual factors unique to each intervention [@steenbergen_theory_2021-1; @lam_scaling_2020; @lund_when_2018].
 Below, we make a quantitative argument for using DOI to understand the spread of community-based conservation interventions.
@@
 
-# Quantitative applications
-In this section we explain how we can fit an observed pattern of diffusion mathematically to estimate the rates of adoption and dropout, as well as to predict the speed and degree to which an intervention will be adopted by a population.
+## 1.4 - Quantitative applications
+In this section we explain how we can fit an observed pattern of diffusion mathematically to estimate rates of adoption and dropout, as well as to predict the speed and degree to which an intervention will be adopted by a population.
 The “S” shaped diffusion pattern of behavioral spread over time mirrors that of community disease transmission [@smaldino_coupled_2020].
@@
 To answer this question, we use an agent-based simulation to generate data on the spread of a conservation behavior via social learning, contingent on feedbacks with the environment.
-Then, using just the first 30% of the data, we probabilistically estimate the rate of intervention spread, rate of dropout, and readoption using a Bayesian instantiation of the SIRS model shown in equation 1.
-We use these estimated parameters to predict the long-term patterns of adoption for the remaining 70% of time steps and compare our predicted values to those observed in the simulation.
+Then, using the five run average of just the first 30% of the data, we probabilistically estimate the average rates of intervention spread, dropout, readoption using a Bayesian instantiation of the SIRS model shown in equation 1.
+We use these estimated parameters to predict the long-term average adoption pattern for the remaining 70% of time steps and compare our predicted values to those observed in the simulations.
 While the purpose of this example is to test whether the SIRS model can capture our core dynamics of interest, it also serves as an explicit demonstration of how a practitioner or other researchers can directly apply the quantitative process described above.
@@
 
-# Agent-based simulation
+# 2 - Modeling 
+## 2.1 - Agent-based simulation
 We seek to test whether the SIRS model shown in equation 1 can produce reliable insights about community uptake of a conservation behavior, especially given behavior-environment feedbacks common in natural resource management institutions.
-To answer this question, we first need complete time series data in-hand which are a product of our dynamics of interest.
-To obtain these data, we develop a minimal agent-based model which we use to simulate data on the spread of a resource safeguarding behavior within a population.
-For comparison, we simulate adoption data both independent of, and contingent on feedbacks with the environment.
-These two dynamics are operationalized using two different social learning mechanisms for behavioral transmission [@kendal_social_2018].
-The first is frequency-dependent copying where agents adopt an alternative behavior with a probability proportional to its frequency in the population, independent of the harvest benefit provided by adopting the behavior or not [@barrett_equifinality_2019].
-We include a simulation using this mechanism as a baseline for how behavior-environment feedbacks inherent to our second mechanism affect the overall pattern of adoption.
+To answer this question, we first need complete time series data in-hand that are a product of our dynamics of interest.
+To obtain these data, we develop a minimal agent-based model that we use to simulate data on the spread of a resource safeguarding behavior within a population [@muller_describing_2013; @smajgl_empirical_2011].
+In order to assess the impact of feedbacks between behavior and the environment, we simulate data using two different social learning strategies, only one of which involves a behavior-environment feedback [@le_feedback_2012; @kendal_social_2018].
+The first is random copying where agents are paired with another agent at random and have a 25% chance of adopting the strategy of the agent they are paired with, independent of the harvest benefit provided by adopting the behavior or not [@mcelreath_beyond_2008; @barrett_equifinality_2019].
+This strategy does not involve a direct feedback between behavior and the environment.
+We include it as a baseline to compare against our second strategy, which does.
 The second mechanism is success-biased learning, under which agents disproportionately copy the behaviors of other agents with the greatest harvesting success [@baldini_two_2013].
@@
 success); a $\beta$ = 1 in this formulation would thus mean that the probability that one agent copies another is directly proportional to the ratio of their harvests.
+Behavioral change given success-biased learning is therefore dependent on feedbacks with the environment as harvest success is a result of agent behaviors in previous time steps.
 
@@
 We establish a fishery, loosely based on a global average stock integrity, at 20% of its potential carrying capacity [@fao_state_2020].
-In the first time step, we introduce an intervention which delineates 20% of the fishery as a conservation area and just 2% of agents initially change their behavior to forgo harvesting there.
+In the first time step, we introduce an intervention that delineates 20% of the fishery as a conservation area and just 2% of agents initially change their behavior to forgo harvesting there.
 Fish reproduce proportionally to the population in each time period, and move freely between the conservation area and unconserved area.
@@
 Agents not enrolled in the intervention harvest from the entire fishery and forgo the reward.
-After each harvest period, agents are randomly paired, and employ either frequency-dependent or success-biased learning to decide whether or not to adopt the behavior of their partner in the next time step [@baldini_success-biased_2012].
+After each harvest period, agents are randomly paired, and employ either random copying or success-biased learning to decide whether or not to adopt the behavior of their partner in the next time step [@baldini_success-biased_2012].
 The conserved and unconserved resource pools are reduced according to agents’ harvesting behavior, regenerate, and move from one area to the other if the carrying capacity is reached in any respective area.
-A visual description of this minimal model is shown in figure 2 and interested readers can find an interactive web application which allows users to run this simulation given a wide variety of starting conditions at [https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/](https://matthewclark.shinyapps.io/ConservationAdoptionSimulation/).
+A visual description of this minimal model is shown in figure 2 and the full Overview, Design concepts, and Details protocol can be found in the supplemental material (S1) [@grimm_standard_2006; @grimm_odd_2020].
 
@@
 
-We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, frequency-dependent and success-biased.
+We use this model to simulate data on the adoption and persistence of a conservation institution given two different learning mechanisms, random copying and success-biased imitation.
 Aside from the learning mechanism employed by the population, all model conditions are identical for these two scenarios.
-Frequency-dependent transmission is used as a reference as it reliably produces the standard sigmoidal “S” curve of behavioral adoption and does not incorporate any feedbacks between adoption and the environment.
+Random copying is used as a reference as it reliably produces the standard sigmoidal “S” curve of behavioral adoption and does not incorporate any feedbacks between adoption and the environment.
 Under the success-biased learning scenario, agents compare their harvests to others in each time step and disproportionately copy more successful agents.
@@
 The adoption pattern observed in this simulation results in a mix between an “R” and “S” type curve, reinforcing the interpretation of this scenario as a mix of individual and social learning [@henrich_cultural_2001].
-Figure 3 shows the full output of the later simulation where adoption (enrollment) spikes initially, but drops off dramatically as the resources in the unconserved area are reduced following increased harvest pressure there as a result of the conservation intervention.
+Figure 3 shows the full output for five runs of the later simulation where adoption (enrollment) spikes initially, but drops off dramatically as the resources in the unconserved area are reduced following increased harvest pressure there as a result of the conservation intervention.
+This exact dynamic is however highly flexible given the specific parameter combinations as described in section (2.3).
 
-![One run of the agent-based simulation used to produce data for analysis with the SIRS model of behavioral adoption under the success-biased learning scenario.
+![Five runs of the agent-based simulation used to produce data for analysis with the SIRS model of behavioral adoption under the success-biased learning scenario.
 The top figure shows the number of resources in both the conservation and unconserved areas over time, as a percent of the carrying capacity in those areas.
 The middle figure shows the average number of resources harvested (payoff) by each agent at each time step.
-The bottom figure shows the number of agents who have adopted the conservation behavior (enrolled) at every time step.](./SiccessBiasedFullRun.png)
+The bottom figure shows the number of agents who have adopted the conservation behavior (enrolled) at every time step.
+The exact parameter combination used to produce this figure can be found in the supplemental material (S1)](./SiccessBiasedFullRun.png)
 
-# Predicting adoption of conservation behaviors
-With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described in equation 1 to the first 30% of observations.
-We use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation (3) below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time.
+## 2.2 - Fitting data with probabilistic differential equations
+With data on conservation behavior adoption for both learning mechanisms in-hand, we fit the SIRS model described in equation 1 to the first  30% of observations, averaged across five model runs.
+We use multiple model runs to avoid spurious results stemming from model stochasticity alone.
+Five model runs are used as opposed to 100 or 1000, as this leaves some residual stochasticity in the adoption data that is more representative of what might be realistically observed.
+Duplicate analyses to those presented below, but instead run on the average of 100 simulation runs can be found in the supplemental material (S2).
+Given these data, we use probabilistic sampling to infer the parameters $\beta$, $\gamma$, and $\chi$, for the full Bayesian equation (3) below where $\text{Y}_{t}$ represents the total number of individuals who are enrolled in the community-based intervention at any given time.
 Here, "infected" individuals are those that are enrolled in the conservation intervention.
@@
 
-For both learning mechanisms the SIRS model exhibited convergence for all chains and resulted in an $\hat{R}$ of 1 for all estimated parameters (fig 4).
+For both learning mechanisms the SIRS model exhibited Markon chain convergence (fig 4) and resulted in an $\hat{R}$ of 1 for all estimated parameters.
 As the parameters $\beta$, $\gamma$, and $\chi$ are directly interpretable as the rates of adoption, dropout, and willingness to retry a previously abandoned intervention, accurate estimation of these values, especially in the early stages of an intervention, provides practical and straightforward intuition to diagnose areas for improvement in community-based conservation interventions.
-For example, a practitioner might fit this model to real-world participation rates during the early weeks of an intervention in order to estimate whether they should focus more on increasing rates of uptake or retention going forward.
-In addition, this insight may help practitioners better decide which factors from the institutional design literature referenced above may be most impactful for their specific system.
+For example, a practitioner might fit this model to real-world participation rates during the early weeks of an intervention in order to estimate the rates of adoption and dropout.
+If both rates are low, then the practitioner should focus their efforts on increasing initial adoption.
+Conversely, if both rates are high, they should instead focus efforts on retention of enrolled participants.
+In addition, these insights may help practitioners better decide which factors from the institutional design literature may be most impactful for their specific system.
+This might take the form of increasing communication and observability of benefits to boost initial adoption, or investing in conflict resolution mechanisms to reduce dropout.
 
-![Markov chains for each of the three parameters of interest from the SIRS model fit to simulated data.
+![Markov chains for each of the three parameters of interest from the SIRS model fit to simulated data (average of five runs).
 Beta is interpretable as the rate of adoption of the behavior (enrollment).
@@
 Chi represents the rate at which individuals who previously dropped out are added back into the susceptible category.
-The agent-based simulation used to produce the data employed environmentally mediated, success-biased learning.
-A figure showing the Markov chains for the unbiased learning simulation can be found in the supplemental material (S3).](./SuccessPars.png)
+The agent-based simulation used to produce the data employed environmentally mediated, success-biased learning.](./SuccessPars.png)
 
 We can also use these parameter values, estimated from the first 30% of the time series, to predict the long-term patterns of adoption for a community-based conservation initiative.
-In figures 5 & 6, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step.
-Gray and blue points show the initial 30% of the data which was used to estimate the three parameters of interest and the remaining 70% of data, respectively.
-These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both unbiased learning and environmentally mediated success-biased learning.
-R and STAN code for these models can be found in the supplemental material (S3).
+In figures 5 & 6, the translucent black lines show 1000 draws from the posterior predicting the proportion of the population who will be enrolled in the intervention at every time step, averaged across the five model runs.
+Gray and blue points show the initial 30% of the data that was used to estimate the three parameters of interest and the remaining 70% of data, respectively.
+These figures show that the SIRS model accurately captures and predicts the pattern of behavioral adoption for both random copying and environmentally mediated success-biased learning.
+Further, the mean absolute error (MAE) between the simulated “test” data (later 70%) and the median model predictions for those data are 2.98% and 4.52% of the population for the random copying and success-biased learning scenarios respectively, indicating good model fit.
 
-![SIRS model fit of agent-based simulation data under frequency-dependent social learning shown as translucent black lines representing 1000 draws from the posterior.
+![SIRS model fit to the average of five runs of the agent-based simulation data under random copying shown as translucent black lines representing 1000 draws from the posterior.
 Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./DirectCopying.png)
 
-![SIRS model fit of agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior.
+![SIRS model fit to the average of five runs of the agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior.
 Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./SuccessBiasedCopying.png)
 
-Figures 5 & 6 also demonstrate that different scenarios and respective learning mechanisms, while both producing fixation in the adoption rate, can result in different overall rates in adoption of an intervention, even under identical starting conditions.
-In the simulation of a population employing frequency-dependent learning, the population equilibrates at nearly 90% adoption of the intervention behavior.
+Figures 5 & 6 also demonstrate that different scenarios and respective learning mechanisms, while both producing a stable adoption rate, can result in different overall rates in adoption of an intervention, even under identical starting conditions.
+In the simulation of a population employing random copying, the population equilibrates at nearly 90% adoption of the intervention behavior.
 For success-biased learning on the other hand, resource scarcity in the unconserved area caused the population to adopt the intervention at approximately 60% in the long-term.
@@
 
-# Shortcomings for application of diffusion of innovations to community-based conservation
+## 2.3 - Example limitations
+Parallel to the examples above, we can use our agent-based simulation to explore parameter combinations that expose limitations for the application of the SIRS diffusion model for capturing and predicting conservation adoption dynamics under behavior-environment feedbacks.
+Specifically, as the parameters in the SIRS model only directly account for social processes, adoption dynamics resulting from abrupt environmental change are not well predicted.
+For example in figure 7,  we consider a resource pool that is relatively intact at the start of the simulation (80% of carrying capacity), paired with an unsustainable harvest rate.
+In this simulation, all agents initially rapidly enroll in the conservation intervention as there is no cost to doing so (resources in the unconserved area are plentiful).
+However, when the resource pool in the unconserved area is depleted, agents quickly drop out of the conservation institution in favor of harvesting from the conservation area.
+Figure 8 shows the result of fitting these data with the SIRS diffusion model described above.
+The MAE of this model is 43.07%, indicating poor model fit.
+
+![Five runs of the agent-based simulation used to produce data for analysis with the SIRS model of behavioral adoption under the success-biased learning scenario.
+The top figure shows the number of resources in both the conservation and unconserved areas over time, as a percent of the carrying capacity in those areas.
+The middle figure shows the average number of resources harvested (payoff) by each agent at each time step.
+The bottom figure shows the number of agents who have adopted the conservation behavior (enrolled) at every time step.
+The exact parameter combination used to produce this figure can be found in the supplemental material (S1)](./BadFitPanelA.png)
+
+![SIRS model fit to the average of five runs of the agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior.
+Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.
+Dynamics in this figure exemplify poor model fit (MAE = 43.07%) under primarily environmentally driven adoption.](./BadFitPanelB.png)
+
+To ensure that our modeling approach is generally robust against these types of ecologically-driven dynamics, we again ran five iterations of the agent-based model and assessed SIRS model fit for a wide range of parameter combinations.
+As shown in figure 9 and in our supplemental materials (S3), our approach yields reliable model fit, as assessed through MAE, across the spectrum of realistic parameter values.
+It is worth noting however that the SIRS model fails to capture trends in conservation adoption in boundary cases, such as when 80-90% of resources are designated for conservation (figure 9).
+This indicates that the method presented in this paper may not be appropriate for real-world application where adoption dynamics are expected to be heavily driven by ecological change rather than social dynamics.
+
+![Sensitivity analysis for the ability of the SIRS model to capture adoption dynamics under a variety of parameter combinations across a range of values for the resource regeneration rate and the percent of the resource designated for conservation.
+](./Sens1.png)
+
+# 3 - Discussion
+## 3.1 - Shortcomings of diffusion of innovations in community-based conservation
 As demonstrated in figures 5 & 6, the standard SIRS diffusion model can reliably capture the long-term patterns of behavioral adoption for community-based conservation interventions, even when adoption is contingent on feedbacks with ecological processes.
-Fitting this model to observed data on the uptake of a behavior can provide us with estimates of the rates of adoption, dropout, and readoption for a specific intervention and allow us to predict the final stable rate of adoption in the population from early in the project.
+Fitting this model to observed data on the uptake of a behavior can provide us with estimates of the rates of adoption, dropout, and willingness to retry a specific intervention and allow us to predict the final stable rate of adoption in the population from early in the project.
 These insights, while useful, are somewhat limited in their ability to prescribe improvements for an intervention.
-As highlighted by the quote in the beginning of this article, diffusion study can explain what is happening in a system, but gives little intuition as to the mechanisms behind why it is happening [@rogers_diffusion_2010].
+As highlighted by the quote in the beginning of this article, diffusion study can explain *what* is happening in a system, but gives little intuition as to the mechanisms behind *why* it is happening [@rogers_diffusion_2010].
 
-As first described by Henrich [-@henrich_cultural_2001] and demonstrated in figures 5 & 6, we see clear changes in the population patterns of adoption when individuals employ varying levels of social or independent learning.
+As first described by Henrich [-@henrich_cultural_2001] and demonstrated in figures 5 & 6, we see clear changes in the population patterns of adoption when individuals employ differing levels of social or independent learning.
 But when comparing many biased social learning mechanisms, the pattern of adoption may exhibit equifinality, where many different underlying mechanisms may produce similar or even identical patterns over time [@barrett_equifinality_2019].
@@
 
-Lastly, while we have shown here that DOI can inform community-based conservation efforts in silico, the real-world may constrain such productive application.
-In a sharp response to Mascia and Mills [-@mascia_when_2018], authors Lund and Blewstein [-@lund_when_2018] question the narrative of “conservation interventions as innovations that can be freely adopted.” They argue that top-down support and manipulation are common in these contexts and may hinder the applicability of “demand-driven” conceptualizations of how conservation practices spread.
-These criticisms are paralleled by issues raised by Rogers himself when identifying the pro-innovation and individual blame biases of DOI, which portray lack of adoption as a flaw of the individual, opposed to the innovation [@rogers_diffusion_2010].
-Given the problematic history of colonial-type conservation, we urge readers to seriously consider who is providing the narrative when assessing how and why conservation institutions diffuse within populations [@agrawal_enchantment_1999-1].
-
-# Future work
+## 3.2 - Future work
 Understanding the causal mechanisms behind the patterns of behavioral transmission is necessary to better delineate and predict how conservation will affect both people and the environment [@cheng_strengthen_2020].
-Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes based on realistic human cognitions which allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013; @turner_land-useland-cover_2021; @le_polain_de_waroux_role_2021].
-Modeling human behavior in the context of community-based conservation is thus going to require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [@schill_more_2019].
+Researchers are increasingly recognizing that models of human-caused environmental change must explicitly incorporate dynamic social processes based on realistic human cognitions that allow actors to modify their behaviors in response to new information [@meyfroidt_environmental_2013; @turner_land-useland-cover_2021; @le_polain_de_waroux_role_2021].
+Behavioral adoption is not truly binary, as implied by DOI.
+Rather, individuals modify behaviors they adopt in response to their personal experiences and needs [@enquist_modelling_2011].
+Research on cultural evolution has accordingly begun to incorporate these cognitive processes into formal models [@gabora_autocatalytic_2017; @gabora_modeling_2020].
+Hence, modeling human behavior in the context of community-based conservation is in its infancy and going forward will require a complex dynamic systems approach in which behaviors and the biophysical environment coevolve [@schill_more_2019].
 
@@
 We believe this is an exciting new avenue for conservation science as the direct interpretability and applicability of process parameters lend themselves well to tight integration with applied research as is common in the field.
-Future research may be interested in extending the SIRS diffusion model to include a term which represents change in the environment over time and the effect of resource levels on adoption rates to more accurately reflect the social-ecological coupling.
+Future research could extend the SIRS diffusion model to include an exposure term (i.e.
+SEIRS) to allow for individual variation in the exposure to an intervention required before adopting, possibly dependent on individual characteristics.
+
+An additional line of research aimed at better capturing the trends shown in figures 7 & 8 might include a term that represents change in the environment over time and the effect of resource levels on adoption rates to more accurately reflect the social-ecological coupling.
+As exemplified in figures 7 through 9, any model which predicts human behavior as a result of past behavioral dynamics will fail to capture trends driven primarily by environmental conditions.
 An example of such an extension is presented below, where we include another equation for the change in a resource pool over time $\frac{dE}{dt}$.
@@
 
-Coupled human and natural systems commonly exhibit dynamics with thresholds, resilience, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically [@liu_complexity_2007].
-In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems which are difficult or even impossible to account for with process models such the SIRS  or SIRS extension model presented in equation 4 [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
-Agent-based simulations allow modelers to impose a realism in the governance or institutional dynamics that are largely intractable in closed-form mathematical equations [@bravo_agents_2011; @joffre_combining_2015].
+Coupled human and natural systems commonly exhibit dynamics with thresholds, time lags, and heterogeneity not present in social or natural systems in isolation and therefore are difficult to model analytically [@liu_complexity_2007].
+In very complex cases such as these, we argue that agent-based simulations are the preferred approach as they allow for these features of complex systems that are difficult or even impossible to account for with analytical process models such the SIRS  or SIRS extension model presented in equation 4 [@reyers_social-ecological_2018; @railsback_agent-based_2012; @wilensky_introduction_2015].
+Agent-based simulations allow modelers to impose a realism in the governance or institutional dynamics that are largely intractable in closed-form mathematical equations [@kaiser_identifying_2020; @bravo_agents_2011; @joffre_combining_2015].
+Further, agent-based simulations uniquely allow researchers to hypothesize and test causal processes in complex social-ecological systems [@baggio_managing_2018; @schluter_toward_2019].
 As the number of purely subsistence communities shrinks each year worldwide, and multiple levels of institutions are increasingly involved in natural resource management, incorporating these features into our models is paramount for reliable inference [@mckean_common_2000; @armitage_adaptive_2009].
 
-# Conclusions
+# 4 - Conclusions
 Overall, we argue that fitting process models of behavioral diffusion to data on the spread of conservation behaviors is a practical way for researchers and practitioners to apply DOI to community-based conservation.
@@
 We show that this model can be a powerful tool for estimating the natural history characteristics of a particular intervention, such as the rate of spread, and for predicting the long-term patterns of adoption, even under reciprocal behavior-environment interactions.
-We also introduce the broader field of cultural evolution as a more comprehensive set of theories for understanding the underlying mechanisms of behavioral adoption than DOI alone.
+We also introduce the broader field of cultural evolution as a more comprehensive body of theory for understanding the underlying mechanisms of behavioral adoption than DOI alone.
 We hope that this review provides a model for how practitioners might practically apply insights from DOI to community-based conservation interventions, while also clarifying the position of DOI in the current literature on the adoption of sustainable resource management practices.
+
+# 5 - Software and data availability 
+The agent-based model used to simulate data for this paper was produced using the R programming language for statistical computing; R version 4.1.2 [@r_core_team_r_2021].
+These simulated data were analyzed by running the probabilistic programming language STAN within the Rstudio environment; STAN version 2.21.3 [@stan_development_team_stan_2021;@rstudio_team_rstudio_2020].
+All R and STAN code used in this project can be found at the Github repository here: https://github.com/matthewclark1223/A-Quantitative-Application-of-Diffusion-of-Innovations-for-Modeling-the-Spread-of-Conservation-Behav.
+
+# Supplemental 1
+
+ODD model description attached.
+
+# Supplemental 2
+
+![SIRS model fit to the average of 100 runs of the agent-based simulation data under success-biased social learning contingent on behavior-environment feedbacks shown as translucent black lines representing 1000 draws from the posterior.
+Simulated data are overlaid with gray points signifying training data (first 30%) and blue points signifying test (later 70%) data.](./100_RunFit.png)
+
+# Supplemental 3
+
+![Sensitivity analysis for the ability of the SIRS model to capture adoption dynamics under a variety of parameter combinations across a range of values for the starting resource integrity (percent of carrying capacity) and payment amount for agents.](./Sens2.png)
 

### 2022-07-21 DOI_Review/ApplySTANtoABMdat_MultiRun.R

@@ -1,0 +1,272 @@
+source("Binary_Adoption_ABM_MultiRun.R")
+
+# Load some required packages
+##############
+library(deSolve)
+library(dplyr)
+library(ggplot2)
+library(rstan)
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+##############
+
+###fitting Stan data
+
+sample_days<-30#30
+sample_n<-100
+t_max<-100
+sample_y<-ConformBiasdat[1:sample_days,]$Enrolled
+sample_time<-1:sample_days
+# Parameters must be stored in a named list.
+params <- list(beta = NA, #rate of transmission
+               gamma = NA, #rate of dropout
+               chi = NA) #time after dropout before willing to try again
+
+stan_d = list(n_obs = sample_days,
+              n_params = length(params),
+              n_difeq = 3, #number of diff eqs
+              n_sample = sample_n,
+              n_fake = length(1:t_max),
+              y = sample_y,
+              t0 = 0,
+              ts = sample_time,
+              fake_ts = c(1:t_max))
+
+params_monitor = c("y_hat", "y0", "params", "fake_I") #fake_I from generated quantities
+
+# Fit and sample from the posterior
+mod = stan("~/Pemba_Project/DOI_Review/SIRS.stan",
+           data = stan_d,
+           pars = params_monitor,
+           chains = 4,
+           warmup = 500,
+           iter = 1500)
+
+
+# You should do some MCMC diagnostics, including:
+traceplot(mod, pars="lp__")
+names(mod)[94:96]<-c("Beta","Gamma","Chi")
+
+traceplot(mod, pars=c("params"))+mytheme+theme(strip.text.x = element_text(size = 15))
+#summary(mod)$summary[,"Rhat"]
+
+# These all check out for my model, so I'll move on.
+
+# Extract the posterior samples to a structured list:
+posts <- rstan::extract(mod)
+#hist(posts$params[,1])
+#hist(posts$params[,2])
+
+#R0 number of infected people
+#R0<-(posts$params[,1]*0.20)/posts$params[,2]
+
+#length(which(R0<1))/length(R0)
+
+#plot it
+# Proportion infected from the synthetic data:
+sample_prop = sample_y / sample_n
+
+# Model predictions across the sampling time period.
[…]

### 2022-07-21 DOI_Review/Binary_Adoption_ABM_MultiRun.R

@@ -59,7 +59,7 @@ abm<-function(#Specified parameters
   
   ProtectPerDefect<-ifelse(NotEnrollNumStart ==0,as.integer(StartResourceProtected),as.integer(StartResourceProtected/NotEnrollNumStart)) #protected area resources per each individual harvesting there
   WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
-
+  
   
   for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
     agent2<-agents[i,]   #of resources there per individual searching there, and a random draw with a chance of success equal to time spent
@@ -69,7 +69,7 @@ abm<-function(#Specified parameters
     agents[i,4]<-ifelse(agents[i,4]>harvestMax-agents[i,5],harvestMax-agents[i,5],agents[i,4])
   } #see what they get from the working landscape first
   
-
+  
   agents$PayoffTotalLastTime<- agents$PayoffProtectedLastTime + agents$PayoffWorkingLastTime #total payoff is the sum of what they get from both areas
   
   
@@ -113,7 +113,7 @@ abm<-function(#Specified parameters
       if(sum(agentdf$Enrolled)==0){
         payMeanEnrl<-0}
       if(sum(agentdf$Enrolled)!=0){
-      payMeanEnrl<-mean(agentdf[agentdf$Enrolled==1,]$PayoffTotalLastTime)}
+        payMeanEnrl<-mean(agentdf[agentdf$Enrolled==1,]$PayoffTotalLastTime)}
       
       #meanPayoffUnenroll
       if(sum(agentdf$Enrolled)==nrow(agentdf)){
@@ -122,9 +122,9 @@ abm<-function(#Specified parameters
         payMeanNOTEnrl<-mean(agentdf[agentdf$Enrolled==0,]$PayoffTotalLastTime)} 
       
       probEnroll<-exp(BiasStrength*payMeanEnrl)/
-         (exp(BiasStrength*payMeanEnrl)+exp(BiasStrength*payMeanNOTEnrl))
-        
-        agentdf$Enrolled<-rbinom(nrow(agentdf),1, probEnroll)
+        (exp(BiasStrength*payMeanEnrl)+exp(BiasStrength*payMeanNOTEnrl))
+      
+      agentdf$Enrolled<-rbinom(nrow(agentdf),1, probEnroll)
       return(agentdf$Enrolled)
     }
     
[…]

### 2022-07-21 DOI_Review/Binary_Adoption_ABM_NO_Plots.R

@@ -59,7 +59,7 @@ abm<-function(#Specified parameters
   
   ProtectPerDefect<-ifelse(NotEnrollNumStart ==0,as.integer(StartResourceProtected),as.integer(StartResourceProtected/NotEnrollNumStart)) #protected area resources per each individual harvesting there
   WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
-
+  
   
   for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
     agent2<-agents[i,]   #of resources there per individual searching there, and a random draw with a chance of success equal to time spent
@@ -69,7 +69,7 @@ abm<-function(#Specified parameters
     agents[i,4]<-ifelse(agents[i,4]>harvestMax-agents[i,5],harvestMax-agents[i,5],agents[i,4])
   } #see what they get from the working landscape first
   
-
+  
   agents$PayoffTotalLastTime<- agents$PayoffProtectedLastTime + agents$PayoffWorkingLastTime #total payoff is the sum of what they get from both areas
   
   
@@ -113,7 +113,7 @@ abm<-function(#Specified parameters
       if(sum(agentdf$Enrolled)==0){
         payMeanEnrl<-0}
       if(sum(agentdf$Enrolled)!=0){
-      payMeanEnrl<-mean(agentdf[agentdf$Enrolled==1,]$PayoffTotalLastTime)}
+        payMeanEnrl<-mean(agentdf[agentdf$Enrolled==1,]$PayoffTotalLastTime)}
       
       #meanPayoffUnenroll
       if(sum(agentdf$Enrolled)==nrow(agentdf)){
@@ -122,9 +122,9 @@ abm<-function(#Specified parameters
         payMeanNOTEnrl<-mean(agentdf[agentdf$Enrolled==0,]$PayoffTotalLastTime)} 
       
       probEnroll<-exp(BiasStrength*payMeanEnrl)/
-         (exp(BiasStrength*payMeanEnrl)+exp(BiasStrength*payMeanNOTEnrl))
-        
-        agentdf$Enrolled<-rbinom(nrow(agentdf),1, probEnroll)
+        (exp(BiasStrength*payMeanEnrl)+exp(BiasStrength*payMeanNOTEnrl))
+      
+      agentdf$Enrolled<-rbinom(nrow(agentdf),1, probEnroll)
       return(agentdf$Enrolled)
     }
     
[…]

### 2022-07-21 DOI_Review/SensitivityAnalysis.R

@@ -59,7 +59,7 @@ abm<-function(#Specified parameters
   
   ProtectPerDefect<-ifelse(NotEnrollNumStart ==0,as.integer(StartResourceProtected),as.integer(StartResourceProtected/NotEnrollNumStart)) #protected area resources per each individual harvesting there
   WorkingPerTotal<-as.integer(StartResourceWorking/Individuals) #working landscape resources per individual
-
+  
   
   for ( i in 1:nrow(agents)){  #fill the starting df percent payoff from the protected landscape is a function of the amount 
     agent2<-agents[i,]   #of resources there per individual searching there, and a random draw with a chance of success equal to time spent
@@ -69,7 +69,7 @@ abm<-function(#Specified parameters
     agents[i,4]<-ifelse(agents[i,4]>harvestMax-agents[i,5],harvestMax-agents[i,5],agents[i,4])
   } #see what they get from the working landscape first
   
-
+  
   agents$PayoffTotalLastTime<- agents$PayoffProtectedLastTime + agents$PayoffWorkingLastTime #total payoff is the sum of what they get from both areas
   
   
@@ -113,7 +113,7 @@ abm<-function(#Specified parameters
       if(sum(agentdf$Enrolled)==0){
         payMeanEnrl<-0}
       if(sum(agentdf$Enrolled)!=0){
-      payMeanEnrl<-mean(agentdf[agentdf$Enrolled==1,]$PayoffTotalLastTime)}
+        payMeanEnrl<-mean(agentdf[agentdf$Enrolled==1,]$PayoffTotalLastTime)}
       
       #meanPayoffUnenroll
       if(sum(agentdf$Enrolled)==nrow(agentdf)){
@@ -122,9 +122,9 @@ abm<-function(#Specified parameters
         payMeanNOTEnrl<-mean(agentdf[agentdf$Enrolled==0,]$PayoffTotalLastTime)} 
       
       probEnroll<-exp(BiasStrength*payMeanEnrl)/
-         (exp(BiasStrength*payMeanEnrl)+exp(BiasStrength*payMeanNOTEnrl))
-        
-        agentdf$Enrolled<-rbinom(nrow(agentdf),1, probEnroll)
+        (exp(BiasStrength*payMeanEnrl)+exp(BiasStrength*payMeanNOTEnrl))
+      
+      agentdf$Enrolled<-rbinom(nrow(agentdf),1, probEnroll)
       return(agentdf$Enrolled)
     }
     
[…]

### 2022-07-21 DOI_Review/SvsR_CurvePlot.R

@@ -60,7 +60,7 @@ mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_b
                 plot.title = element_text( size=18, color="black",face="bold"),
                 axis.title.x = element_text( size=18),
                 axis.title.y = element_text( size=18),
-                axis.text=(element_text(color="black", size=14)))
+                axis.text=(element_blank()))
 
 
 ggplot(df,aes(x=time,y=I))+geom_line(aes(by=as.character(rate),color=rate),size=1.5,alpha=0.9)+ 
@@ -83,6 +83,7 @@ ggplot(df2,aes(x=time,y=I))+geom_line(color="#3690c0",size=1.5,alpha=0.99)+
   annotate( "text", label = "S curve",x = 45, y = 0.52, size = 8, colour = "#3690c0")+
   annotate( "text", label = "R curve",x = 20, y = 0.75, size = 8, colour = "black")+
   theme(legend.position = "none")+xlab("Time")+ylab("Proportion Adopted")+mytheme+
+  theme(axis.ticks = element_blank())+
   scale_x_continuous(expand = c(0, 2)) + scale_y_continuous(expand = c(0, 0))
 
 

### 2022-11-02 HCRI_Grant/Manuscript.Rmd
Uploading markdown doc for manuscript

---
title: Causal attribution of agricultural expansion in a small island system using approximate Bayesian computation 
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: Max Planck Institute for Evolutionary Anthropology
  name: Jeffrey Andrews
- affiliation: Boise State University - Human Environment Systems
  name: Nicholas Kolarik
- affiliation: Community Forests Pemba
  name: Mbarouk Mussa Omar
- affiliation: Boise State University - Human Environment Systems
  name: Vicken Hillis
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
  word_document: default
bibliography: ./HCRIGrantPaper.bib
abstract: The extent and arrangement of land cover types on our planet directly affects biodiversity, carbon storage, water quality, and many other critical social and ecological conditions at virtually all scales. Given this critical importance of land cover, a key mandate for land system scientists is to understand drivers of change. Identifying causal drivers of change is challenging however, because land systems, such as small-scale agricultural communities, do not lend themselves well to controlled experimentation for logistical and ethical reasons. Even natural experiments in these systems can produce only limited causal inference as they often contain unobserved confounding drivers of land cover change and complex feedbacks between drivers and outcomes. Land system scientists commonly grapple with this complexity by using computer simulations to explicitly delineate hypothesized causal pathways that could have resulted in observed land cover change. Yet, land system science lacks a systematic method for comparing multiple hypothesized pathways and quantifying the probability that a given simulated causal process was in fact responsible for the patterns observed. Here we use a case study of agricultural expansion in Pemba, Tanzania to demonstrate how approximate Bayesian computation provides a straightforward solution to this methodological gap. Specifically, we pair an individual-based simulation of land cover change in Pemba with approximate Bayesian computation to probabilistically estimate the likelihood that observed deforestation from 2018 to 2021 was driven by soil degradation versus external market forces. Using this approach, we can show not only how well a specific hypothesized mechanism fits with empirical data on land cover change, but we also quantify the range of other mechanisms that could have reasonably produced the same outcome (i.e. equifinality). We argue that approximate Bayesian computation is particularly promising as a tool for causal inference for land system science given the wealth of data available in the satellite record. Thus, this paper is a first step toward a robust process for identifying the emergent landscape-level signatures of complex social-ecological mechanisms. 
fontsize: 12pt
link-citations: yes
keywords: Land system science, Individual-based modeling, Complex systems
header-includes:
- \usepackage{pdfpages}
- \usepackage[font={footnotesize}]{caption}
- \usepackage{bm}
---

```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
#
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)
```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)

## Caching code
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')

setwd("C:/Users/Matt/Documents/Pemba_Project/HCRI_Grant")
```

# 1 - Introduction
The mosaic of land cover on the surface of our planet is a product of complex social-ecological dynamics that make up the complete land system [@turner_emergence_2007; @turner_land-useland-cover_2021]. Changes in Earth’s terrestrial surface have profound implications for ecosystem functioning and human wellbeing, and as such, are of critical importance to understand and predict [@steffen_global_2006]. A key challenge in understanding land cover change and designing effective policies is that there are often multiple plausible, and even interacting mechanisms that can cause a switch in land cover from one state to another [@lambin_land_2010]. For example, in the case of agricultural frontier expansion, depletion of soil fertility in existing plots often promotes the conversion of nearby natural vegetation to new cropland, thus pushing the frontier outward [@casetti_formalization_1977]. Agricultural frontier expansion into forested lands is also observed however, as a result of increased market value for crops and increased population pressure [@meyfroidt_middle-range_2018].

When designing policies to limit the conversion of natural areas into more intensive land use types such as rotational agriculture, it is important to determine the drivers of conversion, because the impact of a given policy will depend on the dominant driver. The introduction of new agricultural technologies, for instance, can have drastically different effects on the landscape depending on the primary driver of frontier expansion in a given system [see Kaimowitz et al. -@kaimowitz_economic_1998 for a foundational review]. If loss of soil fertility is the primary driver, new technologies can limit forest conversion as they allow existing plots to be farmed for longer periods of time, minimizing the need for agricultural operations to change location. However, if market forces are the primary driver of frontier expansion, this same intervention is likely to incentivize forest conversion by increasing the returns from any given agricultural plot [@meyfroidt_middle-range_2018]. Despite the importance of identifying causal processes in land cover change, actually doing so in any particular case has often proven difficult given the inherent social-ecological complexity of land systems [@meyfroidt_approaches_2016; @turner_framing_2020].

Land systems are complex adaptive systems characterized by feedbacks between the human and ecological subsystems, where a change in the state of one affects ongoing processes in the other and vice versa [@berkes_linking_2000; @le_feedback_2012; @folke_socialecological_2007]. The possible distribution of land cover types in any one area is also highly path dependent, or constrained by past states and trajectories, sometimes further muddying the relationship between actual drivers and outcomes [@liu_complexity_2007]. Standard statistical tests fail to produce reliable inference in complex systems exhibiting feedbacks and path dependence, given that it is generally impossible to specify likelihood functions for such processes [@levin_social-ecological_2013]. Thus, while critical, identifying causal processes in land systems and social-ecological systems generally has proven difficult.

To begin to build causal theory in complex land systems, researchers commonly use computer simulations to abstract key phenomena and produce ‘what-if’ scenarios [@ahimbisibwe_understanding_2021; @an_challenges_2021]. Simulations allow researchers to code complexities such as feedbacks and path dependence directly into a model of the processes under examination, in order to formally define a hypothesized causal mechanism and check the logical implications and internal validity of their assumptions [@epstein_why_2008; @verburg_simulating_2006; @schluter_toward_2019]. While simulations like this are important for theorizing about social-ecological systems, it can be difficult to relate them back to empirical data and tell where exactly the real-world sits in the multidimensional parameter space of the model [@ren_spatially_2019]. Without this information, we are limited in our knowledge of how well a given simulation accurately distills the processes we are hoping to examine, and how we might use such a model to infer important things about the real-world.

Ecological research has made considerable use of simulation modeling to theorize about how complex interactions among individuals lead to observed patterns at the population level (i.e. individual-based modeling) [@grimm_ten_1999; @deangelis_individual-based_2014; @grimm_individual-based_2013]. Relatively early on in the use of individual-based modeling in ecology, researchers developed the general process of pattern-oriented modeling for relating observed patterns in ecological data to simulated hypothesized causal mechanisms [@wiegand_using_2003; @grimm_pattern-oriented_2005]. This method allowed researchers to match observed trends in population change with plausible rates for various demographic parameters such as pre-breeding survival in woodpeckers, road mortality in lynx, and annual male survival in amphibians, among many others [@rossmanith_pattern-oriented_2007; @kramer-schadt_fragmented_2004; @swanack_projecting_2009, @wiegand_using_2003]. This general method was a significant step toward the development of a systematic process for interrogating the hypothesized causal mechanisms built into computer simulations with observed data on emergent phenomena.

More recently, approximate Bayesian computation (ABC) has emerged as a formal method of pattern-oriented modeling in which researchers run simulation models across many parameter values and systematically accept or reject the outputs of each run as consistent with observed data. All accepted parameter values are then aggregated into a probability distribution of parameter values that are likely to produce the observed data [@hartig_statistical_2011; @troost_how_2022]. ABC has proven effective for identifying the range of simulation model parameters consistent with observed data across a wide breadth of biological fields from ecology to epidemiology [@scranton_approximate_2014;@kosmala_estimating_2016; @van_der_vaart_predicting_2016; @boult_individual-based_2018; @martinez_disentangling_2011; @cipriotti_understanding_2012]. When differing simulation parameters coincide with specific hypothesized causal processes, ABC can then be used to estimate the probability that a given causal process produced a given set of observed data. Importantly, ABC enables researchers to statistically estimate model parameters even for generative models containing complex processes such as the feedbacks and path dependence characteristic of social-ecological systems [@gallagher_theory_2021].

In this paper we demonstrate the utility of ABC for generative inference in complex land systems and social-ecological systems generally. Specifically, we simulate hypothesized patterns of agricultural expansion in a small island system given two possible drivers, declining soil fertility and external market forces. We then filter the range of possible model parameters to just the inputs that produce land use patterns consistent with the observed time-series of agricultural frontier expansion. We show that this method allows us to determine the proportion of each of these drivers in causing the observed agricultural expansion in our study system, Pemba Island, Tanzania. Thus, this paper provides a straightforward demonstration for linking land system simulations with empirical data to draw causal inference in even very complex systems involving feedbacks and path dependence.

# 2 - Study system

## 2.1 - Pemba
The Indian Ocean archipelago of Zanzibar is a semi-autonomous jurisdiction lying off the coast of Tanzania. Pemba, the northernmost island, is densely populated with (XXX) people per square kilometer. While the island has a few main population centers (Wete, Chake Chake, and Mkoani), the vast majority of the island's 400,000 people live in small villages scattered across some 120 wards (shehia), all of which are connected by a dense, relatively modern road network (fig. 1). Based on the 2022 census, we estimate the average growth rate between 2012 and 2022 to be about 2.1% which is more than double the world average of 0.9% per year.

![Caption ](./PaperFigures/OverviewMap.png)

The rural economy of Pemba is primarily subsistence (farming, fishing, and livestock),
with a single important cash crop, cloves. Cloves were introduced to the island in the early 19th century and have come to dominate the economy of the island's western half due to the region's highly productive soils (see below). In contrast, the island's eastern side, characterized by poor, shallow soil, has lower human population density, is less developed, and is generally less economically productive. For example, in an extensive household survey focused on economic production carried out in 2017, we find that the average household income from the sale of crops (excluding cloves) in the east is approximately $25, compared with an estimated $80 in the highly productive western half.

Nevertheless, there is considerable pressure in the east for new farmland. Some 30% of
All eastern households surveyed stated that they had cleared forested land to expand their farming operations in the past 7 years. On average, households report clearing approximately 0.91 acres, primarily to plant staple crops such as cassava. And while cassava is almost exclusively a subsistence crop in Pemba, there has been considerable pressure to develop the Pemban economy in the past decade, as it has lagged behind the rest of Tanzania and Zanzibar. New development initiatives, particularly in agriculture, are a constant of government programming, and new crops such as watermelon and tomatoes are being experimented with on the once underutilized eastern soils. However, Pemba has historically struggled to develop its own internal market for agricultural goods, and the impact that these new cash crops are having on the eastern landscape is unknown.

## 2.2 - Coral rag vegetation and rotational agriculture
Pemba is a narrow island, in many places just 15 kilometers wide, yet most of the environmental variation exists across the narrow east/west span. This is owed to three distinct soil types that run the length of the island and can generally be thought of as going from deep and fertile in the west, to shallow and nutrient poor in the east [@stockley_report_1928]. The easternmost topography is characterized by jagged, fossilized coral beds covered with a shallow soil layer and scrubby vegetation ranging from approximately 1 to 5 meters in height (i.e. coral rag forest) [@burgess_coastal_2000; @burrows_trees_2018]. This forest type has traditionally been overlooked by conservation efforts in Pemba, and Zanzibar generally, yet it is critical habitat for a variety of plant and animal species such as the endemic Pemba flying fox (*Pteropus voeltzkowi*) [@kingdon_east_1988].

![Caption ](./Photos/CoralRagCompositeReal.png)

The shallow soils atop porous coral rag geology characteristic to eastern Pemba cannot be productively farmed for long, continuous periods. Farmers in this region thus typically rely on rotational swidden agriculture where primary forest is cleared and the land is farmed for a short period, then left to recharge for a number of years (fig 2). While this process could potentially be interrupted by regenerative agricultural practices or the introduction of rainwater catchment systems, currently farmers in this region lack resources to escape the cycle of ecologically costly, short-term resource extraction [@leal_filho_using_2020; @biazin_rainwater_2012; @garrity_evergreen_2010]. Thus, in this system, there are considerable feedbacks between the condition of the environment, decisions made by farmers to clear and farm a forested area, and the state of the environment in future time periods. Further, as any one forested area is cleared, it opens new patches to potential clearing through frontier expansion. Hence, specific patches available in any one time period are highly dependent on social-ecological occurrences in previous time steps. Thus, this system displays both the feedbacks and path dependence characteristic of complex social-ecological systems.

# 3 - Methods

## 3.1 - Data

### 3.1.1 - Land cover classification
We produced 20 m land cover maps using top of atmosphere Sentinel-2 time series imagery and ancillary datasets in a data fusion approach using Google Earth Engine (GEE) [@gorelick_google_2017; @mondal_evaluating_2019]. GEE is an open-access cloud computing platform that hosts petabytes of freely available earth observation data, and is ideal for creating land cover maps with built-in classification functions. For this study, we created annual median composite images for conversion into thematic land cover maps. We filtered the time series by date, and used the image metadata to further filter by estimated cloud cover, using a threshold of <20%. We then used the ‘QA60’ quality band to remove any remaining clouds prior to creating the composites of median values. Beyond the native Sentinel-2 spectral bands, we calculated common normalized difference indices useful for distinguishing common land cover types such as the Normalized Difference Vegetation Index (and red edge adaptations) and the Normalized Difference Water Index (NDWI) [@gao_ndwinormalized_1996] (as well as its modified version (MNDWI) [@defries_ndvi-derived_1994; @schuster_testing_2012; @xu_modification_2006]. We also used synthetic aperture radar (SAR) backscatter from the corresponding Sentinel-1 ground range detected time series available in GEE. We used SAR scenes from ascending paths only and incorporated both VV and VH polarizations in our analysis. SAR data are known to be influenced by varying incidence angles, so we normalized these images by multiplying the backscatter by the incidence angle with the understanding that greater incidence angles result in less backscatter returned to the instrument [@banks_wetland_2019; @kaplan_normalizing_2021]. To reduce inherent speckle in the SAR images, we opted for a time for space substitution by using a mean composite of all images for the given year to maintain a 10 m spatial resolution. Lastly, we considered topographic covariates for classification (elevation, slope, and aspect derived from the NASA Shuttle Radar Topography Mission digital elevation model) that dictate locations of land covers of interest relative to sea level and topography (Table S1) [@farr_shuttle_2007].

![Caption ](./PaperFigures/LandCoverMap.png)

From each composite image, we collected representative training samples for all classes of interest (mangrove, high forest, agriculture, urban, bare, coral rag, other woody vegetation, and water). We trained a random forest classifier using training samples from 2018, 2019, & 2021 to account for potentially varying atmospheric and illumination conditions among images [@breiman_random_2001]. The random forest we used for classification had 100 trees, and utilized four variables per split (the square root of the number of covariates), consistent with other remote sensing applications [@belgiu_random_2016]. Due to relative class imbalances, we chose to use a stratified random sampling design to assess the accuracy of our outputs, and computed area adjusted accuracy metrics [@olofsson_making_2013; @stehman_key_2019]. An expert hand labeled these stratified points based on high resolution median composite PlanetScope images with 20 points from each mapped class. Results show estimated overall accuracies of 92.86% (+/-4.21%) and 95.93% (+/- 2.95%) for 2018 and 2021, respectively (Table S2, Table S4). Much of the confusion and sources of error in the maps is found among upland woody vegetation classes (High forest, Mangrove, and Other woody vegetation) that share similar spectral and physical characteristics. Other mentionable errors occur among urban and agriculture in 2018, and bare and agriculture in 2021, leading to relatively high uncertainty for area estimates and producer’s accuracies in the rare urban and bare classes in 2018 and 2021, respectively (Table S3, Table S5). (GEE link)

### 3.1.1 - Interview data

In July 2021, we conducted informal interviews with staff of Community Forests Pemba, a nonprofit aimed at building conservation capacity on the island, and farmers in four shehia with coral rag forests in the east of the island. Researchers asked farmers about how they make decisions regarding when and how long to farm and fallow agricultural plots, as well as how they decide to clear forest vegetation to establish new cropland. There was broad consensus among non-profit staff and farmers that agricultural plots in these areas are typically farmed for two years and then left fallow for three years. Clearing forested land is labor intensive and thus, forested land is only cleared when nearby productive plots reach their two-year limit and no other previously-cleared adjacent plots are available.

## 3.2 - Individual-based simulation

![Caption ](./PaperFigures/ABMABC_ConceptDiagram.png)

## 3.3 - Approximate Bayesian computation (ABC)

![Caption ](./PaperFigures/Prior.png)

# 4 - Results 
## 4.1 - Parameter estimation

![Caption ](./PaperFigures/EstimatesPixels.png)

![Caption ](./PaperFigures/Estimates.png)

## 4.2 - Estimating the contribution of each process

![Caption ](./PaperFigures/Contribution.png)

# 5 - Discussion

## 5.1 - Pemba specific implications

## 5.2 - Implications for land system science and SES

## 5.3 - Limitations and future work

# 6 - Conclusion

# 7- Acknowledgements
We thank the entire Community Forests Pemba staff for continued support on research in Pemba. We also thank the Hazards and Climate Resilience Institute at Boise State University for funding this research. Lastly, we thank the Max Planck Institute for Evolutionary Anthropology, department of Human Behavior, Ecology and Culture for providing the computing resources necessary for this project.

# 8 - Open data and software

# Supplemental 1

ODD model description attached.

# Supplemental 2

# Supplemental 3

\newpage

# References

### 2024-02-27 HCRI_Grant/Manuscript.Rmd
Some updates

@@
 output:
+  word_document: default
   pdf_document:
@@
     template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
-  word_document: default
 bibliography: ./HCRIGrantPaper.bib
 abstract: The extent and arrangement of land cover types on our planet directly affects biodiversity, carbon storage, water quality, and many other critical social and ecological conditions at virtually all scales.
-Given this critical importance of land cover, a key mandate for land system scientists is to understand drivers of change.
+Given the fundamental importance of land cover, a key mandate for land system scientists is to describe the mechanisms by which pertinent cover types spread and shrink.
 Identifying causal drivers of change is challenging however, because land systems, such as small-scale agricultural communities, do not lend themselves well to controlled experimentation for logistical and ethical reasons.
@@
 Yet, land system science lacks a systematic method for comparing multiple hypothesized pathways and quantifying the probability that a given simulated causal process was in fact responsible for the patterns observed.
-Here we use a case study of agricultural expansion in Pemba, Tanzania to demonstrate how approximate Bayesian computation provides a straightforward solution to this methodological gap.
-Specifically, we pair an individual-based simulation of land cover change in Pemba with approximate Bayesian computation to probabilistically estimate the likelihood that observed deforestation from 2018 to 2021 was driven by soil degradation versus external market forces.
-Using this approach, we can show not only how well a specific hypothesized mechanism fits with empirical data on land cover change, but we also quantify the range of other mechanisms that could have reasonably produced the same outcome (i.e.
+Here we use a case study of agricultural expansion in Pemba, Tanzania to demonstrate how approximate Bayesian computation (ABC) provides a straightforward solution to this methodological gap.
+Specifically, we pair an individual-based simulation of land cover change in Pemba with ABC to probabilistically estimate the likelihood that observed deforestation from 2018 to 2021 was driven by soil degradation rather than external market forces.
+Using this approach, we can show not only how well a specific hypothesized mechanism fits with empirical data on land cover change, but we can also quantify the range of other mechanisms that could have reasonably produced the same outcome (i.e.
 equifinality).
-We argue that approximate Bayesian computation is particularly promising as a tool for causal inference for land system science given the wealth of data available in the satellite record.
-Thus, this paper is a first step toward a robust process for identifying the emergent landscape-level signatures of complex social-ecological mechanisms.
+While ABC was developed for use in population genetics, we argue that it is particularly promising as a tool for causal inference for land system science given the wealth of data available in the satellite record.
+Thus, this paper demonstrates a robust process for identifying the emergent landscape-level signatures of complex social-ecological mechanisms.
 fontsize: 12pt
@@
 If loss of soil fertility is the primary driver, new technologies can limit forest conversion as they allow existing plots to be farmed for longer periods of time, minimizing the need for agricultural operations to change location.
-However, if market forces are the primary driver of frontier expansion, this same intervention is likely to incentivize forest conversion by increasing the returns from any given agricultural plot [@meyfroidt_middle-range_2018].
+However, if market forces are the primary driver of frontier expansion, this same intervention is likely to incentivize further forest conversion by increasing the returns from any given agricultural plot [@meyfroidt_middle-range_2018].
 Despite the importance of identifying causal processes in land cover change, actually doing so in any particular case has often proven difficult given the inherent social-ecological complexity of land systems [@meyfroidt_approaches_2016; @turner_framing_2020].
@@
 To begin to build causal theory in complex land systems, researchers commonly use computer simulations to abstract key phenomena and produce ‘what-if’ scenarios [@ahimbisibwe_understanding_2021; @an_challenges_2021].
-Simulations allow researchers to code complexities such as feedbacks and path dependence directly into a model of the processes under examination, in order to formally define a hypothesized causal mechanism and check the logical implications and internal validity of their assumptions [@epstein_why_2008; @verburg_simulating_2006; @schluter_toward_2019].
+Simulations allow researchers to code complexities such as feedbacks and path dependence directly into a model of the processes under examination, in order to formally define a hypothesized causal mechanism and check the logical implications and internal validity of their assumptions [@epstein_why_2008; @verburg_simulating_2006].
 While simulations like this are important for theorizing about social-ecological systems, it can be difficult to relate them back to empirical data and tell where exactly the real-world sits in the multidimensional parameter space of the model [@ren_spatially_2019].
@@
 
-Ecological research has made considerable use of simulation modeling to theorize about how complex interactions among individuals lead to observed patterns at the population level (i.e.
+The biological sciences have largely led the development of methods for comparing simulated, theoretical causal processes to observed data.
+Ecological research in particular has made considerable use of simulation modeling to theorize about how complex interactions among individuals lead to observed patterns at the population level (i.e.
 individual-based modeling) [@grimm_ten_1999; @deangelis_individual-based_2014; @grimm_individual-based_2013].
-Relatively early on in the use of individual-based modeling in ecology, researchers developed the general process of pattern-oriented modeling for relating observed patterns in ecological data to simulated hypothesized causal mechanisms [@wiegand_using_2003; @grimm_pattern-oriented_2005].
+Relatively early on in the use of individual-based modeling in ecology, researchers developed the general process of pattern-oriented modeling in which a given hypothesis is evaluated on its ability to recreate an observed biological pattern at an appropriate temporal and spatial scale [@wiegand_using_2003; @grimm_pattern-oriented_2005].
 This method allowed researchers to match observed trends in population change with plausible rates for various demographic parameters such as pre-breeding survival in woodpeckers, road mortality in lynx, and annual male survival in amphibians, among many others [@rossmanith_pattern-oriented_2007; @kramer-schadt_fragmented_2004; @swanack_projecting_2009, @wiegand_using_2003].
-This general method was a significant step toward the development of a systematic process for interrogating the hypothesized causal mechanisms built into computer simulations with observed data on emergent phenomena.
+While pattern-oriented modeling provides a general structure for interrogating causal hypotheses of complex phenomena with empirical data, it does not adequately account for stochasticity in the outcomes of hypothesized mechanisms.
+In particular, this method fails to quantify the complete range of model parameters that may reasonably reproduce observed patterns and the frequency in which they do so.
 
-More recently, approximate Bayesian computation (ABC) has emerged as a formal method of pattern-oriented modeling in which researchers run simulation models across many parameter values and systematically accept or reject the outputs of each run as consistent with observed data.
+Toward this aim, approximate Bayesian computation (ABC) has emerged as a formal method of pattern-oriented modeling in which researchers run simulation models across many parameter values and systematically accept or reject the outputs of each run as consistent with observed data.
 All accepted parameter values are then aggregated into a probability distribution of parameter values that are likely to produce the observed data [@hartig_statistical_2011; @troost_how_2022].
 ABC has proven effective for identifying the range of simulation model parameters consistent with observed data across a wide breadth of biological fields from ecology to epidemiology [@scranton_approximate_2014;@kosmala_estimating_2016; @van_der_vaart_predicting_2016; @boult_individual-based_2018; @martinez_disentangling_2011; @cipriotti_understanding_2012].
-When differing simulation parameters coincide with specific hypothesized causal processes, ABC can then be used to estimate the probability that a given causal process produced a given set of observed data.
+When different simulation parameters represent specific hypothesized causal processes, ABC can then be used to estimate the probability that a given causal process produced a given set of observed data.
 Importantly, ABC enables researchers to statistically estimate model parameters even for generative models containing complex processes such as the feedbacks and path dependence characteristic of social-ecological systems [@gallagher_theory_2021].
@@
 We show that this method allows us to determine the proportion of each of these drivers in causing the observed agricultural expansion in our study system, Pemba Island, Tanzania.
-Thus, this paper provides a straightforward demonstration for linking land system simulations with empirical data to draw causal inference in even very complex systems involving feedbacks and path dependence.
+In this way, we provide a straightforward demonstration for linking land system simulations with empirical data to draw causal inference in even very complex systems involving feedbacks and path dependence.
 
@@
 The Indian Ocean archipelago of Zanzibar is a semi-autonomous jurisdiction lying off the coast of Tanzania.
-Pemba, the northernmost island, is densely populated with (XXX) people per square kilometer.
-While the island has a few main population centers (Wete, Chake Chake, and Mkoani), the vast majority of the island's 400,000 people live in small villages scattered across some 120 wards (shehia), all of which are connected by a dense, relatively modern road network (fig.
+Pemba, the northernmost island, is densely populated with 428 people per square kilometer.
+While the island has a few main population centers (Wete, Chake Chake, and Mkoani), the vast majority of the island's 400,000 people live in small villages scattered across some 120 wards (*shehia*), all of which are connected by a dense, relatively modern road network (fig.
 1).
-Based on the 2022 census, we estimate the average growth rate between 2012 and 2022 to be about 2.1% which is more than double the world average of 0.9% per year.
+Based on the 2022 census, we estimate the average growth rate between 2012 and 2022 to be about 2.1%, more than double the world average of 0.9% per year.
 
-![Caption ](./PaperFigures/OverviewMap.png)
+![Inset map showing the location of Pemba island relative to the Tanzanian mainland (upper left) and the dense Pemban road network (main figure).](./PaperFigures/OverviewMap.jpg)
 
@@
 In contrast, the island's eastern side, characterized by poor, shallow soil, has lower human population density, is less developed, and is generally less economically productive.
-For example, in an extensive household survey focused on economic production carried out in 2017, we find that the average household income from the sale of crops (excluding cloves) in the east is approximately $25, compared with an estimated $80 in the highly productive western half.
+For example, in an extensive household survey focused on economic production carried out in 2017, we find that the average household income from the sale of crops (excluding cloves) in the east is approximately \$25, compared with an estimated \$80 in the highly productive western half.
 
@@
 
-![Caption ](./Photos/CoralRagCompositeReal.png)
+![Panels A, B, C, and D show the process of coral rag forest conversion to agriculture in Pemba in four distinct stages.
+Panel A shows the cutting of coral rag vegetation, which is then left to dry before burning.
+Panel B shows a freshly burned plot before planting.
+Panel C shows a productive agricultural field.
+Panel D shows a fallow agricultural field.](./Photos/CoralRagCompositeReal.png)
 
@@
 Further, as any one forested area is cleared, it opens new patches to potential clearing through frontier expansion.
-Hence, specific patches available in any one time period are highly dependent on social-ecological occurrences in previous time steps.
+Hence, specific patches available in any one time period are highly dependent on social-ecological events in previous time steps.
 Thus, this system displays both the feedbacks and path dependence characteristic of complex social-ecological systems.
@@
 We then used the ‘QA60’ quality band to remove any remaining clouds prior to creating the composites of median values.
-Beyond the native Sentinel-2 spectral bands, we calculated common normalized difference indices useful for distinguishing common land cover types such as the Normalized Difference Vegetation Index (and red edge adaptations) and the Normalized Difference Water Index (NDWI) [@gao_ndwinormalized_1996] (as well as its modified version (MNDWI) [@defries_ndvi-derived_1994; @schuster_testing_2012; @xu_modification_2006].
+Beyond the native Sentinel-2 spectral bands, we calculated common normalized difference indices useful for distinguishing common land cover types such as the Normalized Difference Vegetation Index (and red edge adaptations) and the Normalized Difference Water Index (NDWI), as well as its modified version (MNDWI) [@defries_ndvi-derived_1994; @schuster_testing_2012; @xu_modification_2006; @gao_ndwinormalized_1996].
 We also used synthetic aperture radar (SAR) backscatter from the corresponding Sentinel-1 ground range detected time series available in GEE.
-We used SAR scenes from ascending paths only and incorporated both VV and VH polarizations in our analysis.
+We used SAR scenes from ascending paths only and incorporated both vertical-vertical and vertical-horizontal polarizations in our analysis.
 SAR data are known to be influenced by varying incidence angles, so we normalized these images by multiplying the backscatter by the incidence angle with the understanding that greater incidence angles result in less backscatter returned to the instrument [@banks_wetland_2019; @kaplan_normalizing_2021].
@@
 
-![Caption ](./PaperFigures/LandCoverMap.png)
+![Land cover estimates for the year 2018 in Pemba.
+The classification for each 20m pixel is distinguished by color, with the class of interest, coral rag vegetation, highlighted in pink.
+Black lines show each of the 19 *shehia* included in this study.](./PaperFigures/LandCoverMap.jpg)
 
@@
 An expert hand labeled these stratified points based on high resolution median composite PlanetScope images with 20 points from each mapped class.
-Results show estimated overall accuracies of 92.86% (+/-4.21%) and 95.93% (+/- 2.95%) for 2018 and 2021, respectively (Table S2, Table S4).
-Much of the confusion and sources of error in the maps is found among upland woody vegetation classes (High forest, Mangrove, and Other woody vegetation) that share similar spectral and physical characteristics.
-Other mentionable errors occur among urban and agriculture in 2018, and bare and agriculture in 2021, leading to relatively high uncertainty for area estimates and producer’s accuracies in the rare urban and bare classes in 2018 and 2021, respectively (Table S3, Table S5).
-(GEE link)
+Results show estimated overall accuracies of 92.86% (+/-4.21%) and 95.93% (+/- 2.95%) for 2018 and 2021, respectively (Table S2, Table S3).
+Much of the confusion and sources of error in the maps is found among upland woody vegetation classes (high forest, mangrove, and other woody vegetation) that share similar spectral and physical characteristics.
+Other mentionable errors occur among urban and agriculture in 2018, and bare and agriculture in 2021, leading to relatively high uncertainty for area estimates and producer’s accuracies in the rare urban and bare classes in 2018 and 2021, respectively (Table S4, Table S5).
 
 ### 3.1.1 - Interview data
-
-In July 2021, we conducted informal interviews with staff of Community Forests Pemba, a nonprofit aimed at building conservation capacity on the island, and farmers in four shehia with coral rag forests in the east of the island.
+In July 2021, we conducted informal interviews with staff of Community Forests Pemba, a nonprofit aimed at building conservation capacity on the island, and farmers in four *shehia* with coral rag forests in the east of the island.
 Researchers asked farmers about how they make decisions regarding when and how long to farm and fallow agricultural plots, as well as how they decide to clear forest vegetation to establish new cropland.
@@
 
-![Caption ](./PaperFigures/ABMABC_ConceptDiagram.png)
+Our individual-based model incorporates two mechanisms of land conversion from coral rag vegetation to productive agricultural land.
+The first follows the decision rules described during the farmer interviews where coral rag vegetation is converted in response to soil degradation and space limitations following the fallowing of cropland.
+Under this mechanism, each pixel (20m area) is autonomous and follows the following basic set of decision rules also described visually in figure 4.
+Each pixel classified as agriculture in the study *shehia* is initialized randomly as either productive or fallow.
+Each productive agricultural pixel is then randomly assigned to either the first or second year of agricultural production.
+Each fallow agricultural pixel is randomly assigned as in the first, second, or third year of fallow time.
+When transitioning to the next year, agricultural pixels in their second year of production go fallow and the lost agricultural production is relocated as follows.
+If there is an adjacent (8 directional) fallow pixel in the final (third) year, the crop production moves there.
+If there is no adjacent third year fallow pixel available and there is adjacent coral rag vegetation, that coral rag pixel is converted to first year productive agriculture for the next year.
+All other land cover types (e.g.
+water, urban, mangrove, etc.) are left alone.
+
+![Visual representation of the individual-based simulation used to model coral rag vegetation conversion to agriculture in each of the 19 *shehia* included in this study.
+The top two boxes show the competing causal forces driving land conversion.
+The bottom portion of the figure shows how those forces affect the pixel-based land cover in each time step.](./PaperFigures/ABMABC_ConceptDiagram.png)
+
+The second mechanism of land conversion from coral rag vegetation to productive agricultural land represents all other factors driving land conversion outside of the soil degradation process.
+In our system, the most prominent other factors include a rapidly growing population and the increasing market value for crops.
+Under this mechanism, some additional percentage of coral rag vegetation is converted to first year agriculture each year, representing a yearly rate of forest loss caused by factors other than soil degradation.
+Coral rag pixels allocated to conversion are those that are adjacent to the greatest number of agricultural pixels.
+When coral rag pixels are adjacent to an equal number of agricultural pixels, then the coral rag pixels closest to a road are selected for conversion to first year agricultural land.
+A visual representation of this simulation can be found in figure 4.
+
+For each run of this simulation, the observed land cover on Pemba Island in 2018 (fig.
+3) is used as the starting year and the model is run for three time steps to yield a predicted land cover for 2021, given a set rate of externally driven agricultural expansion.
+By then comparing this predicted land cover in 2021 to the observed land cover in 2018, we can produce a predicted number of coral rag pixels to be converted to agriculture under different rates of externally driven land conversion.
 
 ## 3.3 - Approximate Bayesian computation (ABC)
+As described in the introduction, simulation modeling allows researchers to formally express complex, hypothesized causal mechanisms in land system science.
+The primary limitation for the use of simulation modeling to enhance our understanding of real-world causality, however, is the absence of a straightforward statistical process for relating simulations to empirical data.
+Approximate Bayesian computation is one way to produce such generative inference [@kandler_generative_2018].
+In this framework, researchers run a simulation model under a wide range of parameter combinations, representing alternative hypotheses, to produce many simulated datasets for which all parameters and outcomes of interest are known.
+All simulated datasets are then systematically accepted or rejected as consistent with the observed data, and the combination of model parameters that provides the best fit is estimated probabilistically [@van_der_vaart_calibration_2015; @beaumont_approximate_2002].
+Hence, ABC allows researchers to quantitatively compare the likelihoods of competing hypothesized causal mechanisms in complex land systems.
 
-![Caption ](./PaperFigures/Prior.png)
+In this study we vary just one model parameter, the externally driven rate of coral rag vegetation conversion to rotational agriculture.
+We first specify a prior distribution that we believe will capture all possible values of this parameter of interest (fig.
+5).
+This prior is based on a combination of calibration with earlier models, and *a priori* understanding of the system from working with local conservation organizations and farmers.
+
+![Two thousand draws from the prior distribution of the rate of externally driven coral rag vegetation conversion to agriculture.
+These draws were used to produce the prior distribution of the expected number of coral rag vegetation pixels to be converted to agriculture from 2018 - 2021 for each *shehia*.
+](./PaperFigures/Prior.png)
+
+We then run our simulation model 2,000 times using draws from this prior distribution as the parameter of interest — the externally driven rate of coral rag forest conversion.
+For each run, the model then produces a synthetic dataset including the number of expected agricultural conversions for each *shehia* in our study.
+Each of these synthetic datasets is then accepted or rejected as consistent with the observed changes in land cover as measured through our 20m land cover classification map for 2021.
+We use an acceptance criteria of predicted agricultural conversions within 10% of the observed conversions.
+Finally, the parameter value for the extrinsic growth rate in each synthetic dataset that is accepted as consistent with the observed data is saved as one “draw” from the posterior for the estimated real-world parameter value.
 
@@
 
-![Caption ](./PaperFigures/EstimatesPixels.png)
+As described in the methods, for each *shehia* we ran 2,000 simulations, using each draw from the prior distribution of externally driven coral rag forest conversion for each simulation (figure 5).
+Each of these simulations converted some number of coral rag vegetation pixels to agriculture in each of the 19 *shehia* from 2018 to 2021 (fig.
+6).
+An average of 231 of these estimates per *shehia* were within 10% of the observed number of converted coral rag vegetation pixels [@sunnaker_approximate_2013].
+Keeping only these synthetic datasets consistent within the 10% error bound, we observe the distribution of externally driven growth parameter values that, based on our model, are likely to have produced the observed 2021 land cover (fig.
+7).
+The median parameter values consistent with the observed data (parameter estimates) for each *shehia* range from 0.0% to 3.9% for yearly coral rag forest cover loss due to external forces (fig.
+7).
+For four of the study *shehia*, the observed number of coral rag forest pixels converted to agriculture was fewer than predicted by the cycles of soil degradation built into our model alone.
+These four *shehia* all generally exhibited overall low rates of observed coral rag forest conversion to agriculture observed through the satellite imagery.
+We discuss contextual factors that may be influencing these trends in the discussion.
 
-![Caption ](./PaperFigures/Estimates.png)
+![Black histograms represent the expected number of coral rag vegetation pixels to be converted to agriculture for each draw from the prior for each *shehia* from 2018 to 2021.
+Red lines show the observed number of conversions for each *shehia* from 2018 to 2021.](./PaperFigures/EstimatesPixels.png)
+
+The width of the posterior parameter estimates shown in figure 7 are indicative of how much information about causal processes we can infer from the observed land cover change from 2018 to 2021.
+Wider estimates indicate a greater degree of equifinality, where a wide range of externally driven deforestation rates could have produced the observed data.
+Conversely, narrow parameter estimates indicate that only a small range of externally driven deforestation rates could have produced the observed number of coral rag vegetation pixels converted to agriculture from 2018 to 2021.
+Thus, when parameter estimates are more narrow, the data carry a stronger underlying causal signature.
+The width of the posterior parameter estimates can then be thought of as the range of processes that could have reasonably produced the observed data [@kandler_generative_2018].
+Across *shehia*, we observe relatively narrow parameter estimates, with an average range of 1.47%; the most narrow being 0.46% and the widest being 2.24% (fig.
+7).
+We can conclude then that based on our model, on average, less than a 1% increase or decrease in the externally driven deforestation rate from the median model estimate is likely is result in the observed land cover change in each *shehia*.
+
+![Black histograms show the prior distribution of externally driven deforestation rates for coral rag vegetation in Pemba.
+The prior values are identical for each *shehia*.
+Purple density plots show the density of draws from the prior in which the prior value resulted in a deforestation rate within 10% of the observed rate for that *shehia*.
+Shading in the density plots shows the tail probability that a particular value from the prior will result in the observed land cover based on our model.](./PaperFigures/Estimates.png)
 
@@
 
-![Caption ](./PaperFigures/Contribution.png)
+With the observed rates of coral rag vegetation conversion from 2018 to 2021 and estimates for the externally driven agricultural expansion (deforestation) rate for each *shehia* in hand, we can assess the proportion of total loss driven by soil degradation versus external influences such as market integration and population growth.
+We subtract the median estimate of the contribution of external forces from the total satellite-observed rates of conversion to yield the point estimates shown in figure 8.
+
+On average, we observe a 5.1% yearly rate of total coral rag forest conversion to agriculture from 2018 to 2021.
+Between study *shehia* this rate of total conversion ranges from 1.1% to 9.8%.
+We estimate that the percent of coral rag vegetation converted to agriculture in each *shehia* driven by soil degradation was between 1.0% and 7.6% per year, with an average rate of 4.0%.
+By comparison, a relatively small proportion of coral rag vegetation in each *shehia* is converted to agriculture each year as a result of external forces.
+We estimate that the average rate of externally driven conversion is 1.1% of total coral rag forest cover, ranging from 0.0% to 3.9% between *shehia*.
+
+While there is considerable variability in both the total rate of coral rag vegetation loss and in the contribution from external drivers between *shehia*, a general and intuitive trend is that *shehia* with a greater proportion of loss caused by external drivers show greater coral rag vegetation loss overall.
+The four *shehia* that showed a total number of observed coral rag pixels lost below that which is expected from soil degradation alone in figure 5, also show relatively little vegetation loss as a percentage of total coral rag cover from 2018 to 2021.
+
+![Barbell plot showing the median estimates of the influence of soil degradation on the observed coral rag vegetation loss in each *shehia*.
+Blue points show the total loss of coral rag vegetation as measured using satellite imagery.
+Black points show the median model estimates for the contribution of soil fertility loss to the observed deforestation.](./PaperFigures/Contribution.png)
 
@@
 
-## 5.2 - Implications for land system science and SES
+Across the *shehia* in our study, we observe a 5.1% yearly rate of coral rag vegetation conversion to agriculture.
+We show here that reported farmer behavior in response to soil degradation, paired with the topography of the study *shehia*, should account for an estimated 4.0% yearly rate of conversion on average.
+Thus, while variable, we can conclude that on average a relatively small proportion of the observed coral rag vegetation conversion to agriculture in the study *shehia* is driven by external forces such as increasing demand driven by increasing subsistence needs or market forces.
+This finding matches our theoretical expectations given that farmers in this region generally clear land in order to plant low value staple crops such as cassava [@meyfroidt_middle-range_2018].
+This suggests that regenerative agriculture programs, along with rainwater catchment systems may considerably reduce the long-term loss of coral rag forest in Pemba, Tanzania.
+These programs will also ease the strain that clearing forested land puts on farmers, and may potentially help farmers break free of the cycle of environmentally damaging agricultural practices in pursuit of short term gains.
+
+While the 1.1% average estimated yearly conversion rate of coral rag vegetation to agriculture in the study *shehia* driven by external forces is relatively small compared to the rate of loss driven by soil degradation, it is not negligible.
+Further, this value could reasonably increase as a result of continued market integration and population growth in Pemba.
+As the impact of external forces on land conversion increases, theory tells us that the effect of improved agricultural technologies on mitigating forest loss may be reduced or even reversed [@kaimowitz_economic_1998].
+Top-down interventions such as designating one or more coral rag forest reserves on the island may help to slow the conversion of primary forest in some areas, but may also be subject to leakage, attenuating their overall efficacy [@meyfroidt_focus_2020; @bastos_lima_leakage_2019].
+Instead, interventions focused on the introduction and establishment of value chains for alternative income sources, aside from rotational agriculture, may have greater success [@akyoo_spice_2007].
+
+## 5.2 - Implications for land system science and social-ecological systems
+Social-ecological systems, and therefore land systems, are inherently very complex.
+They commonly exhibit feedbacks between system components and past and future states.
+Because of this complexity, researchers generally describe phenomena of interest qualitatively, or they break the components of a given causal pathway down into many sub components [@meyfroidt_approaches_2016; @turner_framing_2020].
+Nevertheless, inference from limited time series data is difficult due to issues of equifinality, simultaneous causation, and unobserved complexity[@barrett_design-based_2021; @cumming_advancing_2020].
+We echo the argument of Schlüter and others [-@schluter_toward_2019] that agent/individual-based modeling can allow inference about causality in social-ecological systems and emphasize that this approach is especially powerful when combined with empirical data, as presented here, in order to increase validity and interpretability.
+
+We argue here that individual-based models will be of the greatest utility for land system science when they are combined with a standard, systematic framework for comparing synthetic data to real-world observations.
+Approximate Bayesian computation may fill this niche as it is relatively straightforward and allows for parameter estimation given even very complex generative models.
+A key advantage of this method is that it allows modelers to explicitly confront equifinality in a given simulation and to some extent a given empirical system.
+By explicitly quantifying the range of causal processes that are likely to produce an outcome of interest, researchers will greatly increase the applicability of simulation models to real-world policy decisions [@williams_assessing_2020].
 
@@
 
+A limitation of any process model is that they assume that researchers know and can accurately represent causal processes in silico.
+In the case of models like the individual-based simulation presented here, researchers must abstract down to only key phenomena of interest, eliminating much contextual nuance and again, assuming that we know what matters and what does not.
+This is a big assumption in social-ecological systems considering that emergent and often unexpected phenomena are a defining feature of the field.
+This limitation is exemplified by the four *shehia* (Fundo, Muambe, Jombwe, and Shamiani) that showed fewer coral rag forest pixels converted to agriculture than expected by the model even under no externally driven forest conversion pressure.
+We know through our in-person observations and interviews with Community Forests Pemba staff that these four *shehia* have been the focus of considerable tree planting, particularly of *Casuarina spp.* in woodlots.
+We are likely observing both some confusion between native coral rag vegetation and woodlot vegetation in our satellite observations, as well as a reduction in coral rag deforestation for fuelwood.
+Approximately 95% of Pemban households rely on cutting fuelwood for daily cooking activities.
+Hence, the introduction of woodlots, which we do not account for in our simulation, likely reduces the overall rates of coral rag deforestation and conversion to rotational agriculture.
+Further, Fundo and Shamiani are both islets of Pemba and are both home to luxury resorts.
+The limited connection between these islets and the main population centers of Pemba likely limits the effect of market forces driving agricultural expansion in these two areas.
+Also, our anecdotal experience is that the resort operators intentionally limit local development nearby the properties, possibly reducing the rate of agricultural development below what we expect as a result of soil degradation.
+
+Another contextual limitation of our specific model for drawing inference about agricultural expansion in Pemba Island is that we do not account for the long-term processes of soil degradation that lead to complete land abandonment and eventually the recovery of coral rag vegetation.
+While the two year farm and three year fallow cycle is standard in the coral rag geological areas of Pemba, some agricultural areas are completely abandoned when crop yields are consistently low even after a fallow period.
+Additionally, the agricultural units themselves are independent 20m pixels which is considerably smaller than many agricultural operations.
+Clumping these pixels to better match realistic farm sizes may produce different and more accurate inference than presented here.
+
+Lastly, our model does not allow for heterogeneity or evolution in human behavior.
+All agricultural pixels follow the same scheduling process.
+While this scheduling process is standard in the study *shehia*, an interesting exercise would be to allow for the diffusion of regenerative agricultural practices across farm units to examine direct feedbacks between environmental and cultural change.
+
 # 6 - Conclusion
+Computer simulations are critical to theoretical development in land system science as they allow us to formally define and scrutinize hypothesized mechanisms driving phenomena of interest.
+When we develop competing plausible mechanisms however, it can be difficult to identify the contribution of each hypothesized mechanism in the real world.
+Recent advances in ABC, primarily in population genetics, but also cultural evolution, have provided a structured process to begin to overcome this challenge in other fields [@hartig_statistical_2011; @kandler_generative_2018].
+Until now however, ABC has yet to be applied to land system science.
+In this paper we show how ABC can be used to better leverage the wealth of available satellite data in combination with individual-based models of land system change in order to assess the importance of competing mechanisms.
+
+In particular, we develop an individual-based simulation of agricultural expansion in Pemba, Tanzania under two different mechanisms: soil degradation and external forces such as population growth and increasing market integration.
+We use ABC to systematically compare runs from this model with observed land cover change in 19 *shehia* in Pemba from 2018 to 2021.
+This process allows us to estimate the likelihood that various rates of externally driven agricultural expansion are responsible for the observed land cover change in each *shehia*.
+Importantly, this process also allows us to directly estimate the range of externally driven expansion rates that could have also reasonably resulted in the observed data, or the degree to which the system is equifinal.
 
@@
 We also thank the Hazards and Climate Resilience Institute at Boise State University for funding this research.
-Lastly, we thank the Max Planck Institute for Evolutionary Anthropology, department of Human Behavior, Ecology and Culture for providing the computing resources necessary for this project.
+We thank Dr.
+Monique Borgerhoff Mulder and Dr.
+Tim Caro their help in conceptualizing the project and with all things related to fieldwork in Pemba.
+Lastly, we thank the Max Planck Institute for Evolutionary Anthropology, department of Human Behavior, Ecology and Culture for providing the computing resources necessary for this project and Dr.
+Anne Kandler for her support with the ABC and for providing an initial review of the paper.
 
 # 8 - Open data and software
+All data, R code, and GEE code used in this project are available at this github: https://github.com/matthewclark1223/CPR_ABM/tree/master/HCRI_Grant and this GEE link: https://code.earthengine.google.com/3774c9546268839fcbf43176e4d2eb46.
+An open source web app that displays the runs from the agent-based model for specific *shehia* can be found here: https://matthewclark.shinyapps.io/LandUsePredictionsApp/.
 
-# Supplemental 1
-
-ODD model description attached.
-
-# Supplemental 2
-
-# Supplemental 3
+# Supplemental 
+All supplemental materials can be found in the GoogleSheet here: https://docs.google.com/spreadsheets/d/18jGNxQl8uVK2m-jHzR6Pq4H-OmEXWMJFTSswoklx1-8/edit?usp=sharing.
 

### 2024-02-27 HCRI_Grant/ProjectFiles/BarbelChart.R

@@ -63,9 +63,10 @@ ggplot(., aes(y=Shehia, x=EndogenousRate, xend=ConvertedPercentYearly)) +
          axis.title.y = element_text(face = "italic"))
 
 
-mean(ObsData$ConvertedPercentYearly)
+range(ObsData$ConvertedPercentYearly)
 range(ObsData$EndogenousRate)
+mean(ObsData$EndogenousRate)
 mean(ObsData$ExtrinsicRates)
-
+range(ObsData$ExtrinsicRates)
 
 

### 2024-02-27 HCRI_Grant/ProjectFiles/ReproducibleAnalysis/ABM_ABC.R
Reprodicuble Analysis

@@ -1,0 +1,101 @@
+#Code to run the Approximate Bayesian Computation to figure out
+#What is the most likely "intrinsic growth" parameter for the 
+#expansion of agricultural land into coral rag forests in Pemba Island Tanzania
+
+
+#The goal here is to take some data where we know the outcome, in this case
+#the land cover from Pemba in 2019
+
+#And use a generative model of how that data might have been produced (the abm), 
+#To figure out what a likely parameter value in the model was that produced those data
+
+#We also need a prior on what the value(s) we are trying to estimate are
+
+#Lastly, we need a criterion for when the simulated data "matches" the actual data
+#We will likely not often get an exact match, so we need some other metric
+
+
+#First let's make the prior
+library(raster)
+library(sf)
+library(tidyverse)
+set.seed(1)
+
+
+prior<-rgamma(2000,5.5,0.8)  #Prior on the rate of intrinsic growth of ag land
+#hist(prior)
+
+#Shehia with a considerable amount of coral rag
+studyShehia<-c("Fundo","Kangagani","Kojani","Shumba Mjini","Maziwa Ng'ombe","Mjini Ole","Ole",
+               "Uwandani","Vitongoji","Kibokoni","Mvumoni","Pujini","Dodo","Chambani",
+               "Jombwe","Shamiani","Muambe","Kiwani","Mjini Wingwi")
+
+##### For Testing use these #############
+## Comment out if not testing ####
+priorTEST<-rgamma(5,5.5,0.8) 
+
+studyShehiaTEST<-c("Fundo","Kangagani")
+abc_data<-expand.grid(studyShehiaTEST,priorTEST)
+
+#################### End of testing ##################################
+
+
+#Create data for each run
+abc_data<-expand.grid(studyShehia,prior)
+names(abc_data)[1:2]<-c("Shehia","GrowthRate")
+
+#Create an empty column to fill with the estimated number of conversion events 
+abc_data$numberConversions<-rep(NA,nrow(abc_data))
+abc_data$observedConversions<-rep(NA,nrow(abc_data))
+abc_data$numberCorrectPixelConversions=rep(NA,length(prior))
+abc_data$percentCorrectPixelConversions=rep(NA,length(prior))
+
+#create the LU_abm function
+source("./Full_LandUse_ABM.R")
+
+
+###Add data for pixel/pixel validation
[…]

### 2024-02-27 HCRI_Grant/ProjectFiles/ReproducibleAnalysis/Full_LandUse_ABM.R

@@ -1,0 +1,326 @@
+library(raster)
+library(sf)
+library(tidyverse)
+
+
+
+LU_AMB<-function(
+  YearsPast2018 = 3, #years (timesteps) to run model
+  Wards = "Kojani",  #character vector or wards to model. Default is full model
+  FallowTime = 3, #time (in years) it takes for fallow land to recharge
+  AgLimit = 2, #Time (in years) land can be farmed in-between fallow periods
+  IntrinsicExp = 3 #Intrinsic rate (percentage) of agricultural expansion into the CR forest (pop growth, market integration, etc)
+){
+  
+  Years<-(1:YearsPast2018)+2018
+  
+  #specifics to this dataset
+  stackRS<-raster::stack("./PembaFiresAndPredictors.tif")
+  names(stackRS)<-c("roads_proximity","slope","soil_cat","fires2019")
+  #rProbBurn<-raster::raster("./ProjectFiles/PredFire2019.tif")
+  
+  #These will be used to assign working/fallow ag randomly
+  FallowRechargeTime<-FallowTime
+  AgLimit<-AgLimit
+  
+  propstrt<-FallowRechargeTime/(AgLimit+FallowRechargeTime) #probability that any ag pixel is fallow at the start
+  
+  IntrinsicExp=IntrinsicExp/100 # get this on a scale we can multiply by
+  
+  
+  #Import 2018 Lc Data
+  Pem2018LC<-raster::raster("./pemmyLC2018updateSAR.tif")
+  
+  rLndCvr2018<-Pem2018LC
+  
+  
+  rLndCvr2018[which(rLndCvr2018[]==2)]<-sample(c(100,2),size=length(which(rLndCvr2018[]==2)), #assign 100 to Fallow 
+                                               replace=T,prob=c(propstrt,1-propstrt)) #2 is productive ag
+  #0 = Mangrove
+  #1 = HF
+  #2 = agriculture
+  #3 = Urban
+  #4 = Bare
+  #5 = Coral rag
+  #6 = OWV/agroforestry
+  #7 = water
+  ### ONLY CORAL RAG BURNABLE
+  rLndCvr2018[]<-as.factor(ifelse(rLndCvr2018[] %in% c(0,1,3,4,6,7),"Unburnable", #everything besides ag, fallow, and CR
+                                  ifelse(rLndCvr2018[] == 2,"Agriculture", #productive ag
[…]

### 2024-02-27 ParticipatoryMapping/MakeGriddedBasemaps.R
Update MakeGriddedBasemaps.R

@@ -2,7 +2,7 @@ library(ggmap)
 library(sf)
 library(tidyverse)
 library(ggspatial)
-register_google(key="AIzaSyCh-KGA6T9qaPF-wjfnBV8hb7zzGYMlGTw")
+register_google(key="REMOVED")
 
 Pemba <- read_sf("~/Pemba_Project/PembaShapeFile.shp")
 

### 2024-02-27 HCRI_Grant/ProjectFiles/ReproducibleAnalysis/Full_LandUse_ABM.R
updated with smaller roads layer

@@ -16,7 +16,7 @@ LU_AMB<-function(
   
   #specifics to this dataset
   stackRS<-raster::stack("./PembaFiresAndPredictors.tif")
-  names(stackRS)<-c("roads_proximity","slope","soil_cat","fires2019")
+  names(stackRS)<-c("roads_proximity","soil_cat")
   #rProbBurn<-raster::raster("./ProjectFiles/PredFire2019.tif")
   
   #These will be used to assign working/fallow ag randomly
